### 2010-11-27 Text/paper.tex
Started gathering information for tests.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Recently we investigated the applicability and parameters choice for the two-dimensional case
Bretherton problem. This paper is focused on the three-dimensional case to validate the continuous
interface methods to simulate phenomena happening in the Bretherton/Taylor flow for rectangular
and square capillaries. 
\end{abstract}
\section{Introduction}
While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-planar},
vast number of experimental results and simulations are available for the three-dimensional case.
It was found that the for rectangular and/or square shaped capillaries there is a transition for
certain capillary number, where the flow changes from the non axi symmetric to axis symmetric case.
The number is indicated in a number of works of as $Ca=0.04$ \cite{cerro-bubble-train}, $Ca=0.1$
\cite{cerro-space}, $Ca=0.033$ and for $Ca>4$ \cite{heil-threedim} measured at 11 half-widths behind
the finger tip.

\section{Variation bubble length}
The work of \cite{heil-threedim} shows the variation of the bubble thickness through the length of
bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten}
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}
We want to obtain the same characteristics as indicated in Fig.
\ref{fig:thickness:variation:ca:ten}. One can see that the difference between diagonal radius and
usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
simulation is conducted by the grid $50\mathrm{x}50$ in the cross section that means that we will
see the difference in one grid node. Therefore the grid is $52\mathrm{x}52\mathrm{750}$. Here we
conduct a scaling:
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}
\end{equation}
The parameters for the binary liquid problem are chosen as $A=0.04$ and $k=0.01$. Therefore the
surface tension is:
\begin{equation}
\sigma = \sqrt{\frac{8 k A}{9}}=0.0596
\end{equation}
The relaxation times for liquid and gas are taken as, $\tau_{gas}=0.7$ and $\tau_{liq}=2.5$,
therefore the viscosities ratio is $10$.
\begin{equation}
\begin{aligned}
&10=\frac{\frac{2}{3} U_{bubble}}{0.0596},
&U_{bubble}=0.894
\end{aligned}
\end{equation}
which is quite high for the lattice Boltzmann method. We can decrease the ratio by scaling
parameters $k$ and $A$ while keeping the interface thickness. Let us choose parameters as $k=0.001$
and $A=0.004$, therefore the mobility should be adjusted to address the issue with small parameters
and small surface tension in this case. It parameters are chosen as $A=0.004$ and $k=0.001$ then
the velocity of the bubble is  $0.0894$. Therefore we can adjust the pressure gradient to plug it
into the bubble initialization:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca=0.0001907,
\end{equation}
which is quite high for lattice Boltzmann simulations. Therefore probably one more time the surface
tension needs to be reduced. {\color{red} Many authors state that increasing mobility kills the
proper results of simulations}. 
The film thickness as it is seen from the picture is taken as $70$ percent of the half channel.
Therefore the initalization width is $0.15*50=7.5$.

\section{Non-axisymmetric case}
That could be fucking difficult. For example according to \cite{heil-threed} for the capillary
number $Ca=0.01$, the radiuses converge as $r_h=0.99$ and $r_d=1.1$. Therefore one needs to resolve
as the $0.5\%$ of whole channel width to see the difference. That certainly implies so large grid
that can kill the simulation at least $200x200x3000$. However, we can refer to artificially made
interface as $1-2$ lattice Boltzmann units but keeping the capillary number as it is required. In
this case we can refer like we can't resolve the interface but we can certainly to attribute it to
the wall gradient as extreme case. The dimnesionalization case is therefore as follows:
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
\end{equation}
therefore let us pick up the same parameters as for the previous case and just change use the
pressure gradient reduced by $1000$ as $Ca=10$ and $Ca=0.01$ are different. 
Let us wait for simulation.

\section{Capillary number region}

\section{Simulation setup}

\bibliographystyle{unsrtnat}
\bibliography{paper}
\end{document}

### 2010-11-27 Code/Python/read.py

@@ -41,5 +41,36 @@ def read(name):
     mlab.pipeline.vector_cut_plane(v, mask_points=2, scale_factor=3)
     mlab.show()
 
+def show(name,slice):
+    from enthought.tvtk.api import tvtk
+    import pylab
+    import numpy
+
+
+    gridreader = tvtk.XMLStructuredGridReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    points=grid.points
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("Phase"))
+    velocity=numpy.array(data.get_array("Velocity"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
+    phase_numpy=phase.reshape(dims)
+    velx_numpy =velx.reshape(dims)
+    vely_numpy =vely.reshape(dims)
+    velz_numpy =velz.reshape(dims)
+    
+    pylab.imshow(phase_numpy[slice,:,:])
+    pylab.show()
+
 if __name__=="__main__":
-    read("../phase00000.vts")
+    name="../../Temp/equili.vts"
+    read(name)
+    #show(name,300)

### 2010-11-27 Text/paper.tex
Minor changes with Latex file.

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{bm}
@@
 While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
-only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-planar},
+only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-bretherton},
 vast number of experimental results and simulations are available for the three-dimensional case.
@@
 \section{Non-axisymmetric case}
-That could be fucking difficult. For example according to \cite{heil-threed} for the capillary
+That could be fucking difficult. For example according to \cite{heil-threedim} for the capillary
 number $Ca=0.01$, the radiuses converge as $r_h=0.99$ and $r_d=1.1$. Therefore one needs to resolve

### 2010-12-09 Text/paper.tex
Adding literature review. Need to add the patch of Michal.

@@
 Recently we investigated the applicability and parameters choice for the two-dimensional case
-Bretherton problem. This paper is focused on the three-dimensional case to validate the continuous
-interface methods to simulate phenomena happening in the Bretherton/Taylor flow for rectangular
-and square capillaries. 
+Bretherton problem using the lattice Boltzmann method (LBM) {\color{red} Give citation here}. This
+paper is the continuation of the previous work and is focused on the three-dimensional case to
+validate the LBM for simulation of phenomena for rectangular and square capillaries in the
+moderate range of Capillary number parameters. 
 \end{abstract}
+\section{What to examine}
+I suggest to examine first of all the application of the LBM for really small capillary number,
+where it's not possible to resolve the interface. In this case one should find the asymmetric
+radiuses for the droplet, the difference should be around $10$ percent. While applying the forcing
+it should come to the spheric shape. As well we need to examine the capillary number dependancy on
+the length and suggest to examine some axicentric factors for rectangular channels.
+
 \section{Introduction}
@@
 It was found that the for rectangular and/or square shaped capillaries there is a transition for
-certain capillary number, where the flow changes from the non axi symmetric to axis symmetric case.
-The number is indicated in a number of works of as $Ca=0.04$ \cite{cerro-bubble-train}, $Ca=0.1$
-\cite{cerro-space}, $Ca=0.033$ and for $Ca>4$ \cite{heil-threedim} measured at 11 half-widths behind
-the finger tip.
+certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
+The number is indicated in a number of works of as $\hat{Ca}=0.04$ \cite{cerro-bubble-train},
+$\hat{Ca}=0.1$
+\cite{cerro-space}, $\hat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger than
+the critical capillary number, i.e. $Ca>\hat{Ca}$, then the flow becomes axisimmetric with the
+radius of the droplet dependant on the capillary number.
+
+The following pictures summarize the radius of deposited liquid on the capillary length.
+\citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
+square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
+square capillaries are shown in Fig.\ref{fig:heil:three:dim}. On another hand authors performed
+numerical simulations for rectangular shaped capillaries. It was found that aspect ratio $\alpha$
+allows to put all the different curves on one purely empirical curve through the following parameter
+$s_{\infty}=r_{\infty} \alpha^{-1/2}$, where $r_{\infty}$ is the aximsymmetric bubble radius
+infinitely far behind the bubble tip and can be calculated from the flow rate in the computational
+domain:
+\begin{equation}
+r_{\infty}=2\sqrt{(a-Q)/\pi}.
+\end{equation}
+It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
+axisymmetric case. The $s_{\infty}$ dependency on $\alpha$ is shown in Fig.
+\ref{fig:heil:three:dim}.
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/capillary_width_heil.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/capillary_rectangular_width_heil.eps}\\
+\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
+capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
+number $Ca\leq\hat{Ca}=0.04$.\label{fig:heil:three:dim}}
+\end{figure}
+There is also interesting thing to track down is the recirculating regime for flow fields. It is
+known that for certain capillary number the streamlines either go around the bubble or create the
+circulating regime in front of the bubble. The calculation ({\color{red} or analytical data? Look
+at White 1991, pp. 119-120. As well put figures if there is a space from Heil et al.}) are presented
+in Table \ref{table:recirculation:data}.
+\begin{table}
+\begin{tabular}{c|c}
+$\alpha$& $Ca_{T}$\\
+\hline
+1& 0.691\\
+1.1 & 0.688\\
+1.3 & 0.666\\
+1.5 & 0.631\\
+\end{tabular}
+\caption{The results for the recirculation region.\label{table:recirculation:data}}
+\end{table}
+
+\citet{shikazono-square} obtained a lot of different results for the deposition length versus the
+capillary number variation for FC-40/air, ethanol/air and water/air experiments. All of them are
+presented in Fig. \ref{fig:deposition:han}.
+\begin{figure}
+\includegraphics[width=0.33\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
+\includegraphics[width=0.33\textwidth]{Figures/han_capillary_water_air.eps}\hfill
+\includegraphics[width=0.33\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
+\caption{Han results. \label{fig:deposition:han}}
+\end{figure}
 

### 2010-12-09 Sailfish/bmc_3d.py

@@ -47,7 +47,8 @@ class FESim(lbm.BinaryFluidFreeEnergy):
 
         lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)
 
-        self.add_body_force((0.0, 0.0, self.options.force), grid=0, accel=False)
+        self.add_body_force((0.0,0.0,0.0), grid=0, accel=False)
+        #self.add_body_force((0.0, 0.0, self.options.force), grid=0, accel=False)
 
         # Use the fluid velocity in the relaxation of the order parameter field,
         # and the molecular velocity in the relaxation of the density field.

### 2010-12-10 SailfishPatch/bmc_3d.py
Added comparison results, but haven't done it till the end.

@@ -1,0 +1,60 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo3D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width),
+                                  np.logical_or(hz < self.wall_width, hz >= self.lat_nz-self.wall_width))
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(np.logical_and(
+                       np.logical_and(hz >= width-1 + self.wall_width, hz <= self.lat_nz-width-self.wall_width),
+                       np.logical_and(hy >= width-1 + self.wall_width, hy <= self.lat_ny-width-self.wall_width)),
+                       np.logical_and(hx >= self.lat_nx/3.0, hx<=2*self.lat_nx/3.0))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_3d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 750,'grid': 'D3Q19',
+                    'lat_ny': 52, 'lat_nz' : 52 ,
+                    'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True,'periodic_z': True}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=6.0e-6))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
+
[…]

### 2010-12-10 SailfishPatch/compare.py

@@ -1,0 +1,119 @@
+#!/usr/bin/python
+def read_my(name):
+    from enthought.tvtk.api import tvtk
+    from enthought.mayavi import mlab
+    import numpy
+
+
+    gridreader = tvtk.XMLStructuredGridReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    points=grid.points
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("Phase"))
+    velocity=numpy.array(data.get_array("Velocity"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
+    phase_numpy=phase.reshape(dims)
+    velx_numpy =velx.reshape(dims)
+    vely_numpy =vely.reshape(dims)
+    velz_numpy =velz.reshape(dims)
+    
+    fig=mlab.figure()
+    src = mlab.pipeline.scalar_field(phase_numpy)
+    v= mlab.pipeline.vector_field(velx_numpy,vely_numpy,velz_numpy)
+    vx=mlab.pipeline.scalar_field(velx_numpy)
+    vy=mlab.pipeline.scalar_field(vely_numpy)
+    vz=mlab.pipeline.scalar_field(velz_numpy)
+    
+    extract=mlab.pipeline.extract_grid(src)
+    extract.set(z_min=1,z_max=dims[2]-2,y_min=1,y_max=dims[1]-2)
+    surf = mlab.pipeline.contour_surface(extract)
+ 
+    mlab.pipeline.image_plane_widget(vx, plane_orientation='x_axes', slice_index=250)
+    #mlab.pipeline.vectors(v, mask_points=20, scale_factor=3.)
+    mlab.pipeline.vector_cut_plane(v, mask_points=2, scale_factor=3)
+    mlab.show()
+
+def read_michal(name):
+    from enthought.tvtk.api import tvtk
+    from enthought.mayavi import mlab
+    import numpy
+
+
+    gridreader = tvtk.XMLImageDataReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    #points=grid.points
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
[…]

### 2010-12-10 Text/paper.tex
Modified code and compare.py - the machine accurate results. Really strange.

@@
 
+\citet{cerro-bubble-train} performed a lot of experiments for a buble-train flow in capillaries of
+circular and square cross section for horizontal, upward and downward flows. The main results are
+presented in Fig. \ref{fig:thulasidas:results}.
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/thulasidas_capillary.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/thulasidas_relative_velocity.eps}\\
+\caption{\citet{cerro-bubble-train} results for the film thickness
+variation and the relative bubble velocity versus the capillary
+number.\label{fig:thulasidas:results}}
+\end{figure}
+
+\citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
+shaped capillaries as square and triangle shaped capillaries. Some interesting results for bubble
+shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
+numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
+\ref{fig:wang:critical:capillary}. 
+\begin{figure}
+\includegraphics[width=\textwidth]{Figures/wang_bubble_shape.eps}
+\caption{The bubble shape across the length for different capillary numbers.
+\label{fig:wong:bubble:shape}}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
+\caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
+velocity for th  \label{fig:wang:critical:capillary}}
+\end{figure}
+
 \section{Variation bubble length}

### 2010-12-10 Code/dynamics_BGK.cpp

@@ -177,14 +177,15 @@ void DynamicsBGK::collide_stream(int iX,int iY,int iZ)
 
 	//Omega calculation
 	double tau_temp=tau_gas+(phase_temp+1.0)/2.0*(tau_liq-tau_gas);
-	double omega_temp=1.0/tau_temp;
-
+	double omega_rho=1.0/tau_temp;
+	double omega_phi=1.0/tau_phi;
+	
     //Force calculation
     double sum_force=0.0;
     double force_pop[NPOP];
     for (int k=1;k<NPOP;k++)
     {
-        force_pop[k]=(1.0-0.5*omega_temp)*weights[k]*(force_x*((cx[k]-ux_temp)+3.0*cx[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
+        force_pop[k]=(1.0-0.5*omega_rho)*weights[k]*(force_x*((cx[k]-ux_temp)+3.0*cx[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
                 force_y*((cy[k]-uy_temp)+3.0*cy[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
                 force_z*((cz[k]-uz_temp)+3.0*cz[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp)));
 
@@ -196,8 +197,8 @@ void DynamicsBGK::collide_stream(int iX,int iY,int iZ)
 	for (int k=0; k<NPOP; k++)
 	{
 		//cout<<"Increment="<<-omega*(lattice->f[counter*NPOP+k]-feq[k])<<"\n";
-		lattice->f[counter*NPOP+k]+=-omega_temp*(lattice->f[counter*NPOP+k]-feq[k])+force_pop[k];
-		lattice->g[counter*NPOP+k]+=-omega_temp*(lattice->g[counter*NPOP+k]-geq[k])+force_pop[k];
+		lattice->f[counter*NPOP+k]+=-omega_rho*(lattice->f[counter*NPOP+k]-feq[k])+force_pop[k];
+		lattice->g[counter*NPOP+k]+=-omega_phi*(lattice->g[counter*NPOP+k]-geq[k]);
 	}
 
 	//Streaming procedure (separate loop for normal)
[…]

### 2010-12-10 Code/dynamics_BGK.h

@@ -22,6 +22,7 @@ class DynamicsBGK:public Dynamics
     double aconst;
     double kconst;
     double gammaconst;
+    double tau_phi;
     double tau_liq;
     double tau_gas;
     double force_x;

### 2010-12-10 Code/dynamics_special.cpp

@@ -373,14 +373,15 @@ void DynamicsSpecial::collide_stream(int iX,int iY, int iZ)
     geq[0]=phase_temp-sum_phase;
 
 	double tau_temp=tau_gas+(phase_temp+1.0)/2.0*(tau_liq-tau_gas);
-    double omega_temp=1.0/tau_temp;
+    double omega_rho=1.0/tau_temp;
+    double omega_phi=1.0/tau_phi;
 
     //Force calculation
     double sum_force=0.0;
     double force_pop[NPOP];
     for (int k=1;k<NPOP;k++)
     {
-        force_pop[k]=(1.0-0.5*omega_temp)*weights[k]*(force_x*((cx[k]-ux_temp)+3.0*cx[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
+        force_pop[k]=(1.0-0.5*omega_rho)*weights[k]*(force_x*((cx[k]-ux_temp)+3.0*cx[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
  					force_y*((cy[k]-uy_temp)+3.0*cy[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp))+
  					force_z*((cz[k]-uz_temp)+3.0*cz[k]*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp)));
 
@@ -392,8 +393,8 @@ void DynamicsSpecial::collide_stream(int iX,int iY, int iZ)
 	for (int k=0; k<NPOP; k++)
 	{
 		//cout<<"Increment="<<-omega*(lattice->f[counter*NPOP+k]-feq[k])<<"\n";
-		lattice->f[counter*NPOP+k]+=-omega_temp*(lattice->f[counter*NPOP+k]-feq[k])+force_pop[k];
-		lattice->g[counter*NPOP+k]+=-omega_temp*(lattice->g[counter*NPOP+k]-geq[k])+force_pop[k];
+		lattice->f[counter*NPOP+k]+=-omega_rho*(lattice->f[counter*NPOP+k]-feq[k])+force_pop[k];
+		lattice->g[counter*NPOP+k]+=-omega_phi*(lattice->g[counter*NPOP+k]-geq[k]);
     }
 
 	//Streaming procedure

### 2010-12-10 Code/main.cpp

@@ -16,18 +16,18 @@
 #include "params_list.h"
 
 //Lattice Boltzmann initialization and parameters
-const int NX=30;
-const int NY=30;
-const int NZ=451;
+const int NX=52;
+const int NY=52;
+const int NZ=750;
 const int NPOP=19;
 const int NUM=NX*NY*NZ;
 const int NUMTOTAL=NUM*NPOP;
-const int NUMTIME=10000;
-const int NUMOUTPUT=1000;
+const int NUMTIME=1001;
+const int NUMOUTPUT=100;
 const int NUMSIGNAL=20;
 
 //Binary-liquid initialization
-const int width=4;
+const int width=6;
 const int radius=6;
 const double rhol=1.0;
 const double rhog=1.0;
@@ -87,17 +87,19 @@ int main(int argc,char* argv[])
     //Specify parameters
  	ParamsList params;
 	params.add("aconst", 0.04);
-	params.add("kconst", 0.01);
+	params.add("kconst", 0.04);
 	params.add("gammaconst",1.0);
 	params.add("omega",1.0);
 	params.add("phase_gradient",0.0);
 	params.add("phase_wall",0.0);
 	params.add("rho_wall",0.5);
-    params.add("tau_liq",2.7);
+	params.add("tau_phi",2.0);
+    params.add("tau_liq",2.5);
     params.add("tau_gas",0.7);
     params.add("force_x",0.0);
     params.add("force_y",0.0);
-    params.add("force_z",0.0001);
+    params.add("force_z",0.000015);
+	//params.add("force_z",0.0);
 
     //Useless parameters to be deleted after
     params.add("rho_press_inlet",1.0);
@@ -118,7 +120,9 @@ int main(int argc,char* argv[])
     Solver solver(geometry,params);
 
     //Solver initialization from the file
-    solver.load_file("equili");
+    //solver.load_file("equili");
+
+	macro_init(solver);
 
     //Initialization of the populations
     solver.init();

### 2010-12-10 Code/solver.cpp

@@ -1,10 +1,7 @@
 #include <vtkXMLStructuredGridWriter.h>
 #include <vtkXMLStructuredGridReader.h>
-#include <vtkXMLImageDataWriter.h>
-#include <vtkXMLImageDataReader.h>
 #include <vtkSmartPointer.h>
 #include <vtkStructuredGrid.h>
-#include <vtkImageData.h>
 #include <vtkIdList.h>
 #include <vtkDoubleArray.h>
 #include <vtkPointData.h>
@@ -639,48 +636,6 @@ void Solver::writeWholeDensityPhaseVelocity(std::string name)
     }
 
 }
-
-void Solver::writeWholeImageDensityPhaseVelocity(std::string name)
-{
-
-    Solver::getWholeDensity();
-    Solver::getWholePhase();
-    Solver::getWholeVelocity();
-    if (rank==0)
-    {
-        name=name+".vti";
-
-        vtkSmartPointer<vtkImageData> imageData = vtkSmartPointer<vtkImageData>::New();
-        imageData->SetDimensions(NX,NY,NZ);
-        imageData->SetOrigin(0,0,0);
-        imageData->SetSpacing(1,1,1);
-        imageData->SetNumberOfScalarComponents(5);
-        imageData->SetScalarTypeToDouble();
-
-
-        for(vtkIdType i = 0; i < NX*NY*NZ; i++)
-        {
-            int iZ=i/(NX*NY);
-            int iY=(i%(NX*NY))/NX;
-            int iX=(i%(NX*NY))%NX;
-
-            double* pixel = static_cast<double*>(imageData->GetScalarPointer(iX,iY,iZ));
-            pixel[0] = density[i];
-            pixel[1] = phase[i];
-            pixel[2] = velocity[3*i];
-            pixel[3] = velocity[3*i+1];
-            pixel[4] = velocity[3*i+2];
-        }
-
-        vtkSmartPointer<vtkXMLImageDataWriter> writer = vtkSmartPointer<vtkXMLImageDataWriter>::New();
-        writer->SetFileName(name.c_str());
-        writer->SetInput(imageData);
-        writer->Write();
-    }
-
-}
-
-
 
 void Solver::writeTextXZVelocity(std::string name)
 {

### 2010-12-10 Code/solver.h

@@ -43,7 +43,6 @@ class Solver
     void writeWholePhase(std::string name);
     void writeWholeVelocity(std::string name);
     void writeWholeDensityPhaseVelocity(std::string name);
-    void writeWholeImageDensityPhaseVelocity(std::string name);
 
     void writeTextXZVelocity(std::string name);
     void writeTextXYVelocity(std::string name);

### 2010-12-10 SailfishPatch/bmc_3d.py

@@ -36,13 +36,13 @@ class FESim(lbm.BinaryFluidFreeEnergy):
                     'lat_ny': 52, 'lat_nz' : 52 ,
                     'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
                     'scr_scale': 1,
-                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 2.0,
                     'bc_wall': 'halfbb',
                     'periodic_x': True, 'periodic_y': True,'periodic_z': True}
         settings.update(defaults)
 
         opts = []
-        opts.append(optparse.make_option('--force', dest='force', type='float', default=6.0e-6))
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=0.000015))
         opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
 
         lbm.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)

### 2010-12-10 SailfishPatch/compare.py

@@ -113,7 +113,62 @@ def show(name,slice):
     pylab.imshow(phase_numpy[slice,:,:])
     pylab.show()
 
+def compare(name_michal,name_alex,slice):
+    from enthought.tvtk.api import tvtk
+    import pylab
+    import numpy
+
+    gridreader_alex = tvtk.XMLStructuredGridReader()
+    gridreader_alex.file_name = name_alex
+    gridreader_alex.update()
+
+    gridreader_michal = tvtk.XMLImageDataReader()
+    gridreader_michal.file_name = name_michal
+    gridreader_michal.update()
+
+    grid_alex  = gridreader_alex.output
+    data_alex  = grid_alex.point_data
+    dims_alex  = grid_alex.dimensions
+    dims_alex  = dims_alex.tolist()
+    dims_alex.reverse()
+    phase_alex = numpy.array(data_alex.get_array("Phase"))
+    phase_alex = phase_alex.reshape(dims_alex)
+
+    grid_michal  = gridreader_michal.output
+    data_michal  = grid_michal.point_data
+    dims_michal  = grid_michal.dimensions
+    dims_michal  = dims_michal.tolist()
+    dims_michal.reverse()
+    phase_michal = numpy.array(data_michal.get_array("phi"))
+    phase_michal = phase_michal.reshape(dims_michal)
+    
+    print phase_alex.shape
+    print phase_michal.shape
+    
+    surf_alex = phase_alex[slice,:,:]
+    surf_michal=phase_michal[:,:,slice]
+    pylab.figure()
+    pylab.imshow(surf_alex)
+    pylab.figure()
+    pylab.imshow(surf_michal)
+    
+    pylab.figure()
+
+    prof_michal=surf_michal[:,surf_michal.shape[1]/2]
+    prof_alex=surf_michal[:,surf_alex.shape[1]/2]
+    pylab.plot(prof_michal[1:-1])
+    pylab.plot(prof_alex[1:-1])
+    print numpy.max(numpy.abs(prof_michal-prof_alex))
+    print prof_michal-prof_alex
+    pylab.show()
+
+
+ 
 if __name__=="__main__":
-    name="test1000_0.vti"
-    read_michal(name)
+    name_michal="test1000_0.vti"
+    name_alex="phase01000.vts"
+    #read_michal(name_michal)
+    read_my(name_alex)
+    slice=350
+    #compare(name_michal,name_alex,slice)
     #show(name,300)

### 2010-12-11 Text/paper.tex
Minor revisions. Started obtaining asymmetric cap results.

@@
 numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
-\ref{fig:wang:critical:capillary}. 
+\ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
+different capillary numbers. \ref{fig:wang:critical:capillary}.
 \begin{figure}
@@
 \caption{The bubble shape across the length for different capillary numbers.
-\label{fig:wong:bubble:shape}}
+\label{fig:wang:bubble:shape}}
 \end{figure}
@@
 \includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/wang_relative_velocity.eps}
 \caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
-velocity for th  \label{fig:wang:critical:capillary}}
+velocity depending on the capillary number  \label{fig:wang:critical:capillary}}
 \end{figure}

### 2010-12-14 Text/paper.tex
Added another patch from Michal. Started obtaining Asymmetric cap region. Pushing the continuous interface to the limit. The Poiseuille profile just doesn't work in this case. Let us see results.

@@
 \end{equation}
-therefore let us pick up the same parameters as for the previous case and just change use the
-pressure gradient reduced by $1000$ as $Ca=10$ and $Ca=0.01$ are different. 
-Let us wait for simulation.
+therefore let us pick up the same parameters as for the previous case. To perform simulations we
+started with the following parameters as
+$k=0.04$,$A=0.04$,$\Gamma=1.0$,$N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$.The
+initialized density was chosen as $w=2.5$ lattice Boltzmann units. For underresolved schemes, the
+pressure gradient doesn't work in the way we expect. Therefore, we take pressure gradient in the
+following way and later on establish the associated capillary number. In this case we can check the
+extreme values for the pressure gradients and the continuous interface methods:
+\begin{equation}
+\frac{\mathrm{d}P}{\mathrm{d}x}=
+\begin{cases}
+1\dots10\mathrm{x}10^{-6}\\
+10,15\dots 100 \mathrm{x}10^{-6}
+\end{cases}
+\end{equation}
 
 \section{Capillary number region}
-
+The same as in the previous example can help us to identify the necessary capillary regime. The
+most important thing is to establish different accentricity factors.
 \section{Simulation setup}

### 2010-12-14 Sailfish/Asymmetric/asymmetric.py

@@ -1,0 +1,61 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo3D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width),
+                                  np.logical_or(hz < self.wall_width, hz >= self.lat_nz-self.wall_width))
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(np.logical_and(
+                       np.logical_and(hz >= width-1 + self.wall_width, hz <= self.lat_nz-width-self.wall_width),
+                       np.logical_and(hy >= width-1 + self.wall_width, hy <= self.lat_ny-width-self.wall_width)),
+                       np.logical_and(hx >= self.lat_nx/3.0, hx<=2*self.lat_nx/3.0))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_3d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 750,'grid': 'D3Q19',
+                    'lat_ny': 52, 'lat_nz' : 52 ,
+                    'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 2.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True,'periodic_z': True,
+                    'bc_wall_grad_order': 1}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=1.0e-6))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=2))
[…]

### 2010-12-14 Sailfish/Asymmetric/govern.py

@@ -1,0 +1,225 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+import math
+ 
+def Run_Simulations():
+    print os.getcwd()
+    
+    force_init=1e-6;
+    #os.mkdir("temp")
+    for i in range(1,11):
+        dir_temp=str(i)
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','asymmetric.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        force=force_init*i
+        subprocess.call(['./asymmetric.py','--force='+str(force),'--batch','--every=50000','--max_iters=200001','--output=asym','--output_format=vtk'])
+        os.chdir("..")
+    for i in range(15,105,5):
+        dir_temp=str(i)
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','asymmetric.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        force=force_init*i
+        subprocess.call(['./asymmetric.py','--force='+str(force),'--batch','--every=50000','--max_iters=200001','--output=asym','--output_format=vtk'])
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return (zero-0.5)/(len(prof)-2)
+
+def Analyze_Simulations():
+    from numpy import genfromtxt
+    print os.getcwd()
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
+    capillary_str=["3","5","8","10","20","40","60","80"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
+
+    exam=[2000,2200,2600,2500,1100,950,600,350]
+    good=[0,1,2,3,4,5,6,7]
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
[…]

### 2010-12-14 SailfishPatch/compare.py

@@ -83,6 +83,14 @@ def read_michal(name):
     #mlab.pipeline.vector_cut_plane(v, mask_points=2, scale_factor=3)
     mlab.show()
 
+def get_zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero-0.5
 
 def show(name,slice):
     from enthought.tvtk.api import tvtk
@@ -90,18 +98,17 @@ def show(name,slice):
     import numpy
 
 
-    gridreader = tvtk.XMLStructuredGridReader()
+    gridreader = tvtk.XMLImageDataReader()
     gridreader.file_name = name
     gridreader.update()
 
     grid  = gridreader.output
     data  = grid.point_data
-    points=grid.points
     dims  =grid.dimensions
     dims=dims.tolist()
     dims.reverse()
-    phase= numpy.array(data.get_array("Phase"))
-    velocity=numpy.array(data.get_array("Velocity"))
+    phase= numpy.array(data.get_array("phi"))
+    velocity=numpy.array(data.get_array("v"))
     velx=velocity[:,0]
     vely=velocity[:,1]
     velz=velocity[:,2]
@@ -109,8 +116,16 @@ def show(name,slice):
     velx_numpy =velx.reshape(dims)
     vely_numpy =vely.reshape(dims)
     velz_numpy =velz.reshape(dims)
+    arr=phase_numpy[:,:,slice]
     
-    pylab.imshow(phase_numpy[slice,:,:])
+    pylab.figure()
+    pylab.imshow(arr)
+    pylab.figure()
+    pylab.plot(numpy.diag(arr))
+    print get_zero(numpy.diag(arr))
+    print velz_numpy[27,27,slice]
+    print velx_numpy[27,27,slice]
+    print vely_numpy[27,27,slice]
     pylab.show()
 
 def compare(name_michal,name_alex,slice):
@@ -155,7 +170,7 @@ def compare(name_michal,name_alex,slice):
     pylab.figure()
 
     prof_michal=surf_michal[:,surf_michal.shape[1]/2]
-    prof_alex=surf_michal[:,surf_alex.shape[1]/2]
[…]

### 2010-12-16 Sailfish/Asymmetric/govern.py
Started analyzing 3D simulations. Something wrong with the phase gradient. Need to finish radiussed versus capillaries.

@@ -4,20 +4,23 @@ import subprocess
 import pylab
 import numpy
 import math
+from enthought.tvtk.api import tvtk
+from enthought.mayavi import mlab
+
  
 def Run_Simulations():
     print os.getcwd()
     
     force_init=1e-6;
     #os.mkdir("temp")
-    for i in range(1,11):
-        dir_temp=str(i)
-        os.mkdir(dir_temp)
-        subprocess.call(['cp','asymmetric.py',dir_temp+"/"])
-        os.chdir(dir_temp)
-        force=force_init*i
-        subprocess.call(['./asymmetric.py','--force='+str(force),'--batch','--every=50000','--max_iters=200001','--output=asym','--output_format=vtk'])
-        os.chdir("..")
+    #for i in range(1,11):
+    #    dir_temp=str(i)
+    #    os.mkdir(dir_temp)
+    #    subprocess.call(['cp','asymmetric.py',dir_temp+"/"])
+    #    os.chdir(dir_temp)
+    #    force=force_init*i
+    #    subprocess.call(['./asymmetric.py','--force='+str(force),'--batch','--every=50000','--max_iters=200001','--output=asym','--output_format=vtk'])
+    #    os.chdir("..")
     for i in range(15,105,5):
         dir_temp=str(i)
         os.mkdir(dir_temp)
@@ -34,7 +37,108 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    return (zero-0.5)/(len(prof)-2)
+    return zero-0.5
+
+def Read_Phase(name):
+    gridreader = tvtk.XMLImageDataReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("phi"))
+    velocity=numpy.array(data.get_array("v"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
[…]

### 2010-12-17 Text/paper.tex
Added a few simulations picture. Need to resolve the interface

@@
 \section{Non-axisymmetric case}
-That could be fucking difficult. For example according to \cite{heil-threedim} for the capillary
+For example according to \cite{heil-threedim} for the capillary
 number $Ca=0.01$, the radiuses converge as $r_h=0.99$ and $r_d=1.1$. Therefore one needs to resolve
@@
 interface as $1-2$ lattice Boltzmann units but keeping the capillary number as it is required. In
-this case we can refer like we can't resolve the interface but we can certainly to attribute it to
-the wall gradient as extreme case. The dimnesionalization case is therefore as follows:
-\begin{equation}
-Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
-\end{equation}
-therefore let us pick up the same parameters as for the previous case. To perform simulations we
-started with the following parameters as
-$k=0.04$,$A=0.04$,$\Gamma=1.0$,$N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$.The
-initialized density was chosen as $w=2.5$ lattice Boltzmann units. For underresolved schemes, the
-pressure gradient doesn't work in the way we expect. Therefore, we take pressure gradient in the
-following way and later on establish the associated capillary number. In this case we can check the
-extreme values for the pressure gradients and the continuous interface methods:
+this case we can refer like we can't resolve the interface. However, one can check whether
+completely underresolved interfaces can give something good with the continuous interface methods.
+The dimensionalization case is therefore as follows. Instead of starting with the capillary number
+simulations we initialize the initial thickness with $2.5$ lattice boltzmann units, change the
+pressure gradients through the following numbers:
 \begin{equation}
@@
 \end{equation}
+After performing the simulations we determine the $R_{axis}$, $R_{diag}$ versus calculated
+capillary number
+\begin{equation}
+Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}.
+\end{equation}
+To perform simulations we
+started with the following parameters as
+$k=0.04$,$A=0.04$,$\Gamma=1.0$,$N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$.
+For underresolved schemes, the pressure gradient doesn't work in the way we expect. Therefore, we
+take pressure gradient in the following way and later on establish the associated capillary number.
+We want to keep track of $R_{diag}$ when the $R_{axis}$ is underresolved. The preliminary results
+for the film thickness measured in the middle of the bubble are shown in Fig.
+\ref{fig:underresolved:capillaries}. 
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/underresolved_capillaries.eps}
+\caption{The underresolved profiles axis and diagonal radiuses versus capillary numbers. $50$
+percent resolution starts from $Ca=1.0$, which is quite high. One can see it's exactly where
+radiuses converge instead of $Ca=0.04$. However, one can see from simulations Fig.
+\ref{fig:wang:critical:capillary} and Fig.\ref{fig:thickness:variation:ca:ten} the thickness,
+especially for Fig.\ref{fig:thickness:variation:ca:ten}, can significantly variate over a bubble
+length. \label{fig:underresolved:capillaries}}
+\end{figure}
+
+One can see the variation of thickness over the bubble length for the given capillary $Ca=1.06$ in
+Fig. \ref{fig:variation:bubble:ca:one}. One can see that variation over the length can be neglected
+for relatively high $Ca$ numbers.
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/bubble_length_ca_one.eps}
+\caption{The bubble length variation over for given $Ca=1.06$.
+One can see that there is no such high variation over the bubble
+length. \label{fig:variation:bubble:ca:one}}
+\end{figure}
+
+\section{Steady-state case}
+We want to perform understanding of the steady state regime. As well probably it's good idea to
+change domain as well, to reduce the length but to insure enough length of the bubble. The
+steady-state calculations are performed on $52\mathrm{x}52\mathrm{x}750$ with large enough the
+interface.
 

### 2010-12-17 Sailfish/Asymmetric/govern.py

@@ -6,6 +6,9 @@ import numpy
 import math
 from enthought.tvtk.api import tvtk
 from enthought.mayavi import mlab
+import matplotlib.pyplot as plt
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
 
  
 def Run_Simulations():
@@ -112,220 +115,166 @@ def Analyze_Phase(name):
     prof_diag=numpy.diag(slice)
     
     axis_zero=(1.0-Get_Zero(prof_axis)/(0.5*(shape[1]-2)))
-    diag_zero=(math.sqrt(2)-Get_Zero(prof_diag)/(0.5*(shape[1]-2)))
+    diag_zero=(math.sqrt(2.0)-math.sqrt(2.0)*Get_Zero(prof_diag)/(0.5*(shape[1]-2)))
     
-    print "Velx",velx_numpy[dims[0]/2,dims[1]/2,mid]
-    print "Vely",vely_numpy[dims[0]/2,dims[1]/2,mid]
-    print "Velz",velz_numpy[dims[0]/2,dims[1]/2,mid]
+    #print "Velx",velx_numpy[dims[0]/2,dims[1]/2,mid]
+    #print "Vely",vely_numpy[dims[0]/2,dims[1]/2,mid]
+    #print "Velz",velz_numpy[dims[0]/2,dims[1]/2,mid]
     
     #Calculation of the capillary number
-    capillary=velx_numpy[dims[0]/2,dims[1]/2,mid]*2.0/3.0/sqrt(8.0*k*a/9.0)    
-    pylab.figure()
-    pylab.imshow(phase_numpy[:,:,mid])
-    pylab.figure()
-    pylab.plot(prof_axis)
-    pylab.figure()
-    pylab.plot(prof_diag)
+    capillary=velx_numpy[dims[0]/2,dims[1]/2,mid]*(2.0/3.0)/math.sqrt(8.0*k*a/9.0)    
+    
+    #pylab.figure()
+    #pylab.imshow(phase_numpy[:,:,mid])
+    #pylab.figure()
+    #pylab.plot(prof_axis)
+    #pylab.figure()
+    #pylab.plot(prof_diag)
+
+
+    print axis_zero,diag_zero
+    print "Capillary=",capillary
+
+    return axis_zero,diag_zero,capillary
     #mlab.show()
     
 def Analyze_Consequence():
-    print os.getcwd()
-    for i in range(1,3):
[…]

### 2010-12-17 Sailfish/Steady/govern.py

@@ -1,0 +1,217 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+import math
+ 
+def Run_Simulations():
+    print os.getcwd()
+    
+    force_init=30e-6;
+    for i in range(1,5):
+        dir_temp=str(i)
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','steady.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        force=force_init*i
+        subprocess.call(['./steady.py','--force='+str(force_init),'--batch','--every='+str(50000*i),'--max_iters='+str(200000*i+1),
+        '--iwidth=6','--output=steady','--output_format=vtk','--bc_wall_grad_order=1'])
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return (zero-0.5)/(len(prof)-2)
+
+def Analyze_Simulations():
+    from numpy import genfromtxt
+    print os.getcwd()
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
+    capillary_str=["3","5","8","10","20","40","60","80"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
+
+    exam=[2000,2200,2600,2500,1100,950,600,350]
+    good=[0,1,2,3,4,5,6,7]
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    giavedoni=genfromtxt("planarcasesolution.csv",delimiter=',',dtype=float)[1:]
+
+
+    for i in range(0, len(capillary_theor)):
+        dir_temp="Results/"+capillary_str[i]
+        #ratio=float(ny[i]-2)/100.0
+        os.chdir(dir_temp)
+        #os.chdir("Force")
+        #pylab.figure()
[…]

### 2010-12-17 Sailfish/Steady/steady.py

@@ -1,0 +1,61 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo3D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width),
+                                  np.logical_or(hz < self.wall_width, hz >= self.lat_nz-self.wall_width))
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(np.logical_and(
+                       np.logical_and(hz >= width-1 + self.wall_width, hz <= self.lat_nz-width-self.wall_width),
+                       np.logical_and(hy >= width-1 + self.wall_width, hy <= self.lat_ny-width-self.wall_width)),
+                       np.logical_and(hx >= self.lat_nx/3.0, hx<=2*self.lat_nx/3.0))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_3d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 750,'grid': 'D3Q19',
+                    'lat_ny': 52, 'lat_nz' : 52 ,
+                    'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True,'periodic_z': True,
+                    'bc_wall_grad_order': 1}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=1.0e-6))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=2))
[…]

### 2011-01-19 Text/paper.tex
Added descriptor but doesn't work till the end.

@@
 \section{Steady-state case}
+We performed different simulations to understand a number of steps required for the problem to come
+on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ with the initial
+width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=3
+\mathrm{x}10^{-5}$. The simulation was performed for $200000$,$400000$,$600000$ and $800000$ time
+iterations. The results are summarised in Table \ref{table:steady:state}. One of the profiles in
+the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}.
+\begin{table}
+\begin{tabular}{|c|c|c|c|}
+$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
+$200000$&$0.285$&$0.873$&$0.921$\\
+$400000$&$0.274$&$0.873$&$0.920$\\
+$600000$&$0.263$&$0.872$&$0.919$\\
+$800000$&$0.252$&$0.873$&$0.917$\\
+\end{tabular}
+\caption{Results for the steady-state case. One can see that $200000$ steps are enough to structure
+the bubble. However, the interface is underresolved. That could be the influence of the bubble
+length.\label{table:steady:state}}
+\end{table}
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
+\caption{Slice in the middle of the bubble and
+three-dimensional example. \label{fig:steady:state:profile:example}}
+\end{figure}
+
+To understand the influence of the steady-state case
+For the steady-state regime the following parameters were obtained
 We want to perform understanding of the steady state regime. As well probably it's good idea to

### 2011-01-19 Code/descriptor.h

@@ -1,0 +1,77 @@
+#ifndef DESCRIPTOR_H
+#define DESCRIPTOR_H
+#include <vector>
+#include <cmath>
+#include <iostream>
+class Descriptor
+{
+	public:
+	static const int d=3;
+	static const int NPOP=19;
+
+	static const char cx[NPOP];
+    static const char cy[NPOP];
+    static const char cz[NPOP];
+
+    //Stencils
+    static const double wxx[NPOP];
+    static const double wyy[NPOP];
+    static const double wzz[NPOP];
+    static const double wxy[NPOP];
+    static const double wyz[NPOP];
+    static const double wzx[NPOP];
+    static const double weights[NPOP];
+
+    static const double laplace_stencil[NPOP];
+    static const double gradx_stencil[NPOP];
+    static const double grady_stencil[NPOP];
+    static const double gradz_stencil[NPOP];
+    static const int compliment[NPOP];
+
+
+	public:
+	Descriptor(){}
+};
+
+const char Descriptor::cx[19]={0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1};
+const char Descriptor::cy[19]={0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0};
+const char Descriptor::cz[19]={0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1};
+const double Descriptor::wxx[19]={0.0,5.0/12.0,5.0/12.0,-1.0/3.0,-1.0/3.0,-1.0/3.0,-1.0/3.0,
+		-1.0/24.0,-1.0/24.0,-1.0/24.0,-1.0/24.0,1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,
+		-1.0/24.0,-1.0/24.0,-1.0/24.0,-1.0/24.0};
[…]

### 2011-01-19 Code/dynamics.h

@@ -3,13 +3,13 @@
 
 #include "params_list.h"
 //#include "solver.h"
-class Solver;
+template<typename D> class Solver;
 
-class Dynamics
+template<typename D> class Dynamics
 {
     public:
 
-    explicit Dynamics(Solver * _solver,ParamsList _params):
+    explicit Dynamics(Solver<D> * _solver,ParamsList _params):
         solver(_solver),params(_params)
     {}
 
@@ -22,7 +22,7 @@ class Dynamics
     virtual void updateFields(int iX,int iY,int iZ)=0;
 
     public:
-    Solver * solver;
+    Solver<D> * solver;
     ParamsList params;
 
 };

### 2011-01-19 Code/dynamics_BGK.cpp

@@ -2,9 +2,9 @@
 #include "lattice.h"
 #include "solver.h"
 
-void DynamicsBGK::init(int iX,int iY, int iZ)
+template<typename D> void DynamicsBGK<D>::init(int iX,int iY, int iZ)
 {
-	Lattice * lattice=this->solver->getLattice();
+	Lattice<D> * lattice=this->solver->getLattice();
 
 	int zbegin=lattice->getZbegin();
 	int zend=lattice->getZend();
@@ -15,12 +15,12 @@ void DynamicsBGK::init(int iX,int iY, int iZ)
 	int counter=iZ*NX*NY+iY*NX+iX;
 
 	double gradx_temp=0.0,grady_temp=0.0,gradz_temp=0.0,laplace_temp=0.0;
-	for (int k=0;k<NPOP;k++)
+	for (int k=0;k<D::NPOP;k++)
 	{
 		double phase_temp;
-		int iZ2=iZ+cz[k];
-		int iY2=(iY+cy[k]+NY)%NY;
-		int iX2=(iX+cx[k]+NX)%NX;
+		int iZ2=iZ+D::cz[k];
+		int iY2=(iY+D::cy[k]+NY)%NY;
+		int iX2=(iX+D::cx[k]+NX)%NX;
 		if (iZ2==-1)
 			phase_temp=lattice->phase_bottom[iY2*NX+iX2];
 		else
@@ -29,10 +29,10 @@ void DynamicsBGK::init(int iX,int iY, int iZ)
 			else
 				phase_temp=lattice->phase[iZ2*NX*NY+iY2*NX+iX2];
 
-		gradx_temp+=gradx_stencil[k]*phase_temp;
-		grady_temp+=grady_stencil[k]*phase_temp;
-		gradz_temp+=gradz_stencil[k]*phase_temp;
-		laplace_temp+=laplace_stencil[k]*phase_temp;
+		gradx_temp+=D::gradx_stencil[k]*phase_temp;
+		grady_temp+=D::grady_stencil[k]*phase_temp;
+		gradz_temp+=D::gradz_stencil[k]*phase_temp;
+		laplace_temp+=D::laplace_stencil[k]*phase_temp;
 	}
 
 	double phase_temp=lattice->phase[counter];
[…]

### 2011-01-19 Code/dynamics_BGK.h

@@ -4,11 +4,12 @@
 #include "dynamics.h"
 #include "params_list.h"
 
-class DynamicsBGK:public Dynamics
+template<typename D>
+class DynamicsBGK:public Dynamics<D>
 {
     public:
 
-    explicit DynamicsBGK(Solver * _solver,ParamsList _params);
+    explicit DynamicsBGK(Solver<D> * _solver,ParamsList _params);
     virtual ~DynamicsBGK()
     {}
 
@@ -18,7 +19,6 @@ class DynamicsBGK:public Dynamics
 
     public:
 
-    double omega;
     double aconst;
     double kconst;
     double gammaconst;
@@ -28,26 +28,6 @@ class DynamicsBGK:public Dynamics
     double force_x;
     double force_y;
     double force_z;
-
-    //Constants for the calculations
-    double wxx[19];
-    double wyy[19];
-    double wzz[19];
-    double wxy[19];
-    double wyz[19];
-    double wzx[19];
-    double weights[19];
-
-    char cx[19];
-    char cy[19];
-    char cz[19];
-    int NPOP;
-
-    //Parameters for stencils
-    double laplace_stencil[19];
-    double gradx_stencil[19];
-    double grady_stencil[19];
-    double gradz_stencil[19];
 
 };
 #endif

### 2011-01-19 Code/dynamics_none.h

@@ -3,22 +3,16 @@
 #include "lattice.h"
 #include "solver.h"
 
-class DynamicsNONE:public Dynamics
+template<typename D> class DynamicsNONE:public Dynamics<D>
 {
     public:
 
-    explicit DynamicsNONE(Solver * _solver,ParamsList _params):
-        Dynamics(_solver,_params),
-        phase_wall(params("phase_wall").value<double>()),
-        rho_wall(params("rho_wall").value<double>()),
-        NPOP(19)
+    explicit DynamicsNONE(Solver<D> * _solver,ParamsList _params):
+        Dynamics<D>(_solver,_params),
+        phase_wall(_params("phase_wall").value<double>()),
+        rho_wall(_params("rho_wall").value<double>()),
+        NPOP(D::NPOP)
     {
-        double weights_temp[19]={0,1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0,
-		1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,
-		1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0,1.0/12.0};
-
-        for (int counter=0;counter<19;counter++)
-            weights[counter]=weights_temp[counter];
     }
 
     virtual ~DynamicsNONE()
@@ -26,7 +20,7 @@ class DynamicsNONE:public Dynamics
 
     void collide_stream(int iX,int iY,int iZ)
     {
-       	Lattice * lattice=this->solver->getLattice();
+       	Lattice<D> * lattice=this->solver->getLattice();
         int NX=this->solver->geom->getNX();
         int NY=this->solver->geom->getNY();
 
@@ -42,7 +36,7 @@ class DynamicsNONE:public Dynamics
     {
         int NX=this->solver->geom->getNX();
         int NY=this->solver->geom->getNY();
-        Lattice * lattice=this->solver->getLattice();
+        Lattice<D> * lattice=this->solver->getLattice();
 
         int counter=iZ*NX*NY+iY*NX+iX;
 
@@ -51,9 +45,9 @@ class DynamicsNONE:public Dynamics
         {
             dense_temp+=lattice->f[counter*NPOP+k];
             phase_temp+=lattice->g[counter*NPOP+k];
[…]

### 2011-01-19 Code/dynamics_simple_BGK.cpp

@@ -2,9 +2,9 @@
 #include "lattice.h"
 #include "solver.h"
 
-void DynamicsSimpleBGK::init(int iX,int iY, int iZ)
+template<typename D> void DynamicsSimpleBGK<D>::init(int iX,int iY, int iZ)
 {
-	Lattice * lattice=this->solver->getLattice();
+	Lattice<D> * lattice=this->solver->getLattice();
 
 	int zbegin=lattice->getZbegin();
 	int zend=lattice->getZend();
@@ -25,40 +25,40 @@ void DynamicsSimpleBGK::init(int iX,int iY, int iZ)
 	double sum=0.0;
 	double sum_phase=0.0;
 
-	for (int k=1; k<NPOP; k++)
+	for (int k=1; k<D::NPOP; k++)
 	{
-		feq=weights[k]*(dense_temp/3.0+dense_temp*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp)
-						+1.5*dense_temp*((cx[k]*cx[k]-1.0/3.0)*ux_temp*ux_temp+(cy[k]*cy[k]-1.0/3.0)*uy_temp*uy_temp+(cz[k]*cz[k]-1.0/3.0)*uz_temp*uz_temp
-										 +2.0*(ux_temp*uy_temp*cx[k]*cy[k]+ux_temp*uz_temp*cx[k]*cz[k]+uy_temp*uz_temp*cy[k]*cz[k])));
-		geq=weights[k]*(phase_temp/3.0+phase_temp*(cx[k]*ux_temp+cy[k]*uy_temp+cz[k]*uz_temp)
-						+1.5*phase_temp*((cx[k]*cx[k]-1.0/3.0)*ux_temp*ux_temp+(cy[k]*cy[k]-1.0/3.0)*uy_temp*uy_temp+(cz[k]*cz[k]-1.0/3.0)*uz_temp*uz_temp
-										 +2.0*(ux_temp*uy_temp*cx[k]*cy[k]+ux_temp*uz_temp*cx[k]*cz[k]+uy_temp*uz_temp*cy[k]*cz[k])));
+		feq=D::weights[k]*(dense_temp/3.0+dense_temp*(D::cx[k]*ux_temp+D::cy[k]*uy_temp+D::cz[k]*uz_temp)
[…]

### 2011-01-19 Code/dynamics_simple_BGK.h

@@ -4,11 +4,11 @@
 #include "dynamics.h"
 #include "params_list.h"
 
-class DynamicsSimpleBGK:public Dynamics
+template<typename D> class DynamicsSimpleBGK:public Dynamics<D>
 {
     public:
 
-    explicit DynamicsSimpleBGK(Solver * _solver,ParamsList _params);
+    explicit DynamicsSimpleBGK(Solver<D> * _solver,ParamsList _params);
     virtual ~DynamicsSimpleBGK()
     {}
 
@@ -22,13 +22,6 @@ class DynamicsSimpleBGK:public Dynamics
     double force_x;
     double force_y;
     double force_z;
-
-    double weights[19];
-
-    char cx[19];
-    char cy[19];
-    char cz[19];
-    int NPOP;
 
 };
 #endif

### 2011-01-19 Code/dynamics_simple_special.cpp

@@ -1,7 +1,8 @@
 #include "dynamics_simple_special.h"
 #include "solver.h"
-DynamicsSimpleSpecial::DynamicsSimpleSpecial(Solver * _solver, ParamsList _params_list):
-    DynamicsSimpleBGK(_solver,_params_list)
+template<typename D>
+DynamicsSimpleSpecial<D>::DynamicsSimpleSpecial(Solver<D> * _solver, ParamsList _params_list):
+    DynamicsSimpleBGK<D>(_solver,_params_list)
 {
 	int NX=this->solver->geom->getNX();
 	int NY=this->solver->geom->getNY();
@@ -12,11 +13,11 @@ DynamicsSimpleSpecial::DynamicsSimpleSpecial(Solver * _solver, ParamsList _param
     iY=this->solver->iterY;
     iZ=this->solver->iterZ;
 
-    for(int k=0; k<NPOP; k++)
+    for(int k=0; k<D::NPOP; k++)
     {
-        int iX2=(iX+Solver::cx[k]+NX)%NX;
-        int iY2=(iY+Solver::cy[k]+NY)%NY;
-        int iZ2=(iZ+Solver::cz[k]+NZ)%NZ;
+        int iX2=(iX+D::cx[k]+NX)%NX;
+        int iY2=(iY+D::cy[k]+NY)%NY;
+        int iZ2=(iZ+D::cz[k]+NZ)%NZ;
         Geometry * geom=this->solver->geom;
         if (geom->getType(iX2,iY2,iZ2)==SolidNode)
         {
@@ -25,14 +26,12 @@ DynamicsSimpleSpecial::DynamicsSimpleSpecial(Solver * _solver, ParamsList _param
     }
 
 
-    int compliment_temp[19]={0, 2, 1, 4, 3, 6, 5, 10, 9, 8, 7, 14, 13, 12, 11, 18, 17, 16, 15};
-    for (int k=0;k<NPOP;k++)
-        compliment[k]=compliment_temp[k];
 }
 
-void DynamicsSimpleSpecial::init(int iX,int iY,int iZ)
+template<typename D>
+void DynamicsSimpleSpecial<D>::init(int iX,int iY,int iZ)
 {
[…]

### 2011-01-19 Code/dynamics_simple_special.h

@@ -3,11 +3,12 @@
 #include <vector>
 #include "dynamics_simple_BGK.h"
 #include <iostream>
-class DynamicsSimpleSpecial : public DynamicsSimpleBGK
+template<typename D>
+class DynamicsSimpleSpecial : public DynamicsSimpleBGK<D>
 {
     public:
 
-    explicit DynamicsSimpleSpecial(Solver * _solver,ParamsList _params);
+    explicit DynamicsSimpleSpecial(Solver<D> * _solver,ParamsList _params);
 
     virtual ~DynamicsSimpleSpecial()
     {
@@ -25,7 +26,7 @@ class DynamicsSimpleSpecial : public DynamicsSimpleBGK
     int iZ;
 
     std::vector<int> directions;
-    int compliment[19];
+
 };
 
 #endif

### 2011-01-19 Code/dynamics_symmetric.cpp

@@ -1,9 +1,9 @@
 #include "dynamics_special.h"
 #include "solver.h"
 template<typename D>
-DynamicsSpecial<D>::DynamicsSpecial(Solver<D> * _solver, ParamsList _params_list):
+DynamicsSpecial<D>::DynamicsSpecial(Solver * _solver, ParamsList _params_list):
     DynamicsBGK<D>(_solver,_params_list),
-    phase_gradient(_params_list("phase_gradient").value<double>())
+    phase_gradient(this->params("phase_gradient").value<double>())
 {
     normx=0.0;
     normy=0.0;
@@ -17,17 +17,17 @@ DynamicsSpecial<D>::DynamicsSpecial(Solver<D> * _solver, ParamsList _params_list
     iY=this->solver->iterY;
     iZ=this->solver->iterZ;
 
-    for(int k=0; k<D::NPOP; k++)
+    for(int k=0; k<NPOP; k++)
     {
-        int iX2=(iX+D::cx[k]+NX)%NX;
-        int iY2=(iY+D::cy[k]+NY)%NY;
-        int iZ2=(iZ+D::cz[k]+NZ)%NZ;
+        int iX2=(iX+Solver::cx[k]+NX)%NX;
+        int iY2=(iY+Solver::cy[k]+NY)%NY;
+        int iZ2=(iZ+Solver::cz[k]+NZ)%NZ;
         Geometry * geom=this->solver->geom;
         if (geom->getType(iX2,iY2,iZ2)==SolidNode)
         {
-            normx+=double(D::cx[k]);
-            normy+=double(D::cy[k]);
-            normz+=double(D::cz[k]);
+            normx+=double(cx[k]);
+            normy+=double(cy[k]);
+            normz+=double(cz[k]);
             directions.push_back(k);
         }
     }
@@ -38,14 +38,16 @@ DynamicsSpecial<D>::DynamicsSpecial(Solver<D> * _solver, ParamsList _params_list
     normy/=norm;
     normz/=norm;
 
-    wall_densities.resize(D::NPOP);
+    wall_densities.resize(NPOP);
 
[…]

### 2011-01-19 Code/dynamics_symmetric.h

@@ -3,12 +3,11 @@
 #include <vector>
 #include "dynamics_BGK.h"
 #include <iostream>
-template<typename D>
-class DynamicsSpecial : public DynamicsBGK<D>
+class DynamicsSpecial : public DynamicsBGK
 {
     public:
 
-    explicit DynamicsSpecial(Solver<D> * _solver,ParamsList _params);
+    explicit DynamicsSpecial(Solver * _solver,ParamsList _params);
 
     virtual ~DynamicsSpecial()
     {
@@ -38,6 +37,7 @@ class DynamicsSpecial : public DynamicsBGK<D>
 
     std::vector<int> directions;
     std::vector<double> wall_densities;
+    int compliment[19];
 };
 
 #endif

### 2011-01-19 Code/geometry.h

@@ -1,7 +1,7 @@
 #ifndef GEOMETRY_H
 #define GEOMETRY_H
 
-enum NodeType {FluidNode, SolidNode, PressureNode};
+enum NodeType {FluidNode, SolidNode, SymmetricNode};
 
 class Geometry
 {

### 2011-01-19 Code/lattice.cpp

@@ -2,13 +2,9 @@
 #include "dynamics_BGK.h"
 #include "mpi_singleton.h"
 
-const char Lattice::cx[19]={0,1,-1,0,0,0,0,1,-1,1,-1,0,0,0,0,1,-1,1,-1};
-const char Lattice::cy[19]={0,0,0,1,-1,0,0,1,1,-1,-1,1,-1,1,-1,0,0,0,0};
-const char Lattice::cz[19]={0,0,0,0,0,1,-1,0,0,0,0,1,1,-1,-1,1,1,-1,-1};
 
-
-
-void Lattice::writePhase(std::string name)
+template<typename D>
+void Lattice<D>::writePhase(std::string name)
 {
 
     name=name+".vts";
@@ -44,7 +40,8 @@ void Lattice::writePhase(std::string name)
     writer->Write();
 }
 
-void Lattice::writeTextPhase(std::string name)
+template<typename D>
+void Lattice<D>::writeTextPhase(std::string name)
 {
 
     name=name+".dat";
@@ -56,7 +53,8 @@ void Lattice::writeTextPhase(std::string name)
 
 }
 
-void Lattice::writeTextDensity(std::string name)
+template<typename D>
+void Lattice<D>::writeTextDensity(std::string name)
 {
 
     name=name+".dat";
@@ -68,11 +66,8 @@ void Lattice::writeTextDensity(std::string name)
 
 }
 
-
-
-
-
-void Lattice::writeDensity(std::string name)
+template<typename D>
+void Lattice<D>::writeDensity(std::string name)
 {
 
     name=name+".vts";
@@ -109,7 +104,8 @@ void Lattice::writeDensity(std::string name)
     writer->Write();
 }
 
-void Lattice::init()
+template<typename D>
+void Lattice<D>::init()
 {
 	for (int counter=0;counter<NUMLOCAL;counter++)
 	{
@@ -120,7 +116,8 @@ void Lattice::init()
 	}
 }
 
-void Lattice::collide_stream()
+template<typename D>
+void Lattice<D>::collide_stream()
 {
 	for (int counter=0;counter<NUMLOCAL;counter++)
 	{
@@ -131,45 +128,47 @@ void Lattice::collide_stream()
 	}
 }
 
[…]

### 2011-01-19 Code/lattice.h

@@ -11,19 +11,20 @@
 #include <vtkPointData.h>
 
 
+template<typename D>
 class Lattice
 {
     public:
     Lattice(int _zbegin,int _zend,int _NX,int _NY,int _NZ,int _NUMLOCAL):
         zbegin(_zbegin),zend(_zend),
         NX(_NX),NY(_NY),NZ(_NZ),
-        NUMLOCAL(_NUMLOCAL), NPOP(19)
+        NUMLOCAL(_NUMLOCAL)
     {
         //Populations
-        f = new double[NUMLOCAL*NPOP];
-        f2= new double[NUMLOCAL*NPOP];
-        g = new double[NUMLOCAL*NPOP];
-        g2= new double[NUMLOCAL*NPOP];
+        f = new double[NUMLOCAL*D::NPOP];
+        f2= new double[NUMLOCAL*D::NPOP];
+        g = new double[NUMLOCAL*D::NPOP];
+        g2= new double[NUMLOCAL*D::NPOP];
 
         //Macro variables
         phase = new double[NUMLOCAL];
@@ -35,11 +36,11 @@ class Lattice
         //Helpful materials
         phase_top = new double[NX*NY];
         phase_bottom = new double[NX*NY];
-        dynamics_list = new Dynamics*[NUMLOCAL];
-        layer_top_f = new double[NX*NY*NPOP];
-		layer_bottom_f = new double[NX*NY*NPOP];
-		layer_top_g = new double[NX*NY*NPOP];
-		layer_bottom_g = new double[NX*NY*NPOP];
+        dynamics_list = new Dynamics<D>*[NUMLOCAL];
+        layer_top_f = new double[NX*NY*D::NPOP];
+		layer_bottom_f = new double[NX*NY*D::NPOP];
+		layer_top_g = new double[NX*NY*D::NPOP];
+		layer_bottom_g = new double[NX*NY*D::NPOP];
 
         geom_top=new NodeType[NX*NY];
         geom_bottom=new NodeType[NX*NY];
@@ -150,11 +151,10 @@ class Lattice
     private:
         int zbegin;
         int zend;
-        int NX;
-        int NY;
-        int NZ;
+        const int NX;
+        const int NY;
+        const int NZ;
         int NUMLOCAL;
-        const int NPOP;
 
     public:
         double * f;
@@ -178,9 +178,6 @@ class Lattice
         NodeType * geom_top;
         NodeType * geom_bottom;
 
-        Dynamics ** dynamics_list;
[…]

### 2011-01-19 Code/main.cpp

@@ -14,11 +14,12 @@
 #include "geometry.h"
 #include "solver.h"
 #include "params_list.h"
+#include "descriptor.h"
 
 //Lattice Boltzmann initialization and parameters
 const int NX=52;
 const int NY=52;
-const int NZ=750;
+const int NZ=51;
 const int NPOP=19;
 const int NUM=NX*NY*NZ;
 const int NUMTOTAL=NUM*NPOP;
@@ -32,7 +33,7 @@ const int radius=6;
 const double rhol=1.0;
 const double rhog=1.0;
 
-void macro_init(Solver & solver)
+void macro_init(Solver<Descriptor> & solver)
 {
     //Density and phase field initialization
 	double rho_temp;
@@ -86,10 +87,10 @@ int main(int argc,char* argv[])
 
     //Specify parameters
  	ParamsList params;
+	params.add("omega",1.0);
 	params.add("aconst", 0.04);
 	params.add("kconst", 0.04);
 	params.add("gammaconst",1.0);
-	params.add("omega",1.0);
 	params.add("phase_gradient",0.0);
 	params.add("phase_wall",0.0);
 	params.add("rho_wall",0.5);
@@ -99,13 +100,6 @@ int main(int argc,char* argv[])
     params.add("force_x",0.0);
     params.add("force_y",0.0);
     params.add("force_z",0.000015);
-	//params.add("force_z",0.0);
-
-    //Useless parameters to be deleted after
-    params.add("rho_press_inlet",1.0);
-    params.add("rho_press_outlet",1.0);
-    params.add("phase_press_inlet",1.0);
-    params.add("phase_press_outlet",1.0);
 
     //Specify geometry
     Geometry * geometry=new Geometry(NX,NY,NZ);
@@ -117,7 +111,11 @@ int main(int argc,char* argv[])
     geometry->setType(0,NY-1,0,NX-1,NY-1,NZ-1,SolidNode);
     geometry->setType(NX-1,0,0,NX-1,NY-1,NZ-1,SolidNode);
 
-    Solver solver(geometry,params);
+    //SymmetricNode specification
+    //geometry->setType(0,0,0,NX-1,0,NZ-1,SymmetricNode);
+    //geometry->setType(0,0,0,0,NY-1,NZ-1,SymmetricNode);
+
+    Solver<Descriptor> solver(geometry,params);
 
     //Solver initialization from the file
     //solver.load_file("equili");

### 2011-01-19 Code/solver.cpp

@@ -16,17 +16,13 @@
 #include <sstream>
 #include <string>
 
-const char Solver::cx[19]={0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1};
-const char Solver::cy[19]={0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0};
-const char Solver::cz[19]={0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1};
-
-Solver::Solver(Geometry * _geom,ParamsList _params_list):
+template<typename D>
+Solver<D>::Solver(Geometry * _geom,ParamsList _params_list):
 	geom(_geom),NX(geom->getNX()),
 	NY(geom->getNY()),NZ(geom->getNZ()),
 	rank(mpi_wrapper().get_rank()),
 	size(mpi_wrapper().get_size()),
-	params_list(_params_list),
-	NPOP(19)
+	params_list(_params_list)
 {
 	//Initialization of ranges
 	int zstep=NZ/size;
@@ -37,15 +33,16 @@ Solver::Solver(Geometry * _geom,ParamsList _params_list):
 	int NUMLOCAL=NX*NY*(zend-zbegin+1);
 
 	//Creation the necessary objects
-	lattice = new Lattice(zbegin,zend,NX,NY,NZ,NUMLOCAL);
-	Dynamics * dynamics_bgk;
+	lattice = new Lattice<D>(zbegin,zend,NX,NY,NZ,NUMLOCAL);
+	Dynamics<D> * dynamics_bgk;
 	#ifdef HYDRO
-        dynamics_bgk = destroyer_list.take_ownership(new DynamicsSimpleBGK(this,params_list));
+        dynamics_bgk = destroyer_list.take_ownership(new DynamicsSimpleBGK<D>(this,params_list));
 	#else
-        dynamics_bgk = destroyer_list.take_ownership(new DynamicsBGK(this,params_list));
+        dynamics_bgk = destroyer_list.take_ownership(new DynamicsBGK<D>(this,params_list));
     #endif
-    Dynamics * dynamics_special;
-    Dynamics * dynamics_none = destroyer_list.take_ownership(new DynamicsNONE(this,params_list));
+    Dynamics<D> * dynamics_special;
+    Dynamics<D> * dynamics_none = destroyer_list.take_ownership(new DynamicsNONE<D>(this,params_list));
[…]

### 2011-01-19 Code/solver.h

@@ -9,6 +9,7 @@
 #include "destroyer.h"
 #include "params_list.h"
 
+template<typename D>
 class Solver
 {
     public:
@@ -75,7 +76,7 @@ class Solver
     void exchangeMatrices();
 
 
-    Lattice * getLattice()
+    Lattice<D> * getLattice()
     {
         return lattice;
     }
@@ -88,11 +89,6 @@ class Solver
     int iterY;
     int iterZ;
 
-    //Velocity set. TODO: put it in the Descriptor Object
-    static const char cx[19];
-    static const char cy[19];
-    static const char cz[19];
-
 	private:
 
     const int size;
@@ -100,15 +96,14 @@ class Solver
     const int NX;
     const int NY;
     const int NZ;
-    const int NPOP;
 
 
-    Lattice * lattice;
+    Lattice<D> * lattice;
     double * density;
     double * phase;
     double * velocity;
 
-    Destroyer<Dynamics> destroyer_list;
+    Destroyer<Dynamics<D> > destroyer_list;
     ParamsList params_list;
 };
 

### 2011-01-19 Sailfish/Steady/govern.py

@@ -4,6 +4,12 @@ import subprocess
 import pylab
 import numpy
 import math
+import math
+from enthought.tvtk.api import tvtk
+from enthought.mayavi import mlab
+import matplotlib.pyplot as plt
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
  
 def Run_Simulations():
     print os.getcwd()
@@ -26,191 +32,245 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    return (zero-0.5)/(len(prof)-2)
+    return zero-0.5
 
-def Analyze_Simulations():
-    from numpy import genfromtxt
+def Read_Phase(name):
     print os.getcwd()
-    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
-    capillary_str=["3","5","8","10","20","40","60","80"]
-    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
+    gridreader = tvtk.XMLImageDataReader()
+    gridreader.file_name = name
+    gridreader.update()
 
-    exam=[2000,2200,2600,2500,1100,950,600,350]
-    good=[0,1,2,3,4,5,6,7]
-    #style=["bo","rH","c<","y>","bs","g^"]
-    #color=["b","r","c","y","b","g"]
-    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    grid  = gridreader.output
+    data  = grid.point_data
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("phi"))
+    velocity=numpy.array(data.get_array("v"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
+    phase_numpy=phase.reshape(dims)
+    velx_numpy =velx.reshape(dims)
+    vely_numpy =vely.reshape(dims)
+    velz_numpy =velz.reshape(dims)
+    return phase_numpy,velx_numpy,vely_numpy,velz_numpy
+
+def Show_Phase_Slice(phase_numpy,slice):
+    fig=pylab.figure()
[…]

### 2011-01-19 Sailfish/Train/govern.py

@@ -1,0 +1,216 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+import math
+ 
+def Run_Simulations():
+    print os.getcwd()
+    
+    force_init=30e-6;
+    for i in numpy.arange(3,7,0.5):
+        dir_temp=str(int(10*i))
+        os.mkdir(dir_temp)
+        subprocess.call(['cp','train.py',dir_temp+"/"])
+        os.chdir(dir_temp)
+        subprocess.call(['./train.py','--force='+str(force_init),'--batch','--every='+str(50000),'--max_iters=200001',\
+        '--iwidth=10','--output=train','--start='+str(375-int(i*50)),'--finish='+str(375+int(i*50)),'--output_format=vtk','--bc_wall_grad_order=1'])
+        os.chdir("..")
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return (zero-0.5)/(len(prof)-2)
+
+def Analyze_Simulations():
+    from numpy import genfromtxt
+    print os.getcwd()
+    capillary_theor=[0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8]
+    capillary_str=["3","5","8","10","20","40","60","80"]
+    width_theor=[0.04,0.06,0.08,0.1,0.12,0.13,0.15,0.16]
+
+    exam=[2000,2200,2600,2500,1100,950,600,350]
+    good=[0,1,2,3,4,5,6,7]
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    giavedoni=genfromtxt("planarcasesolution.csv",delimiter=',',dtype=float)[1:]
+
+
+    for i in range(0, len(capillary_theor)):
+        dir_temp="Results/"+capillary_str[i]
+        #ratio=float(ny[i]-2)/100.0
+        os.chdir(dir_temp)
[…]

### 2011-01-19 Sailfish/Train/train.py

@@ -1,0 +1,66 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo3D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        wall_map = np.logical_or(np.logical_or(hy < self.wall_width, hy >= self.lat_ny-self.wall_width),
+                                  np.logical_or(hz < self.wall_width, hz >= self.lat_nz-self.wall_width))
+        self.set_geo(wall_map, self.NODE_WALL)
+
+    def init_fields(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+        start=self.sim.options.start
+        finish=self.sim.options.finish
+
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(np.logical_and(
+                       np.logical_and(hz >= width-1 + self.wall_width, hz <= self.lat_nz-width-self.wall_width),
+                       np.logical_and(hy >= width-1 + self.wall_width, hy <= self.lat_ny-width-self.wall_width)),
+                       np.logical_and(hx >= start, hx<=finish))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lbm.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_3d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 750,'grid': 'D3Q19',
+                    'lat_ny': 102, 'lat_nz' : 102 ,
+                    'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 1.0,
+                    'bc_wall': 'halfbb',
+                    'periodic_x': True, 'periodic_y': True,'periodic_z': True,
+                    'bc_wall_grad_order': 1}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=30e-6))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=10))
[…]

### 2011-01-19 Code/descriptor.h
Rewrote with description templates. Don't include cpp files but do implementation in hpp files. Next step is to introduce symmetric nodes.

@@ -13,6 +13,13 @@ class Descriptor
     static const char cy[NPOP];
     static const char cz[NPOP];
 
+    static const double qxx[NPOP];
+    static const double qyy[NPOP];
+    static const double qzz[NPOP];
+    static const double qxy[NPOP];
+    static const double qxz[NPOP];
+    static const double qyz[NPOP];
+
     //Stencils
     static const double wxx[NPOP];
     static const double wyy[NPOP];
@@ -36,6 +43,23 @@ class Descriptor
 const char Descriptor::cx[19]={0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1};
 const char Descriptor::cy[19]={0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0};
 const char Descriptor::cz[19]={0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1};
+
+const double Descriptor::qxx[19]={-1.0/3.0, 2.0/3.0, 2.0/3.0, -1.0/3.0, -1.0/3.0, -1.0/3.0,
+    -1/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, -1.0/3.0, -1.0/3.0, -1.0/3.0,
+    -1.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0};
+
+const double Descriptor::qyy[19]={-1.0/3.0, -1.0/3.0, -1.0/3.0, 2.0/3.0, 2.0/3.0, -1.0/3.0,
+    -1.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0, 2.0/3.0,
+    -1.0/3.0, -1.0/3.0, -1.0/3.0, -1.0/3.0};
+
+const double Descriptor::qzz[19]={-1.0/3.0, -1.0/3.0, -1.0/3.0, -1.0/3.0, -1.0/3.0, 2.0/3.0,
[…]

### 2011-01-19 Code/dynamics_BGK.hpp

@@ -82,14 +82,29 @@ void DynamicsBGK<D>::updateFields(int iX,int iY,int iZ)
 
 
     double phase_temp=0.0, dense_temp=0.0, ux_temp=0.0, uy_temp=0.0, uz_temp=0.0;
-	for (int k=0;k<D::NPOP;k++)
-	{
-		dense_temp+=lattice->f[counter*D::NPOP+k];
-		phase_temp+=lattice->g[counter*D::NPOP+k];
-		ux_temp+=lattice->f[counter*D::NPOP+k]*D::cx[k];
-		uy_temp+=lattice->f[counter*D::NPOP+k]*D::cy[k];
-		uz_temp+=lattice->f[counter*D::NPOP+k]*D::cz[k];
-	}
+	double * f_temp=&lattice->f[counter*D::NPOP];
+	double * g_temp=&lattice->g[counter*D::NPOP];
+
+	dense_temp=f_temp[0]+f_temp[1]+f_temp[2]+f_temp[3]+f_temp[4]
+                +f_temp[5]+f_temp[6]+f_temp[7]+f_temp[8]+f_temp[9]
+                +f_temp[10]+f_temp[11]+f_temp[12]+f_temp[13]+f_temp[14]+
+                +f_temp[15]+f_temp[16]+f_temp[17]+f_temp[18];
+	phase_temp=g_temp[0]+g_temp[1]+g_temp[2]+g_temp[3]+g_temp[4]
+                +g_temp[5]+g_temp[6]+g_temp[7]+g_temp[8]+g_temp[9]
+                +g_temp[10]+g_temp[11]+g_temp[12]+g_temp[13]+g_temp[14]+
+                +g_temp[15]+g_temp[16]+g_temp[17]+g_temp[18];
+
+    ux_temp=f_temp[1]-f_temp[2]+f_temp[7]-f_temp[8]+f_temp[9]-f_temp[10]+f_temp[15]-f_temp[16]+f_temp[17]-f_temp[18];
+	uy_temp=f_temp[3]-f_temp[4]+f_temp[7]+f_temp[8]-f_temp[9]-f_temp[10]+f_temp[11]-f_temp[12]+f_temp[13]-f_temp[14];
[…]

### 2011-01-19 Code/main.cpp

@@ -12,7 +12,17 @@
 #include <mpi.h>
 #include "mpi_singleton.h"
 #include "geometry.h"
-#include "solver.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+//Inclusion of dynamics
+#include "dynamics_BGK.hpp"
+#include "dynamics_simple_BGK.hpp"
+#include "dynamics_special.hpp"
+#include "dynamics_simple_special.hpp"
+
 #include "params_list.h"
 #include "descriptor.h"
 
@@ -32,46 +42,6 @@ const int width=6;
 const int radius=6;
 const double rhol=1.0;
 const double rhog=1.0;
-
-void macro_init(Solver<Descriptor> & solver)
-{
-    //Density and phase field initialization
-	double rho_temp;
-	double phase_temp;
-	double u_temp[3];
-	for (int counter=0; counter<NUM; counter++)
-	{
-		int iZ=counter/(NX*NY);
-		int iY=(counter%(NX*NY))/NX;
-		int iX=(counter%(NX*NY))%NX;
-
-        //Initialization of the part of the channel
-
-		if ((iZ>=(NZ-1)/3)&&(iZ<=2*(NZ-1)/3)&&(iX>=width)&&(iX<=NX-width-1)&&(iY>=width)&&(iY<=NY-width-1))
-		{
-			rho_temp=rhog;
-			phase_temp=-1.0;
-            u_temp[0]=0.0;
-            u_temp[1]=0.0;
-            u_temp[2]=0.0;
-
-        }
-		else
-        {
-			rho_temp=rhol;
-			phase_temp=1.0;
-            u_temp[0]=0.0;
-            u_temp[1]=0.0;
-            u_temp[2]=0.0;
-
-		}
-
-		solver.putDensity(iX,iY,iZ,iX,iY,iZ,rho_temp);
-		solver.putPhase(iX,iY,iZ,iX,iY,iZ,phase_temp);
-		solver.putVelocity(iX,iY,iZ,iX,iY,iZ,u_temp);
-    }
-
-}
 
 int main(int argc,char* argv[])
 {
@@ -120,7 +90,42 @@ int main(int argc,char* argv[])
     //Solver initialization from the file
     //solver.load_file("equili");
 
-	macro_init(solver);
+	//Initialization
+	    //Density and phase field initialization
[…]

### 2011-01-19 Code/solver.hpp

@@ -6,6 +6,7 @@
 #include <vtkDoubleArray.h>
 #include <vtkPointData.h>
 #include "solver.h"
+#include "lattice.h"
 #include "dynamics_BGK.h"
 #include "dynamics_special.h"
 #include "dynamics_simple_BGK.h"
@@ -15,6 +16,7 @@
 #include <iostream>
 #include <sstream>
 #include <string>
+#include "descriptor.h"
 
 template<typename D>
 Solver<D>::Solver(Geometry * _geom,ParamsList _params_list):

### 2011-01-21 QuarterCode/lattice.h
Simplify the code in the QuarterCode to have symmetric boundary conditions. Need to find mistakes.

@@ -1,24 +1,35 @@
 #ifndef LATTICE_H
 #define LATTICE_H
 
-#include "dynamics.h"
-#include "geometry.h"
 #include <vtkXMLStructuredGridWriter.h>
 #include <vtkSmartPointer.h>
 #include <vtkStructuredGrid.h>
 #include <vtkIdList.h>
 #include <vtkDoubleArray.h>
 #include <vtkPointData.h>
+#include "params_list.h"
 
 
 template<typename D>
 class Lattice
 {
     public:
-    Lattice(int _zbegin,int _zend,int _NX,int _NY,int _NZ,int _NUMLOCAL):
+    Lattice(int _zbegin,int _zend,int _NX,int _NY,int _NZ,int _NUMLOCAL,ParamsList params):
         zbegin(_zbegin),zend(_zend),
         NX(_NX),NY(_NY),NZ(_NZ),
-        NUMLOCAL(_NUMLOCAL)
+        NUMLOCAL(_NUMLOCAL),
+        aconst(params("aconst").value<double>()),
+        gammaconst(params("gammaconst").value<double>()),
+        kconst(params("kconst").value<double>()),
+        force_x(params("force_x").value<double>()),
+        force_y(params("force_y").value<double>()),
+        force_z(params("force_z").value<double>()),
+        tau_gas(params("tau_gas").value<double>()),
+        tau_liq(params("tau_liq").value<double>()),
+        tau_phi(params("tau_phi").value<double>()),
+        rho_wall(params("rho_wall").value<double>()),
+        phase_wall(params("phase_wall").value<double>()),
+        phase_gradient(params("phase_gradient").value<double>())
     {
         //Populations
         f = new double[NUMLOCAL*D::NPOP];
@@ -36,15 +47,43 @@ class Lattice
         //Helpful materials
         phase_top = new double[NX*NY];
         phase_bottom = new double[NX*NY];
-        dynamics_list = new Dynamics<D>*[NUMLOCAL];
         layer_top_f = new double[NX*NY*D::NPOP];
 		layer_bottom_f = new double[NX*NY*D::NPOP];
 		layer_top_g = new double[NX*NY*D::NPOP];
 		layer_bottom_g = new double[NX*NY*D::NPOP];
 
-        geom_top=new NodeType[NX*NY];
-        geom_bottom=new NodeType[NX*NY];
-        geometry=new NodeType[NUMLOCAL];
+        //Symmetric populations
[…]

### 2011-01-24 Code/main.cpp
Made a comparison with another code. Need to check parallel version. Close results.

@@ -28,8 +28,8 @@
 #include "descriptor.h"
 
 //Lattice Boltzmann initialization and parameters
-const int NX=53;
-const int NY=53;
+const int NX=52;
+const int NY=52;
 const int NZ=52;
 const int NPOP=19;
 const int NUM=NX*NY*NZ;
@@ -70,7 +70,7 @@ int main(int argc,char* argv[])
     params.add("tau_gas",0.7);
     params.add("force_x",0.0);
     params.add("force_y",0.0);
-    params.add("force_z",0.0);
+    params.add("force_z",0.00003);
 
     //Specify geometry
     Geometry * geometry=new Geometry(NX,NY,NZ);

### 2011-01-24 Python/comparison.py

@@ -26,7 +26,10 @@ def read(name):
     velz_numpy =velz.reshape(dims)
     
     fig=mlab.figure()
+ 
     src = mlab.pipeline.scalar_field(phase_numpy)
+    mlab.axes()
+
     v= mlab.pipeline.vector_field(velx_numpy,vely_numpy,velz_numpy)
     vx=mlab.pipeline.scalar_field(velx_numpy)
     vy=mlab.pipeline.scalar_field(vely_numpy)
@@ -36,7 +39,6 @@ def read(name):
     extract.set(z_min=1,z_max=dims[2]-2,y_min=1,y_max=dims[1]-2)
     surf = mlab.pipeline.contour_surface(extract)
  
-    mlab.axes()
     mlab.pipeline.image_plane_widget(vx, plane_orientation='x_axes', slice_index=250)
     #mlab.pipeline.vectors(v, mask_points=20, scale_factor=3.)
     mlab.pipeline.vector_cut_plane(v, mask_points=2, scale_factor=3)
@@ -163,16 +165,22 @@ def compare(name,name_quarter):
     velz_quarter_numpy =velz_quarter.reshape(dims_quarter)
 
     pylab.figure()
-    pylab.imshow(phase_numpy[25,0:26,0:26])
+    pylab.imshow(phase_numpy[25,0:27,26:])
     pylab.figure()
-    pylab.imshow(phase_quarter_numpy[25,:,:],origin="lower")
+    pylab.imshow(phase_quarter_numpy[25,1:,1:],origin="lower")
     pylab.figure()
     pylab.plot(phase_quarter_numpy[25,:,0])
     pylab.plot(phase_quarter_numpy[25,:,1])
+    pylab.figure()
+    pylab.plot(phase_numpy[25,26:,26])
+    pylab.plot(phase_quarter_numpy[25,1:,0])
     
-    print velx_quarter_numpy[25,:10,0], velx_quarter_numpy[25,:10,1]
-    print vely_quarter_numpy[25,:10,0], vely_quarter_numpy[25,:10,1]
-    print velz_quarter_numpy[25,:10,0], velz_quarter_numpy[25,:10,1]
+    print "The difference is", phase_quarter_numpy[25,:,0]-phase_quarter_numpy[25,:,1]
+    print "Another difference is", phase_numpy[25,26:,26]-phase_quarter_numpy[25,1:,0]
+    
+    #print velx_quarter_numpy[25,:10,0], velx_quarter_numpy[25,:10,1]
[…]

### 2011-01-24 QuarterCode/lattice.hpp

@@ -550,14 +550,13 @@ void Lattice<D>::collide_stream()
     {
         for(int iX=0;iX<NX-1;iX++)
         {
-            int counter=iZ*NX*NY+(NY-1)*NX+iX;
+            int counter=iZ*NX*NY+(NY-2)*NX+iX;
             int iY=NY-2;
             for(int k=0;k<D::NPOP;k++)
             {
                 int iZ2=(iZ+D::cz[k]+zend-zbegin+1)%(zend-zbegin+1);
                 int iY2=(iY+D::cy[k]+NY)%NY;
                 int iX2=(iX+D::cx[k]+NX)%NX;
-                int counter2=iZ2*NX*NY+iY2*NX+iX2;
                 if (iY2==NY-1)
                 {
                     f2[counter*D::NPOP+D::compliment[k]]=f[counter*D::NPOP+k];
@@ -569,7 +568,7 @@ void Lattice<D>::collide_stream()
 
         for(int iY=0;iY<NY-1;iY++)
         {
-            int counter=iZ*NX*NY+iY*NX+NX-1;
+            int counter=iZ*NX*NY+iY*NX+NX-2;
             int iX=NX-2;
 
             for(int k=0;k<D::NPOP;k++)
@@ -577,7 +576,6 @@ void Lattice<D>::collide_stream()
                 int iZ2=(iZ+D::cz[k]+zend-zbegin+1)%(zend-zbegin+1);
                 int iY2=(iY+D::cy[k]+NY)%NY;
                 int iX2=(iX+D::cx[k]+NX)%NX;
-                int counter2=iZ2*NX*NY+iY2*NX+iX2;
                 if (iX2==NX-1)
                 {
                     f2[counter*D::NPOP+D::compliment[k]]=f[counter*D::NPOP+k];

### 2011-01-24 QuarterCode/main.cpp

@@ -15,8 +15,8 @@
 #include "descriptor.h"
 
 //Lattice Boltzmann initialization and parameters
-const int NX=28;
-const int NY=28;
+const int NX=27;
+const int NY=27;
 const int NZ=52;
 const int NPOP=19;
 const int NUM=NX*NY*NZ;
@@ -57,7 +57,7 @@ int main(int argc,char* argv[])
     params.add("tau_gas",0.7);
     params.add("force_x",0.0);
     params.add("force_y",0.0);
-    params.add("force_z",0.00);
+    params.add("force_z",0.00003);
     params.add("NX",NX);
     params.add("NY",NY);
     params.add("NZ",NZ);
@@ -150,6 +150,7 @@ int main(int argc,char* argv[])
  			time_string<<time_counter;
 
 			file_name<<"phase"<<std::string(5-time_string.str().size(),'0')<<time_counter;
+            //solver.writeTextWholeVelocity(file_name.str());
             solver.writeWholeDensityPhaseVelocity(file_name.str());
             cout<<"Output is done on the step "<<time_counter<<"\n";
 		}

### 2011-01-26 Text/paper.tex
Added graphs for train simulations. Something is wrong with the steady state.

@@
 most important thing is to establish different accentricity factors.
+
+\section{Train simulation}
+As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
+examine how the distance between bubbles influences on the thickness. For this purpose the
+simulation was conducted on the grid 82x82x750. We examined different initialized distances between
+bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
+lattice units. Therefore, the corresponding initial distances between bubbles are
+$450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become $386$,
+$317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based on
+the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
+calculated radiuses (diagonal radius equal to the axes radius) are as follows
+$0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
+on each other velocity for the given body force. However, in terms of the film thicknesses those
+numbers look adequate. One can see the film thicknesses dependency on the capillary number for the
+\citeauthor{heil-threedim} and bubble train simulations data in Fig. \ref{fig:capillaries:train}. 
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
+\caption{Some more information on representing \label{fig:capillaries:train}}
+\end{figure}
+
 \section{Simulation setup}

### 2011-01-26 Python/velocity.py

@@ -1,0 +1,32 @@
+#!/usr/bin/python
+
+def read(name):
+    import numpy
+    import pylab
+    
+    arr=numpy.load(name)
+    
+    dims=arr['phi'].shape
+    velx=arr['v'][0,:,:]
+    vely=arr['v'][1,:,:]
+    
+    print dims
+    
+    x,y=numpy.mgrid[0:dims[0],0:dims[1]]
+    x_short=x[::30,::150]
+    y_short=y[::30,::150]
+    velx_short=velx[::30,::150]
+    vely_short=vely[::30,::150]
+  
+    pylab.figure()
+    pylab.imshow(arr['phi'])
+    
+    pylab.figure(figsize=[31,2.5])
+    pylab.quiver(y_short,x_short,velx_short,vely_short,scale=0.1)
+    
+    pylab.show()
+
+if __name__=="__main__":
+    name="../../binary_microchannel/Sailfish/Grid/Results/202/grid400000.npz"
+    
+    read(name)

### 2011-01-26 Sailfish/Steady/govern.py

@@ -4,7 +4,6 @@ import subprocess
 import pylab
 import numpy
 import math
-import math
 from enthought.tvtk.api import tvtk
 from enthought.mayavi import mlab
 import matplotlib.pyplot as plt
@@ -66,6 +65,7 @@ def Show_Phase(phase_numpy):
     
     fig=mlab.figure()
     src = mlab.pipeline.scalar_field(phase_numpy)
+ 
     mlab.outline()
     mlab.orientation_axes()
     #v= mlab.pipeline.vector_field(velx_numpy,vely_numpy,velz_numpy)
@@ -88,9 +88,8 @@ def Analyze_Phase(name):
     a=0.04
     
     phase_numpy,velx_numpy,vely_numpy,velz_numpy=Read_Phase(name)
-    #Show_Phase_Slice(phase_numpy,175)
+    Show_Phase(phase_numpy)
     dims=phase_numpy.shape
-
 
 
     center=phase_numpy[dims[0]/2,dims[1]/2,:]
@@ -99,6 +98,8 @@ def Analyze_Phase(name):
     if z1==0:
         z2=numpy.min(numpy.where(center>0.0))+dims[2]
         z1=numpy.max(numpy.where(center>0.0))
+    pylab.figure()
+    pylab.plot(center)
     print z1,z2
     
     mid =((z1+z2)/2)%dims[2]
@@ -111,19 +112,19 @@ def Analyze_Phase(name):
     axis_zero=(1.0-Get_Zero(prof_axis)/(0.5*(shape[1]-2)))
     diag_zero=(math.sqrt(2.0)-math.sqrt(2.0)*Get_Zero(prof_diag)/(0.5*(shape[1]-2)))
     
-    #print "Velx",velx_numpy[dims[0]/2,dims[1]/2,mid]
-    #print "Vely",vely_numpy[dims[0]/2,dims[1]/2,mid]
-    #print "Velz",velz_numpy[dims[0]/2,dims[1]/2,mid]
+    print "Velx",velx_numpy[dims[0]/2,dims[1]/2,mid]
+    print "Vely",vely_numpy[dims[0]/2,dims[1]/2,mid]
+    print "Velz",velz_numpy[dims[0]/2,dims[1]/2,mid]
     
     #Calculation of the capillary number
     capillary=velx_numpy[dims[0]/2,dims[1]/2,mid]*(2.0/3.0)/math.sqrt(8.0*k*a/9.0)    
     
-    pylab.figure()
-    pylab.imshow(phase_numpy[:,:,mid])
[…]

### 2011-01-26 Sailfish/Train/govern.py

@@ -4,7 +4,23 @@ import subprocess
 import pylab
 import numpy
 import math
- 
+from enthought.tvtk.api import tvtk
+from enthought.mayavi import mlab
+
+import matplotlib.pyplot as plt
+plt.rcParams["xtick.major.size"] = 6
+plt.rcParams["xtick.minor.size"] = 4
+plt.rcParams["ytick.major.size"] = 6
+plt.rcParams["ytick.minor.size"] = 4
+
+#plt.rcParams["axes.linewidth"]=2
+#plt.rcParams["ytick.linewidth"]=2
+
+
+#xtick.major.size     : 4      # major tick size in points
+#xtick.minor.size     : 2      # minor tick size in points
+
+
 def Run_Simulations():
     print os.getcwd()
     
@@ -25,17 +41,179 @@ def Get_Zero(prof):
     for counter in range(0, len(prof)/2):
         if prof[counter]>=0 and prof[counter+1]<0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
-    return (zero-0.5)/(len(prof)-2)
+    return zero-0.5
 
-def Analyze_Simulations():
+def Read_Phase(name):
+    print os.getcwd()
+    gridreader = tvtk.XMLImageDataReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("phi"))
+    velocity=numpy.array(data.get_array("v"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
+    phase_numpy=phase.reshape(dims)
+    velx_numpy =velx.reshape(dims)
+    vely_numpy =vely.reshape(dims)
+    velz_numpy =velz.reshape(dims)
+    return phase_numpy,velx_numpy,vely_numpy,velz_numpy
+
+def Show_Phase_Slice(phase_numpy,slice):
+    fig=pylab.figure()
+    pylab.imshow(phase_numpy[:,:,slice])
+    
+    
+def Show_Phase(phase_numpy):
+    dims=phase_numpy.shape
+    
+    fig=mlab.figure()
[…]

### 2011-01-26 Text/paper.tex
Repaired the quarter code for parallel version. Right now totally coincides with the serial version.

@@
 on each other velocity for the given body force. However, in terms of the film thicknesses those
-numbers look adequate. One can see the film thicknesses dependency on the capillary number for the
-\citeauthor{heil-threedim} and bubble train simulations data in Fig. \ref{fig:capillaries:train}. 
+numbers look adequate. One can see the film thicknesses dependency on the capillary number for
+simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
+\ref{fig:capillaries:train}. One can see that the results are consistent and in a good agreement.
+Therefore, though the bubble train mutual motion influences on the film thickness by increasing the
+associated capillary number,  however, the film thickness is in accordance with the capillary
+number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
+obtain the data. However, one should be cautious about the initialization, because the Poiseuille
+flow initialization is deviated even further due to mutual motion of bubbles.
 \begin{figure}
 \includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
-\caption{Some more information on representing \label{fig:capillaries:train}}
+\caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
+simulations. Though, one can see that the decrease of the distance between bubbles causes the
+increase in the associated capillary number, however, the correlation dependance between film
+thickness and capillary number stays the same. This can help for simulations to reduce the
+numerical domain. \label{fig:capillaries:train}}
 \end{figure}

### 2011-01-26 Code/main.cpp

@@ -155,7 +155,8 @@ int main(int argc,char* argv[])
  			time_string<<time_counter;
 
 			file_name<<"phase"<<std::string(5-time_string.str().size(),'0')<<time_counter;
-            solver.writeWholeDensityPhaseVelocity(file_name.str());
+            //solver.writeWholeDensityPhaseVelocity(file_name.str());
+            solver.writeTextWholeVelocity(file_name.str());
             cout<<"Output is done on the step "<<time_counter<<"\n";
 		}
 	}

### 2011-01-26 QuarterCode/lattice.h

@@ -165,6 +165,7 @@ class Lattice
     void init();
     void updateWall();
     void updateSymmetric();
+    void updateSymmetricPopulations();
 
     //Collide and stream
     void collide_stream();

### 2011-01-26 QuarterCode/lattice.hpp

@@ -321,30 +321,54 @@ void Lattice<D>::init()
         for(int iX=0;iX<NX-1;iX++)
         {
             int counter=iZ*NX*NY+(NY-1)*NX+iX;
-            for(int k=0;k<D::NPOP;k++)
+            double sum=0.0;
+            double sum_phase=0.0;
+
+            for(int k=1;k<D::NPOP;k++)
             {
-                f[counter*D::NPOP+k]=D::weights[k]*rho[counter];
-                g[counter*D::NPOP+k]=D::weights[k]*phase[counter];
+                f[counter*D::NPOP+k]=D::weights[k]*rho[counter]/3.0;
+                g[counter*D::NPOP+k]=D::weights[k]*phase[counter]/3.0;
+                sum=sum+f[counter*D::NPOP+k];
+                sum_phase=sum_phase+g[counter*D::NPOP+k];
             }
+            f[counter*D::NPOP]=rho[counter]-sum;
+            g[counter*D::NPOP]=phase[counter]-sum_phase;
         }
 
         for(int iY=0;iY<NY-1;iY++)
         {
             int counter=iZ*NX*NY+iY*NX+NX-1;
+            double sum=0.0;
+            double sum_phase=0.0;
 
-            for(int k=0;k<D::NPOP;k++)
+            for(int k=1;k<D::NPOP;k++)
             {
-                f[counter*D::NPOP+k]=D::weights[k]*rho[counter];
-                g[counter*D::NPOP+k]=D::weights[k]*phase[counter];
+                f[counter*D::NPOP+k]=D::weights[k]*rho[counter]/3.0;
+                g[counter*D::NPOP+k]=D::weights[k]*phase[counter]/3.0;
+                sum=sum+f[counter*D::NPOP+k];
+                sum_phase=sum_phase+g[counter*D::NPOP+k];
             }
+
+            f[counter*D::NPOP]=rho[counter]-sum;
+            g[counter*D::NPOP]=phase[counter]-sum_phase;
+
         }
 
         int counter=iZ*NX*NY+(NY-1)*NX+NX-1;
-        for(int k=0;k<D::NPOP;k++)
+        double sum=0.0;
+        double sum_phase=0.0;
+
+        for(int k=1;k<D::NPOP;k++)
         {
             f[counter*D::NPOP+k]=D::weights[k]*rho[counter];
[…]

### 2011-01-26 QuarterCode/solver.h

@@ -55,6 +55,7 @@ class Solver
 
     void writeTextWholeVelocity(std::string name);
     void writeTextWholePhase(std::string name);
+    void writeTextWholeDensity(std::string name);
     void writeVTKWholePhase(std::string name);
 
     void writeTextXZPhase(std::string name);

### 2011-01-26 QuarterCode/solver.hpp

@@ -632,6 +632,23 @@ template<typename D> void Solver<D>::writeTextWholePhase(std::string name)
     }
 }
 
+template<typename D> void Solver<D>::writeTextWholeDensity(std::string name)
+{
+    Solver::getWholeDensity();
+    if (rank==0)
+    {
+    	name=name+".dat";
+        std::ofstream fout(name.c_str());
+        fout.precision(10);
+        for(int counterZ=0;counterZ<NZ;counterZ++)
+            for(int counterY=0;counterY<NY;counterY++)
+                for(int counterX=0;counterX<NX;counterX++)
+                    fout << counterX << " " << counterY << " " << counterZ << " "<<density[counterZ*NX*NY+counterY*NX+counterX]<<"\n";
+    }
+}
+
+
+
 //Write planes of phase
 template<typename D> void Solver<D>::writeTextXYPhase(std::string name)
 {
@@ -728,6 +745,7 @@ template<typename D> void Solver<D>::writeTextWholeVelocity(std::string name)
     {
     	name=name+".dat";
         std::ofstream fout(name.c_str());
+        fout.precision(12);
         for(int counterZ=0;counterZ<NZ;counterZ++)
             for(int counterY=0;counterY<NY;counterY++)
                 for(int counterX=0;counterX<NX;counterX++)

### 2011-01-26 Sailfish/Steady/govern.py

@@ -153,7 +153,32 @@ def Analyze_Consequence():
     pylab.plot(capillaries,diag_zeros)
     numpy.savetxt("steady.txt",zip(capillaries,ax_zeros,diag_zeros))
     pylab.show()
-    mlab.show()    
+    mlab.show()
+    
+def Analyze_NoForce_Consequence():
+    ax_zeros=[]
+    diag_zeros=[]
+    capillaries=[]
+    os.chdir("NoForce")
+    for i in range(1,5):
+        print os.getcwd()
+        name="steady"+str(2*i)+"00000_0.vti"
+        print name
+        axis_zero, diag_zero, capillary=Analyze_Phase(name)
+        ax_zeros.append(axis_zero)
+        diag_zeros.append(diag_zero)
+        capillaries.append(capillary)
+        #Calculate the capillary number
+        os.chdir("..")
+    pylab.plot(capillaries,ax_zeros)
+    pylab.plot(capillaries,diag_zeros)
+    #numpy.savetxt("steady.txt",zip(capillaries,ax_zeros,diag_zeros))
+    pylab.show()
+    mlab.show()
+
+
+    
+        
 
 def Draw_Consequence():
     #from numpy import genfromtxt
@@ -273,5 +298,6 @@ if __name__=="__main__":
     #Analyze_Simulations()    
     #Analyze_Velocities()
     #Analyze_Bubble()
-    Compare_Plots()
+    #Compare_Plots()
+    Analyze_NoForce_Consequence()
     #pylab.show()

### 2011-02-08 Python/streamlines.m
Found how to execute data with VTK. Still don't know how to limit vectors to the certain place easily without construction of the data. Added matlab script to produce the streamlines. Looks more informative.

@@ -1,0 +1,30 @@
+%% cleaning
+clear all
+clc
+
+%% read necessary files
+vy=dlmread('vy.txt',' ');
+vz=dlmread('vz.txt',' ');
+phase=dlmread('phase.txt',' ');
+%% draw streamlines
+
+% specifying the starting poitns
+
+%[sx,sy] = meshgrid(0:20:1500,0:10:50);
+
+sy=1:52;
+sz=ones(1,52);
+
+fig=figure()
+set(gcf,'PaperUnits','centimeters')
+xSize = 8; ySize = 12;
+%xLeft = (21-xSize)/2; yTop = (30-ySize)/2;
+%set(gcf,'PaperPosition',[xLeft yTop xSize ySize])
+set(gcf,'Position',[0 0 1500 400])
+
+%axis=gca()
+%get(axis)
+h=streamline(vz,vy,sz,sy,[0.1, 15000])
+set(h,'Color','red')
+figure()
+imshow(phase)

### 2011-02-08 Python/streamlines.py

@@ -56,6 +56,143 @@ def read_paraview(name):
     Render()
     #Show()
     Start()
+
+def read_vtk_2d(name):
+    import vtk
+    import numpy
+
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    #gridreader.SetPointArrayStatus("Density",0)
+    selection=gridreader.GetPointDataArraySelection()
+    selection.DisableArray("Density")
+    #selection.DisableArray("Velocity")
+    #selection.DisableArray("Phase")
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    #data.SetActiveScalars("Phase"); 
+    data.SetActiveVectors("Velocity")
+    velocity=data.GetArray("Velocity")
+    phase = data.GetArray("Phase")
+    
+    vel_image=vtk.vtkImageData()
+    vel_image.SetDimensions(dims[1],dims[2],1)
+    vel_image.SetSpacing(1.0, 1.0, 1.0)
+    vel_image.SetOrigin(0.0, 0.0, 0.0)
+
+    
+    Array = vtk.vtkDoubleArray()
+    Array.SetNumberOfComponents(3)
+    Array.SetNumberOfTuples(points.GetNumberOfPoints())
+    Array.Initialize()
+    Array.SetName("VelocitySlice")
+
+    ArrayPhase = vtk.vtkDoubleArray()
+    ArrayPhase.SetNumberOfComponents(1)
+    ArrayPhase.SetNumberOfTuples(points.GetNumberOfPoints())
+    ArrayPhase.Initialize()
+    ArrayPhase.SetName("PhaseSlice")
+    
+    for counter in range(0,10):
+        print points.GetPoint(counter)
+    raw_input()
+    exam=1
+    print dims[0],dims[1],dims[2]
+    #raw_input()
+    for counter in range(0,vel_image.GetNumberOfPoints()):
+        x,y,z=vel_image.GetPoint(counter)
+        counter_big=y*dims[0]*dims[1]+x*dims[0]+exam
+        #print x,y,z
+        #print counter_big
+        
+        vz=velocity.GetTuple3(counter_big)[2]
+        vy=velocity.GetTuple3(counter_big)[1]
[…]

### 2011-02-09 Text/paper.tex
Started the analysis of my files. Found mistake with the determination of diagonal issues

@@
 \section{Non-axisymmetric case}
-For example according to \cite{heil-threedim} for the capillary
-number $Ca=0.01$, the radiuses converge as $r_h=0.99$ and $r_d=1.1$. Therefore one needs to resolve
-as the $0.5\%$ of whole channel width to see the difference. That certainly implies so large grid
-that can kill the simulation at least $200x200x3000$. However, we can refer to artificially made
-interface as $1-2$ lattice Boltzmann units but keeping the capillary number as it is required. In
-this case we can refer like we can't resolve the interface. However, one can check whether
-completely underresolved interfaces can give something good with the continuous interface methods.
-The dimensionalization case is therefore as follows. Instead of starting with the capillary number
+As it was indicated before for a quite low capillary number the flow becomes axisymmetric. For
+example, according to \cite{heil-threedim} for the capillary
+number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
+Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
+difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
+ resolved around $40-50$ percent for calculations to be convergent and grid independent. This
+condition certainly implies a large grid with the computational high demand. For instance, to
+properly simulate the abovementioned example the grid should be as 
+$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
+interface whether it can produce the consistent film thicknesses for low capillary numbers. The
+interface occupies around $1-2$ lattice units.
+
+Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
+the quarter of the channel using the mirror boundary conditions (Section \ref{append:sym}). That
+results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
+out with the pressure gradient as:
+\begin{equation}
+\frac{\mathrm{d}P}{\mathrm{d}x}=1.6\,10^{-6}.
+\end{equation}
+The initial film thickness was taken as $6$ nodes.
+
+One can see the phase profile in the middle of the bubble in
+Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
+the calculation of the bubble velocity in the center of the bubble. The radiuses are as follows:
+$R_{axis}=0.99$ and $R_{diag}=1.400$. While th  
+1.4000714267
+
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/}
+\end{figure}
+
+\section{Capillary number}
+ The
+dimensionalization case is therefore as follows. Instead of starting with the capillary number
 simulations we initialize the initial thickness with $2.5$ lattice boltzmann units, change the
@@
 
+\appendix{Symmetric boundary conditions}
+\label{append:sym}
+For the memory save we performed the simulation of the quarter of the channel while using the
+antisymmetric boundary conditions. In the macroscopic variables the mirror boundary conditions
+require to have the same scalar fields, but for velocity only tangential velocities are the same
+but perpendicular have the opposite directions. This implies the following in terms of the
+macroscopic variables:
+\begin{equation}
+\rho_B = \rho_F, \quad \phi_B = \phi_F \quad U_{B\tau}=U_{F\tau}, U_{B n}=U_{F n}, 
+\end{equation}
+ where $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.
+
+In terms of the lattice Boltzmann populations this conditions require the same set of the
+populations for the boundary node as for the fluid node. However, the set is mirrored against the
+plane of the boundary. This case allows to conserve scalar fields, to have the same tangential
+velocities and to have opposite normal velocities. The grid was chosen as 52x52x1500, which is the
+effective calculated domain as 100x100x1500 grid nodes.
+
 \bibliographystyle{unsrtnat}

### 2011-02-09 Results/Capillary01/steady.py

@@ -1,0 +1,303 @@
+#!/usr/bin/python
+import os
+import subprocess
+import pylab
+import numpy
+import math
+from enthought.tvtk.api import tvtk
+from enthought.mayavi import mlab
+import matplotlib.pyplot as plt
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
+ 
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)/2):
+        if prof[counter]>=0 and prof[counter+1]<0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+def Read_Phase(name):
+    print os.getcwd()
+    gridreader = tvtk.XMLStructuredGridReader()
+    gridreader.file_name = name
+    gridreader.update()
+
+    grid  = gridreader.output
+    data  = grid.point_data
+    points= grid.points
+    dims  =grid.dimensions
+    dims=dims.tolist()
+    dims.reverse()
+    phase= numpy.array(data.get_array("Phase"))
+    velocity=numpy.array(data.get_array("Velocity"))
+    velx=velocity[:,0]
+    vely=velocity[:,1]
+    velz=velocity[:,2]
+    phase_numpy=phase.reshape(dims)
+    velx_numpy =velx.reshape(dims)
+    vely_numpy =vely.reshape(dims)
+    velz_numpy =velz.reshape(dims)
+    
+    #right way to reshape
+    #phase_numpy=numpy.zeros(dims)
+    #velx_numpy=numpy.zeros(dims)
+    #vely_numpy =numpy.zeros(dims)
+    #velz_numpy =numpy.zeros(dims)
+    
+    #for counter in range(0,phase.number_of_tuples):
+    #    x,y,z=grid.get_point(counter)
+    #    x=int(x)
+    #    y=int(y)
+    #    z=int(z)
+    #    #print x,y,z
+    #    phase_numpy[x,y,z]=phase[counter]
+    #    #print velocity[counter]
+    #    velx_numpy[x,y,z]=velocity[counter][0]
+    #    vely_numpy[x,y,z]=velocity[counter][1]
[…]

### 2011-02-10 Text/paper.tex
Added one profile. The diagonal number doesn't look too bad.

@@
 Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
-the calculation of the bubble velocity in the center of the bubble. The radiuses are as follows:
-$R_{axis}=0.99$ and $R_{diag}=1.400$. While th  
-1.4000714267
-
+the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
+are as $R_{axis}=0.93$ and $R_{diag}=1.06$. While $R_{axis}=0.93$ is underestimated parameter and
+we can't take into the consideration apriori, one can consider parameter $R_{diag}=1.06$ the
+parameters for the capillary number $0.053$ should be according to \citet{heil-threedim} as $0.97$
+and $1.05$ according to the VOF simulations of \citet{wang-non-circular}. One can conclude that
+even though the axis film thickness is not properly resolved the diagonal values are consistent and
+can give an approximate information. 
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/}
+\includegraphics[width=0.97\textwidth]{Figures/quarter.eps}\\ %\hfill
+\includegraphics[width=0.97\textwidth]{Figures/stream_cap01.eps}\\
+\caption{Streamlines and phase profile of the quarter of the channel. The capillary number is
+$0.053$ and the film thickness is underresolved. \label{fig:quarter:capillary:capillary01}}
 \end{figure}
+{\color{red} Add more data here either with better grid or with the less velocity that we can
+track the numbers.}
 
@@
 \section{Steady-state case}
+phase140000.vts
+679 1225
+952
+Shape[1]= 52
+46.8951843799
+38.0090192997
+Velx 1.83514289361e-07
+Vely 1.83514289434e-07
+Velz 0.0030119820173
+0.937903687597 1.07505741172
+Capillary= 0.0532448227311
+phase150000.vts
+693 1240
+966
+Shape[1]= 52
+46.8788857523
+37.9120765245
+Velx 1.69288494763e-07
+Vely 1.69288494735e-07
+Velz 0.00302054511694
+0.937577715046 1.07231545597
+Capillary= 0.0533961983767
+phase160000.vts
+707 1255
+981
+Shape[1]= 52
+46.8620677255
+37.8057241388
+Velx 1.52729098327e-07
+Vely 1.52729098372e-07
+Velz 0.00302874387384
+0.93724135451 1.06930735625
+Capillary= 0.0535411332917
+phase170000.vts
+721 1269
+995
+Shape[1]= 52
+46.8470974943
+37.699205304
+Velx 1.36013430193e-07
+Vely 1.36013430134e-07
+Velz 0.00303548291501
+0.936941949886 1.06629454863
+Capillary= 0.0536602638346
+phase180000.vts
+735 1284
+1009
+Shape[1]= 52
+46.8329488966
+37.5823654668
+Velx 1.20677254065e-07
+Vely 1.20677254023e-07
+Velz 0.00304140201352
+0.936658977932 1.06298981898
+Capillary= 0.0537648997019
+
 We performed different simulations to understand a number of steps required for the problem to come

### 2011-02-10 Results/Capillary01/steady.py

@@ -14,8 +14,8 @@ def Get_Zero(prof):
     zero=0
     #pylab.figure()
     #pylab.plot(prof)
-    for counter in range(0, len(prof)/2):
-        if prof[counter]>=0 and prof[counter+1]<0:
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
             zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
     return zero+0.5
 
@@ -114,8 +114,11 @@ def Analyze_Phase(name):
     prof_axis=slice[0,1:]
     prof_diag=numpy.diag(slice[1:,1:])
     
-    axis_zero=1.0-Get_Zero(prof_axis)/(shape[1]-2.0)
-    diag_zero=math.sqrt(2.0)*(1.0-Get_Zero(prof_diag)/(shape[1]-2.0))
+    print "Shape[1]=",shape[1]
+    print Get_Zero(prof_axis)
+    print Get_Zero(prof_diag)
+    axis_zero=Get_Zero(prof_axis)/(shape[1]-2.0)
+    diag_zero=math.sqrt(2.0)*Get_Zero(prof_diag)/(shape[1]-2.0)
     
     print "Velx",velx_numpy[mid,0,0]
     print "Vely",vely_numpy[mid,0,0]
@@ -132,33 +135,37 @@ def Analyze_Phase(name):
     
     pylab.figure()
     pylab.imshow(slice[1:,1:])
-    pylab.show()
+    #pylab.show()
+    numpy.savetxt("velz.dat",velz_numpy[:,:,0].transpose())
+    numpy.savetxt("vely.dat",vely_numpy[:,:,0].transpose())
+    numpy.savetxt("phase.dat",phase_numpy[:,:,0].transpose())
 
-
-    #return axis_zero,diag_zero,capillary
+    return axis_zero,diag_zero,capillary
     #mlab.show()
     
 def Analyze_Consequence():
     ax_zeros=[]
     diag_zeros=[]
     capillaries=[]
-    for i in range(1,5):
-        dir_temp=str(i)
-        os.chdir(dir_temp)
+    file_list=[]
+    for root,dirs,files in os.walk(os.getcwd()):
+        for file in files:
+            if file[0:5]=="phase":   
+                file_list.append(file)
+    for file in sorted(file_list):
         print os.getcwd()
-        name="steady"+str(2*i)+"00000_0.vti"
-        print name
[…]

### 2011-02-10 Results/Capillary01/streamlines.m

@@ -1,0 +1,56 @@
+%% cleaning
+clear all
+clc
+
+%% read necessary files
+vy=dlmread('vely.dat',' ');
+vz=dlmread('velz.dat',' ');
+phase=dlmread('phase.dat',' ');
+%% draw streamlines
+
+% specifying the starting poitns
+
+%[sx,sy] = meshgrid(0:20:1500,0:10:50);
+
+sy1=1.0/(size(vz,1))*(2:size(vz,1)-1);
+sz1=15*ones(1,size(vz,1)-2)/size(vz,2);
+
+sy2=1.0/(size(vz,1))*(2:size(vz,1)-1);
+sz2=11*ones(1,size(vz,1)-2);
+%sy=1:size(vz,1)
+%sz=ones(1,size(vz,1))
+
+%sy=[0.1 0.2]
+%sz=[1.0 1.0]
+
+fig=figure()
+
+%subplot(211)
+set(gcf,'PaperUnits','centimeters')
+xSize = 8; ySize = 12;
+set(gcf,'Position',[0 0 1500 400])
+[x,y]=meshgrid(1.0*15/size(vz,2)*(1:1500),1.0/size(vz,1)*(1:52));
+h=streamline(x,y,vz,vy,sz1,sy1,[0.1 15000]);
+hold on
+v=0.00001
+cont=contour(x,y,phase,v,'LineWidth',2,'Color','black')
+%hold on
+xlabel('X')
+ylabel('Y')
+%subplot(212)
+%axis([0,15,0,1])
+%xscale([0 15])
+%set(l,'Linewidth',3)
+%quiver(vz,vy)
+%set(axis,'XLabel','Coorx')
+%set(axis,'YLabel','Coory')
+%h=streamline(x,y,vz,vy,sz2,sy2,[0.1 15000])
+%xlim([0 15])
+%set(axis,'DataAspectRatio',[10 1 1])
+%set(h,'Color','red')
+%axes('Position',[0 15 0 1])
+%figure()
+%imshow(vy)
+legend('Streamlines','Phase contour')
+print -depsc2 -tiff streamlines.eps
+%saveas(fig,'streamlines.eps',format='eps')

### 2011-02-11 Text/paper.tex
Copied results. Next step is to obtain steady solutions, non-diagonal dependency and capillary curve.

--- 2011-02-09 Text/paper.tex
@@
 Recently we investigated the applicability and parameters choice for the two-dimensional case
-Bretherton problem using the lattice Boltzmann method (LBM) {\color{red} Give citation here}. This
+Bretherton problem using the lattice Boltzmann method (LBM) \cite{kuzmin-binary2d}. This
 paper is the continuation of the previous work and is focused on the three-dimensional case to
@@
 \end{abstract}
+
 \section{What to examine}
-I suggest to examine first of all the application of the LBM for really small capillary number,
-where it's not possible to resolve the interface. In this case one should find the asymmetric
-radiuses for the droplet, the difference should be around $10$ percent. While applying the forcing
-it should come to the spheric shape. As well we need to examine the capillary number dependancy on
-the length and suggest to examine some axicentric factors for rectangular channels.
+My plan is to investigate preferably the low capillary number regims as far we have interesting
+phenomena occuring where: the axisymmetric radiuses for axes and diagonals. As far as it's
+computationally very expensive to resolve the film thickness, it will be nice to investigate
+whether underresolved films simulations can produce consistent results. The second thing to study
+is the velocity pattern which is much easier to examine because the threshold is located for
+$Ca\approx 0.6$. The third thing to study is the film thickness dependancy for the range of
+capillary numbers. The last thing to study is the applicability of derived by \citet{heil-threedim}
+correlations for rectangular channels of different shapes.
 
@@
 certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
-The number is indicated in a number of works of as $\hat{Ca}=0.04$ \cite{cerro-bubble-train},
-$\hat{Ca}=0.1$
-\cite{cerro-space}, $\hat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger than
-the critical capillary number, i.e. $Ca>\hat{Ca}$, then the flow becomes axisimmetric with the
-radius of the droplet dependant on the capillary number.
+Non axisymmetric case is attributed to the case where the axial radius is different from the
+diagonal radius and the bubble has the non-circular shape in the channel crossection, see the
+example in Fig. \ref{fig:quarter:capillary:capillary01} for a non-axisymmetric case.
 
-The following pictures summarize the radius of deposited liquid on the capillary length.
+The transition number between non-axisymmetric case and symmetric is indicated in a number of
+works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
+$\widehat{Ca}=0.1$
+\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
+than
+the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
+radius of the droplet dependant on the capillary number. One of the example for the film
+thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+
 \citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
 square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
-square capillaries are shown in Fig.\ref{fig:heil:three:dim}. On another hand authors performed
-numerical simulations for rectangular shaped capillaries. It was found that aspect ratio $\alpha$
-allows to put all the different curves on one purely empirical curve through the following parameter
-$s_{\infty}=r_{\infty} \alpha^{-1/2}$, where $r_{\infty}$ is the aximsymmetric bubble radius
-infinitely far behind the bubble tip and can be calculated from the flow rate in the computational
-domain:
+square capillaries are shown in Fig.\ref{fig:heil:three:dim}. Also, authors performed
+numerical simulations for rectangular shaped capillaries. It was found that for different shaped
+capillaries there is an empirical correlation allowing to put all the different curves for film
+thickness dependency on the capillary number on one curve through introduction of the following
+parameter
+$s_{\infty}=r_{\infty} \alpha^{-1/2}$. The parameter $r_{\infty}$ is the aximsymmetric bubble radius
+infinitely far behind the bubble tip and can be calculated as:
 \begin{equation}
-r_{\infty}=2\sqrt{(a-Q)/\pi}.
+r_{\infty}=2\sqrt{(\alpha-Q)/\pi},
 \end{equation}
+where $Q$ is the net flow rate in the computational domain. {\color{red} What is it? - how to
+calculate? Why velocity is involved?}
+
 It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
-axisymmetric case. The $s_{\infty}$ dependency on $\alpha$ is shown in Fig.
+axisymmetric case for the range of capillary numbers. The $s_{\infty}$ dependency on $\alpha$ is
+shown in Fig.
 \ref{fig:heil:three:dim}.
@@
 capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
-number $Ca\leq\hat{Ca}=0.04$.\label{fig:heil:three:dim}}
+number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
 \end{figure}
-There is also interesting thing to track down is the recirculating regime for flow fields. It is
+
+\citet{heil-threedim} indicate the interesting occuring thing in square and rectangular channels
+as the recirculating regime for flow fields.  It is
 known that for certain capillary number the streamlines either go around the bubble or create the
-circulating regime in front of the bubble. The calculation ({\color{red} or analytical data? Look
-at White 1991, pp. 119-120. As well put figures if there is a space from Heil et al.}) are presented
-in Table \ref{table:recirculation:data}.
+circulating regime in front of the bubble, Fig. \ref{fig:heil:recirculating}.  
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/recirculating.eps}
+\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
+Streamlines in the moving frame of reference for a rectangular tube $\alpha=1.5$ (a) Ca = 0.2,
+(b) Ca = 0.5, (c) Ca = 0.55, (d) Ca = 2.0. Streamlines are shown on the surface of the bubble and
+in the plane along the shorter semi-axis.
+\label{fig:heil:recirculating}}
+\end{figure}
+The critical capillary numbers for changing velocity pattern depending on the eccentricity
+parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
 \begin{table}
@@
 \citet{shikazono-square} obtained a lot of different results for the deposition length versus the
-capillary number variation for FC-40/air, ethanol/air and water/air experiments. All of them are
+capillary number variation ethanol/air and water/air experiments. All of them are
 presented in Fig. \ref{fig:deposition:han}.
 \begin{figure}
-\includegraphics[width=0.33\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
-\includegraphics[width=0.33\textwidth]{Figures/han_capillary_water_air.eps}\hfill
-\includegraphics[width=0.33\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
-\caption{Han results. \label{fig:deposition:han}}
+%\includegraphics[width=0.47\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/han_capillary_water_air.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
+\caption{The axial and diagonal radiuses dependency on the capillary number. Courtesy of
+\citet{shikazono-square} \label{fig:deposition:han}}
 \end{figure}
@@
 \citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
-shaped capillaries as square and triangle shaped capillaries. Some interesting results for bubble
+shaped capillaries as square and triangle shaped capillaries. The VOF technique is the continuous
+interface method based on the non-uniform grid. It is of particular interest for us to compare it
+with the continuous interface method by the LBM implementation. Some interesting results are
+obtained for a bubble
 shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
@@
 \ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
-different capillary numbers. \ref{fig:wang:critical:capillary}.
+different capillary numbers.
 \begin{figure}
@@
 \caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
-velocity depending on the capillary number  \label{fig:wang:critical:capillary}}
+velocity depending on the capillary number.  \label{fig:wang:critical:capillary}}
 \end{figure}
 
+\section{Non-axisymmetric case}
+We first start with the non-axisymmetric case. As it was indicated before for a quite low capillary
+number the flow becomes axisymmetric. For
+example, according to \cite{heil-threedim} for the capillary
+number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
+Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
+difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
+ resolved around $40-50$ percent for calculations to be convergent and grid independent. This
+condition certainly implies a large grid with the computational high demand. For instance, to
+properly simulate the abovementioned example the grid should be as 
+$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
+interface whether it can produce the consistent film thicknesses for low capillary numbers. The
+interface occupies around $1-2$ lattice units.
+
+Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
+the quarter of the channel using the mirror boundary conditions (Section \ref{append:sym}). That
+results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
+out with the pressure gradient as:
+\begin{equation}
+\frac{\mathrm{d}P}{\mathrm{d}x}=1.6\,10^{-6}.
+\end{equation}
+The initial film thickness was taken as $6$ nodes.
+
+One can see the phase profile in the middle of the bubble in
+Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
+the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
+are as $R_{axis}=0.93$ and $R_{diag}=1.06$. While $R_{axis}=0.93$ is underestimated parameter and
+we can't take into the consideration apriori, one can consider parameter $R_{diag}=1.06$ the
+parameters for the capillary number $0.053$ should be according to \citet{heil-threedim} as $0.97$
+and $1.05$ according to the VOF simulations of \citet{wang-non-circular}. One can conclude that
+even though the axis film thickness is not properly resolved the diagonal values are consistent and
+can give an approximate information. 
+\begin{figure}
+\includegraphics[width=0.97\textwidth]{Figures/quarter.eps}\\ %\hfill
+\includegraphics[width=0.97\textwidth]{Figures/stream_cap01.eps}\\
+\caption{Streamlines and phase profile of the quarter of the channel. The capillary number is
+$0.053$ and the film thickness is underresolved. \label{fig:quarter:capillary:capillary01}}
+\end{figure}
+{\color{red} Add more data here either with better grid or with the less velocity that we can
+track the numbers.}
+
 \section{Variation bubble length}
-The work of \cite{heil-threedim} shows the variation of the bubble thickness through the length of
-bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten}
+The works of \citet{heil-threedim} of \citet{wang-non-circular} show the variation of the bubble
+thickness through the length of
+bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten} and
+Fig.\ref{fig:wang:bubble:shape}. One can see that it takes a certain distance for the diagonal
+radius to be equal to the axial even in the symmetric regime.
 \begin{figure}
@@
 Therefore the initalization width is $0.15*50=7.5$.
-
-\section{Non-axisymmetric case}
-As it was indicated before for a quite low capillary number the flow becomes axisymmetric. For
-example, according to \cite{heil-threedim} for the capillary
-number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
-Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
-difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
- resolved around $40-50$ percent for calculations to be convergent and grid independent. This
-condition certainly implies a large grid with the computational high demand. For instance, to
-properly simulate the abovementioned example the grid should be as 
-$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
-interface whether it can produce the consistent film thicknesses for low capillary numbers. The
-interface occupies around $1-2$ lattice units.
-
-Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
-the quarter of the channel using the mirror boundary conditions (Section \ref{append:sym}). That
-results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
-out with the pressure gradient as:
-\begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=1.6\,10^{-6}.
-\end{equation}
-The initial film thickness was taken as $6$ nodes.
-
-One can see the phase profile in the middle of the bubble in
-Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
-the calculation of the bubble velocity in the center of the bubble. The radiuses are as follows:
-$R_{axis}=0.99$ and $R_{diag}=1.400$. While th  
-1.4000714267
-
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/}
-\end{figure}
 
@@
 We performed different simulations to understand a number of steps required for the problem to come
-on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ with the initial
-width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=3
-\mathrm{x}10^{-5}$. The simulation was performed for $200000$,$400000$,$600000$ and $800000$ time
-iterations. The results are summarised in Table \ref{table:steady:state}. One of the profiles in
+on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
+quarter of the channel with the initial
+width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
+\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
+$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
+capillary number varies only in the 4th digit, radiuses are one percent accurate.
+
+One of the
+profiles in
 the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}.
@@
 $N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
-$200000$&$0.285$&$0.873$&$0.921$\\
-$400000$&$0.274$&$0.873$&$0.920$\\
-$600000$&$0.263$&$0.872$&$0.919$\\
-$800000$&$0.252$&$0.873$&$0.917$\\
+$140000$&$0.0532$&$0.9379$&$1.075$\\
+$150000$&$0.0533$&$0.9375$&$1.072$\\
+$160000$&$0.0535$&$0.9372$&$1.069$\\
+$170000$&$0.0536$&$0.9369$&$1.066$\\
+$180000$&$0.0537$&$0.9366$&$1.062$\\
 \end{tabular}

### 2011-02-11 Results/Force0000002/main.cpp

@@ -1,0 +1,167 @@
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+#include <mpi.h>
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=52;
+const int NY=52;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=10000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=6;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",atof(argv[1])*0.000002);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",2.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-02-11 Results/Force0000002x82/main.cpp

@@ -1,0 +1,168 @@
+#include <mpi.h>
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=82;
+const int NY=82;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=10000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=10;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",atof(argv[1])*0.000002);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",2.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-02-11 Results/Force0000005x52/main.cpp

@@ -1,0 +1,168 @@
+#include <mpi.h>
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=52;
+const int NY=52;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=10000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=10;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",atof(argv[1])*0.000005);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",2.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-02-12 Text/paper.tex
Added part of the results. Need to perform bunch of simulations with smaller capillary numbers and better grid.

@@
 thickness through the length of
-bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten} and
-Fig.\ref{fig:wang:bubble:shape}. One can see that it takes a certain distance for the diagonal
-radius to be equal to the axial even in the symmetric regime.
+bubble. The length variation is presented in Fig.\ref{fig:thickness:variation:ca:ten} and
+Fig.\ref{fig:wang:bubble:shape}. While Fig. \ref{fig:wang:bubble:shape} indicates that the diagonal
+radius has a variation for the rather small capillary numbers as $Ca<0.1$, another profile is
+presented for the $Ca=10$ and shows the variation for the diagonal radius. We examined a bunch of
+different capillaries numbers to present the diagonal radius variations over the length:
+The axis and diagonal radiuses variation over the bubble are shown in Figure
+\ref{fig:bubble:variation:capillaries}. The rather small capillary numbers one can see the increase
+in the rear menisci of the diagonal radius which coincides with the results of
+\citet{wang-non-circular}.
+\begin{figure}[ht]
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_36.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_61.eps}\\
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}\\
+\caption{The radiuses variations for the capillary number $Ca=0.46$.
+One can see the diagonal radius different from the axial radius near the front of the tip which is
+somehow coincides with the results of
+\citet{heil-threedim}. The results are presented for the
+capillary numbers (from left to right, top to bottom)
+$0.36$, $0.61$, $0.88$, $1.17$. \label{fig:bubble:variation:capillaries}}
+\end{figure}
 \begin{figure}
@@
 \end{figure}
-We want to obtain the same characteristics as indicated in Fig.
-\ref{fig:thickness:variation:ca:ten}. One can see that the difference between diagonal radius and
+
+Concerning results indicated in Fig.
+\ref{fig:thickness:variation:ca:ten} simulations are unstable that means one needs to work with
+small surface tension ({\color{red} See results of Pagonabarraga}). One can see that the difference
+between diagonal radius and
 usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the

### 2011-02-12 Paraview/paraview.sh

@@ -1,0 +1,5 @@
+set PVBUILD=/home/shurik/Programs/ParaView/
+export PATH=${PATH}:${PVBUILD}/bin
+export LD_LIBRARY_PATH=${LD_LIBRARY_PATH}:${PVBUILD}/bin
+export PYTHONPATH=${PVBUILD}/bin:
+export PYTHONPATH=${PYTHONPATH}:${PVBUILD}/Utilities/VTKPythonWrapping

### 2011-02-12 Python/fast_extract_profiles.py

@@ -1,0 +1,300 @@
+#!/usr/bin/python
+import vtk
+import numpy
+import matplotlib
+#matplotlib.use('GTKAgg')
+import pylab
+import math
+import matplotlib.pyplot as plt
+
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
+#plt.rcParams["text.usetex"]=True
+
+def extract_files(name):
+    #very long and expensive procedure
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[0],dims[1],dims[2]])
+    vy=numpy.zeros([dims[0],dims[1],dims[2]])
+    vx=numpy.zeros([dims[0],dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[0],dims[1],dims[2]])
+    
+    print vz.shape
+    print vy.shape
+
+    for counter in range(0,points.GetNumberOfPoints()):
+        coorx,coory,coorz=points.GetPoint(counter)
+        velx,vely,velz=velocity.GetTuple3(counter)
+        vz[int(coorx),int(coory),int(coorz)]=velz
+        vy[int(coorx),int(coory),int(coorz)]=vely
+        vx[int(coorx),int(coory),int(coorz)]=velx
+        phase_numpy[int(coorx),int(coory),int(coorz)]=phase.GetTuple1(counter)
+
+    numpy.savetxt("vz.txt",vz[0,:,:])
+    numpy.savetxt("vy.txt",vy[0,:,:])
+    numpy.savetxt("phase.txt",phase_numpy[0,:,:])
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+
+def extract_profiles(name):
+    
[…]

### 2011-02-12 Results/Force0000002/fast_extract_profiles.py

@@ -1,0 +1,440 @@
+#!/usr/bin/python
+import vtk
+import numpy
+import matplotlib
+#matplotlib.use('GTKAgg')
+import pylab
+import math
+import matplotlib.pyplot as plt
+
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
+#plt.rcParams["text.usetex"]=True
+
+def extract_files(name):
+    #very long and expensive procedure
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[0],dims[1],dims[2]])
+    vy=numpy.zeros([dims[0],dims[1],dims[2]])
+    vx=numpy.zeros([dims[0],dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[0],dims[1],dims[2]])
+    
+    print vz.shape
+    print vy.shape
+
+    for counter in range(0,points.GetNumberOfPoints()):
+        coorx,coory,coorz=points.GetPoint(counter)
+        velx,vely,velz=velocity.GetTuple3(counter)
+        vz[int(coorx),int(coory),int(coorz)]=velz
+        vy[int(coorx),int(coory),int(coorz)]=vely
+        vx[int(coorx),int(coory),int(coorz)]=velx
+        phase_numpy[int(coorx),int(coory),int(coorz)]=phase.GetTuple1(counter)
+
+    numpy.savetxt("vz.txt",vz[0,:,:])
+    numpy.savetxt("vy.txt",vy[0,:,:])
+    numpy.savetxt("phase.txt",phase_numpy[0,:,:])
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
[…]

### 2011-02-15 Text/paper.tex
Added some capillary number comparison. Need to add streamlines comparison and explanations along with steady state simulations.

@@
 \end{figure}
-{\color{red} Add more data here either with better grid or with the less velocity that we can
-track the numbers.}
+For underresolved schemes we performed a number of simulations on GPU code with the grid of size
+$50\mathrm{x}50\mathrm{750}$. While changing the applied force in a manner indicated in Appendix
+\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
+see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
+and axial radiuses are quite significant and they coincide only when the interface is resolved
+($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
+axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
+resembles already published date of \citet{heil-threedim} in a better way, see Fig.
+\ref{fig:capillary:comparison}. 
+\begin{figure}[ht]
+\includegraphics[width=0.97\textwidth]{Figures/underresolved_capillaries.eps}
+\caption{The underresolved profiles axis and diagonal radiuses versus capillary numbers. $50$
+percent resolution starts from $Ca=1.0$, which is quite high. One can see it's exactly where
+radiuses converge instead of $Ca=0.04$. {\color{red} To delete? However, one can see from
+simulations Fig.
+\ref{fig:wang:critical:capillary} and Fig.\ref{fig:thickness:variation:ca:ten} the thickness,
+especially for Fig.\ref{fig:thickness:variation:ca:ten}, can significantly variate over a bubble
+length.} \label{fig:underresolved:capillaries}}
+\end{figure}
 
@@
 in the rear menisci of the diagonal radius which coincides with the results of
-\citet{wang-non-circular}.
+\citet{wang-non-circular}. One can see that variation over the length can be neglected
+for relatively high $Ca$ numbers. 
 \begin{figure}[ht]
 \includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_36.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_61.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}\\
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_46.eps}\\
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_61.eps}\hfill
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\\
+\includegraphics[width=0.97\textwidth]{Figures/bubble_length_ca_one.eps}
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}
 \caption{The radiuses variations for the capillary number $Ca=0.46$.
@@
 capillary numbers (from left to right, top to bottom)
-$0.36$, $0.61$, $0.88$, $1.17$. \label{fig:bubble:variation:capillaries}}
+$0.36$,$0.46$, $0.61$, $0.88$,$1.06$,$1.17$. \label{fig:bubble:variation:capillaries}}
 \end{figure}
@@
 usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
-simulation is conducted by the grid $50\mathrm{x}50$ in the cross section that means that we will
-see the difference in one grid node. Therefore the grid is $52\mathrm{x}52\mathrm{750}$. Here we
-conduct a scaling:
+simulation is conducted by the grid $100\mathrm{x}100$ in the cross section that means that we will
+see the difference in two grid nodes. Therefore the grid is $102\mathrm{x}102\mathrm{1500}$. It is
+known from simulations that for the grid $102\mathrm{x}102\mathrm{x}1500$ the largest force gradient
+is $2\mathrm{x}10^{-5}$ with free energy binary liquid parameters $A=0.04$,$k=0.04$ and $Ca\approx
+1.0$. Therefore, we want to keep the same body force but to obtain larger capillary number. That
+can be done through the scaling indicated in Appendix \ref{append:scaling}:
 \begin{equation}
-Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}
+\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
 \end{equation}
-The parameters for the binary liquid problem are chosen as $A=0.04$ and $k=0.01$. Therefore the
-surface tension is:
+Therefore the same body force can stand for capillary number ten times larger if the surface
+tension is ten times smaller:
 \begin{equation}
-\sigma = \sqrt{\frac{8 k A}{9}}=0.0596
+\gamma=\sqrt{\frac{8 k A}{9}}.
 \end{equation}
-The relaxation times for liquid and gas are taken as, $\tau_{gas}=0.7$ and $\tau_{liq}=2.5$,
-therefore the viscosities ratio is $10$.
-\begin{equation}
-\begin{aligned}
-&10=\frac{\frac{2}{3} U_{bubble}}{0.0596},
-&U_{bubble}=0.894
-\end{aligned}
-\end{equation}
-which is quite high for the lattice Boltzmann method. We can decrease the ratio by scaling
-parameters $k$ and $A$ while keeping the interface thickness. Let us choose parameters as $k=0.001$
-and $A=0.004$, therefore the mobility should be adjusted to address the issue with small parameters
-and small surface tension in this case. It parameters are chosen as $A=0.004$ and $k=0.001$ then
-the velocity of the bubble is  $0.0894$. Therefore we can adjust the pressure gradient to plug it
-into the bubble initialization:
-\begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca=0.0001907,
-\end{equation}
-which is quite high for lattice Boltzmann simulations. Therefore probably one more time the surface
-tension needs to be reduced. {\color{red} Many authors state that increasing mobility kills the
+The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
+increasing the mobility of the interface to be able to adjust to the interface
+\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
+parameters and $\Gamma=20$ for the interface to be stable. 
+{\color{red} However, many authors state that increasing mobility kills
+the
 proper results of simulations}.
-The film thickness as it is seen from the picture is taken as $70$ percent of the half channel.
-Therefore the initalization width is $0.15*50=7.5$.
+{\color{red} Need to wait results from bugaboo under LargeCap simulation}
 
 \section{Capillary number}
- The
-dimensionalization case is therefore as follows. Instead of starting with the capillary number
-simulations we initialize the initial thickness with $2.5$ lattice boltzmann units, change the
-pressure gradients through the following numbers:
-\begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=
-\begin{cases}
-1\dots10\mathrm{x}10^{-6}\\
-10,15\dots 100 \mathrm{x}10^{-6}
-\end{cases}
-\end{equation}
-After performing the simulations we determine the $R_{axis}$, $R_{diag}$ versus calculated
+We performed a number of simulations with quarter CPU code and GPU train simulations and compared
+them with data obtained by \cite{heil-threedim}. After performing the simulations we determine the
+$R_{axis}$, $R_{diag}$ versus calculated
 capillary number
@@
 started with the following parameters as
-$k=0.04$,$A=0.04$,$\Gamma=1.0$,$N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$.
-For underresolved schemes, the pressure gradient doesn't work in the way we expect. Therefore, we
-take pressure gradient in the following way and later on establish the associated capillary number.
-We want to keep track of $R_{diag}$ when the $R_{axis}$ is underresolved. The preliminary results
-for the film thickness measured in the middle of the bubble are shown in Fig.
-\ref{fig:underresolved:capillaries}. 
+$k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
+The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
+quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
+grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
+length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
+chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
+comparison presented in Fig. \ref{fig:capillary:comparison}.
 \begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/underresolved_capillaries.eps}
-\caption{The underresolved profiles axis and diagonal radiuses versus capillary numbers. $50$
-percent resolution starts from $Ca=1.0$, which is quite high. One can see it's exactly where
-radiuses converge instead of $Ca=0.04$. However, one can see from simulations Fig.
-\ref{fig:wang:critical:capillary} and Fig.\ref{fig:thickness:variation:ca:ten} the thickness,
-especially for Fig.\ref{fig:thickness:variation:ca:ten}, can significantly variate over a bubble
-length. \label{fig:underresolved:capillaries}}
-\end{figure}
-
-One can see the variation of thickness over the bubble length for the given capillary $Ca=1.06$ in
-Fig. \ref{fig:variation:bubble:ca:one}. One can see that variation over the length can be neglected
-for relatively high $Ca$ numbers.
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/bubble_length_ca_one.eps}
-\caption{The bubble length variation over for given $Ca=1.06$.
-One can see that there is no such high variation over the bubble
-length. \label{fig:variation:bubble:ca:one}}
+\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison.eps}
+\caption{The comparison between different runs of codes for the axial and the diagonal radiuses
+versus capillary numbers. \ref{fig:capillary:comparison}}. One can see that the code mimics
+behavior.
 \end{figure}
@@
 \begin{tabular}{|c|c|c|c|}
+\hline
 $N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
@@
 $180000$&$0.0537$&$0.9366$&$1.062$\\
+\hline
 \end{tabular}
@@
 \section{Train simulation}
+\label{section:bubble:train}
 As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
@@
 
+\appendix{Scaling procedure}
+\label{append:scaling}
+For scaling purposes one need to obtain setup a body force as $\frac{\mathrm{d}P}{\mathrm{d}x}$. If
+one obtained a certain velocity of the bubble with the associated gradient then another gradient
+can be obtained through the porpotionality condition as:
+\begin{equation}
+\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca,
+\end{equation}
+or in other words th functionality can be extended as:
+\begin{equation}
+\dfrac{\dfrac{\mathrm{d}P_1}{\mathrm{d}x}}{\dfrac{\mathrm{d}P_2}{\mathrm{d}x}}=\frac{Ca_1}{Ca_2}
+\frac{N_{
+y2}^2 }{N_{y1}^2}
+\end{equation}
+
 \bibliographystyle{unsrtnat}

### 2011-02-15 Results/Force0000002x82/fast_extract_profiles.py

@@ -1,0 +1,440 @@
+#!/usr/bin/python
+import vtk
+import numpy
+import matplotlib
+#matplotlib.use('GTKAgg')
+import pylab
+import math
+import matplotlib.pyplot as plt
+
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
+#plt.rcParams["text.usetex"]=True
+
+def extract_files(name):
+    #very long and expensive procedure
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[0],dims[1],dims[2]])
+    vy=numpy.zeros([dims[0],dims[1],dims[2]])
+    vx=numpy.zeros([dims[0],dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[0],dims[1],dims[2]])
+    
+    print vz.shape
+    print vy.shape
+
+    for counter in range(0,points.GetNumberOfPoints()):
+        coorx,coory,coorz=points.GetPoint(counter)
+        velx,vely,velz=velocity.GetTuple3(counter)
+        vz[int(coorx),int(coory),int(coorz)]=velz
+        vy[int(coorx),int(coory),int(coorz)]=vely
+        vx[int(coorx),int(coory),int(coorz)]=velx
+        phase_numpy[int(coorx),int(coory),int(coorz)]=phase.GetTuple1(counter)
+
+    numpy.savetxt("vz.txt",vz[0,:,:])
+    numpy.savetxt("vy.txt",vy[0,:,:])
+    numpy.savetxt("phase.txt",phase_numpy[0,:,:])
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
[…]

### 2011-02-15 Results/Force0000005x52/fast_extract_profiles.py

@@ -1,0 +1,440 @@
+#!/usr/bin/python
+import vtk
+import numpy
+import matplotlib
+#matplotlib.use('GTKAgg')
+import pylab
+import math
+import matplotlib.pyplot as plt
+
+plt.rcParams["xtick.major.pad"] = 10
+plt.rcParams["ytick.major.pad"] = 10
+#plt.rcParams["text.usetex"]=True
+
+def extract_files(name):
+    #very long and expensive procedure
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[0],dims[1],dims[2]])
+    vy=numpy.zeros([dims[0],dims[1],dims[2]])
+    vx=numpy.zeros([dims[0],dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[0],dims[1],dims[2]])
+    
+    print vz.shape
+    print vy.shape
+
+    for counter in range(0,points.GetNumberOfPoints()):
+        coorx,coory,coorz=points.GetPoint(counter)
+        velx,vely,velz=velocity.GetTuple3(counter)
+        vz[int(coorx),int(coory),int(coorz)]=velz
+        vy[int(coorx),int(coory),int(coorz)]=vely
+        vx[int(coorx),int(coory),int(coorz)]=velx
+        phase_numpy[int(coorx),int(coory),int(coorz)]=phase.GetTuple1(counter)
+
+    numpy.savetxt("vz.txt",vz[0,:,:])
+    numpy.savetxt("vy.txt",vy[0,:,:])
+    numpy.savetxt("phase.txt",phase_numpy[0,:,:])
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
[…]

### 2011-02-15 Results/capillaries.py

@@ -1,0 +1,109 @@
+#!/usr/bin/python
+
+import numpy
+import os
+import pylab
+from numpy import genfromtxt
+ 
+def draw_capillaries():
+    dirs=["Force0000002","Force0000002x82","Force0000005x52"]
+    style=["<","^",">"]
+    fig=pylab.figure()
+
+    for counter,dir in enumerate(dirs):
+        os.chdir(dir)
+        print os.getcwd()
+        
+        mat=numpy.loadtxt("capillaries.txt")
+        
+        #print mat
+        pylab.plot(mat[:,0],0.5*(mat[:,1]+mat[:,2]),style[counter],markersize=10,linewidth=2)
+        #pylab.plot(mat[:,0],mat[:,2],"+")
+        os.chdir("..")
+
+    
+    capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
+    radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+   
+   
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    heil=genfromtxt("heil.csv",delimiter=';',dtype=float)
+
+    #print heil
+    #print heil.shape
+    
+    ax=fig.add_subplot(111)
+    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    
+    pylab.semilogx(heil[:,0],heil[:,1],linewidth=2)
+    pylab.semilogx(capillary_theor,radiusses,'o-',markersize=7,linewidth=2)
+    #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
+    #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
+    #pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
+
+    pylab.xlim(xmin=0.1,xmax=5)
+    #pylab.ylim(ymin=0.01)
+    #numpy.savetxt("capillary.dat",zip(capillaries,widths))
+    
+    pylab.xticks(fontsize=20)
[…]

### 2011-02-15 Text/paper.tex
Small writeup is finished. Need to perform simulations with smaller grids.

@@
 
+\section{Steady-state case}
+We performed different simulations to understand a number of steps required for the problem to come
+on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
+quarter of the channel with the initial
+width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
+\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
+$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
+capillary number varies only in the 4th digit, radiuses are one percent accurate.
+
+As well we performed longer simulations in the range of moderate capillary numbers after
+$140000$, $160000$, $180000$, $200000$, $220000$, $240000$. The results for the capillary number
+based on the bubble velocity in the middle crossection are shown in Table \ref{table:steady:state}. 
+%{\color{red} What is it One of the
+%profiles in
+%the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}. }
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
+$140000$&$0.0532$&$0.9379$&$1.075$\\
+$150000$&$0.0533$&$0.9375$&$1.072$\\
+$160000$&$0.0535$&$0.9372$&$1.069$\\
+$170000$&$0.0536$&$0.9369$&$1.066$\\
+$180000$&$0.0537$&$0.9366$&$1.062$\\
+\hline
+\end{tabular}\\
+\begin{tabular}{|c|c|c|c|}
+\hline
+$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
+$140000$&$0.3578$&$0.8657$&$0.8708$\\
+$160000$&$0.3590$&$0.8650$&$0.8702$\\
+$180000$&$0.3598$&$0.8647$&$0.8700$\\
+$200000$&$0.3603$&$0.8645$&$0.8698$\\
+$220000$&$0.3607$&$0.8644$&$0.8697$\\
+$240000$&$0.3609$&$0.8645$&$ 0.8697$\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Underresolved capillary number results and resolved Results for the steady-state case. One
+can see that $200000$ steps are enough to structure
+the bubble. \label{table:steady:state}}
+\end{table}
+
+%\begin{figure}
+%\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
+%\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
+%\caption{Slice in the middle of the bubble and
+%three-dimensional example. \label{fig:steady:state:profile:example}}
+%\end{figure}
+
 \section{Non-axisymmetric case}
@@
 Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
-the quarter of the channel using the mirror boundary conditions (Section \ref{append:sym}). That
+the quarter of the channel using the mirror boundary conditions (See Appendix \ref{append:sym}).
+That
 results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
@@
 One can see the phase profile in the middle of the bubble in
-Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
+Fig.\ref{fig:quarter:capillary:capillary01}. For comparison reason the axisymmetric phase profile
+for $Ca=0.6$ is presented in Fig. \ref{fig:bubble:axisymmetric}. The associated capillary number is
+$0.053$ based on
 the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
@@
 \end{figure}
-For underresolved schemes we performed a number of simulations on GPU code with the grid of size
+\begin{figure}[ht]
+\includegraphics[width=0.97\textwidth]{Figures/square.eps}
+\caption{The middle bubble phase profile. One can see that the diagonal radius coincides with the
+axial radius. {\color{red} $Ca=0.5$} \label{fig:bubble:axisymmetric}}
+\end{figure}
+
+For underresolved schemes we performed a number of simulations on a GPU code with the grid of size
 $50\mathrm{x}50\mathrm{750}$. While changing the applied force in a manner indicated in Appendix
@@
 length.} \label{fig:underresolved:capillaries}}
+\end{figure}
+
+\section{Velocity pattern}
+We examined two different velocity patterns to identify change of pattern. Two capillaries numbers
+were taken below and above the numbers indicated in Table \ref{table:recirculation:data}, as
+$Ca=0.36$ and $Ca=0.88$.Velocity streamlines are
+indicated in Fig. \ref{fig:streamlines:pattern}. One can see that the vortex indicated in
+Fig.\ref{fig:heil:recirculating} doesn't appear in present calculations. This fact can be
+explained because the simulation is carried our for the binary-liquid case but not free-surface
+interface presented in the classical Bretherton problem. In the latter case the void phase pushes
+the liquid towards the walls. In the present simulation there is no a void phase. Then the
+streamlines have the ability to go through the interface and we exactly see this situation occuring.
+\begin{figure}[ht]
+\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca036.eps}\\
+\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca088.eps}
+\caption{Streamlines for $Ca=0.36$ and $Ca=0.88$. One can see that there are no vortexes created
+before the bubble. It can be explained by the fact that the binary-liquid model is simulated
+instead of free-surface interface gas air finger flow.\label{fig:streamlines:pattern}}
 \end{figure}
@@
 \includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\\
-\includegraphics[width=0.97\textwidth]{Figures/bubble_length_ca_one.eps}
+\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_one.eps}
 \includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}
@@
 proper results of simulations}.
-{\color{red} Need to wait results from bugaboo under LargeCap simulation}
+{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}
 
@@
 \caption{The comparison between different runs of codes for the axial and the diagonal radiuses
-versus capillary numbers. \ref{fig:capillary:comparison}}. One can see that the code mimics
-behavior.
+versus capillary numbers. \label{fig:capillary:comparison}}. One can see that the code mimics
+behavior of the earlier published results.
 \end{figure}
-
-\section{Steady-state case}
-We performed different simulations to understand a number of steps required for the problem to come
-on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
-quarter of the channel with the initial
-width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
-$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
-capillary number varies only in the 4th digit, radiuses are one percent accurate.
-
-One of the
-profiles in
-the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}.
-\begin{table}
-\begin{tabular}{|c|c|c|c|}
-\hline
-$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
-$140000$&$0.0532$&$0.9379$&$1.075$\\
-$150000$&$0.0533$&$0.9375$&$1.072$\\
-$160000$&$0.0535$&$0.9372$&$1.069$\\
-$170000$&$0.0536$&$0.9369$&$1.066$\\
-$180000$&$0.0537$&$0.9366$&$1.062$\\
-\hline
-\end{tabular}
-\caption{Results for the steady-state case. One can see that $200000$ steps are enough to structure
-the bubble. However, the interface is underresolved. That could be the influence of the bubble
-length.\label{table:steady:state}}
-\end{table}
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
-\caption{Slice in the middle of the bubble and
-three-dimensional example. \label{fig:steady:state:profile:example}}
-\end{figure}
-
-To understand the influence of the steady-state case
-For the steady-state regime the following parameters were obtained
-We want to perform understanding of the steady state regime. As well probably it's good idea to
-change domain as well, to reduce the length but to insure enough length of the bubble. The
-steady-state calculations are performed on $52\mathrm{x}52\mathrm{x}750$ with large enough the
-interface.
-
-\section{Capillary number region}
-The same as in the previous example can help us to identify the necessary capillary regime. The
-most important thing is to establish different accentricity factors.
 
@@
 simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
-\ref{fig:capillaries:train}. One can see that the results are consistent and in a good agreement.
+\ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
+results are consistent and in a good agreement.
 Therefore, though the bubble train mutual motion influences on the film thickness by increasing the
@@
 
-\section{Simulation setup}
+\section{Conclusion}
+The present work conducts the binary-liquid simulations with the lattice Boltzman method. The
+method is a continuous interface method. Along with similarity of results with already published
+results of \citet{heil-threedim} and \citet{wang-non-circular}, the differencies were found,
+especially in the velocity pattern. Results evedince that the lattice Boltzmann binary liquids
+simulations can be used for simulation of gas bubbles in the microchannel.
 
-\appendix{Symmetric boundary conditions}
+\appendix
+\section{Symmetric boundary conditions}
 \label{append:sym}
@@
 
-\appendix{Scaling procedure}
+\section{Scaling procedure}
 \label{append:scaling}

### 2011-02-15 Results/capillaries.py

@@ -4,6 +4,8 @@ import numpy
 import os
 import pylab
 from numpy import genfromtxt
+import vtk
+import math
  
 def draw_capillaries():
     dirs=["Force0000002","Force0000002x82","Force0000005x52"]
@@ -104,6 +106,165 @@ def draw_capillaries():
     pylab.savefig("capillaries_comparison.eps",format="EPS",dpi=300)
     pylab.show()
 
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+def Analyze_Consequence():
+    ax_zeros=[]
+    diag_zeros=[]
+    capillaries=[]
+    file_list=[]
+    os.chdir("Force0000002/4")
+    for root,dirs,files in os.walk(os.getcwd()):
+        for file in files:
+            if file[0:5]=="phase":   
+                file_list.append(file)
+    for file in sorted(file_list):
+        print os.getcwd()
+        print file
+        axis_zero, diag_zero, capillary=extract_profiles(file)
+        ax_zeros.append(axis_zero)
+        diag_zeros.append(diag_zero)
+        capillaries.append(capillary)
+        #Calculate the capillary number
+        #os.chdir("..")
+    #pylab.plot(capillaries,ax_zeros)
+    #pylab.plot(capillaries,diag_zeros)
+    #numpy.savetxt("steady.txt",zip(capillaries,ax_zeros,diag_zeros))
+    pylab.show()
+    #mlab.show()
+
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[1],dims[2]])
+    vy=numpy.zeros([dims[1],dims[2]])
[…]

### 2011-02-15 Results/streamlines.m

@@ -1,0 +1,56 @@
+%% cleaning
+clear all
+clc
+
+%% draw streamlines
+
+% specifying the starting poitns
+
+%[sx,sy] = meshgrid(0:20:1500,0:10:50);
+
+names={'Force0000002/4/','Force0000002/8/'}
+
+for counter=1:2  
+     %% read necessary files
+     filenamey=strcat(names{counter},'vy.txt')
+     filenamez=strcat(names{counter},'vz.txt')
+     filenamephase=strcat(names{counter},'phase.txt')
+     vy=dlmread(filenamey,' ');
+     vz=dlmread(filenamez,' ');
+     phase=dlmread(filenamephase,' ');
+     size(vy)
+     size(vz)
+     size(phase)
+     %sy=1:52;
+     %sz=ones(1,52);
+     
+     sy=1.0/(2*size(vz,1))*(2:2*size(vz,1)-1); 
+     sz=15*ones(1,2*size(vz,1)-2)/size(vz,2);
+
+     
+     fig=figure()
+     set(gcf,'PaperUnits','centimeters')
+     xSize = 8; ySize = 12;
+     %xLeft = (21-xSize)/2; yTop = (30-ySize)/2;
+     %set(gcf,'PaperPosition',[xLeft yTop xSize ySize])
+     set(gcf,'Position',[0 0 1500 400])
+
+     %axis=gca()
+     %get(axis)
+     %% Visualization part
+     [x,y]=meshgrid(1.0*15/size(vz,2)*(1:1500),1.0/size(vz,1)*(1:52));
+     h=streamline(x,y,vz,vy,sz,sy,[0.1, 15000])
+     set(h,'Color','red')
+     hold on
+     v=0.00001
+     cont=contour(x,y,phase,v,'LineWidth',2,'Color','black')
+     %hold on
+     xlabel('X') 
+     ylabel('Y')
+
+     %figure()
+     %imshow(phase)
+
+
+end
+

### 2011-03-16 Conference/CFD2011/Template/ExampleFile.tex
Added the CFD2011 initial paper.

\documentclass{CFD2011}
\usepackage{CFD2011}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This template is created for complying with the author instructions for the 
%8th International Conference on CFD in Oil & Gas, Metallurgical and Process Industries
%hosted by SINTEF/NTNU, Trondheim Norway
%21-23 June 2011
%
%Sverre G. Johnsen (sverre.g.johnsen@sintef.no)
%SINTEF Materials and Chemistry

%Required files:
%   ExampleFile.tex
%   xampleFile.nls
%   CFD2011.bst
%   CFD2011.cls
%   CFD2011.sty
%   References.bib

%Example-specific files:
%   figure.eps
%   Table.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Industrial Applications of CFD}
\paperID{CFD11-52}
\author{Ola}{Nordmann} %{forename}{surname}
\presenting  %the previous author is presenting the paper (name becomes underlined)
\address{SINTEF Materials and Chemistry, 7465 Trondheim, NORWAY}%affiliation of the previous author
\email{ola.nordmann@sintef.no}%e-mail address of the previous author
\author{Zhi}{L. Xie}
\address{NTNU Department of Physics, 7491 Trondheim, NORWAY}
\email{lin.xie@ntnu.no}
%\author{Sverre}{G. Johnsen}
%\address{SINTEF Materials and Chemistry, 7465 Trondheim, NORWAY}
%\email{sverre.g.johnsen@sintef.no}

\begin{document}
\maketitle  %create the title page
\headers   %create the page headers and footers

\abstract{
  This file is an example \LaTeX file for submission to CFD2011.  A
  limit of 10 pages applies.
}
\keywords{
  CFD, hydrodynamics, chemical reactors
}
\normalfont\normalsize

%\section{Nomenclature}
A complete list of symbols used, with dimensions, is required.
%the nomenclature needs to be entered into the file ExampleFile.nls
\printnomenclature[0.7cm]  
\vskip .1em

\section{Introduction}
The introduction goes here.

\newpage

\section{Model Description}
You should give a thorough description of your model.
\vspace{4cm}
\subsection{Example of Subheading}
Here is how to produce a numbered equation under a second level
heading \cite{James1988}.
\vspace{2cm}\\
\emph{Continuity equation}
\begin{equation}
  \frac{\partial \rho_G}{\partial t}+\nabla\left(\rho_G\mathbf
    u\right)=0
\end{equation}

\subsubsection*{Example of Sub-subheading}
This is how \cite{Luke1988} produced an unnumbered equation under a
third level heading.
\begin{equation}
  \mathbf J=\sigma(\mathbf E+\mathbf u\times\mathbf B)
\end{equation}
\newpage

\InsFig{figure}{Schematic diagram of geometry.}{label1}
\InsFigRot{figure}{Rotated schematic diagram of geometry.}{label2}

\section{Results}
The results of using the \LaTeX template is a great looking paper.
In Figures \ref{fig:label1} and \ref{fig:label2} it can be seen how figures are easily included.
In Table \ref{tab:label} it is seen how we can include a table.
The table is constructed in the file table.tex, where also the table caption and label are defined.

\newpage
\InsTab{Table}

\section{Conclusion}
The conclusions are:
\begin{enumerate}
  \item Trondheim is a nice city.
  \item CFD is great fun, and useful too.
\end{enumerate}

\newpage

%% %---------------------------
%% %BIBLIOGRAPHY
%% %---------------------------
%The bibliography is created using BiBTeX
\bibliographystyle{CFD2011}
\bibliography{References}

\newpage
\section{Appendix A}
List of animation files:
s10.avi Motion of the spherical particles.
v10.mped Flow field around each particle.

\end{document}

### 2011-03-16 Conference/CFD2011/paper.tex
Added the CFD2011 initial paper.

\documentclass{CFD2011}
\usepackage{CFD2011}
\usepackage{tabularx}
\usepackage{color}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This template is created for complying with the author instructions for the 
%8th International Conference on CFD in Oil & Gas, Metallurgical and Process Industries
%hosted by SINTEF/NTNU, Trondheim Norway
%21-23 June 2011
%
%Sverre G. Johnsen (sverre.g.johnsen@sintef.no)
%SINTEF Materials and Chemistry

%Required files:
%   ExampleFile.tex
%   xampleFile.nls
%   CFD2011.bst
%   CFD2011.cls
%   CFD2011.sty
%   References.bib

%Example-specific files:
%   figure.eps
%   Table.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\Ca}{\mathrm{Ca}}

\title{Multiphase microchannel flows simulations via the lattice Boltzmann scheme}
\paperID{CFD11-52}
\author{Alexandr}{Kuzmin} %{forename}{surname}
\presenting  %the previous author is presenting the paper (name becomes underlined)
\email{shurik.kuzmin@gmail.com}%e-mail address of the previous author
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author
\author{Dmitry}{Eskin}
\email{deskin@slb.com}
\address{DBR Schlumberger Technology Center, 9450 17 Ave NW, T6N 1M9 Edmonton,CANADA}
\author{Jos}{Derksen}
%\address{NTNU Department of Physics, 7491 Trondheim, CANADA}
\email{jos@ualberta.ca}
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author

\begin{document}
\maketitle  %create the title page
\headers   %create the page headers and footers

\abstract{
In this work a free-energy binary liquid lattice Boltzmann scheme is used to simulate
Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
thin liquid films between the bubbles and the channel walls. The main focus of our work is a
feasibility study of the binary-liquid model lattice Botlzmann model to simulate the
Bretherton/Taylor phenomena. The numerical scheme has a diffuse
interface, and we thoroughly assess resolution requirements for correctly resolving
the liquid film and bubble motion. The simulations are two-dimensional and three-dimensional and
span a capillary number
range of $0.03$ to $0.8$ (for 2D) and range of $0.1$ to $1.0$ where the capillary number is based
on the liquid's viscosity, the
velocity of the bubble's interface, and the interfacial tension. The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. We obtain grid independent
results
as long as the liquid film thickness is at least two times the width of the diffuse interface. The
results show that the lattice Boltzmann free energy binary liquid model can be used to simulate the
Bretherton problem with a good accuracy.
}
\keywords{CFD, Multiphase Flows, Lattice Boltzmann Method, Microchannel simulations, Bretherton
Problem, Binary liquid, Free energy method}
\normalfont\normalsize

%\section{Nomenclature}
%A complete list of symbols used, with dimensions, is required.
%the nomenclature needs to be entered into the file ExampleFile.nls
\printnomenclature[0.7cm]  
\vskip .1em

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. Depending on the geometry it was found that the deposited film thickness
is a complicated function of the capillary number $Ca$, see Eq. (\ref{capillary:number:definition}).
For example, the deposited film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers for circular channels
\cite{bretherton,heil-bretherton}. 
\vskip .4em
The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
is of significant interests for chemical industry as far as it is widely used in chemical monolith
microreactors \cite{kreutzer-pressure-drop}. While it
is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
not possible to extrapolate it to the wider range of capillary numbers used in the chemical
industry. Thus, the problem of consistent numerical simulations arises.
\vskip .4em
For what follows we distinguish two- and three-dimensional cases and give a detailed description
of the literature review. 
\begin{description}
\item[2D case]
Bretherton made analytical assymptotic analysis and predicted the $\Ca^{2/3}$ dependency of the film
thickness for the circular shaped microchannels in the range of the small capillary numbers
(typically $\Ca<0.003$), where capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
\Ca=\frac{\mu_{\mathrm{liq}} U_\mathrm{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
$\gamma$ is the interfacial tension between gas and liquid. For the case of the continuous
interface models,  $U_{bubble}$ is the interface velocity.

Later on, it was realized
\cite{wong-films,wong-pressure} that the film thickness
is proportional to $\Ca^{2/3}$ only in the certain region behind the front meniscus. Overall, the
film thickness
varies over the bubble length for bubbles of finite or infinite length
\cite{heil-threedim}. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule and Reynolds number 
effects on the film thickness for capillary numbers larger than $0.003$.

To consistently  predict a flow pattern for capillaries in
different ranges of the capillary number, simulations have been conducted. The simulations are
typically validated with analytical solutions in the range of the small capillary numbers
 and then extende to more practical regimes.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} also used the finite volume method to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. 
\citet{heil-bretherton} and \citet{ingham-plates} studied gas finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using the finite element method. \citet{giavedoni-numerical} performed cross validation of
the
finite element solution with previously published results.
The solutions were obtained for circular and planar cases. 
%The planar case results for the film
%thickness dependency on the capillary number are indicated in Fig. \ref{fig:giavedoni:planar}.
%\begin{figure}[ht]
%\includegraphics[width=0.5\textwidth]{Figures/giavedoni_planar.eps}
%\caption{\citet{giavedoni-numerical} gathered results across the
%literature for different capillary numbers (reprinted from \cite{giavedoni-numerical}). Results are
%scaled to the half-width of the
%channel. $h_{\infty}$ is the film thickness of gas finger at
%the infinite distance from the front meniscus. \label{fig:giavedoni:planar}}
%\end{figure}

\item[3D case]
In comparison with the two-dimensional Bretherton flow there is a vast number of experimental
results available for the three-dimensional case. For instance, \citet{shikazono-square} obtained
experimental
results for the deposition length dependency on the
capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
shaped mirochannels. \citet{cerro-bubble-train} performed a lot of experiments for a buble-train
flow in capillaries of
circular and square cross section for horizontal, upward and downward flows. The interested reader
is referred to these works for comprehensive experimental correlations.

The experimental works supported by numerical simulations reveal interesting phenomena happening in
three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for the
certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case,
Fig. \ref{fig:sym:antisym}.
From herein, we limit ourselves to the case of microchannels with square crosssections.
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like as the rounded square. The dependance
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the film
thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} also indicated another interesting phenomena. They indicated the transition
capillary number where the streamlines pattern
changes from having vortex in front of the bubble and not having it for larger capilllary number,
Fig. \ref{fig:streamlines:pattern}.
For the square channel the critical number is $Ca=0.691$.   
%\begin{figure}
%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
%(bottom). 
%\label{fig:heil:recirculating}}
%\end{figure}

There are also a number of numerical works for the three-dimensional case. For instance, 
\citet{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes for different
slug and bubble cross sections and menisci appearance.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
square and rectangular shaped capillaries. Also, they investigated gravitational effects for
circular shaped capillaries. They also found the empirical correlation which allows to put
different  aspect ratio rectangular channels simulation results on one
curve. \citet{wang-non-circular} performed numerical simulations by the VOF technique for the
non-circular
shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
interface method based on the non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation. 
They also performed
numerical simulations for the determination of the critical capillary number $\hat{Ca}$.
\end{description}
Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
are performed with the help of the boundary value approach
\cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
applicability for
problems
involving complex geometries, free interface motion, or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kind of simulations.
However, if the interface is smeared out over several grid nodes, the question of
proper film resolution in comparison with the interface resolution arises.  Therefore, the
{\emph goal} of this work is to reproduce the numerical results and phenomena presented above
with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We do
a particular focus on exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers. A thorough
parametric study for this approach and the ability of the
method to reconstruct consistent with other methods results has not been done to the best of the
authors' knowledge.
\vskip .5em
The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and allows incorporation of
physical phenomena on the microscopic level, as in the case of multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film.
\vskip .5em
The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects can be 
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
numbers.
Therefore, the major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. The goal of this work is to do a feasibility assessment study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for th Bretherton/Taylor
problem. The work results are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton,heil-threedim}.
\vskip .5em
One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for a relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}.
\vskip .5em

The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model. Then, the preliminary results for the
two-dimensional and for the three-dimensional case are presented in the results section. The paper
is
concluded with a summary of the main findings.

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.

The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and the velocity fields, and another to represent the
phase field $\phi$ indicating gas or liquid.
The equilibrium populations \cite{pooley-contact} are defined as:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, i=1\div Q-1\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div Q-1\\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependent on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
to change viscosity from the gas viscosity
$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
A}{9}}$.

\section{Numerical benchmark}
To be able to properly simulate and compare simulation results with data
published in the literature, one needs to design a numerical benchmark addressing
all the challenges for the lattice Boltzmann method. That includes the diffuse nature of the
interface which needs to have a negligible effect on the results. Also, as far as the density is
uniform, the binary liquid model can address only different viscosities of gas and liquid.
Another challenge is boundary conditions because one needs to couple inlet and outlet
boundary conditions \cite{heil-threedim} for the film thickness to establish.  The geometry of
the problem is also of concern. Most literature numerical simulations are done for a channel with
circular cross section,
which is quite difficult to address in terms of lattice Boltzmann framework. We thoroughly discuss
all the
challenges associated with the lattice Boltzmann microchannel simulations in our recent work
\cite{kuzmin-binary2d}.

Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow between plates and three-dimensional square
shaped microchannels, both driven by a body force. We limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. This
implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
periodic boundary
conditions are applied, not a single bubble but a bubble train is simulated. In this case one
needs
to ensure that the distance between bubbles is large enough to exclude mutual bubble influence. 
Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
to be sufficient to give results consistent with the theory. The
channel length is taken to be $3$ bubble lengths (or $15$ channel heights)
to minimize influence of one bubble on another,
because of the periodicity of boundary conditions. The geometry
dimensions are represented in Fig. \ref{fig:benchmark:sketch}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/benchmark.eps}
\caption{The two-dimensional benchmark sketch. $\delta$ corresponds to film thickness. $5\xi$
corresponds to the interface thickness. We take the microchannel length to be $15$ times larger than
its height. The threedimensional benchmark has the same geometry extending one dimension to have a
square shaped crosssection.
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{2D results}
Two-dimensional binary-liquid simulations are presented in our recent work
\cite{kuzmin-binary2d}. Here we outline two most important results. The first one addresses the
question of how the interface should be resolved to obtain grid independent result. Though the
simulations are performed for the lattice Boltzmann model, the results can be applied for any
uniform grid continuous interface method. The second part of results are comparisons with the
literature results for the film thickness dependency on the capillary number.

\subsubsection{Grid refinement}
To properly estimate the interface resolution one needs to study the convergence as a function of
the grid resolution. To do that the grid resolution is varied while all remaining parameters,
including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
ratio of the interface thickness to the 
film thickness at which results are no longer dependent on the grid resolution.

We performed a number of simulations for grids indicated in
Table \ref{table:parameters:grid:refinement}. Other parameters are defined as $k=0.04$, $A=0.04$ and
$\Gamma=1.0$. Note that in the case of the half-way bounce-back walls \cite{yu} which are used in
the
simulations one needs to calculate the film thickness as:
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2)=(\phi_0-0.5)/H_{\mathrm{eff}},
\end{equation}
where $\phi_0$ is the grid coordinate where phase field is $0$, $H_{\mathrm{eff}}=N_y-2$
is the effective channel height.
If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
which represent
the physical domain, top and bottom nodes are bounce-back nodes. Then the effective wall location is
in the middle between bounce-back and fluid
nodes. Overall there are $N_y-2$ nodes representing the fluid. Note that it is a big
simplification to impose the boundary in the middle between the bounce-back
node and the fluid node. The location of the wall for bounce-back nodes is
viscosity dependent \cite{ginzburg-multireflection}.  However, the effective location of the
wall for the
multiphase models to the best authors' knowledge is not yet derived.
The bubble is initialized as a rectangular box with coordinates
$y=7\times\frac{H_\mathrm{eff}}{100}\dots N_y-7\times\frac{H_\mathrm{eff}}{100}-1$,
$x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The initialization procedure is done to keep the self-similarity. The initial width and
body force were chosen to aim for the capillary number $\Ca=0.05$ \cite{kuzmin-binary2d}.
After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant.  It is easy to check
that it yields the following quantity to
be independent of grid size
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. For the grid $H_{\mathrm{eff}}=100$ we chose the body force as
$\frac{\mathrm{d}P}{\mathrm{d}x}=1.508 \times 10^{-6}$ lattice units.
The
simulation results in terms of the film thickness $\delta$, interface bubble
velocity $U_{bubble}$ and associated capillary numbers are summarized in Table
\ref{table:parameters:grid:refinement}. 
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see that results converge for $H_{\mathrm{eff}}\geq 175$. To calculate how well the interface is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$ is shown in Table \ref{table:parameters:grid:refinement}.

Based on these results one can conclude
that
 the interface needs to be resolved as $40-50$ percent of the
expected film thickness for simulations to be grid independent. We further examine the
velocities in the center of the bubble to calculate the capillary number. One can see from Table
\ref{table:parameters:grid:refinement} that the bubble velocities are consistent and the
calculated capillary number corresponding to these velocities is $0.07$.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/norm_grid_profs.eps}
\caption{Grid-refined profiles for the effective
channel widths
$H_{\mathrm{eff}}=100,125,150,175,200,225$. $\delta$ is scaled on $H_{\mathrm{eff}}$ and $\delta=0$
corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
Capillary number we aim for is $\Ca=0.05$. The capillary number obtained from simulations
$\Ca=0.07$. 
\label{fig:grid:profiles}}
\end{figure}
\begin{table}
\begin{tabularx}{0.5\textwidth}{|X|X|X|X|X|X|}
\hline
$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$\\
\hline
$1500$&$102$&$0.0694$&$0.0041$&$0.824$&$200000$\\
\hline
$1875$&$127$&$0.0688$&$0.0041$&$0.646$&$250000$\\
\hline
$2250$&$152$&$0.0676$&$0.0040$&$0.539$&$300000$\\
\hline
$2625$&$177$&$0.0679$&$0.0040$&$0.453$&$350000$\\
\hline
$3000$&$202$&$0.0668$&$0.0041$&$0.400$&$400000$\\
\hline
$3375$&$227$&$0.0663$&$0.0039$&$0.355$&$450000$\\
\hline
\end{tabularx}
\caption{The parameters and results for grid resolution. The simulated domain is
of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the interface velocity of the bubble center
measured at the front meniscus.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta
(N_y-2)$ is the size of the film in
lattice Boltzmann units.
\label{table:parameters:grid:refinement}}
\end{table}

\subsubsection{Capillary number region}
The purpose of this section is to validate the correlations of
\citet{giavedoni-numerical} and
\citet{heil-bretherton} for a film thickness dependancy on the range of capillary
numbers. Because of limited computational resources, we skip the
small capillary numbers and make calculations for the range of capillary numbers $0.03-0.8$, which
is a computationally reasonable task.  To be consistent with the grid independant results, we choose
the grid to be
$202 \times 3001$. Then $5$ lattice Boltzmann units do not occupy more than $60$
percent of the effective film thickness. Because the simulation gets unstable with
smaller grids and larger gradients, all the capillary number simulations were
performed on the same grid. To properly initialize the body force, the
proportionality law was utilized. The forcing
$6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
number $0.05$. The forcing for other grids can be obtained using
simple proportionality relationships:
\begin{equation}
\begin{aligned}
&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
\end{aligned}
\end{equation}
where the subscript ,,lit'' stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
involved because all simulations are conducted on the same grid.
The pressure gradient can be obtained through the capillary number
ratio:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation}
The film thickness is initialized through the ratio of capillary numbers as well:
\begin{equation*}
w=12 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation*}
The results obtained after $2\cdot10^5$ steps are presented in Table
\ref{table:parameters:capillary:number}. One can see that the calculated
capillary numbers are overpredicted but the obtained thicknesses are overpredicted
as well due to the body force setting caused by the Poiseuille flow assumption. If one
needs to obtain the desired capillary number, the shooting method for the forcing is
necessary, starting with the Poiseuille pressure gradient force as the initial condition. We
extracted
the data of 
\citet{giavedoni-numerical} and \citet{heil-bretherton} with the help of software ``Engauge
Digitizer'' and compared them with our results (see 
Fig. \ref{fig:capillary:comparison}).
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison.eps}
\caption{A comparison between simulation results and results of
\citet{giavedoni-numerical} and \citet{heil-bretherton}. One can see a
reasonable agreement. The plots depict the film thickness as a function of the
the capillary number.\label{fig:capillary:comparison}}
\end{figure}

\begin{table}
\begin{tabularx}{0.5\textwidth}{|X|X|X|X|X|}
\hline
$Ca_{\mathrm{lit}}$&$\delta_{\mathrm{lit}}$&$\delta$&$U_{bubble}$&$Ca$\\
\hline
$0.03$&$0.04$&$0.042$&$0.0022$&$0.039$\\
\hline
$0.05$&$0.06$&$0.066$&$0.0041$&$0.072$\\
\hline
$0.08$&$0.08$&$0.085$&$0.0068$&$0.120$\\
\hline
$0.1$&$0.1$&$0.077$&$0.0055$&$0.098$\\
\hline
$0.2$&$0.12$&$0.123$&$0.0185$&$0.328$\\
\hline
$0.4$&$0.13$&$0.153$&$0.0388$&$0.686$\\
\hline
$0.6$&$0.15$&$0.164$&$0.0592$&$1.046$\\
\hline
$0.8$&$0.16$&$0.173$&$0.0782$&$1.383$\\
\hline
\end{tabularx}
\caption{The parameters and results for capillary number region simulations.
$\delta_{\mathrm{lit}}$ is the film thickness with corresponding $\Ca_{\mathrm{lit}}$ taken from
literature. $\delta_{\mathrm{sim}}$ is the simulation film thickness with corresponding $\Ca$.
$U_{bubble}$ is the bubble interface velocity. $\Ca$ is based on the interface bubble velocity
$U_{bubble}$.
\label{table:parameters:capillary:number}}
\end{table}

\subsection{3D results}
One can see from Fig. \ref{fig:heil:three:dim} that the separation of the axial and diagonal radii
happens at the capillary numbers less than $0.1$. However the axial radius has a value around $0.99
H_{\mathrm{eff}}$. Therefore, the film occupies $0.01 H_{\mathrm{eff}}$ which needs to be resolved
with at minimum $10$ lattice units (see results for grid independency). This condition implies grids
to be of at least size $1000\times1000\times 15000$, which is not a computationally feasible
task. Even the symmetric and antisymmetric profiles are obtained by current simulations, Fig.
\ref{fig:sym:antisym}, the grid independence is not guaranteed.
Therefore our target simulations are from $\Ca\geq 0.1$, where the axial radius equals to the
diagonal radius. In what follows we will show two particular simulations. One is for the dependance
of the bubble radius on the capillary number. Another one is for the velocity patterns. In
comparison with simulations of \citet{heil-threedim} the transition capillary
number is located in the range $0.47<\Ca<0.63$. 
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/sym.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/antisym.eps}\\
\caption{Symmetric ($\Ca=0.63$) and antisymmetric ($\Ca=0.026$) crosssections in the middle of the
bubble. One can see that for the symmetric profile $R_h=R_d$ and for the antisymmetric $R_h\neq
R_d$.  \label{fig:sym:antisym} }
\end{figure}

We also simulate only the quarter of the channel by imposing the mirror boundary conditions. This
gives the improvement in memory size and time by factor of four.  
\subsubsection{Capillary region}
We performed a number of simulations with CPU code of grid size $52\times52\times 1500$ for the
quarter of geometry and of refined grid size $82\times82\times 1500$. All other parameters were
chosen as
$k=0.04$,$A=0.04$, and $\Gamma=1.0$.

The results are presented in Fig.
\ref{fig:capillary:comparison:3d} along with reference data obtained by \citet{heil-threedim}. One
can see that refined grids produce the same results. Thus, the grid of size $52\times52\times 1500$
for the quarter of geometry is enough to obtain grid independent results for $\Ca\geq0.1$. The
capillary number is calculated based on the interface velocity along the center axis. The
simulations results as radii values are underestimated in comparison with the reference data by
\citet{heil-threedim}. This can be explained by the influence of inertia. For example, one can refer
to experimental data obtained by \citet{shikazono-square} with different liquids where real liquid
results can be located below or above the reference curve.
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_3d.eps}
\caption{The comparison between reference data by \citet{heil-threedim} and simulation for the axial
and the diagonal radiuses
versus capillary numbers. One can see that the code mimics
behavior of the earlier published results. The difference can be attributed to the simulation of
binary-liquid instead of the free-surface classical Bretherton
problem. \label{fig:capillary:comparison:3d}}
\end{figure}

\subsubsection{Velocity pattern}
We examined velocity patterns to identify the moment of the streamlines pattern change. We chose two
representative capillary numbers as $Ca=0.47$ and $Ca=0.61$. Scaled velocity profiles in the center
plane are shown  in Fig. \ref{fig:streamlines:pattern}. For demonstration purposes, we restricted
the size of shown velocity vector maps to apply certain scaling and to be able to show the vortex in
the slug. Velocity vector maps are generated in the reference frame of the moving bubble. One can
see a clear transition between associated patterns. The transition capillary number
\cite{heil-threedim} is slightly higher than simulations results. We outline that this fact can be
attributed to the binary-liquid simulations instead of the free-surface classical Bretherton
problem. 
\begin{figure}[!]
\includegraphics[width=0.5\textwidth]{Figures/vortex_ca47_png.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/vortex_ca63_png.eps}\\
\caption{Velocity vector maps for $\Ca=0.47$ and $\Ca=0.63$. One can see that there are no vortexes
created
before the bubble for $\Ca=0.63$. The transition happens between $\Ca=0.47$ and $\Ca=0.63$, which
is a different value than $\Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
\end{figure}

\section{Conclusion}
This work presents numerical studies of the
Bretherton/Taylor problem using the binary liquid lattice Boltzmann method. The bubble was chosen
sufficiently long for the film
thickness to
stabilize, and periodic boundary conditions were used to keep the simulations robust. The
distance between bubbles was taken large enough to minimize
the mutual influence of neighboring bubbles. The computational
results in terms of capillary number dependence and shape of the bubbles show consistency with the
previously published data in two and three dimensions. The simulations show that in the range of low
capillary numbers the
viscous effects dominate over the inertia. The binary liquid model can be used to
simulate gas finger/bubbles propagation in the microchannel. The model is able to catch not only
the film widths but also the velocity patterns change. 
 An examination of the influence of grid resolution on the
results allowed
us to determine that the phase interface should be resolved as at least $50$ percent of the film
thickness
in order for the simulations to be grid independent. Though our results are specific to the binary
liquid lattice
Boltzmann method, the numerical hints and procedures can be used for any
continuous interface method.

\section{Acknowledgement}
A.~Kuzmin wants to thank Schlumberger for their financial support.
%\newpage

%% %---------------------------
%% %BIBLIOGRAPHY
%% %---------------------------
%The bibliography is created using BiBTeX
\bibliographystyle{CFD2011}
%\bibliography{References}   
\bibliography{paper}
%\newpage
\section{Appendix A}
\label{app:models}
Specific realization of $D2Q9$ and $D3Q15$ models.
\subsection{$D2Q9$ model}
Velocity set is defined as $c_{ix}=\{0,1,0,-1,0,1,-1,-1,1\}$ and $c_{iy}=\{0,0,1,0,-1,1,1,-1,-1\}$.
Parameters specific to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$.

\subsection{$D3Q15$ model}
Velocity set is defined as:
\begin{equation} 
\begin{aligned}
&c_{ix}=\{0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1\}\\
&c_{iy}=\{0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0\}\\
&c_{iz}=\{0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1\}.
\end{aligned}
\end{equation}
The weights are $w_0=0$, $w_{1-6}=\frac{1}{6}$ and $w_{7-18}=\frac{1}{12}$. The weights
related to the inclusion of the surface tension are:
\begin{equation}
\begin{aligned}
&w^{xx}_{1-2}=w^{yy}_{3-4}=w^{zz}_{5-6}=\frac{5}{12}\\
&w^{xx}_{3-6}=w^{yy}_{1-2,5-6}=w^{zz}_{1-4}=-\frac{1}{3}\\
&w^{xx}_{7-10,15-18}=w^{yy}_{7-14}=w^{zz}_{11-18}=-\frac{1}{24}\\
&w^{xx}_{11-15,}=w^{yy}_{15-18}=w^{zz}_{7-10}=\frac{1}{12}\\
&w^{xy}_{7,10}=-w^{xy}_{8,9}=w^{yz}_{11,14}=-w^{yz}_{12,13}=w^{zx}_{15,18}=-w^{zx}_{16,17}=\frac{1}{
4}.
\end{aligned}
\end{equation}

\end{document}

### 2011-03-16 Python/vtk_streamlines.py

@@ -169,36 +169,88 @@ def extract_profiles(name):
     #fig=pylab.figure()
     #pylab.imshow(phase_numpy,origin="lower")
     x,y=numpy.mgrid[0:dims[0],0:dims[2]]
+    deltax=0.5/(dims[0]-1)
+    deltay=15.0/(dims[2]-1)
     positive=numpy.where(phase_numpy>0.0)
     negative=numpy.where(phase_numpy<0.0)
     large=numpy.where(x>25)
-    x_short=x[::3,::15]
-    y_short=y[::3,::15]
+    bounds=numpy.where(numpy.logical_and(y>z1-5,y<z2%dims[2]+10))
+    print z1,z2 #,bounds
+    #print y<z1-30
+    x_short=x[::3,::25]*deltay
+    y_short=y[::3,::25]*deltax
     vz_diff_mask=vz_diff #[numpy.where(phase_numpy>0.0)]
     vz_diff_mask[negative]=None
     vz_diff_mask[large]=None
+    vz_diff_mask[bounds]=None
     
-    velx_short=vz_diff_mask[::3,::15]
+    velx_short=vz_diff_mask[::3,::25]
     vy_mask=vy #[numpy.where(phase_numpy>0.0)]
     vy_mask[negative]=None
     vy_mask[large]=None
+    vz_diff_mask[bounds]=None
     
-    vely_short=vy_mask[::3,::15]
+    vely_short=vy_mask[::3,::25]
 
     #pylab.figure()
     #phase_neg=phase_numpy
     #phase_neg[negative]=None
     #pylab.imshow(phase_neg, origin="lower")
     
-    fig=pylab.figure(figsize=[23,2.5])
+    fig=pylab.figure(figsize=[10,10])
     fig=fig.subplots_adjust(bottom=0.15,top=0.8)
-    pylab.quiver(y_short,x_short,velx_short,vely_short,scale=0.1,width=0.0023,headwidth=2)# ,scale=0.1)
-    pylab.title(r'''$Ca='''+str(capillary)[0:4]+r'''$''',size=30)
-    pylab.contour(phase_numpy,[0.0],linewidths=[4])
+    #axes=fig.xaxis
+    pylab.quiver(y_short,x_short,velx_short,vely_short,scale=0.08,width=0.0033,headwidth=10,headlength=5,headaxislength=5)# ,scale=0.1)
[…]

### 2011-03-16 Python/whole_cap.py

@@ -2,6 +2,104 @@
 import numpy
 import pylab
 from numpy import genfromtxt
+
+def draw_capillaries_paper():
+    names=["force0000002","force0000002x82"]
+    style=["<","^",">"]
+    fig=pylab.figure()
+
+    for counter,name in enumerate(names):
+        mat=numpy.loadtxt("Capillaries/capillaries_"+name+".txt")
+        
+        #print mat
+        pylab.plot(mat[:,0],0.5*(mat[:,1]+mat[:,2]),style[counter],markersize=10,linewidth=2)
+        #pylab.plot(mat[:,0],mat[:,2],"+")
+
+    
+    #capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
+    #radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+   
+   
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    heil=genfromtxt("Capillaries/heil.csv",delimiter=';',dtype=float)
+
+    #print heil
+    #print heil.shape
+    
+    ax=fig.add_subplot(111)
+    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    
+    pylab.semilogx(heil[:,0],heil[:,1],linewidth=2)
+    #pylab.semilogx(capillary_theor,radiusses,'o-',markersize=7,linewidth=2)
+    #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
+    #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
+    #pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
+
+    pylab.xlim(xmin=0.1,xmax=5)
+    #pylab.ylim(ymin=0.01)
+    #numpy.savetxt("capillary.dat",zip(capillaries,widths))
+    
+    pylab.xticks(fontsize=20)
+    pylab.yticks(fontsize=20)
+    
+    pylab.ylabel(r'''$R_{diag},R_{axes}$''',fontsize=30)
[…]

### 2011-04-06 Conference/CFD2011/paper.tex
Started writing it as a paper. Need to do the introduction.

@@
 
-\title{Multiphase microchannel flows simulations via the lattice Boltzmann scheme}
+\title{Lattice Boltzmann Microchannel Simulations via Binary Liquid Model}
 \paperID{CFD11-52}
@@
 \presenting  %the previous author is presenting the paper (name becomes underlined)
-\email{shurik.kuzmin@gmail.com}%e-mail address of the previous author
+\email{kuzmin@ualberta.ca}%e-mail address of the previous author
 \address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
@@
 thin liquid films between the bubbles and the channel walls. The main focus of our work is a
-feasibility study of the binary-liquid model lattice Botlzmann model to simulate the
+feasibility study of the binary-liquid model lattice Boltzmann model to simulate the
 Bretherton/Taylor phenomena. The numerical scheme has a diffuse
@@
 results show that the lattice Boltzmann free energy binary liquid model can be used to simulate the
-Bretherton problem with a good accuracy.
+Bretherton problem with good accuracy.
 }
-\keywords{CFD, Multiphase Flows, Lattice Boltzmann Method, Microchannel simulations, Bretherton
-Problem, Binary liquid, Free energy method}
+\keywords{CFD, Multiphase Flows, Lattice Boltzmann Method, Microchannel flow, Bretherton
+problem, Binary liquid, Free energy method}
 \normalfont\normalsize
@@
 narrow channels. Depending on the geometry it was found that the deposited film thickness
-is a complicated function of the capillary number $Ca$, see Eq. (\ref{capillary:number:definition}).
+is a complicated function of the capillary number $Ca$, Eq. (\ref{capillary:number:definition}).
 For example, the deposited film thickness
@@
 The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
-is of significant interests for chemical industry as far as it is widely used in chemical monolith
+is of significant interests for chemical industry as it is widely used in chemical monolith
 microreactors \cite{kreutzer-pressure-drop}. While it
@@
 not possible to extrapolate it to the wider range of capillary numbers used in the chemical
-industry. Thus, the problem of consistent numerical simulations arises.
+industry. Thus, the desire of consistent numerical simulations arises.
 \vskip .4em
-For what follows we distinguish two- and three-dimensional cases and give a detailed description
-of the literature review. 
+For what follows we distinguish two- and three-dimensional cases and give a detailed acount
+of the literature. 
 \begin{description}
@@
 \end{equation}
-where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
+where $\mu_{liq}$ is the liquid's kinematic viscosity, $U_{bubble}$ is the bubble velocity, and
 $\gamma$ is the interfacial tension between gas and liquid. For the case of the continuous
@@
 typically validated with analytical solutions in the range of the small capillary numbers
- and then extende to more practical regimes.
+ and then extended to more practical regimes.
 
@@
 three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
-or square shaped capillaries there is a transition for the
-certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case,
+or square shaped capillaries there is a transition for a
+certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
+symmetric case,
 Fig. \ref{fig:sym:antisym}.
-From herein, we limit ourselves to the case of microchannels with square crosssections.
+From hereon, we limit ourselves to the case of microchannels with square crosssections.
 Non axisymmetric case is attributed to the case where the axial radius is different from the
 diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
-the bubble shape mimics the shape of the square and looks like as the rounded square. The dependance
+the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
 of the diagonal and axial radii on the capillary number is shown in Fig. 
@@
 than
-the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. One of the example for the film
-thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
+radius of the droplet dependant on the capillary number. One of the example for the bubble radii
+dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
 
-\citet{heil-threedim} also indicated another interesting phenomena. They indicated the transition
+\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
 capillary number where the streamlines pattern
-changes from having vortex in front of the bubble and not having it for larger capilllary number,
+changes from having a vortex in front of the bubble and not having it for larger capilllary
+numbers,
 Fig. \ref{fig:streamlines:pattern}.
@@
 square and rectangular shaped capillaries. Also, they investigated gravitational effects for
-circular shaped capillaries. They also found the empirical correlation which allows to put
+circular shaped capillaries. They also found an empirical correlation which allows to put
 different  aspect ratio rectangular channels simulation results on one
@@
 shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
-interface method based on the non-uniform grid. It is of particular interest for us to compare it
+interface method based on a non-uniform grid. It is of particular interest for us to compare it
 with the continuous interface method by the LBM implementation. 
 They also performed
-numerical simulations for the determination of the critical capillary number $\hat{Ca}$.
+numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.
 \end{description}
 Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
-are performed with the help of the boundary value approach
+are performed with the help of the finite element method
 \cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
@@
 involving complex geometries, free interface motion, or for problems involving coalescence and/or
-droplet breakup. The continuous interface models are more flexible for those kind of simulations.
+droplet breakup. The continuous interface models are more flexible for this kind of simulations.
 However, if the interface is smeared out over several grid nodes, the question of
 proper film resolution in comparison with the interface resolution arises.  Therefore, the
-{\emph goal} of this work is to reproduce the numerical results and phenomena presented above
-with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We do
-a particular focus on exploring the parameter range of the binary liquid lattice
-Boltzmann method for correctly resolving the flow in a range of capillary numbers. A thorough
-parametric study for this approach and the ability of the
-method to reconstruct consistent with other methods results has not been done to the best of the
-authors' knowledge.
+\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
+with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
+focus on exploring the parameter range of the binary liquid lattice
+Boltzmann method for correctly resolving the flow in a range of capillary numbers. 
 \vskip .5em
@@
 numbers.
-Therefore, the major governing factor for microchannel flows is not the density ratio, but the
-viscosity ratio. The goal of this work is to do a feasibility assessment study of the LBM
-binary-liquid model to simulate and correctly predict flow patterns for th Bretherton/Taylor
+Therefore, a major governing factor for microchannel flows is not the density ratio, but the
+viscosity ratio. The goal of this work is to do a feasibility study of the LBM
+binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
 problem. The work results are in good agreement with other simulations
@@
 -\partial_{\beta}P_{\alpha \beta} +
-\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
+\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
 &\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
@@
 \end{equation}
-where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
+where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
 $M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
@@
 shaped microchannels, both driven by a body force. We limit ourselves to the study
-of body force driven flows, because of their simplicity and better numerical stability. This
+of body force driven flows, because of their simplicity and better numerical stability. 
+This
 implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
@@
 needs
-to ensure that the distance between bubbles is large enough to exclude mutual bubble influence. 
-Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
-to be sufficient to give results consistent with the theory. The
-channel length is taken to be $3$ bubble lengths (or $15$ channel heights)
-to minimize influence of one bubble on another,
-because of the periodicity of boundary conditions. The geometry
+to ensure that the channel length is large enough to exclude mutual bubble influence.  The
+channel length is taken to be $15$ channel heights with the bubble length to be $5$ channel
+heights. The chosen numbers are shown to be sufficient to give results consistent with the theory,
+i.e. for the the film thickness to stabilize and to exclude mutual bubble influence. The geometry
 dimensions are represented in Fig. \ref{fig:benchmark:sketch}.
@@
 \includegraphics[width=0.5\textwidth]{Figures/benchmark.eps}
-\caption{The two-dimensional benchmark sketch. $\delta$ corresponds to film thickness. $5\xi$
+\caption{The $x-y$ benchmark crosssection. $\delta$ corresponds to the film thickness. $5\xi$
 corresponds to the interface thickness. We take the microchannel length to be $15$ times larger than
-its height. The threedimensional benchmark has the same geometry extending one dimension to have a
+its height. The threedimensional benchmark has the same geometry extending it in the $z$ dimension
+to have a
 square shaped crosssection.
@@
 Two-dimensional binary-liquid simulations are presented in our recent work
-\cite{kuzmin-binary2d}. Here we outline two most important results. The first one addresses the
+\cite{kuzmin-binary2d}. Here we outline the two most important results. The first one addresses the
 question of how the interface should be resolved to obtain grid independent result. Though the
@@
 uniform grid continuous interface method. The second part of results are comparisons with the
-literature results for the film thickness dependency on the capillary number.
+literature for the film thickness dependency on the capillary number.
 
@@
 is the effective channel height.
-If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
-which represent
-the physical domain, top and bottom nodes are bounce-back nodes. Then the effective wall location is
+If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions representing
+the physical domain with top and bottom nodes being bounce-back nodes. Then the effective wall
+location is
 in the middle between bounce-back and fluid
-nodes. Overall there are $N_y-2$ nodes representing the fluid. Note that it is a big
-simplification to impose the boundary in the middle between the bounce-back
+nodes. Overall there are $N_y-2$ nodes representing the fluid. Note that it is a simplification to
+impose the boundary in the middle between the bounce-back
 node and the fluid node. The location of the wall for bounce-back nodes is
@@
 \ref{table:parameters:capillary:number}. One can see that the calculated
-capillary numbers are overpredicted but the obtained thicknesses are overpredicted
-as well due to the body force setting caused by the Poiseuille flow assumption. If one
-needs to obtain the desired capillary number, the shooting method for the forcing is
-necessary, starting with the Poiseuille pressure gradient force as the initial condition. We
+capillary numbers are overpredicted (the actual capillary number is larger than the capillar number
+we aim for, $\Ca>\Ca_{\mathrm{lit}}$). We
 extracted
@@
 \subsection{3D results}
-One can see from Fig. \ref{fig:heil:three:dim} that the separation of the axial and diagonal radii
+One can see from Fig. \ref{fig:heil:three:dim} that the difference between the axial and diagonal
+radii
 happens at the capillary numbers less than $0.1$. However the axial radius has a value around $0.99
@@
 with at minimum $10$ lattice units (see results for grid independency). This condition implies grids
-to be of at least size $1000\times1000\times 15000$, which is not a computationally feasible
-task. Even the symmetric and antisymmetric profiles are obtained by current simulations, Fig.
-\ref{fig:sym:antisym}, the grid independence is not guaranteed.
-Therefore our target simulations are from $\Ca\geq 0.1$, where the axial radius equals to the
+to be at least size $1000\times1000\times 15000$, which is not a computationally feasible
+task. Even the symmetric and non-symmetric profiles are obtained by current simulations, Fig.
+\ref{fig:sym:antisym}, the grid independence is only guaranteed for symmetric cases.
+Therefore our target simulations are for $\Ca\geq 0.1$, where the axial radius equals to the
 diagonal radius. In what follows we will show two particular simulations. One is for the dependance
@@
 \includegraphics[width=0.5\textwidth]{Figures/antisym.eps}\\
-\caption{Symmetric ($\Ca=0.63$) and antisymmetric ($\Ca=0.026$) crosssections in the middle of the
+\caption{Symmetric ($\Ca=0.63$) and non-symmetric ($\Ca=0.026$) crosssections in the middle of the
 bubble. One can see that for the symmetric profile $R_h=R_d$ and for the antisymmetric $R_h\neq
@@
 We also simulate only the quarter of the channel by imposing the mirror boundary conditions. This
-gives the improvement in memory size and time by factor of four.  
+gives an improvement in memory requirements and simulution time by factor of four.  
 \subsubsection{Capillary region}
-We performed a number of simulations with CPU code of grid size $52\times52\times 1500$ for the
-quarter of geometry and of refined grid size $82\times82\times 1500$. All other parameters were
+We performed a number of simulations of grid size $52\times52\times 1500$ for the
+quarter of geometry and of refined grid size $82\times82\times 1500$. Other parameters were
 chosen as
@@
 and the diagonal radiuses
-versus capillary numbers. One can see that the code mimics
-behavior of the earlier published results. The difference can be attributed to the simulation of
-binary-liquid instead of the free-surface classical Bretherton
-problem. \label{fig:capillary:comparison:3d}}
+versus capillary numbers. \label{fig:capillary:comparison:3d}}
 \end{figure}
@@
 see a clear transition between associated patterns. The transition capillary number
-\cite{heil-threedim} is slightly higher than simulations results. We outline that this fact can be
+\cite{heil-threedim} is slightly higher than the simulation results. We outline that this fact can
+be
 attributed to the binary-liquid simulations instead of the free-surface classical Bretherton
@@
 previously published data in two and three dimensions. The simulations show that in the range of low
-capillary numbers the
-viscous effects dominate over the inertia. The binary liquid model can be used to
+capillary numbers the binary liquid model can be used to
 simulate gas finger/bubbles propagation in the microchannel. The model is able to catch not only

### 2011-04-06 Text/paper.tex
Started writing it as a paper. Need to do the introduction.

@@
 \begin{abstract}
-Recently we investigated the applicability and parameters choice for the two-dimensional case
-Bretherton problem using the lattice Boltzmann method (LBM) \cite{kuzmin-binary2d}. This
-paper is the continuation of the previous work and is focused on the three-dimensional case to
-validate the LBM for simulation of phenomena for rectangular and square capillaries in the
-moderate range of Capillary number parameters. 
+Classical Bretherton problem describes the propagation of gas fingers through liquid media with
+thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
+complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
+the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
+between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
+\cite{kuzmin-binary2d}. This
+work is the continuation of the previous work. It is focused on the three-dimensional capillaries
+simulations with rectangular and square crosssections to
+validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
+number $0.1\leq Ca \leq 1.0$.  
+The flow is driven by a body force,
+and
+periodic boundary conditions are applied in the streamwise direction. The results show that the
+binary liquid model is able to capture a number of phenomena happening in three-dimensional
+capillaries , as the existance of the vortex in front of the bubble and the bubble radii dependency
+on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
+used to simulate the Bretherton problem with good accuracy.
 \end{abstract}
 
-\section{What to examine}
-My plan is to investigate preferably the low capillary number regims as far we have interesting
-phenomena occuring where: the axisymmetric radiuses for axes and diagonals. As far as it's
-computationally very expensive to resolve the film thickness, it will be nice to investigate
-whether underresolved films simulations can produce consistent results. The second thing to study
-is the velocity pattern which is much easier to examine because the threshold is located for
-$Ca\approx 0.6$. The third thing to study is the film thickness dependancy for the range of
-capillary numbers. The last thing to study is the applicability of derived by \citet{heil-threedim}
-correlations for rectangular channels of different shapes.
+\section{Introduction}
+The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
+narrow channels. Depending on the geometry it was found that the deposited film thickness
+is a complicated function of the capillary number $Ca$:
+\begin{equation}
+\label{capillary:number:definition}
+Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
+\end{equation}
+where $U_{\mathrm{bubble}}$ is the propagation bubble velocity, $\mu_{\mathrm{liq}}$ is the
+kinematic liquid viscosity and $\gamma$ is the interfacial surface tension between gas and liquid.
 
-\section{Introduction}
-While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
-only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-bretherton},
-vast number of experimental results and simulations are available for the three-dimensional case.
-It was found that the for rectangular and/or square shaped capillaries there is a transition for
-certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
+For example, the deposited film thickness
+is proportional to $Ca^{2/3}$ in the range of small capillary numbers for circular channels
+\cite{bretherton,heil-bretherton}. 
+The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
+is of significant interests for chemical industry as it is widely used in chemical monolith
+microreactors \cite{kreutzer-pressure-drop}. The large mass transfer can be achieved because of the
+large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. The heat transfer is
+also increased in comparison with one phase flows \cite{fukugata-levelset}. While it
+is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
+not possible to extrapolate it to the wider range of capillary and Reynolds numbers used in the
+chemical
+industry. Thus, the desire of consistent numerical simulations arises.
+
+In comparison with the two-dimensional Bretherton flow (circular tubes, flow between parallel
+plates) covered in our previous work \cite{kuzmin-binary2d}, there is a vast number of experimental
+results available for the three-dimensional case as microchannels with square, triangular and
+rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
+experiments for a
+bubble-train flow in capillaries of
+circular and square cross section for horizontal, upward and downward flows. They found a
+significant difference in mass transfer between upward and downward flows for the small capillary
+number region.
+
+\citet{shikazono-square} obtained
+experimental
+results for the deposition length dependency on the
+capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
+shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
+based on the capillary number $Ca$ and the Weber number $We=Re Ca$. Thus, there is the effect of
+inertia on the bubble formation. It is indicated by \citet{gupta-review} that in comparison with
+one phase flows the flows in microchannels depend on the complicated interplay of the gravitational,
+interfacial, inertial and viscous forces.
+{\color{red} Historically,  it was indicated \cite{bretherton} that inertia effects can be 
+neglicted. Later on, the results of \citet{giavedoni-numerical} and
+\citet{heil-bretherton} show
+{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
+range of Reynolds
+numbers. 
+For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
+are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
+number from $0$ to $70$ for the flow between plates. Later on, \citet{heil-bretherton} extended
+results for the flow between plates upto Reynolds number
+$300$. The indicate that while the influence of the established film thickness is insignificant
+($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
+significantly the pressure distrubution and the flow field near the front bubble tip. Recent
+experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
+following dependency on the capillary number $Ca$ and the Weber number $We$:
+\begin{equation}
+\begin{aligned}
+&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
+&R_{axis}=
+\begin{cases}
+1, &R_{diag}>1\\
+R_{diag}, &R_{diag}\leq 1.
+\end{cases}
+\end{aligned}
+\end{equation}
+In the present
+simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
+order of unity. Thus, it allows to neglect 
+inertia effects. 
+Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
+but the
+viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
+simulation of the square shaped capillaries.
+
+ Moreover, the results based on uniform density, as in the case of the LBM
+binary liquid model, are in good agreement with other simulations
+\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
+same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
+gas and liquid, which are of
+significantly
+different densities and viscosities.
+}
+
+The interested reader
+is referred to these works for comprehensive experimental correlations.
+
+The experimental works supported by numerical simulations reveal interesting phenomena happening in
+three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
+or square shaped capillaries there is a transition for a
+certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
+symmetric case,
+Fig. \ref{fig:sym:antisym}.
+{\color{red} From hereon, we limit ourselves to the case of microchannels with square
+crosssections.}
 Non axisymmetric case is attributed to the case where the axial radius is different from the
-diagonal radius and the bubble has the non-circular shape in the channel crossection, see the
-example in Fig. \ref{fig:quarter:capillary:capillary01} for a non-axisymmetric case.
-
+diagonal radius and the bubble has the non-circular shape in the channel crossection. 
+{\color{red}
+\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
+flow in the side of the capillary thus changing the qualitative differencies depending on the
+gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
+larger fro upflow than for downflow''. 
+}
+In
+this case
+the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
+of the diagonal and axial radii on the capillary number is shown in Fig. 
+\ref{fig:heil:three:dim}.
+\begin{figure}[ht]
+\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
+\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
+capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
+number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
+\end{figure}
 The transition number between non-axisymmetric case and symmetric is indicated in a number of
@@
 than
-the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. One of the example for the film
-thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
+radius of the droplet dependant on the capillary number. One of the example for the bubble radii
+dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+
+\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
+capillary number where the streamlines pattern
+changes from having a vortex in front of the bubble and not having it for larger capilllary
+numbers,
+Fig. \ref{fig:streamlines:pattern}.
+For the square channel the critical number is $Ca=0.691$.   
+%\begin{figure}
+%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
+%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
+%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
+%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
+%(bottom). 
+%\label{fig:heil:recirculating}}
+%\end{figure}
+
+There are also a number of numerical works for the three-dimensional case. For instance, 
+\citet{wong-films,wong-pressure} studied
+three-dimensional bubbles in
+polygonal capillaries and calculated bubble shapes for different
+slug and bubble cross sections and menisci appearance.
+\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
+square and rectangular shaped capillaries. Also, they investigated gravitational effects for
+circular shaped capillaries. They also found an empirical correlation which allows to put
+different  aspect ratio rectangular channels simulation results on one
+curve. \citet{wang-non-circular} performed numerical simulations by the VOF technique for the
+non-circular
+shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
+interface method based on a non-uniform grid. It is of particular interest for us to compare it
+with the continuous interface method by the LBM implementation. 
+They also performed
+numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.
+
+Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
+are performed with the help of the finite element method
+\cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
+applicability for
+problems
+involving complex geometries, free interface motion, or for problems involving coalescence and/or
+droplet breakup. The continuous interface models are more flexible for this kind of simulations.
+However, if the interface is smeared out over several grid nodes, the question of
+proper film resolution in comparison with the interface resolution arises.  Therefore, the
+\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
+with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
+focus on exploring the parameter range of the binary liquid lattice
+Boltzmann method for correctly resolving the flow in a range of capillary numbers.
+
+The lattice Boltzmann method has emerged as a successful method to simulate
+a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
+\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
+ferrofluids \cite{kuzmin-aniso}, and multiphase flows
+\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
+method easily tackles complex geometries and allows incorporation of
+physical phenomena on the microscopic level, as in the case of multiphase models. Most
+multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
+the interface using continuous interface methods where the interface spans over several grid nodes.
+Such a representation
+brings issues of the film thickness resolution versus interface
+resolution -- the diffuse interface should be dealt with in such a way as to have a
+negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
+resolution was examined. It was found that the film thickness should be at least as twice as the
+interface for the simulation to be accurate.
+
+The binary liquid free-energy LB model due to \citet{swift} we used
+simulates two liquids with the assumption of uniform overall
+density. While the classical Bretherton problem is stated for gas and liquid, which are of
+significantly
+different densities and viscosities, it was indicated \cite{bretherton} that inertia effects can be 
+neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
+negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
+numbers.
+Therefore, a major governing factor for microchannel flows is not the density ratio, but the
+viscosity ratio. The goal of this work is to do a feasibility study of the LBM
+binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
+problem. The work results are in good agreement with other simulations
+\cite{giavedoni-numerical,heil-bretherton,heil-threedim}.
+
+One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
+in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
+simulations of two-dimensional channel flows for a relatively large capillary numbers, and
+found discrepancies with the classical Bretherton theory, which
+is limited to the low capillary number regime \cite{giavedoni-numerical}.
+
+The paper is organized as follows.  First, we briefly
+explain the binary liquid lattice Boltzmann model. Then, the preliminary results for the
+two-dimensional and for the three-dimensional case are presented in the results section. The paper
+is
+concluded with a summary of the main findings.
 

### 2011-04-06 Python/whole_cap.py

@@ -2,7 +2,7 @@
 import numpy
 import pylab
 from numpy import genfromtxt
-<<<<<<< HEAD
+import math
 
 def draw_capillaries_paper():
     names=["force0000002","force0000002x82"]
@@ -99,11 +99,6 @@ def draw_capillaries_paper():
     pylab.savefig("capillaries_comparison.eps",format="EPS",dpi=300)
     pylab.show()
 
-
-
-=======
-import math
->>>>>>> 847167bd5b558c0919de9ec707d4bd3cb4ff091f
 def draw_capillaries():
     names=["force0000002","force0000002x82","force0000005x52"]
     style=["<","^",">"]

### 2011-04-06 Results/capillaries.py

@@ -27,7 +27,6 @@ def draw_capillaries():
     capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
     radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
    
-   
     #style=["bo","rH","c<","y>","bs","g^"]
     #color=["b","r","c","y","b","g"]
     #style_diff=["b-","r:","c-.","y--","b^","g<"]

### 2011-04-07 Text/paper.tex
Added the introductions. Need to prepare 3D graphs using Sketch3D.

@@
 \usepackage{color}
+\usepackage{tikz}
 \begin{document}
@@
 binary liquid model is able to capture a number of phenomena happening in three-dimensional
-capillaries , as the existance of the vortex in front of the bubble and the bubble radii dependency
+capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
 on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
@@
 
-In comparison with the two-dimensional Bretherton flow (circular tubes, flow between parallel
-plates) covered in our previous work \cite{kuzmin-binary2d}, there is a vast number of experimental
+The two-dimensional geometry (circular tubes, parallel plates) flows are studied extensively in
+experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
+\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the Bretherton
+analysis is valid only for small capillaries numbers $Ca\leq 0.003$ and deviates for larger ones.
+That is caused by complex interplay of the gravitational,
+interfacial, inertial and viscous forces \cite{gupta-review}. Historically, 
+\citet{bretherton} indicated that inertia effects can be 
+neglected. \citet{giavedoni-numerical} suggested
+that
+inertia effects
+are negligible for $Ca \leq 0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
+number from $0$ to $70$ for two-dimensional geometry flows. Later on, \citet{heil-bretherton}
+extended
+results for the flow between plates upto Reynolds number
+$300$. They indicate that while the influence of the established film thickness is insignificant
+($7$ percent from the film thickness measured at $Re=0$), the change in Reynolds number 
+influences significantly the pressure distrubution and the flow field near the front bubble tip.
+\citet{cerro-bubble-train} show that the mass transfer is strongly affected by flow direction for
+small capillary numbers in case of upward and downward flows. Thus, to fully describe the
+propagation of the semi-infinite finger in liquid media one needs to take into the account the
+viscous, gravitational, surface and inertia forces \cite{gupta-review}.
+
+In comparison with the two-dimensional Bretherton flow  (circular tubes, flow between
+parallel
+plates), there is a vast number of experimental
 results available for the three-dimensional case as microchannels with square, triangular and
@@
 bubble-train flow in capillaries of
-circular and square cross section for horizontal, upward and downward flows. They found a
-significant difference in mass transfer between upward and downward flows for the small capillary
-number region.
-
+circular and square cross section for horizontal, upward and downward flows.
 \citet{shikazono-square} obtained
@@
 shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
-based on the capillary number $Ca$ and the Weber number $We=Re Ca$. Thus, there is the effect of
-inertia on the bubble formation. It is indicated by \citet{gupta-review} that in comparison with
-one phase flows the flows in microchannels depend on the complicated interplay of the gravitational,
-interfacial, inertial and viscous forces.
-{\color{red} Historically,  it was indicated \cite{bretherton} that inertia effects can be 
-neglicted. Later on, the results of \citet{giavedoni-numerical} and
-\citet{heil-bretherton} show
-{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
-range of Reynolds
-numbers. 
-For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
-are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
-number from $0$ to $70$ for the flow between plates. Later on, \citet{heil-bretherton} extended
-results for the flow between plates upto Reynolds number
-$300$. The indicate that while the influence of the established film thickness is insignificant
-($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
-significantly the pressure distrubution and the flow field near the front bubble tip. Recent
-experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
+based on the capillary number $Ca$ and the Weber number $We=Re\,Ca$.
+
+In comparison with two-dimensional geometries, the experimental works
+\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
+wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
+\cite{heil-threedim,wong-films} that for rectangular
+or square shaped capillaries there is a transition for a
+certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
+symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
+different from the
+diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
+the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
+of the diagonal and axial radii on the capillary number is shown in Fig. 
+\ref{fig:heil:three:dim}.
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
+\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
+capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
+number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
+\end{figure}
+The transition number between non-axisymmetric case and symmetric is indicated in a number of
+works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
+$\widehat{Ca}=0.1$
+\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
+number is larger
+than
+the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
+radius of the droplet dependant on the capillary number. One of the example for the bubble radii
+dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+
+There are also a number of numerical works for the three-dimensional case. For instance, 
+\citet{wong-films,wong-pressure} studied
+three-dimensional bubbles in
+polygonal capillaries and calculated bubble shapes for different
+slug and bubble cross sections and menisci appearance.
+\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
+square and rectangular shaped capillaries. They indicated the transition
+capillary number where the streamlines pattern
+changes from having a vortex in front of the bubble and not having it for larger capilllary
+numbers, {\color{red} Fig. \ref{fig:streamlines:pattern} }. For the square channel the critical
+capillary number is $Ca=0.691$.  Authors also found an empirical correlation which
+allows to put
+different aspect ratio rectangular channels simulation results on one
+curve. It is also indicated that for certain aspect ratio microchannels $\alpha=\frac{a}{b}\geq
+2.04$ the interface can not make up the
+axisymmetric case for any capillary number. \citet{wang-non-circular} performed numerical
+simulations by the VOF technique for the
+non-circular
+shaped capillaries as square and triangular shaped capillaries. They also investigated the relative
+velocity of the bubble for
+different capillary numbers.
+
+As it is mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
+phenomena are the Volume of Fluid (VOF) method \cite{wang-non-circular}, the level-set method
+\cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It is
+also indicated \cite{gupta-review}, that the new techniques which are still in the development stage
+are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
+While the finite element methods solve the Bretherton flow as a free-surface problem with a sharp
+interface but without underlying details for gas, the lattice Boltzmann method which is a continuous
+interface method arises as a promising alternative for simulation of gas finger propagation. The
+continuous interface has more flexebility in simulations involving coalescence and/or droplet
+breakup.
+
+In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between plates.
+It was shown that the free-energy binary liquid model, which is a phase field method, simulates
+with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine the
+three-dimensional geometry simulations conducted by the free-energy binary liquid lattice Boltzmann
+method. We focus on the feasibility study of the numerical method for the microchannels with
+square crosssections.
+
+The lattice Boltzmann method has emerged as a successful method to simulate
+a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
+\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
+ferrofluids \cite{kuzmin-aniso}, and multiphase flows
+\cite{swift,Shan-chen:extended}. The method is a particle method which allows to simulate physical
+phenomena on the microscopic level. For instance, the introduction of the force or potential on the
+microscopic level allows to restore multiphase macroscopic equations \cite{swift,
+Shan-chen:extended}.
+
+The binary liquid free-energy LB model due to \citet{swift} we used
+simulates two liquids with the assumption of uniform overall
+density. While the classical Bretherton problem is stated for gas and liquid, which are of
+significantly
+different densities and viscosities, the parameters were carefully chosen as to neglect the inertia
+effects, see Section \ref{sec:numerical:benchmark}. Moreover, the results of
+\citet{giavedoni-numerical} and \citet{heil-bretherton} show
+negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
+numbers. Thus, a major governing factor for microchannel flows is not the density ratio, but
+the
+viscosity ratio. That is why the uniform density binary liquid model is suitable for this kind of
+simulations. The goal of this work is to do a feasibility study of the LBM
+binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
+problem. The work results are in good agreement with other simulations
+\cite{heil-threedim,wang-non-circular}.
+
+One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
+in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
+simulations of two-dimensional channel flows for relatively large capillary numbers, and
+found discrepancies with the classical Bretherton theory, which
+is limited to the low capillary number regime \cite{giavedoni-numerical}. The Shan-Chen model was
+used to simulate the Bretherton problem, which is sometimes referred to contain thermodynamically
+inconsistent interface \cite{nourgaliev-breakup}.
+
+The paper is organized as follows.  First, we explain the simulation benchmark construction
+mainly based on our previous work \cite{kuzmin-binary2d}. Then, the binary liquid lattice
+Boltzmann model is outlined. The simulation results for the three-dimensional case are presented in
+the results section. The results cover the film thickness dependency on the capillary number, the
+relative velocity of the bubble and the vortex profiles. The paper
+is concluded with a summary of the main findings.
+%The critical capillary numbers for changing velocity pattern depending on the eccentricity
+%parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
+%\begin{table}
+%\begin{tabular}{c|c}
+%$\alpha$& $Ca_{T}$\\
+%\hline
+%1& 0.691\\
+%1.1 & 0.688\\
+%1.3 & 0.666\\
+%1.5 & 0.631\\
+%\end{tabular}
+%\caption{The results for the recirculation region.\label{table:recirculation:data}}
+%\end{table}
+
+\section{Numerical benchmark approach}
+\label{sec:numerical:benchmark}
+The main discussion here follows and is based on the two-dimensional benchmark approach
+\cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
+dimensions. Instead of using
+
+\begin{tikzpicture}
+%\draw (-2,-1,0,0)--(0,0,1);
+%\draw (0,0,1)--(0,1,1);
+\draw (-2,-1,0)--(2,-1,0)--(2,1,0)--(-2,1,0)--cycle;
+\draw (-2,-1,0)--(-2,-1,1)--(2,-1,1)--(2,-1,0)--cycle;
+\draw (-2,1,0)--(-2,1,1)--(2,1,1)--(2,1,0)--cycle;
+\draw (-2,-1,1)--(2,-1,1)--(2,1,1)--(-2,1,1)--cycle;
+\draw[rotate=90] (-2,0,0) circle (0.5);
+%(Label(XY()*"$x$",align=-3Y));
+%//surface s=surface(f,(0,0),(pi,2pi),100,Spline);
+%draw("$z$",project(O--X),Arrow);
+%draw("$y$",project(O--Y),Arrow);
+%draw("$x$",project(O--Z),Arrow);
+%//draw(YZ);
+%label(project("W",Y,Z,(2,0,-0.3)));
+%label(project("L",X,Z,(0,-1,-0.3)));
+%label(project("H",Y,Z,(-2,-0.8,0.5)));
+%draw(face1);
+%draw(face2);
+%draw(face3);
+%draw(face4);
+\end{tikzpicture}
+\begin{tikzpicture}[line join=round]
+\filldraw[draw=none,fill=white](-2,0)--(2,0)--(2,1)--(-2,1)--cycle;
+\draw(2,0)--(2,1)--(-2,1)--(-2,0);
+\filldraw[fill=white](-2,-.985)--(2,-.985)--(2,.985)--(-2,.985)--cycle;
+\filldraw[fill=white](-2,-.985)--(-2,-1.158)--(2,-1.158)--(2,-.985)--cycle;
+\filldraw[draw=none,fill=white](-2,0)--(-2,-1)--(2,-1)--(2,0)--cycle;
+\draw(-2,0)--(-2,-1)--(2,-1)--(2,0);
+\filldraw[fill=white](-2,1)--(-2,1)--(2,1)--(2,1)--cycle;
+\filldraw[draw=none,fill=white](-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087)--cycle;
+\draw(-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087);
+\filldraw[fill=white](-2,-1)--(-2,-1)--(2,-1)--(2,-1)--cycle;
+\filldraw[draw=none,fill=white](-2,.839)--(-2,-1)--(2,-1)--(2,.839)--cycle;
+\draw(-2,.839)--(-2,-1)--(2,-1)--(2,.839);
+\filldraw[fill=white](-2,.985)--(-2,.811)--(2,.811)--(2,.985)--cycle;
+\filldraw[draw=none,fill=white](-2,.839)--(2,.839)--(2,1)--(-2,1)--cycle;
+\draw(2,.839)--(2,1)--(-2,1)--(-2,.839);
+\filldraw[draw=none,fill=white](-2,-.087)--(2,-.087)--(2,.811)--(-2,.811)--cycle;
+\draw(2,-.087)--(2,.811)--(-2,.811)--(-2,-.087);
+\end{tikzpicture}% End sketch output
+
+ In order to avoid the mutual
+
+It was indicated that the film thickness should be at least as twice large
+as the interface thickness. The interface thickness in the continuous formulation of the
+free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
+resolved at least as $6-10$ nodes.
+
+As it was indicated before \cite{heil-threedim}, the diagonal bubble radius is different from the
+axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
+$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 a$, where $a$ is the side length of
+the square microchannel. Therefore, taking the minimal requirement for the film thickness to be
+resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times $ to
+simulate
+
+Non axisymmetric case is attributed to the case where the axial radius is different from the
+diagonal radius and the bubble has the non-circular shape in the channel crossection. 
+{\color{red}
+\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
+flow in the side of the capillary thus changing the qualitative differencies depending on the
+gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
+larger fro upflow than for downflow''. 
+}
+
+Most multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
+the interface using continuous interface methods where the interface spans over several grid nodes.
+Such a representation
+brings issues of the film thickness resolution versus interface
+resolution -- the diffuse interface should be dealt with in such a way as to have a
+negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
+resolution was examined. It was found that the film thickness should be at least as twice as the
+interface for the simulation to be accurate.
+
+Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
+but the
+viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
+simulation of the square shaped capillaries.
+
+ Moreover, the results based on uniform density, as in the case of the LBM
+binary liquid model, are in good agreement with other simulations
+\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
+same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
+gas and liquid, which are of
+significantly
+different densities and viscosities.
+
+Resolving the interface on a fine level for capillary numbers smaller than
+$0.003$ is computationally expensive even in the two-dimensional case
+(the amount of memory necessary to perform the simulation would be of the order
+of tens or hundreds of gigabytes).  This work is therefore based on a
+comparison of the results obtained with the known data
+validated for the Bretherton bubble flow.  In this paper, we present
+techniques for initialization of the simulations and discuss optimal
+parameter ranges for the binary liquid LB model, such as the binary liquid parameters, the viscosity
+ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
+of film thicknesses.
+
+The VOF technique is the continuous
+interface method based on a non-uniform grid. It is of particular interest for us to compare it
+with the continuous interface method by the LBM implementation.
+
+Thus, there is the effect of
+inertia on the bubble formation.
+
+Recent experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
 following dependency on the capillary number $Ca$ and the Weber number $We$:
@@
 inertia effects.
-Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
-but the
-viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
-simulation of the square shaped capillaries.
-
- Moreover, the results based on uniform density, as in the case of the LBM
-binary liquid model, are in good agreement with other simulations
-\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
-same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
-gas and liquid, which are of
-significantly
-different densities and viscosities.
-}
-
-The interested reader
-is referred to these works for comprehensive experimental correlations.
-
-The experimental works supported by numerical simulations reveal interesting phenomena happening in
-three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
-or square shaped capillaries there is a transition for a
-certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
-symmetric case,
-Fig. \ref{fig:sym:antisym}.
-{\color{red} From hereon, we limit ourselves to the case of microchannels with square
-crosssections.}
-Non axisymmetric case is attributed to the case where the axial radius is different from the
-diagonal radius and the bubble has the non-circular shape in the channel crossection. 
-{\color{red}
-\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
-flow in the side of the capillary thus changing the qualitative differencies depending on the
-gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
-larger fro upflow than for downflow''. 
-}
-In
-this case
-the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
-of the diagonal and axial radii on the capillary number is shown in Fig. 
-\ref{fig:heil:three:dim}.
-\begin{figure}[ht]
-\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
-\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
-capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
-number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
-\end{figure}
-The transition number between non-axisymmetric case and symmetric is indicated in a number of
-works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
-$\widehat{Ca}=0.1$
-\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
-than
-the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. One of the example for the bubble radii
-dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
-
-\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
-capillary number where the streamlines pattern
-changes from having a vortex in front of the bubble and not having it for larger capilllary
-numbers,
-Fig. \ref{fig:streamlines:pattern}.
-For the square channel the critical number is $Ca=0.691$.   
-%\begin{figure}
-%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
-%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
-%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
-%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
-%(bottom). 
-%\label{fig:heil:recirculating}}
-%\end{figure}
-
-There are also a number of numerical works for the three-dimensional case. For instance, 
-\citet{wong-films,wong-pressure} studied
-three-dimensional bubbles in
-polygonal capillaries and calculated bubble shapes for different
-slug and bubble cross sections and menisci appearance.
-\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
-square and rectangular shaped capillaries. Also, they investigated gravitational effects for
-circular shaped capillaries. They also found an empirical correlation which allows to put
-different  aspect ratio rectangular channels simulation results on one
-curve. \citet{wang-non-circular} performed numerical simulations by the VOF technique for the
-non-circular
-shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
-interface method based on a non-uniform grid. It is of particular interest for us to compare it
-with the continuous interface method by the LBM implementation. 
-They also performed
-numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.
-
-Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
-are performed with the help of the finite element method
-\cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
-applicability for
-problems
-involving complex geometries, free interface motion, or for problems involving coalescence and/or
-droplet breakup. The continuous interface models are more flexible for this kind of simulations.
-However, if the interface is smeared out over several grid nodes, the question of
-proper film resolution in comparison with the interface resolution arises.  Therefore, the
-\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
-with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
-focus on exploring the parameter range of the binary liquid lattice
-Boltzmann method for correctly resolving the flow in a range of capillary numbers.
-
-The lattice Boltzmann method has emerged as a successful method to simulate
-a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
-\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
-ferrofluids \cite{kuzmin-aniso}, and multiphase flows
-\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
-method easily tackles complex geometries and allows incorporation of
-physical phenomena on the microscopic level, as in the case of multiphase models. Most
-multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface using continuous interface methods where the interface spans over several grid nodes.
-Such a representation
-brings issues of the film thickness resolution versus interface
-resolution -- the diffuse interface should be dealt with in such a way as to have a
-negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
-resolution was examined. It was found that the film thickness should be at least as twice as the
-interface for the simulation to be accurate.
-
-The binary liquid free-energy LB model due to \citet{swift} we used
-simulates two liquids with the assumption of uniform overall
-density. While the classical Bretherton problem is stated for gas and liquid, which are of
-significantly
-different densities and viscosities, it was indicated \cite{bretherton} that inertia effects can be 
-neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
-negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
-numbers.
-Therefore, a major governing factor for microchannel flows is not the density ratio, but the
-viscosity ratio. The goal of this work is to do a feasibility study of the LBM
-binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
-problem. The work results are in good agreement with other simulations
-\cite{giavedoni-numerical,heil-bretherton,heil-threedim}.
-
-One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
-in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
-simulations of two-dimensional channel flows for a relatively large capillary numbers, and
-found discrepancies with the classical Bretherton theory, which
-is limited to the low capillary number regime \cite{giavedoni-numerical}.
-
-The paper is organized as follows.  First, we briefly
-explain the binary liquid lattice Boltzmann model. Then, the preliminary results for the
-two-dimensional and for the three-dimensional case are presented in the results section. The paper
-is
-concluded with a summary of the main findings.
-
-\citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
-square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
-square capillaries are shown in Fig.\ref{fig:heil:three:dim}. Also, authors performed
-numerical simulations for rectangular shaped capillaries. It was found that for different shaped
-capillaries there is an empirical correlation allowing to put all the different curves for film
-thickness dependency on the capillary number on one curve through introduction of the following
-parameter
-$s_{\infty}=r_{\infty} \alpha^{-1/2}$. The parameter $r_{\infty}$ is the aximsymmetric bubble radius
-infinitely far behind the bubble tip and can be calculated as:
-\begin{equation}
-r_{\infty}=2\sqrt{(\alpha-Q)/\pi},
-\end{equation}
-where $Q$ is the net flow rate in the computational domain. {\color{red} What is it? - how to
-calculate? Why velocity is involved?}
-
-It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
-axisymmetric case for the range of capillary numbers. The $s_{\infty}$ dependency on $\alpha$ is
-shown in Fig.
-\ref{fig:heil:three:dim}.
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/capillary_width_heil.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/capillary_rectangular_width_heil.eps}\\
-\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
-capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
-number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
-\end{figure}
-
-\citet{heil-threedim} indicate the interesting occuring thing in square and rectangular channels
-as the recirculating regime for flow fields.  It is
-known that for certain capillary number the streamlines either go around the bubble or create the
-circulating regime in front of the bubble, Fig. \ref{fig:heil:recirculating}.  
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/recirculating.eps}
-\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
-Streamlines in the moving frame of reference for a rectangular tube $\alpha=1.5$ (a) Ca = 0.2,
-(b) Ca = 0.5, (c) Ca = 0.55, (d) Ca = 2.0. Streamlines are shown on the surface of the bubble and
-in the plane along the shorter semi-axis.
-\label{fig:heil:recirculating}}
-\end{figure}
-The critical capillary numbers for changing velocity pattern depending on the eccentricity
-parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
-\begin{table}
-\begin{tabular}{c|c}
-$\alpha$& $Ca_{T}$\\
-\hline
-1& 0.691\\
-1.1 & 0.688\\
-1.3 & 0.666\\
-1.5 & 0.631\\
-\end{tabular}
-\caption{The results for the recirculation region.\label{table:recirculation:data}}
-\end{table}
-
-\citet{shikazono-square} obtained a lot of different results for the deposition length versus the
-capillary number variation ethanol/air and water/air experiments. All of them are
-presented in Fig. \ref{fig:deposition:han}.
-\begin{figure}
-%\includegraphics[width=0.47\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/han_capillary_water_air.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
-\caption{The axial and diagonal radiuses dependency on the capillary number. Courtesy of
-\citet{shikazono-square} \label{fig:deposition:han}}
-\end{figure}
-
-\citet{cerro-bubble-train} performed a lot of experiments for a buble-train flow in capillaries of
-circular and square cross section for horizontal, upward and downward flows. The main results are
-presented in Fig. \ref{fig:thulasidas:results}.
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/thulasidas_capillary.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/thulasidas_relative_velocity.eps}\\
-\caption{\citet{cerro-bubble-train} results for the film thickness
-variation and the relative bubble velocity versus the capillary
-number.\label{fig:thulasidas:results}}
-\end{figure}
-
-\citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
-shaped capillaries as square and triangle shaped capillaries. The VOF technique is the continuous
-interface method based on the non-uniform grid. It is of particular interest for us to compare it
-with the continuous interface method by the LBM implementation. Some interesting results are
-obtained for a bubble
-shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
-numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
-\ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
-different capillary numbers.
-\begin{figure}
-\includegraphics[width=\textwidth]{Figures/wang_bubble_shape.eps}
-\caption{The bubble shape across the length for different capillary numbers.
-\label{fig:wang:bubble:shape}}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/wang_relative_velocity.eps}
-\caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
-velocity depending on the capillary number.  \label{fig:wang:critical:capillary}}
-\end{figure}
 

### 2011-04-11 Text/paper.tex
Finished with the benchmark pictures.
Second section editing.

@@
 \cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
-dimensions. Instead of using
+dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in Fig.
+\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
+The film thickness in this case is measured at the inlet. In lattice Boltzmann framework such
+formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
+dynamic coupling of the inlet and outlet conditions \cite{giavedoni-numerical}. Instead, we
+proposed the numerical benchmark indicated in Fig. \ref{fig:lbm:benchmark}. The dimensions of the
+channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
+initialized bubble length is taken as $5 H_{\mathrm{eff}}$.\citet{giavedoni-numerical} showed that
+the film stabilizes at the distances of $2.6-4.0$ diameters
+from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
+middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
+bubble tip. Moreover, the film thickness examination, see Section
+\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
+$0.05\leq Ca \leq 1$, the standard deviation of the film thickness from the bubble middle film
+thickness is around $7$
+percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones ($Ca
+\geq 0.1$). %at the distances one channel height from the rear tip upto one channel
 
-\begin{tikzpicture}
-%\draw (-2,-1,0,0)--(0,0,1);
-%\draw (0,0,1)--(0,1,1);
-\draw (-2,-1,0)--(2,-1,0)--(2,1,0)--(-2,1,0)--cycle;
-\draw (-2,-1,0)--(-2,-1,1)--(2,-1,1)--(2,-1,0)--cycle;
-\draw (-2,1,0)--(-2,1,1)--(2,1,1)--(2,1,0)--cycle;
-\draw (-2,-1,1)--(2,-1,1)--(2,1,1)--(-2,1,1)--cycle;
-\draw[rotate=90] (-2,0,0) circle (0.5);
-%(Label(XY()*"$x$",align=-3Y));
-%//surface s=surface(f,(0,0),(pi,2pi),100,Spline);
-%draw("$z$",project(O--X),Arrow);
-%draw("$y$",project(O--Y),Arrow);
-%draw("$x$",project(O--Z),Arrow);
-%//draw(YZ);
-%label(project("W",Y,Z,(2,0,-0.3)));
-%label(project("L",X,Z,(0,-1,-0.3)));
-%label(project("H",Y,Z,(-2,-0.8,0.5)));
-%draw(face1);
-%draw(face2);
-%draw(face3);
-%draw(face4);
-\end{tikzpicture}
-\begin{tikzpicture}[line join=round]
-\filldraw[draw=none,fill=white](-2,0)--(2,0)--(2,1)--(-2,1)--cycle;
-\draw(2,0)--(2,1)--(-2,1)--(-2,0);
-\filldraw[fill=white](-2,-.985)--(2,-.985)--(2,.985)--(-2,.985)--cycle;
-\filldraw[fill=white](-2,-.985)--(-2,-1.158)--(2,-1.158)--(2,-.985)--cycle;
-\filldraw[draw=none,fill=white](-2,0)--(-2,-1)--(2,-1)--(2,0)--cycle;
-\draw(-2,0)--(-2,-1)--(2,-1)--(2,0);
-\filldraw[fill=white](-2,1)--(-2,1)--(2,1)--(2,1)--cycle;
-\filldraw[draw=none,fill=white](-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087)--cycle;
-\draw(-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087);
-\filldraw[fill=white](-2,-1)--(-2,-1)--(2,-1)--(2,-1)--cycle;
-\filldraw[draw=none,fill=white](-2,.839)--(-2,-1)--(2,-1)--(2,.839)--cycle;
-\draw(-2,.839)--(-2,-1)--(2,-1)--(2,.839);
-\filldraw[fill=white](-2,.985)--(-2,.811)--(2,.811)--(2,.985)--cycle;
-\filldraw[draw=none,fill=white](-2,.839)--(2,.839)--(2,1)--(-2,1)--cycle;
-\draw(2,.839)--(2,1)--(-2,1)--(-2,.839);
-\filldraw[draw=none,fill=white](-2,-.087)--(2,-.087)--(2,.811)--(-2,.811)--cycle;
-\draw(2,-.087)--(2,.811)--(-2,.811)--(-2,-.087);
-\end{tikzpicture}% End sketch output
+This length of the bubble was shown as
+sufficient for the film thickness to stabilize.
+
+\begin{figure}[ht]
+\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
+\caption{The classical benchmark. The semiinfinite gas bubble
+propagates through the liquid media. \label{fig:classical:benchmark}}
+\end{figure}[ht]
+\begin{figure}
+\includegraphics*[bb=152 480 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
+\caption{The classical benchmark. The semiinfinite gas bubble
+propagates through the liquid media. \label{fig:lbm:benchmark}}
+\end{figure}
 

### 2011-04-12 Text/paper.tex
Wrote the numerical benchmark section. Need to obtain more results. Probably for the aspect ratio as well.

@@
 channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
-initialized bubble length is taken as $5 H_{\mathrm{eff}}$.\citet{giavedoni-numerical} showed that
+initialized bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
 the film stabilizes at the distances of $2.6-4.0$ diameters
-from the front tip depending on the Reynolds number.  We chose to measure the film thickness in the
+from the front tip depending on the Reynolds number.
+\citet{heil-threedim} measure the bubble radii at the distance as $5.5 H_{\mathrm{eff}}$ to be sufficient for $Ca\leq 10$.
+In this work the conducted simulations capillary number is not larger than unity $Ca\leq 1$.
+Thus, the film thickness is chosen to be measured in the
 middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
-bubble tip. Moreover, the film thickness examination, see Section
-\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
-$0.05\leq Ca \leq 1$, the standard deviation of the film thickness from the bubble middle film
-thickness is around $7$
-percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones ($Ca
-\geq 0.1$). %at the distances one channel height from the rear tip upto one channel
+bubble tip.%at the distances one channel height from the rear tip upto one channel
 
-This length of the bubble was shown as
-sufficient for the film thickness to stabilize.
-
+For simplicity, the periodic boundary conditions are applied.
+Thus, the bubble train is simulated.
+To avoid the mutual influence of bubbles on each other the channel length was chosen as $15 H_{\mathrm{eff}}$.
+Periodic boundary conditions imply as well that the fluid is accelerated by a body force in the framework of the LBM but not with the pressure difference.
+Therefore, we can not address simulations with upward and downward flows \citet{cerro-bubble}.
+It is assumed that the gravity effects are negligible and the flow is in the horizontal direction driven by the body force.
 \begin{figure}[ht]
@@
 
- In order to avoid the mutual
-
-It was indicated that the film thickness should be at least as twice large
+It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least as twice large
 as the interface thickness. The interface thickness in the continuous formulation of the
@@
 
-As it was indicated before \cite{heil-threedim}, the diagonal bubble radius is different from the
+\cite{heil-threedim} specify that the diagonal bubble radius is different from the
 axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
-$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 a$, where $a$ is the side length of
-the square microchannel. Therefore, taking the minimal requirement for the film thickness to be
-resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times $ to
-simulate
+$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal requirement for the film thickness to be
+resolved, i.e.
+$6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to properly resolve the binary liquid microchannels bubbles motion.
+This size implies highly computational resource to be available for simulations.
 
-Non axisymmetric case is attributed to the case where the axial radius is different from the
-diagonal radius and the bubble has the non-circular shape in the channel crossection. 
-{\color{red}
-\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
-flow in the side of the capillary thus changing the qualitative differencies depending on the
-gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
-larger fro upflow than for downflow''. 
-}
+One can argue that the non-uniform grid can be used.
+However, the deposited film thickness is the function of the curvatures of the front bubble tip \cite{bretherton}.
+Therefore, the grid needs to be refined just at the interface between gas and liquid.
+It is complicated to achieve for the dynamic systems where a bubble moves in the streamwise direction.
 
-Most multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
-the interface using continuous interface methods where the interface spans over several grid nodes.
-Such a representation
-brings issues of the film thickness resolution versus interface
-resolution -- the diffuse interface should be dealt with in such a way as to have a
-negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
-resolution was examined. It was found that the film thickness should be at least as twice as the
-interface for the simulation to be accurate.
-
-Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
-but the
-viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
-simulation of the square shaped capillaries.
-
- Moreover, the results based on uniform density, as in the case of the LBM
-binary liquid model, are in good agreement with other simulations
-\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
-same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
-gas and liquid, which are of
-significantly
-different densities and viscosities.
-
-Resolving the interface on a fine level for capillary numbers smaller than
-$0.003$ is computationally expensive even in the two-dimensional case
-(the amount of memory necessary to perform the simulation would be of the order
-of tens or hundreds of gigabytes).  This work is therefore based on a
-comparison of the results obtained with the known data
-validated for the Bretherton bubble flow.  In this paper, we present
-techniques for initialization of the simulations and discuss optimal
-parameter ranges for the binary liquid LB model, such as the binary liquid parameters, the viscosity
-ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
-of film thicknesses.
-
-The VOF technique is the continuous
-interface method based on a non-uniform grid. It is of particular interest for us to compare it
-with the continuous interface method by the LBM implementation.
-
-Thus, there is the effect of
-inertia on the bubble formation.
-
-Recent experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
+Also, the binary liquid model is used.
+The model has an uniform density.
+A part of discussion is addressed in the introduction that to fully address the Bretherton problem one needs to take into the account gravitational, viscous, inertial, and surface tension forces.
+However,  \citet{shikazono-square} indicated that for square shaped capillaries radii have the
 following dependency on the capillary number $Ca$ and the Weber number $We$:
 \begin{equation}
+\label{radii:experimental}
 \begin{aligned}
@@
 \end{equation}
-In the present
-simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
-order of unity. Thus, it allows to neglect 
-inertia effects.
+Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$ implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from $R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$.
+The error of $3$ percent of the channel height is acceptable for current simulations where the Reynolds number is less than {\color{red}$100$}.
 
-\section{Steady-state case}
+Thus, having limitations of the computational memory and performance requirements the focus of the present work is a feasibility study of the binary liquid lattice Boltzmann method for the microchannel simulations with square crossections conducted in the moderate capillary number range $0.1 \leq Ca \leq 1.0$.
+We base our work on comparison with the already available studies.
+
+To further reduce the computational overhead only quarter of the channel is simulated, see Appendix \ref{appendix}.
+The next section further examines the steady-state approach, the dependency of the radii on the moderate capillary number values $0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio.
+
+\section{Results}
+
+\section{Steady-state approach}
 We performed different simulations to understand a number of steps required for the problem to come

### 2011-04-13 Text/paper.tex
Writing results. Need to play with VTK.

@@
 \usepackage{color}
-\usepackage{tikz}
+\usepackage{tabularx}
 \begin{document}
@@
 
-For simplicity, the periodic boundary conditions are applied.
-Thus, the bubble train is simulated.
-To avoid the mutual influence of bubbles on each other the channel length was chosen as $15 H_{\mathrm{eff}}$.
-Periodic boundary conditions imply as well that the fluid is accelerated by a body force in the framework of the LBM but not with the pressure difference.
-Therefore, we can not address simulations with upward and downward flows \citet{cerro-bubble}.
-It is assumed that the gravity effects are negligible and the flow is in the horizontal direction driven by the body force.
+For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
+To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
+H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
+body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
+address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
+the gravity effects are negligible and the flow is in the horizontal direction driven by the body
+force. 
 \begin{figure}[ht]
@@
 \begin{figure}
-\includegraphics*[bb=152 480 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
+\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
 \caption{The classical benchmark. The semiinfinite gas bubble
@@
 
-To further reduce the computational overhead only quarter of the channel is simulated, see Appendix \ref{appendix}.
-The next section further examines the steady-state approach, the dependency of the radii on the moderate capillary number values $0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio.
+To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
+\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
+the steady-state approach, the dependency of the radii on the moderate capillary number values
+$0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio are
+examined, see Section \ref{sec:results}.
+
+\section{Binary liquid lattice Boltzmann model}
+The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
+physical domain. It utilizes
+probability distribution functions (also known as particle populations)
+containing information about
+macroscopic variables, such as fluid density and momentum. LBE consists of
+two parts: a local collision step, and a propagation step which transports
+information from one node to another in certain
+directions specified by the discrete velocity set.
+The LBE is typically implemented as follows:
+\begin{equation}
+\label{standard:implementation}
+\begin{aligned}
+&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
+F_i,&&\text{collision step}\\
+&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
+\end{aligned}
+\end{equation}
+where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
+relaxation parameter, and $F_i$ is the external force population.
+
+The binary fluid LB model is
+based on a free-energy functional \cite{swift,landau}, and operates with two
+sets of populations: one to track the pressure and the velocity fields, and another to represent the
+phase field $\phi$ indicating gas or liquid.
+The equilibrium populations \cite{pooley-contact} are defined as:
+\begin{equation}
+\label{set:equilibrium:binary}
+\begin{aligned}
+&f_i^{eq}&&=w_i 
+\biggl(3
+p_0 - k \phi \Delta \phi
++\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
+\beta}}{2 c_s^4}\biggr)\\
+&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, i=1\div Q-1\\
+&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
+&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div Q-1\\
+&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
+\end{aligned}
+\end{equation}
+where $\Gamma$ is the mobility parameter; the chemical potential
+$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
+tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
+$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
+$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
+the sound speed parameter $c_s^2=1/3$. The bulk pressure
+is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
+sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.
+
+The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
+fluid equations as:
+\begin{equation}
+\begin{aligned}
+&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
+-\partial_{\beta}P_{\alpha \beta} +
+\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
+\end{aligned}
+\label{binary:fluid:system}
+\end{equation}
+where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
+are the relaxation parameters of density and phase fields.
+
+The system allows the separation of the liquid
+phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
+relaxation time is taken as linearly dependent on the relaxation
+times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
+$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
+to change viscosity from the gas viscosity
+$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
+$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
+accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
+A}{9}}$.
+
+Velocity set is defined as:
+\begin{equation} 
+\begin{aligned}
+&c_{ix}=\{0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1\}\\
+&c_{iy}=\{0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0\}\\
+&c_{iz}=\{0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1\}.
+\end{aligned}
+\end{equation}
+The weights are $w_0=0$, $w_{1-6}=\frac{1}{6}$ and $w_{7-18}=\frac{1}{12}$. The weights
+related to the inclusion of the surface tension are:
+\begin{equation}
+\begin{aligned}
+&w^{xx}_{1-2}=w^{yy}_{3-4}=w^{zz}_{5-6}=\frac{5}{12}\\
+&w^{xx}_{3-6}=w^{yy}_{1-2,5-6}=w^{zz}_{1-4}=-\frac{1}{3}\\
+&w^{xx}_{7-10,15-18}=w^{yy}_{7-14}=w^{zz}_{11-18}=-\frac{1}{24}\\
+&w^{xx}_{11-15,}=w^{yy}_{15-18}=w^{zz}_{7-10}=\frac{1}{12}\\
+&w^{xy}_{7,10}=-w^{xy}_{8,9}=w^{yz}_{11,14}=-w^{yz}_{12,13}=w^{zx}_{15,18}=-w^{zx}_{16,17}=\frac{1}{
+4}.
+\end{aligned}
+\end{equation}
 
 \section{Results}
+\label{sec:results}
+This section describes numerical simulations. First, we examine when the steady-state approach is
+achieved. Then, the critical capillary number is identified when the transition from the
+non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
+number is presented for the range of moderate capillary numbers $0.1 \leq Ca 1.0$. Results are
+concluded with studies of the velocity pattern,{\color{red} the aspect ratio, and the influence of
+inertia}.
 
-\section{Steady-state approach}
+\subsection{Steady-state approach}
 We performed different simulations to understand a number of steps required for the problem to come
@@
 quarter of the channel with the initial
-width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
-$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
+width as $6.5$ lattice Boltzmann units together with the body force as
+$\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
+\mathrm{x}10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
+$220000$ and
+$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
 capillary number varies only in the 4th digit, radiuses are one percent accurate. 
-
-As well we performed longer simulations in the range of moderate capillary numbers after
-$140000$, $160000$, $180000$, $200000$, $220000$, $240000$. The results for the capillary number
-based on the bubble velocity in the middle crossection are shown in Table \ref{table:steady:state}. 
-%{\color{red} What is it One of the
-%profiles in
-%the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}. }
 \begin{table}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabularx}{\textwidth}{|X|X|X|X|}
 \hline
 $N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
-$140000$&$0.0532$&$0.9379$&$1.075$\\
-$150000$&$0.0533$&$0.9375$&$1.072$\\
-$160000$&$0.0535$&$0.9372$&$1.069$\\
-$170000$&$0.0536$&$0.9369$&$1.066$\\
-$180000$&$0.0537$&$0.9366$&$1.062$\\
 \hline
-\end{tabular}\\
-\begin{tabular}{|c|c|c|c|}
+$140000$&$0.1879$&$0.8657$&$0.8708$\\
+$160000$&$0.1878$&$0.8650$&$0.8702$\\
+$180000$&$0.1882$&$0.8647$&$0.8700$\\
+$200000$&$0.1893$&$0.8645$&$0.8698$\\
+$220000$&$0.1925$&$0.8644$&$0.8697$\\
+$240000$&$0.1961$&$0.8645$&$0.8697$\\
 \hline
-$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
-$140000$&$0.3578$&$0.8657$&$0.8708$\\
-$160000$&$0.3590$&$0.8650$&$0.8702$\\
-$180000$&$0.3598$&$0.8647$&$0.8700$\\
-$200000$&$0.3603$&$0.8645$&$0.8698$\\
-$220000$&$0.3607$&$0.8644$&$0.8697$\\
-$240000$&$0.3609$&$0.8645$&$ 0.8697$\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Underresolved capillary number results and resolved Results for the steady-state case. One
+\end{tabularx}
+\caption{Results for the steady-state case. One
 can see that $200000$ steps are enough to structure
-the bubble. \label{table:steady:state}}
+the bubble and approach the steady state. The small noise in the capillary number is connected to
+the identification of the interface. One needs to address the interpolation and the spurious
+currents existing in the system. All other parameters are relaxed to the
+steady-state.\label{table:steady:state}}
 \end{table}
 
-%\begin{figure}
-%\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
-%\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
-%\caption{Slice in the middle of the bubble and
-%three-dimensional example. \label{fig:steady:state:profile:example}}
-%\end{figure}
+The bubble velocity is the velocity of the interface. One can see it with ...
+
+\subsection{Radii transition}
+Due to performance issues we can only access the
+transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
+is guaranteed. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
+were conducted. Surprisingly the transition occurs in the region computationally accessible for
+current simulations. In comparison with the results of the \citet{heil-threedim}
+($\widehat{Ca}=0.04$) our
+results are more close to the VOF continuous interface method simulation by
+\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and experimental dependency of
+\citet{shikazono-square}, see Table
+\ref{table:transition:results}. The transition happens for $Ca\approx 0.09$.  One can see
+two examples of non-symmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see Fig.
+\ref{fig:crosssections:sym}.  
+\begin{table}
+\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
+\hline
+$Ca$&$R_{axis}$&$R_{diag}$&$Re$&$R_{axis,Wang}$&$R_{diag,Wang}$\\
+\hline
+$0.053$&$0.95$&$1.01$&$0.907$&$0.99$&$0.99$\\
+$0.078$&$0.93$&$0.95$&$1.062$&$0.96$&$0.96$\\
+$0.132$&$0.89$&$0.89$&$1.80$&$0.93$&$0.93$\\
+\hline
+\end{tabularx}
+\caption{Simulation results in terms of $R_{axis}$ and $R_{diag}$ for the transition region between
+non-symmetric and axisymmetric cases. $R_{axis,Wang}$ and $R_{diag,Wang}$ are experimental
+correlations by \citet{wang-non-circular}. \label{table:transition:results}}
+\end{table}
+\begin{figure}[ht]
+\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca5.eps}
+\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca13.eps}\\
+\caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
+(left) and $Ca=1.13$ (right). Crosssections are rescaled on the half channel height. Other
+parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
+is asymmetric in comparison
+with the right one. The transition happens for $Ca\approx 0.09$.\label{fig:crosssections:sym}}
+\end{figure}
 

### 2011-04-13 Results/Force0000002/fast_extract_profiles.py

@@ -74,14 +74,6 @@ def extract_profiles(name):
     print vz.shape
     print vy.shape
 
-    #for counter in range(0,points.GetNumberOfPoints()):
-    #    coorx,coory,coorz=points.GetPoint(counter)
-    #    if coorx==0:
-    #        velx,vely,velz=velocity.GetTuple3(counter)
-    #       vz[int(coory),int(coorz)]=velz
-    #        vy[int(coory),int(coorz)]=vely
-    #        vx[int(coory),int(coorz)]=velx
-    #       phase_numpy[int(coory),int(coorz)]=phase.GetTuple1(counter)
 
     for coory in range(0,dims[1]):
         for coorz in range(0,dims[2]):
@@ -93,9 +85,9 @@ def extract_profiles(name):
             phase_numpy[coory,coorz]=phase.GetTuple1(counter)
 
 
-    numpy.savetxt("vz.txt",vz)
-    numpy.savetxt("vy.txt",vy)
-    numpy.savetxt("phase.txt",phase_numpy[1:,1:])
+    #numpy.savetxt("vz.txt",vz)
+    #numpy.savetxt("vy.txt",vy)
+    #numpy.savetxt("phase.txt",phase_numpy[1:,1:])
 
     #parameters of the binary liquid model
     k=0.04
@@ -113,10 +105,6 @@ def extract_profiles(name):
     print mid
     
     phase_mid=numpy.zeros([dims[0],dims[1]])    
-    #for counter in range(0,points.GetNumberOfPoints()):
-    #    coorx,coory,coorz=points.GetPoint(counter)
-    #    if coorz==mid:
-    #       phase_mid[int(coorx),int(coory)]=phase.GetTuple1(counter)
     for coorx in range(0,dims[0]):
         for coory in range(0,dims[1]):
             counter=mid*dims[0]*dims[1]+coory*dims[0]+coorx
@@ -133,12 +121,9 @@ def extract_profiles(name):
     axis_zero=Get_Zero(prof_axis)/(dims[0]-2.0)
     diag_zero=math.sqrt(2.0)*Get_Zero(prof_diag)/(dims[0]-2.0)
     
-    print "Velx",vx[0,mid]
-    print "Vely",vy[0,mid]
-    print "Velz",vz[0,mid]
     
     #Calculation of the capillary number
[…]

### 2011-04-13 Results/Moderate/main.cpp

@@ -1,0 +1,168 @@
+#include <mpi.h>
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=82;
+const int NY=82;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=25000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=6;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",0.000001+atof(argv[1])*0.0000002);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",2.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-04-13 Results/Velocity/main.cpp

@@ -1,0 +1,168 @@
+#include <mpi.h>
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=52;
+const int NY=52;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=20000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=8;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",0.000008+atof(argv[1])*0.0000008);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",1.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-04-13 Results/velocity.py

@@ -1,0 +1,385 @@
+#!/usr/bin/python
+
+import numpy
+import os
+import pylab
+from numpy import genfromtxt
+import vtk
+import math
+
+
+def draw_capillaries():
+    dirs=["Force0000002","Force0000002x82","Force0000005x52"]
+    style=["<","^",">"]
+    fig=pylab.figure()
+
+    for counter,dir in enumerate(dirs):
+        os.chdir(dir)
+        print os.getcwd()
+        
+        mat=numpy.loadtxt("capillaries.txt")
+        
+        #print mat
+        pylab.plot(mat[:,0],0.5*(mat[:,1]+mat[:,2]),style[counter],markersize=10,linewidth=2)
+        #pylab.plot(mat[:,0],mat[:,2],"+")
+        os.chdir("..")
+
+    
+    capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
+    radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+   
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    heil=genfromtxt("heil.csv",delimiter=';',dtype=float)
+
+    #print heil
+    #print heil.shape
+    
+    ax=fig.add_subplot(111)
+    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    
+    pylab.semilogx(heil[:,0],heil[:,1],linewidth=2)
+    pylab.semilogx(capillary_theor,radiusses,'o-',markersize=7,linewidth=2)
+    #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
+    #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
+    #pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
+
+    pylab.xlim(xmin=0.1,xmax=5)
+    #pylab.ylim(ymin=0.01)
+    #numpy.savetxt("capillary.dat",zip(capillaries,widths))
+    
+    pylab.xticks(fontsize=20)
[…]

### 2011-04-14 Text/paper.tex
Added interface velocities. When work with vtk don't use twodimensional images.

@@
 force. 
-\begin{figure}[ht]
+\begin{figure}[!]
 \includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
@@
 propagates through the liquid media. \label{fig:classical:benchmark}}
-\end{figure}[ht]
+\end{figure}[!]
 \begin{figure}
@@
 
-The bubble velocity is the velocity of the interface. One can see it with ...
+To further examine the steady-state the velocity in the streamwise direction is plotted. The values
+of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
+ to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
+component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
+towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
+clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
+examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
+to each other, then it's exactly the velocity with which the bubble propagates in the microchannel. 
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
+\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
+\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
+interface. Directions $y$ and $z$ are scaled to the half of the channel
+$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
+same velocities, which is actually the bubble velocity. For more details on velocity patterns
+see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
+\end{figure}
 
@@
 \section{Velocity pattern}
+\label{sec:velocity}
 We examined two different velocity patterns to identify change of pattern. Two capillaries numbers

### 2011-04-14 Results/velocity.py

@@ -264,78 +264,152 @@ def read_vtk_2d(name):
     data  = grid.GetPointData()
     points=grid.GetPoints()
     dims  =grid.GetDimensions()
-    #data.SetActiveScalars("Phase"); 
+    data.SetActiveScalars("Phase"); 
     data.SetActiveVectors("Velocity")
     velocity=data.GetArray("Velocity")
     phase = data.GetArray("Phase")
     
-    vel_image=vtk.vtkImageData()
-    vel_image.SetDimensions(dims[1],dims[2],1)
-    vel_image.SetSpacing(1.0, 1.0, 1.0)
-    vel_image.SetOrigin(0.0, 0.0, 0.0)
-
+    image=vtk.vtkImageData()
+    image.SetSpacing(1.0,1.0,1.0)
+    image.SetOrigin(0.0,0.0,0.0)
+    image.SetDimensions(dims[0],dims[1],dims[2])
+    image.GetPointData().SetScalars(phase)
+    image.GetPointData().SetVectors(velocity)
+    image.Update()
+    print "image=",image
     
-    Array = vtk.vtkDoubleArray()
-    Array.SetNumberOfComponents(3)
-    Array.SetNumberOfTuples(points.GetNumberOfPoints())
-    Array.Initialize()
-    Array.SetName("VelocitySlice")
-
-    ArrayPhase = vtk.vtkDoubleArray()
-    ArrayPhase.SetNumberOfComponents(1)
-    ArrayPhase.SetNumberOfTuples(points.GetNumberOfPoints())
-    ArrayPhase.Initialize()
-    ArrayPhase.SetName("PhaseSlice")
-    
-    #for counter in range(0,10):
-    #    print points.GetPoint(counter)
-    #raw_input()
-    exam=1
-    #print dims[0],dims[1],dims[2]
-    #raw_input()
-    for counter in range(0,vel_image.GetNumberOfPoints()):
-        x,y,z=vel_image.GetPoint(counter)
-        counter_big=y*dims[0]*dims[1]+x*dims[0]+exam
-        #print x,y,z
-        #print counter_big
-        
-        vz=velocity.GetTuple3(counter_big)[2]
-        vy=velocity.GetTuple3(counter_big)[1]
-        phase_value=phase.GetTuple1(counter_big)
-        Array.InsertNextTuple3(vy,vz,0.0)
-        ArrayPhase.InsertNextTuple1(phase_value)
[…]

### 2011-04-15 Text/paper.tex
Added velocity profiles plus bubble shapes.

@@
 \ref{fig:heil:three:dim}.
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
@@
 force. 
-\begin{figure}[!]
+\begin{figure}[h]
 \includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
-\caption{The classical benchmark. The semiinfinite gas bubble
-propagates through the liquid media. \label{fig:classical:benchmark}}
-\end{figure}[!]
-\begin{figure}
+\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
+propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
+free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
+\label{fig:classical:benchmark}}
+\end{figure}
+\begin{figure}[h]
 \includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
-\caption{The classical benchmark. The semiinfinite gas bubble
-propagates through the liquid media. \label{fig:lbm:benchmark}}
+\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
+the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
+length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
+Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
+force. \label{fig:lbm:benchmark}}
 \end{figure}
@@
 directions specified by the discrete velocity set.
-The LBE is typically implemented as follows:
+The LBE is typically implemented as follows \cite{ginzburg-boundary-main}:
 \begin{equation}
@@
 where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
-relaxation parameter, and $F_i$ is the external force population.
+relaxation parameter, and $F_i$ is the external force population. The external force population
+mimics the force behavior, Eq. \ref{full:navier:stokes}.
 
@@
 \beta}}{2 c_s^4}\biggr)\\
-&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, i=1\div Q-1\\
+&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, 1\leq i \leq Q-1\\
 &f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
 &g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
-\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div Q-1\\
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), \leq i \leq Q-1\\
 &g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
@@
 \begin{equation}
+\label{full:navier:stokes}
 \begin{aligned}
@@
 \caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
-interface. Directions $y$ and $z$ are scaled to the half of the channel
+interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
 $H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
@@
 
-\section{Non-axisymmetric case}
-We first start with the non-axisymmetric case. As it was indicated before for a quite low capillary
-number the flow becomes axisymmetric. For
-example, according to \cite{heil-threedim} for the capillary
-number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
-Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
-difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
- resolved around $40-50$ percent for calculations to be convergent and grid independent. This
-condition certainly implies a large grid with the computational high demand. For instance, to
-properly simulate the abovementioned example the grid should be as 
-$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
-interface whether it can produce the consistent film thicknesses for low capillary numbers. The
-interface occupies around $1-2$ lattice units.
-
-Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
-the quarter of the channel using the mirror boundary conditions (See Appendix \ref{append:sym}).
-That
-results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
-out with the pressure gradient as:
-\begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=1.6\,10^{-6}.
-\end{equation}
-The initial film thickness was taken as $6$ nodes.
-
-One can see the phase profile in the middle of the bubble in
-Fig.\ref{fig:quarter:capillary:capillary01}. For comparison reason the axisymmetric phase profile
-for $Ca=0.6$ is presented in Fig. \ref{fig:bubble:axisymmetric}. The associated capillary number is
-$0.053$ based on
-the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
-are as $R_{axis}=0.93$ and $R_{diag}=1.06$. While $R_{axis}=0.93$ is underestimated parameter and
-we can't take into the consideration apriori, one can consider parameter $R_{diag}=1.06$ the
-parameters for the capillary number $0.053$ should be according to \citet{heil-threedim} as $0.97$
-and $1.05$ according to the VOF simulations of \citet{wang-non-circular}. One can conclude that
-even though the axis film thickness is not properly resolved the diagonal values are consistent and
-can give an approximate information. 
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/quarter.eps}\\ %\hfill
-\includegraphics[width=0.97\textwidth]{Figures/stream_cap01.eps}\\
-\caption{Streamlines and phase profile of the quarter of the channel. The capillary number is
-$0.053$ and the film thickness is underresolved. \label{fig:quarter:capillary:capillary01}}
-\end{figure}
-\begin{figure}[ht]
-\includegraphics[width=0.97\textwidth]{Figures/square.eps}
-\caption{The middle bubble phase profile. One can see that the diagonal radius coincides with the
-axial radius. {\color{red} $Ca=0.5$} \label{fig:bubble:axisymmetric}}
-\end{figure}
-
-For underresolved schemes we performed a number of simulations on a GPU code with the grid of size
-$50\mathrm{x}50\mathrm{750}$. While changing the applied force in a manner indicated in Appendix
-\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
-see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
-and axial radiuses are quite significant and they coincide only when the interface is resolved
-($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
-axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
-resembles already published date of \citet{heil-threedim} in a better way, see Fig.
-\ref{fig:capillary:comparison}. 
-\begin{figure}[ht]
-\includegraphics[width=0.97\textwidth]{Figures/underresolved_capillaries.eps}
-\caption{The underresolved profiles axis and diagonal radiuses versus capillary numbers. $50$
-percent resolution starts from $Ca=1.0$, which is quite high. One can see it's exactly where
-radiuses converge instead of $Ca=0.04$. {\color{red} To delete? However, one can see from
-simulations Fig.
-\ref{fig:wang:critical:capillary} and Fig.\ref{fig:thickness:variation:ca:ten} the thickness,
-especially for Fig.\ref{fig:thickness:variation:ca:ten}, can significantly variate over a bubble
-length.} \label{fig:underresolved:capillaries}}
-\end{figure}
-
 \section{Velocity pattern}
 \label{sec:velocity}
-We examined two different velocity patterns to identify change of pattern. Two capillaries numbers
-were taken below and above the numbers indicated in Table \ref{table:recirculation:data}, as
-$Ca=0.36$ and $Ca=0.88$.Velocity streamlines are
-indicated in Fig. \ref{fig:streamlines:pattern}. One can see that the vortex indicated in
-Fig.\ref{fig:heil:recirculating} doesn't appear in present calculations. This fact can be
-explained because the simulation is carried our for the binary-liquid case but not free-surface
-interface presented in the classical Bretherton problem. In the latter case the void phase pushes
-the liquid towards the walls. In the present simulation there is no a void phase. Then the
-streamlines have the ability to go through the interface and we exactly see this situation occuring.
-\begin{figure}[ht]
-\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca036.eps}\\
-\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca088.eps}
-\caption{Streamlines for $Ca=0.36$ and $Ca=0.88$. One can see that there are no vortexes created
-before the bubble. It can be explained by the fact that the binary-liquid model is simulated
-instead of free-surface interface gas air finger flow.\label{fig:streamlines:pattern}}
+The velocity pattern is known to change its behavior depending on the capillary number. For
+relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
+frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
+vortex in front of the bubble. The transition capillary number is indicated by
+\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
+distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
+pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
+that the pressure is largely influenced by bubble frequency and slug distance. The present
+simulations are conducted for bubble trains and certain differences in terms of changing streamline
+patterns are expected. We examined velocity patterns to identify the moment of the streamlines
+pattern change. We chose two
+representative capillary numbers as $Ca=0.47$ and $Ca=0.63$. 
+One can
+see a clear transition between associated patterns. The transition capillary number
+\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
+transition occurs at $Ca=0.5-0.6$.
+\begin{figure}[h]
+\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
+\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
+\caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
+created
+before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
+is a different value from $Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
 \end{figure}
@@
 \section{Variation bubble length}
-The works of \citet{heil-threedim} of \citet{wang-non-circular} show the variation of the bubble
+The work of \citet{wang-non-circular} shows the variation of the bubble
 thickness through the length of
-bubble. The length variation is presented in Fig.\ref{fig:thickness:variation:ca:ten} and
-Fig.\ref{fig:wang:bubble:shape}. While Fig. \ref{fig:wang:bubble:shape} indicates that the diagonal
-radius has a variation for the rather small capillary numbers as $Ca<0.1$, another profile is
-presented for the $Ca=10$ and shows the variation for the diagonal radius. We examined a bunch of
-different capillaries numbers to present the diagonal radius variations over the length:
-The axis and diagonal radiuses variation over the bubble are shown in Figure
-\ref{fig:bubble:variation:capillaries}. The rather small capillary numbers one can see the increase
-in the rear menisci of the diagonal radius which coincides with the results of
-\citet{wang-non-circular}. One can see that variation over the length can be neglected
-for relatively high $Ca$ numbers. 
-\begin{figure}[ht]
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_36.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_46.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_61.eps}\hfill
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\\
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_one.eps}
-\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}
-\caption{The radiuses variations for the capillary number $Ca=0.46$.
-One can see the diagonal radius different from the axial radius near the front of the tip which is
-somehow coincides with the results of
-\citet{heil-threedim}. The results are presented for the
-capillary numbers (from left to right, top to bottom)
-$0.36$,$0.46$, $0.61$, $0.88$,$1.06$,$1.17$. \label{fig:bubble:variation:capillaries}}
+bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
+capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
+capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
+with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
+the jump.  
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{Figures/bubble_rad_axis.eps}\\
+\includegraphics[width=\textwidth]{Figures/bubble_rad_diag.eps}\\
+\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
+One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
+VOF simulations of \citet{wang-non-circular}. \label{fig:bubble:variation:capillaries}}
 \end{figure}
+
+Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
+their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
+the bubble front tip. We performed the same simulations.
+
 \begin{figure}
@@
 \section{Capillary number}
+
+While changing the applied force in a manner indicated in Appendix
+\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
+see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
+and axial radiuses are quite significant and they coincide only when the interface is resolved
+($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
+axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
+resembles already published date of \citet{heil-threedim} in a better way, see Fig.
+\ref{fig:capillary:comparison}.
+
 We performed a number of simulations with quarter CPU code and GPU train simulations and compared

### 2011-04-15 Results/bubble_length.py

@@ -1,0 +1,219 @@
+#!/usr/bin/python
+
+import numpy
+import os
+import pylab
+from numpy import genfromtxt
+import vtk
+import math
+
+
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+def Analyze_Consequence():
+    ax_zeros=[]
+    diag_zeros=[]
+    capillaries=[]
+    #file_list=["Force0000002/10/phase250000.vts","Velocity/1/phase240000.vts","Moderate/3/phase250000.vts"]
+    file_list=["LargeCap/phase250000.vts"]    
+    for number,file in enumerate(file_list):
+        print file
+        print type(number)
+        axis_zero, diag_zero, capillary=extract_profiles(file,number)
+        ax_zeros.append(axis_zero)
+        diag_zeros.append(diag_zero)
+        capillaries.append(capillary)
+    pylab.figure(1)
+    pylab.ylim(ymin=0.63,ymax=0.7)
+    pylab.xlim(xmax=20)
+    #labels=[r'''$Ca='''+str(ca)[0:4]+r'''$''' for ca in capillaries]
+    ##pylab.plot(rads_axis,"g-.",linewidth=3,markersize=4)
+    ##pylab.plot(rads_diag,"b-",linewidth=3,markersize=4)
+    
+    ##pylab.ylim(ymin=0.0,ymax=0.5)
+    ##pylab.xlim(xmax=5.1)
+    ##labels=[r'''$R_{axis}$''',r'''$R_{diag}$''']
+    #fig=pylab.figure(1)
+    #pylab.title(r'''$R_{axis}$''',fontsize=30)
+    #leg=pylab.legend(labels)
+    #pylab.xticks(fontsize=16)
+    #pylab.yticks(fontsize=16)
+    #pylab.xlabel(r'''$x$''',fontsize=20)
+    #pylab.ylabel(r'''$y$''',fontsize=20)
+    #fig.subplots_adjust(left=0.15,bottom=0.1)  
+    #pylab.savefig("bubble_rad_axis.eps",dpi=300)
+    
+    #fig=pylab.figure(2)
+    #pylab.title(r'''$R_{diag}$''',fontsize=30)
[…]

### 2011-04-15 Results/streamlines.py

@@ -1,0 +1,486 @@
+#!/usr/bin/python
+
+import numpy
+import os
+import pylab
+from numpy import genfromtxt
+import vtk
+import math
+
+
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+def Analyze_Consequence():
+    ax_zeros=[]
+    diag_zeros=[]
+    capillaries=[]
+    file_list=["Moderate/3/phase250000.vts","Moderate/7/phase250000.vts","SmallCapGrid/8/phase250000.vts"]
+    for file in file_list:
+        print file
+        axis_zero, diag_zero, capillary=extract_profiles(file)
+        ax_zeros.append(axis_zero)
+        diag_zeros.append(diag_zero)
+        capillaries.append(capillary)
+    #pylab.plot(capillaries,ax_zeros)
+    #pylab.plot(capillaries,diag_zeros)
+    #numpy.savetxt("steady.txt",zip(capillaries,ax_zeros,diag_zeros))
+    
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[1],dims[2]])
+    vy=numpy.zeros([dims[1],dims[2]])
+    vx=numpy.zeros([dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[1],dims[2]])
+    
+    print vz.shape
+    print vy.shape
+
+    for coory in range(0,dims[1]):
+        for coorz in range(0,dims[2]):
+            counter=coorz*dims[0]*dims[1]+coory*dims[0]
+            velx,vely,velz=velocity.GetTuple3(counter)
+            vz[coory,coorz]=velz
+            vy[coory,coorz]=vely
[…]

### 2011-04-15 Text/paper.tex
Almost added the relative velocity picture. Recalculated it for all simulations. Need to add appendix and conclusion.

@@
 \ref{fig:heil:three:dim}.
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
@@
 force. 
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
@@
 \end{figure}
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
@@
 the steady-state approach, the dependency of the radii on the moderate capillary number values
-$0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio are
-examined, see Section \ref{sec:results}.
+$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity, see Section
+\ref{sec:results}.
 
@@
 \end{aligned}
-\label{binary:fluid:system}
 \end{equation}
@@
 to each other, then it's exactly the velocity with which the bubble propagates in the microchannel.
-\begin{figure}[h!]
+For physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
@@
 transition occurs at $Ca=0.5-0.6$.
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
@@
 the jump.  
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{Figures/bubble_rad_axis.eps}\\
-\includegraphics[width=\textwidth]{Figures/bubble_rad_diag.eps}\\
+\begin{figure}[ht]
+\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_axis.eps}
+\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_diag.eps}\\
 \caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
 One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
-VOF simulations of \citet{wang-non-circular}. \label{fig:bubble:variation:capillaries}}
+VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
+\label{fig:bubble:variation:capillaries}}
 \end{figure}
@@
 their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
-the bubble front tip. We performed the same simulations.
-
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
-\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
-half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
+the bubble front tip which after certain distance decay to have radii equal to each other. The
+bubble shape was calculated for the large capillary number $Ca=6.43$, see Fig.
+\ref{fig:bubble:ca:large}. However, even the diagonal radius shows the jump near the front tip of
+the bubble and it resembles the shape given by \citet{heil-threedim}, but the difference is around
+$1$ percent and can be attributed to the linear interpolation error used for the interface
+calculation. 
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{Figures/bubble_ca_large.eps}
+\caption{The bubble shape for the capillary number $Ca=6.43$ is presented.$z$ increases in the
+streamwise direction. One can see that near the front tip (large $z$) the diagonal radius exhibits
+a jump which is relaxed after to the axis radius. Even this resembles
+the shape of the curves obtained by \citet{heil-threedim}, but the
+difference between radii is around $1$ percent and can be attributed to
+the linear interpolation used in the calculation of the interface.\label{fig:bubble:ca:large}}
 \end{figure}
 
-Concerning results indicated in Fig.
-\ref{fig:thickness:variation:ca:ten} simulations are unstable that means one needs to work with
-small surface tension ({\color{red} See results of Pagonabarraga}). One can see that the difference
-between diagonal radius and
-usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
-simulation is conducted by the grid $100\mathrm{x}100$ in the cross section that means that we will
-see the difference in two grid nodes. Therefore the grid is $102\mathrm{x}102\mathrm{1500}$. It is
-known from simulations that for the grid $102\mathrm{x}102\mathrm{x}1500$ the largest force gradient
-is $2\mathrm{x}10^{-5}$ with free energy binary liquid parameters $A=0.04$,$k=0.04$ and $Ca\approx
-1.0$. Therefore, we want to keep the same body force but to obtain larger capillary number. That
-can be done through the scaling indicated in Appendix \ref{append:scaling}:
-\begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
-\end{equation}
-Therefore the same body force can stand for capillary number ten times larger if the surface
-tension is ten times smaller:
-\begin{equation}
-\gamma=\sqrt{\frac{8 k A}{9}}.
-\end{equation}
-The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
-increasing the mobility of the interface to be able to adjust to the interface
-\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
-parameters and $\Gamma=20$ for the interface to be stable. 
-{\color{red} However, many authors state that increasing mobility kills
-the
-proper results of simulations}.
-{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}
+\section{Capillary number}
+The purpose of this section is to compare radii dependence on the capillary number. The film
+thicknesses were extracted from the middle of the bubble. The data presented in Fig.
+\ref{fig:capillary:comparison} is simulated with different techniques including different
+initialization, different grids as $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
+\times 82 \time 2400$. All the results are consistent. Therefore for the capillary number $Ca\geq
+0.05$ the grid independence is achieved. The results are compared with results of
+\citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
+correspondent references with the help of the program ``Engauge Digitizer''. The lattice Boltzmann
+can catch the transition between non-symmetric to axisymmetric shape of the bubble.
 
-\section{Capillary number}
+Present simulations show that the corresponding critical capillary number is close to
+$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular}. 
+In comparison with
+the continuous method VOF \cite{wang-non-circular} present simulations show less noisy trend and
+more accurate transition.
 
-While changing the applied force in a manner indicated in Appendix
-\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
-see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
-and axial radiuses are quite significant and they coincide only when the interface is resolved
-($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
-axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
-resembles already published date of \citet{heil-threedim} in a better way, see Fig.
-\ref{fig:capillary:comparison}.
-
-We performed a number of simulations with quarter CPU code and GPU train simulations and compared
-them with data obtained by \cite{heil-threedim}. After performing the simulations we determine the
-$R_{axis}$, $R_{diag}$ versus calculated
-capillary number
-\begin{equation}
-Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}.
-\end{equation}
-To perform simulations we
-started with the following parameters as
-$k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
-The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
-quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
-grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
-length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
-chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
-comparison presented in Fig. \ref{fig:capillary:comparison}.
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison.eps}
-\caption{The comparison between different runs of codes for the axial and the diagonal radiuses
-versus capillary numbers. \label{fig:capillary:comparison}}. One can see that the code mimics
-behavior of the earlier published results.
+For moderate capillary number $Ca\geq 0.2$ radii show less value than the literature results. That
+might be corresponding to the bubble train simulation or the influence of the binary-liquid model. 
+\begin{figure}[ht]
+\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_heil.eps}
+\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_wang.eps}\\
+\caption{The comparison for the axial and the diagonal radii
+versus capillary numbers between present simulations and the results of \citet{heil-threedim}
+(top), of \citet{wang-non-circular} (bottom). One can see that
+the code mimics
+behavior of the earlier published results.  \label{fig:capillary:comparison}} 
 \end{figure}
 
-\section{Train simulation}
-\label{section:bubble:train}
-As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
-examine how the distance between bubbles influences on the thickness. For this purpose the
-simulation was conducted on the grid 82x82x750. We examined different initialized distances between
-bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
-lattice units. Therefore, the corresponding initial distances between bubbles are
-$450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become $386$,
-$317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based on
-the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
-calculated radiuses (diagonal radius equal to the axes radius) are as follows
-$0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
-on each other velocity for the given body force. However, in terms of the film thicknesses those
-numbers look adequate. One can see the film thicknesses dependency on the capillary number for
-simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
-\ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
-results are consistent and in a good agreement.
-Therefore, though the bubble train mutual motion influences on the film thickness by increasing the
-associated capillary number,  however, the film thickness is in accordance with the capillary
-number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
-obtain the data. However, one should be cautious about the initialization, because the Poiseuille
-flow initialization is deviated even further due to mutual motion of bubbles.
-\begin{figure}
-\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
-\caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
-simulations. Though, one can see that the decrease of the distance between bubbles causes the
-increase in the associated capillary number, however, the correlation dependance between film
-thickness and capillary number stays the same. This can help for simulations to reduce the
-numerical domain. \label{fig:capillaries:train}}
-\end{figure}
+\section{Relative velocity}
+One of the differencies between the flow in tubes and square cappillaries is that the flow has
+different patterns. In square microchannels the main mass flow goes through corners
+\cite{heil-threedim,wang-non-circular}. This affects the mass flow significantly. One of the
+characteristics of the mass flow is the normalized relative velocity dependence on the capillary
+number. Bubble moves always faster than surrounding liquid. The following quantity is defined as
+relative velocity \cite{cerro-bubble-train}:
+\begin{equation}
+W=\frac{U_{\mathrm{bubble}}-U_{ls}}{U_{\mathrm{bubble}}},
+\end{equation}
+where $U_{\mathrm{bubble}}$ is the bubble interface velocity, $U_{ls}$ is the liquid superficial
+velocity. As far as the steady state is achieved one can take the liquid superficial velocity
+exactly at the distance in the middle between bubbles. Then, the liquid superficial velocity is
+defined as:
+\begin{equation}
+U_{ls}=\frac{\int_{A}{U_{liq}dA}}{A}.
+\end{equation}
+The relative velocity was calculated and compared with the results of \citet{wang-non-circular}.
+
+% \section{Train simulation}
+% \label{section:bubble:train}
+% As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
+% examine how the distance between bubbles influences on the thickness. For this purpose the
+% simulation was conducted on the grid 82x82x750. We examined different initialized distances
+%between
+% bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
+% lattice units. Therefore, the corresponding initial distances between bubbles are
+% $450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become
+%$386$,
+% $317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based
+%on
+% the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
+% calculated radiuses (diagonal radius equal to the axes radius) are as follows
+% $0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
+% on each other velocity for the given body force. However, in terms of the film thicknesses those
+% numbers look adequate. One can see the film thicknesses dependency on the capillary number for
+% simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
+% \ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
+% results are consistent and in a good agreement.
+% Therefore, though the bubble train mutual motion influences on the film thickness by increasing
+%the
+% associated capillary number,  however, the film thickness is in accordance with the capillary
+% number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
+% obtain the data. However, one should be cautious about the initialization, because the Poiseuille
+% flow initialization is deviated even further due to mutual motion of bubbles.
+% \begin{figure}
+% \includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
+% \caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
+% simulations. Though, one can see that the decrease of the distance between bubbles causes the
+% increase in the associated capillary number, however, the correlation dependance between film
+% thickness and capillary number stays the same. This can help for simulations to reduce the
+% numerical domain. \label{fig:capillaries:train}}
+% \end{figure}
 
 \section{Conclusion}
-The present work conducts the binary-liquid simulations with the lattice Boltzman method. The
-method is a continuous interface method. Along with similarity of results with already published
+The present work conducts the binary-liquid three-dimensional simulations with the lattice Boltzman
+method. The
+method is a continuous interface method. The results are similar with the literature  Along with
+similarity of results with already published
 results of \citet{heil-threedim} and \citet{wang-non-circular}, the differencies were found,
@@
 
+can be done through the scaling indicated in Appendix \ref{append:scaling}:
+\begin{equation}
+\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
+\end{equation}
+Therefore the same body force can stand for capillary number ten times larger if the surface
+tension is ten times smaller:
+\begin{equation}
+\gamma=\sqrt{\frac{8 k A}{9}}.
+\end{equation}
+The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
+increasing the mobility of the interface to be able to adjust to the interface
+\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
+parameters and $\Gamma=20$ for the interface to be stable. 
+{\color{red} However, many authors state that increasing mobility kills
+the
+proper results of simulations}.
+{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}
+
+To perform simulations we
+started with the following parameters as
+%k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
+%$The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
+%quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
+%grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
+%length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
+%chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
+%comparison presented in Fig. \ref{fig:capillary:comparison}.
+
 \bibliographystyle{unsrtnat}

### 2011-04-15 Results/bubble_length.py

@@ -22,8 +22,8 @@ def Analyze_Consequence():
     ax_zeros=[]
     diag_zeros=[]
     capillaries=[]
-    #file_list=["Force0000002/10/phase250000.vts","Velocity/1/phase240000.vts","Moderate/3/phase250000.vts"]
-    file_list=["LargeCap/phase250000.vts"]    
+    file_list=["Force0000002/10/phase250000.vts","Velocity/1/phase240000.vts","Moderate/3/phase250000.vts"]
+    #file_list=["LargeCap/phase250000.vts"]    
     for number,file in enumerate(file_list):
         print file
         print type(number)
@@ -31,10 +31,10 @@ def Analyze_Consequence():
         ax_zeros.append(axis_zero)
         diag_zeros.append(diag_zero)
         capillaries.append(capillary)
-    pylab.figure(1)
-    pylab.ylim(ymin=0.63,ymax=0.7)
-    pylab.xlim(xmax=20)
-    #labels=[r'''$Ca='''+str(ca)[0:4]+r'''$''' for ca in capillaries]
+    fig=pylab.figure(1)
+    #pylab.ylim(ymin=0.63,ymax=0.7)
+    #pylab.xlim(xmax=20)
+    labels=[r'''$Ca='''+str(ca)[0:4]+r'''$''' for ca in capillaries]
     ##pylab.plot(rads_axis,"g-.",linewidth=3,markersize=4)
     ##pylab.plot(rads_diag,"b-",linewidth=3,markersize=4)
     
@@ -42,25 +42,26 @@ def Analyze_Consequence():
     ##pylab.xlim(xmax=5.1)
     ##labels=[r'''$R_{axis}$''',r'''$R_{diag}$''']
     #fig=pylab.figure(1)
-    #pylab.title(r'''$R_{axis}$''',fontsize=30)
-    #leg=pylab.legend(labels)
-    #pylab.xticks(fontsize=16)
-    #pylab.yticks(fontsize=16)
-    #pylab.xlabel(r'''$x$''',fontsize=20)
-    #pylab.ylabel(r'''$y$''',fontsize=20)
-    #fig.subplots_adjust(left=0.15,bottom=0.1)  
-    #pylab.savefig("bubble_rad_axis.eps",dpi=300)
+    pylab.ylabel(r'''$R_{axis}$''',fontsize=20)
+    leg=pylab.legend(labels)
+    pylab.xticks(fontsize=16)
+    pylab.yticks(fontsize=16)
[…]

### 2011-04-15 Results/capillaries.py

@@ -8,24 +8,40 @@ import vtk
 import math
  
 def draw_capillaries():
-    dirs=["Force0000002","Force0000002x82","Force0000005x52"]
-    style=["<","^",">"]
+    #dirs=["Force0000002","Force0000002x82","Force0000005x52"]
+    #style=["kv","^",">"]
     fig=pylab.figure()
 
-    for counter,dir in enumerate(dirs):
-        os.chdir(dir)
-        print os.getcwd()
+    #for counter,dir in enumerate(dirs):
+    #    os.chdir(dir)
+    #    print os.getcwd()
+    #    
+    #    mat=numpy.loadtxt("capillaries.txt")
         
-        mat=numpy.loadtxt("capillaries.txt")
-        
-        #print mat
-        pylab.plot(mat[:,0],0.5*(mat[:,1]+mat[:,2]),style[counter],markersize=10,linewidth=2)
+    #    #print mat
+    #    pylab.plot(mat[:,0],0.5*(mat[:,1]+mat[:,2]),style[counter],markersize=10,linewidth=2)
         #pylab.plot(mat[:,0],mat[:,2],"+")
-        os.chdir("..")
+    #    os.chdir("..")
 
+    files=["cap_moderate.txt","cap_velocity.txt","cap_force0000002.txt","cap_force0000005x52.txt","cap_force0000002x82.txt"]
+
+    #read the giavedoni data (not precise though)
+    heil=genfromtxt("../Engauge/wang.csv",delimiter=',',dtype=float)
+
+    ax=fig.add_subplot(111)
+    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
     
-    capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
-    radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+    pylab.semilogx(heil[:,0],heil[:,1],"k>",markersize=7,markerfacecolor="white")
+    pylab.semilogx(heil[:,0],heil[:,2],"k<",markersize=7,markerfacecolor="white")
+   
+    res=numpy.zeros(3)
+    for file in files:
+        mat=numpy.loadtxt(file)        
+        pylab.semilogx(mat[:,0],mat[:,1],"ks",markersize=7)
+        pylab.semilogx(mat[:,0],mat[:,2],"ko",markersize=7)
[…]

### 2011-04-15 Results/relative_velocity.py

@@ -1,0 +1,631 @@
+#!/usr/bin/python
+
+import numpy
+import os
+import pylab
+from numpy import genfromtxt
+import vtk
+import math
+
+
+def draw_velocities_new():
+    files=["vel_moderate.txt","vel_velocity.txt","vel_force0000002.txt"]
+
+    fig=pylab.figure()
+
+    for counter,file in enumerate(files):
+        
+        mat=numpy.loadtxt(file)
+        
+        #print mat
+        pylab.semilogx(mat[:,0],numpy.divide((mat[:,4]-mat[:,5]),mat[:,4]),"kv",markersize=8)
+    
+    
+    #capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
+    #radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+   
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    #fig=pylab.figure()
+    widths=[]
+    velocities=[]
+    
+    #read the giavedoni data (not precise though)
+    wang=genfromtxt("../Engauge/relative_wang.csv",delimiter=',',dtype=float)
+
+    #print heil
+    #print heil.shape
+    
+    
+    #ax=fig.add_subplot(111)
+    #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
+    
+    pylab.semilogx(wang[:,0],wang[:,1],"ko",markersize=8,markerfacecolor="white")
+    #pylab.semilogx(capillary_theor,radiusses,'o-',markersize=7,linewidth=2)
+    #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
+    #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
+    #pylab.loglog(capillaries,widths,"go-",linewidth=3,markersize=10)
+
+    #pylab.xlim(xmin=0.1,xmax=5)
+    #pylab.ylim(ymin=0.01)
+    #numpy.savetxt("capillary.dat",zip(capillaries,widths))
+    
+    #pylab.xticks(fontsize=20)
+    #pylab.yticks(fontsize=20)
+    
[…]

### 2011-04-15 Results/velocity.py

@@ -120,16 +120,22 @@ def Analyze_Consequence():
     ax_zeros=[]
     diag_zeros=[]
     capillaries=[]
-    file_list=["Moderate/3/phase250000.vts","Moderate/7/phase250000.vts","SmallCapGrid/8/phase250000.vts"]
+    reynolds=[]
+    file_list=["Moderate/1/phase250000.vts","Moderate/3/phase250000.vts","Moderate/5/phase250000.vts","Moderate/7/phase250000.vts","Moderate/9/phase250000.vts"]
+               #"SmallCapGrid/8/phase250000.vts"]
+    #file_list=["Velocity/"+str(dir)+"/phase240000.vts" for dir in range(1,11)]    
+    #file_list=["Force0000002x82/2/phase180000.vts","Force0000002x82/4/phase180000.vts"] #,"Force0000005x52/8/phase250000.vts","Force0000005x52/10/phase250000.vts"]    
     for file in file_list:
         print file
-        axis_zero, diag_zero, capillary=extract_profiles(file)
+        axis_zero, diag_zero, capillary,reynolds_value=extract_profiles(file)
         ax_zeros.append(axis_zero)
         diag_zeros.append(diag_zero)
         capillaries.append(capillary)
+        reynolds.append(reynolds_value)
+        
     #pylab.plot(capillaries,ax_zeros)
     #pylab.plot(capillaries,diag_zeros)
-    #numpy.savetxt("steady.txt",zip(capillaries,ax_zeros,diag_zeros))
+    numpy.savetxt("cap_moderate.txt",zip(capillaries,ax_zeros,diag_zeros,reynolds))
     
 def extract_profiles(name):
     
@@ -198,10 +204,11 @@ def extract_profiles(name):
     number=str(capillary*100)[:2]
     if number[1]==".":
         number=number[:1]
-    pylab.savefig("phase_crossection_ca"+number+".eps")    
+    #pylab.savefig("phase_crossection_ca"+number+".eps")    
     #pylab.imshow(array['phi'],cmap="gray",extent=[0.0,15.0,0.0,1.0])
     #pylab.yticks([0.0,0.5,1.0])
 
+    reynolds=vz[0,z2%dims[2]]*2.0*(dims[0]-2)/(2.0/3.0)
     print "Reynolds=",vz[0,z2%dims[2]]*2.0*(dims[0]-2)/(2.0/3.0)
     prof_axis=phase_mid[0,1:]
[…]

### 2011-04-18 Text/paper.tex
Finished the preliminary version of the paper.

@@
 work is the continuation of the previous work. It is focused on the three-dimensional capillaries
-simulations with rectangular and square crosssections to
+simulations with square crosssections to
 validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
@@
 results for the flow between plates upto Reynolds number
-$300$. They indicate that while the influence of the established film thickness is insignificant
+$300$. The author indicates that while the influence of the established film thickness is
+insignificant
 ($7$ percent from the film thickness measured at $Re=0$), the change in Reynolds number 
@@
 the film stabilizes at the distances of $2.6-4.0$ diameters
-from the front tip depending on the Reynolds number.
-\citet{heil-threedim} measure the bubble radii at the distance as $5.5 H_{\mathrm{eff}}$ to be sufficient for $Ca\leq 10$.
-In this work the conducted simulations capillary number is not larger than unity $Ca\leq 1$.
-Thus, the film thickness is chosen to be measured in the
+from the front tip depending on the Reynolds number. \citet{heil-threedim} measure the bubble radii
+at the distance as $5.5 H_{\mathrm{eff}}$ and indicate it to be sufficient for $Ca\leq 10$. In
+this work the conducted simulations capillary number is not larger than unity $Ca\leq 6$.  Thus, the
+film thickness is chosen to be measured in the
 middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
@@
 $Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal requirement for the film thickness to be
-resolved, i.e.
-$6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to properly resolve the binary liquid microchannels bubbles motion.
-This size implies highly computational resource to be available for simulations.
+resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to
+properly resolve the binary liquid microchannels bubbles motion. This size implies highly
+computational resource to be available for simulations.
 
-One can argue that the non-uniform grid can be used.
-However, the deposited film thickness is the function of the curvatures of the front bubble tip \cite{bretherton}.
-Therefore, the grid needs to be refined just at the interface between gas and liquid.
-It is complicated to achieve for the dynamic systems where a bubble moves in the streamwise direction.
+One can argue that the non-uniform grid can be used. However, the deposited film thickness is the
+function of the curvatures of the front bubble tip \cite{bretherton}. Therefore, the grid needs to
+be refined just at the interface between gas and liquid. It is complicated to achieve for the
+dynamic systems where a bubble moves in the streamwise direction.
 
-Also, the binary liquid model is used.
-The model has an uniform density.
-A part of discussion is addressed in the introduction that to fully address the Bretherton problem one needs to take into the account gravitational, viscous, inertial, and surface tension forces.
-However,  \citet{shikazono-square} indicated that for square shaped capillaries radii have the
+Also, the binary liquid model is used. The model has an uniform density. A part of discussion is
+addressed in the introduction that to fully address the Bretherton problem one needs to take into
+the account gravitational, viscous, inertial, and surface tension forces. However, 
+\citet{shikazono-square} indicated that for square shaped capillaries radii have the
 following dependency on the capillary number $Ca$ and the Weber number $We$:
@@
 \end{equation}
-Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$ implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from $R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$.
-The error of $3$ percent of the channel height is acceptable for current simulations where the Reynolds number is less than {\color{red}$100$}.
+Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$
+implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from
+$R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$. The error of $3$ percent of the channel height is
+acceptable for current simulations where the Reynolds number is less than $10$.
 
-Thus, having limitations of the computational memory and performance requirements the focus of the present work is a feasibility study of the binary liquid lattice Boltzmann method for the microchannel simulations with square crossections conducted in the moderate capillary number range $0.1 \leq Ca \leq 1.0$.
-We base our work on comparison with the already available studies.
+Thus, having limitations of the computational memory and performance requirements the focus of the
+present work is a feasibility study of the binary liquid lattice Boltzmann method for the
+microchannel simulations with square crossections conducted in the moderate capillary number range
+$0.1 \leq Ca \leq 1.0$. We base our work on comparison with the already available studies.
 
@@
 F_i,&&\text{collision step}\\
-&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
+&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}\\
+&g_i^{*}(\bm{x},t)=\omega_{\phi} g_i^{eq}(\bm{x},t)-(1-\omega_{\phi})
+g_i(\bm{x},t),&&\text{collision step}\\
+&g_i(\bm{x}+\bm{c_i},t+1)=g_i^{*}(\bm{x},t),&&\text{propagation step},
 \end{aligned}
 \end{equation}
-where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
-relaxation parameter, and $F_i$ is the external force population. The external force population
+where $\{f,g\}_i$ are the probability distribution functions in the direction $\bm{c_i}$, $\omega$
+is the
+relaxation parameter, $\omega_{\phi}$ is the phase relaxation parameter, and $F_i$ is the external
+force population which
 mimics the force behavior, Eq. \ref{full:navier:stokes}.
@@
 $\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
-tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
-$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
+tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($19$
+for the $D3Q19$ model); the tensor
 $Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
 the sound speed parameter $c_s^2=1/3$. The bulk pressure
-is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
-sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.
-
-The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
-fluid equations as:
-\begin{equation}
-\label{full:navier:stokes}
-\begin{aligned}
-&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
-&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
--\partial_{\beta}P_{\alpha \beta} +
-\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
-&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
-\end{aligned}
-\end{equation}
-where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
-$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
-are the relaxation parameters of density and phase fields.
-
-The system allows the separation of the liquid
-phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
-relaxation time is taken as linearly dependent on the relaxation
-times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
-$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
-to change viscosity from the gas viscosity
-$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
-$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
-accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
-A}{9}}$.
-
-Velocity set is defined as:
+is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. Velocity set is defined as:
 \begin{equation} 
@@
 
+The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
+fluid equations as:
+\begin{equation}
+\label{full:navier:stokes}
+\begin{aligned}
+&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=F_{\alpha}
+-\partial_{\beta}P_{\alpha \beta} +
+\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
+\end{aligned}
+\end{equation}
+where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
+are the relaxation parameters of density and phase fields. The first equation of system
+(\ref{standard:implementation}) simulates the continuity equation and the Navier-Stokes equation.
+The lattice Boltzmann update rule for the second set $g_i$ simulates the phase propagation with the
+supplied velocity from the Navier-Stokes equation.
+
+The system allows the separation of the liquid
+phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
+relaxation time is taken as linearly dependent on the relaxation
+times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
+$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
+to change viscosity from the gas viscosity
+$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
+$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
+accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
+A}{9}}$.
+
 \section{Results}
 \label{sec:results}
-This section describes numerical simulations. First, we examine when the steady-state approach is
+This section describes numerical simulations. One can refer to the appendix \ref{append:scaling} for
+initialization and scaling procedures. The simulations were conducted on different grids
+corresponding to fluid domain grids as $100 \times 100 \times 1500$, $160 \times 160 \times 1500$,
+$160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. The body forces were varied as
+$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$. The binary-liquid parameters were kept to be in
+the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
+and $\tau_{\mathrm{gas}}=0.7$ giving the gas-liquid dynamic viscosity ratio $10$. The relaxation
+parameter for the phase field was $\tau_{\phi}=1.0$.
+
+First, we examine when the steady-state approach is
 achieved. Then, the critical capillary number is identified when the transition from the
@@
 number is presented for the range of moderate capillary numbers $0.1 \leq Ca 1.0$. Results are
-concluded with studies of the velocity pattern,{\color{red} the aspect ratio, and the influence of
-inertia}.
+concluded with studies of the velocity pattern.
 
 \subsection{Steady-state approach}
+\label{sec:steady:state}
 We performed different simulations to understand a number of steps required for the problem to come
@@
 before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
-is a different value from $Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
+is a different value from $Ca=0.69$ \cite{heil-threedim}. However, we performed a number of
+different types of simulations (different force, width initialization). The critical transition
+capillary number $\widehat{Ca}$ is not well imposed and varies from $\widehat{Ca}=0.6-0.7$.
+\label{fig:streamlines:pattern}}
 \end{figure}
@@
 \end{equation}
-The relative velocity was calculated and compared with the results of \citet{wang-non-circular}.
+The relative velocity was calculated and compared with the results of \citet{wang-non-circular},
+see Fig. \ref{fig:relative:velocity}. One can see the agreement between simulations indicating that
+the mass transfer coefficients can be captured accurately. 
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{Figures/relative_velocity.eps}
+\caption{Relative velocity comparison between current simulations and simulations of
+\citet{wang-non-circular}. One can see a qualitative agreement. \label{fig:relative:velocity}}
+\end{figure}
+
+The relative velocity is very important quantity in the mass flow characterization
+\cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and the
+inertia effects \citet{heil-bretherton} influence significantly the mass transfer coefficient. The
+scope of this work is the feasibility study of the lattice Boltzmann model to be applied for
+three-dimensional microchannels with square crosssections. The future work is planned to study mass
+transfer coefficient and solve associated advection-diffusion coefficient. However, the preliminary
+studies of the relative velocity show that the binary liquid model can be used for accurate mass
+transfer characterization.
 
@@
 \section{Conclusion}
-The present work conducts the binary-liquid three-dimensional simulations with the lattice Boltzman
-method. The
-method is a continuous interface method. The results are similar with the literature  Along with
-similarity of results with already published
-results of \citet{heil-threedim} and \citet{wang-non-circular}, the differencies were found,
-especially in the velocity pattern. Results evedince that the lattice Boltzmann binary liquids
-simulations can be used for simulation of gas bubbles in the microchannel.
+The present work conducts the binary-liquid simulations for three-dimensional channels with square
+crosssections with the lattice Boltzman method. The
+method is a continuous interface method. By resolving properly the film thickness as twice as the
+interface thickness \cite{kuzmin-binary2d} the results were shown to be in consistence with the
+available in literature results. The film thickness dependency on the capillary number, the
+transition from the axisymmetric to symmetric case, the velocity pattern were investigated. While
+the goal of this work is the feasibility study, the more numerical studies are needed to
+understand the influence of the slug length, inertia effects, pressure distributions
+\cite{kreutzer-taylor,yue-mass} to design the microchannels. Results evedince that the lattice
+Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels.
 
 \appendix
-\section{Symmetric boundary conditions}
-\label{append:sym}
-For the memory save we performed the simulation of the quarter of the channel while using the
-antisymmetric boundary conditions. In the macroscopic variables the mirror boundary conditions
-require to have the same scalar fields, but for velocity only tangential velocities are the same
-but perpendicular have the opposite directions. This implies the following in terms of the
-macroscopic variables:
-\begin{equation}
-\rho_B = \rho_F, \quad \phi_B = \phi_F \quad U_{B\tau}=U_{F\tau}, U_{B n}=U_{F n}, 
-\end{equation}
- where $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.
-
-In terms of the lattice Boltzmann populations this conditions require the same set of the
-populations for the boundary node as for the fluid node. However, the set is mirrored against the
-plane of the boundary. This case allows to conserve scalar fields, to have the same tangential
-velocities and to have opposite normal velocities. The grid was chosen as 52x52x1500, which is the
-effective calculated domain as 100x100x1500 grid nodes.
-
 \section{Scaling procedure}
 \label{append:scaling}
-For scaling purposes one need to obtain setup a body force as $\frac{\mathrm{d}P}{\mathrm{d}x}$. If
-one obtained a certain velocity of the bubble with the associated gradient then another gradient
-can be obtained through the porpotionality condition as:
+The thorough scaling procedure is described in the work \cite{kuzmin-binary2d}. Here only the
+outline is presented on how the simulations are initialized:
+\begin{description}
+ \item[Capillary number] One first needs to set the capillary number and the Reynolds number for
+simulations. 
+ \item[Film thickness] After the capillary number is prescribed, one needs to approximate the axial
+bubble radius $R_{axial}$, the corresponding film thickness is $\delta=1-R_{axial}$. One can use
+correlations of \citet{shikazono-square} or of \citet{kreutzer-taylor}. It can be done either by
+taking correlations from numerical simulations by
+\citet{heil-bretherton} and \citet{giavedoni-numerical}.  
+\item[Grid choice] After specifying the film thickness, one needs to choose the associate
+number of nodes to resolve the film thickness. For binary liquid parameters used in simulations,
+$A=0.004,0.04$ and $k=0.004,0.04$ the
+interface is spread over approximately $5$ lattice units. To obtain grid independent results one
+needs to choose the
+film thickness to be $2-2.5$ times larger than the interface thickness \cite{kuzmin-binary2d}.
+Giving the number of nodes to resolve the film thickness and giving the lattice Boltzmann
+benchmark, see Section \ref{sec:numerical:benchmark}, one can obtain the grid dimensions.
+
+\item[Velocity] The Reynolds and the capillary numbers are supplied from the physical world. We
+accept the following definitions:
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca,
+\label{definitions:numbers}
+\begin{aligned}
+&Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma} \\
+&Re=\frac{U_{\mathrm{bubble}} H_{\mathrm{eff}}}{\mu_{\mathrm{liq}}}.
+\end{aligned}
 \end{equation}
-or in other words th functionality can be extended as:
+In Eq. \eqref{definitions:numbers} the interface tension coefficient is prescribed by parameters of
+the binary liquid model $k$ and $A$, the effective channel height $H_{\mathrm{eff}}$ is obtained
+from the film thickness resolution, $\mu_{\mathrm{liq}}$ is defined through the relaxation rate
+$\tau_{\mathrm{liq}}$. The relaxation rate $\tau_{liq}$ is prescribed from the gas-liquid viscosity
+ratio $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}$, where $\mu_{\mathrm{gas}}$ is defined through
+the gas relaxation rate $\tau_{\mathrm{gas}}$. Parameter $\tau_{\mathrm{gas}}$ have constraints
+$\tau_{gas}\geq 0.5$, therefore constraining the linked parameter $\tau_{\mathrm{liq}}$. Moreover,
+the stability of the lattice Boltzmann scheme is known to deteriorate with the relaxation parameter
+$\tau_{\mathrm{liq},\mathrm{gas}}\approx 0.5$ \cite{kuzmin-trt-stability}. On the other hand the
+accuracy of the lattice Boltzmann liquid model is detiorated as
+$\Bigl(\tau_{\mathrm{liq},\mathrm{gas}}-\frac{1}{2}\Bigr)^2$ \cite{ginzburg-trt-simple-hydro}. Thus,
+there is a compromise of the choice of parameters $\tau_{\mathrm{liq}}$ and $\tau_{\mathrm{gas}}$.
+The gas-liquid viscosity ratio in the current simulations was chosen as
+$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and the relaxation parameters were chosen to be
+in the safe region, i.e. $\tau_{liq}=2.5$ and $\tau_{gas}=0.7$.
+
+Given the considerations of the relaxation parameters, one can obtain the desired bubble velocity
+$U_{\mathrm{bubble}}$ from the equation for the Reynolds number (\ref{definitions:numbers}).
+Substituting the bubble velocity $U_{\mathrm{bubble}}$ to the definition of the capillary number,
+Eq. (\ref{definitions:numbers}), one can obtain the interface tension which is connected with the
+binary liquid parameters $k$ and $A$. The stability of the lattice Boltzmann scheme implies the
+simultaneous change of $\tau_{\phi}$ with the change of parameters $k$ and $A$
+\cite{pagonabarraga-parameters}. Therefore, the recommendation is if Reynolds number is less than
+$50-100$ one can neglect the inertia effects and obtain the bubble velocity straight from the
+capillary number limiting the complex interplay of the parameters for the LBM simulations to be
+stable: 
+\begin{equation}
+\begin{aligned}
+&U_{\mathrm{bubble}}=Ca \frac{\gamma}{\mu_{\mathrm{liq}}}=Ca \frac{\sqrt{\frac{8 k
+A}{9}}}{\frac{1}{3}(\tau_{\mathrm{liq}}-\frac{1}{2})}\\
+\end{aligned}
+\end{equation}
+where usually $k=A=0.004,0.04$.
+
+\item[Body force] 
+It is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in simulations. The
+flow in the simulations is conducted by imposing the body force. Thus, the assumption is needed of
+how the body force is connected with the bubble velocity $U_{\mathrm{bubble}}$. We assume that the
+flow is close to planar Poiseuille flow with $U_{\mathrm{bubble}}$ being the maximum in the
+Poiseuille profile for the square shaped microchannels:
+\begin{equation}
+\label{poiseuille:square:profile}
+\dfrac{\mathrm{d}P}{\mathrm{d}z}=\frac{\mu_{\mathrm{liq}} U_{\mathrm{bubble}}}{\sum_{i\leq 0,j\leq
+0}{\dfrac{16}{\pi^6} \dfrac{(-1)^i (-1)^j}{(2 i+1)(2 j+1)\bigl[\frac{(2 i
++1)^2}{W_{\mathrm{eff}}}+\frac{(2 j+1)^2}{H_{\mathrm{eff}^2}}\bigr]}}}.
+\end{equation}
+Though the Poiseuille profile assumption works reasonably well \cite{kuzmin-binary2d}, instead of
+calculating sums (\ref{poiseuille:square:profile})  one is suggested to start simulations with the
+body force $\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-5}$ for grids of size $50\times 50 \times
+1500$ and the capillary number $Ca\approx0.1-1.0$.  
+\end{description}
+After imposing the body force the simulations can be run. A typical simulation runs approximately
+$100000-300000$ time steps to reach steady-state, see Section \ref{sec:steady:state}. To design the
+simulations one needs to keep proportianality between parameters through equations above. For
+example if one knows the capillary number $Ca_1$ and the corresponding body force
+$\frac{\mathrm{d}P_1}{\mathrm{d}z}$ for simulations conducted on the grid with the characteristic
+number of nodes $N_y^1$ , the body force for another simulation can be calculated as:
 \begin{equation}
@@
 
-can be done through the scaling indicated in Appendix \ref{append:scaling}:
+\section{Symmetric boundary conditions}
+\label{append:sym}
+To reduce a computational load in terms of memory and CPU power we performed the simulation of the
+quarter of the channel using the antisymmetric boundary conditions. In the macroscopic variables the
+mirror boundary conditions require to have same scalar fields. For velocity boundary conditions only
+tangential velocities are the same, perpendicular to boundary velocities have opposite
+signs. This implies the following in terms of the
+macroscopic variables:
 \begin{equation}
-\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
+\rho_B = \rho_F, \quad \phi_B = \phi_F \quad U_{B\tau}=U_{F\tau}, U_{B n}=U_{F n}, 
 \end{equation}
-Therefore the same body force can stand for capillary number ten times larger if the surface
-tension is ten times smaller:
+where subscripts $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.
+
+For the lattice Boltzmann model the incoming populations on the boundary node need to be specified.
+For generality we assume that the plane of the symmetry $\alpha=0$, where $\alpha$ is the
+direction $x$ or $y$. Then, for lattice Boltzmann populations the symmetric boundary conditions can
+be expressed through the following steps:
+\begin{description}
+\item[I] The first step is to find the complimentary directions. The complimentary $j$-th direction
+to the $i$-th direction should satisfy three conditions:
 \begin{equation}
-\gamma=\sqrt{\frac{8 k A}{9}}.
+\begin{aligned}
+&c_{i\alpha}=-c_{j\alpha}, \text{ where $\alpha=$ is the symmetry plane }\\
+&c_{i\beta}=-c_{j\beta}, \text{ where $\beta\neq\alpha$ }
+\end{aligned}
 \end{equation}
-The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
-increasing the mobility of the interface to be able to adjust to the interface
-\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
-parameters and $\Gamma=20$ for the interface to be stable. 
-{\color{red} However, many authors state that increasing mobility kills
-the
-proper results of simulations}.
-{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}
 
-To perform simulations we
-started with the following parameters as
-%k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
-%$The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
-%quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
-%grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
-%length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
-%chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
-%comparison presented in Fig. \ref{fig:capillary:comparison}.
+Note that the complimentary directions for the directions parallel to the symmetry plane coincide
+with the original direction.
+
+\item[II] The second step is taking the distribution populations from the nearest fluid node and
+apply them to the boundary node utilizing complimentary directions:
+\begin{equation}
+f_{i,B}=f_{j,F},
+\end{equation}
+where $j$ is the index of the direction complimentary to the $i$-th direction.
+\end{description}
+
+By applying the procedure described above, one can obtain that the scalar fields and the tangential
+velocity are conserved, the normal flux is reversed. In terms of the lattice Boltzmann populations
+this condition improve the performance and memory requirements by a factor of $4$.
 

### 2011-04-18 Results/VelocityModerate/main.cpp

@@ -1,0 +1,168 @@
+#include <mpi.h>
+#include <iostream>
+#include <math.h>
+#include <fstream>
+#include <string>
+#include <sstream>
+
+#include "mpi_singleton.h"
+
+//Inclusion of lattice implementation
+#include "lattice.hpp"
+#include "solver.hpp"
+
+
+#include "params_list.h"
+#include "descriptor.h"
+
+//Lattice Boltzmann initialization and parameters
+const int NX=52;
+const int NY=52;
+const int NZ=1500;
+const int NPOP=19;
+const int NUM=NX*NY*NZ;
+const int NUMTOTAL=NUM*NPOP;
+const int NUMTIME=250001;
+const int NUMOUTPUT=20000;
+const int NUMSIGNAL=50;
+
+//Binary-liquid initialization
+const int width=8;
+const int radius=6;
+const double rhol=1.0;
+const double rhog=1.0;
+
+int main(int argc,char* argv[])
+{
+
+	//Parallel Debugging
+	#ifdef DEBUG
+     	int DebugWait=1;
+        while (DebugWait);
+    #endif /* DEBUG */
+
+    //Specify main communicator
+	MPISingleton mpi_singleton;
+
+ 	ParamsList params;
+
+	if (argc>1)
+		params.add("force_z",0.000016+atof(argv[1])*0.000001);
+	else
+		params.add("force_z",0.00003);
+
+	std::cout<<params("force_z").value<double>()<<"\n";    
+	//Specify parameters
+	params.add("omega",1.0);
+	params.add("aconst", 0.04);
+	params.add("kconst", 0.04);
+	params.add("gammaconst",1.0);
+	params.add("phase_gradient",0.0);
+	params.add("phase_wall",0.0);
+	params.add("rho_wall",0.5);
+	params.add("tau_phi",1.0);
+    params.add("tau_liq",2.5);
+    params.add("tau_gas",0.7);
+    params.add("force_x",0.0);
+    params.add("force_y",0.0);
+    //params.add("force_z",0.00003);
+    params.add("NX",NX);
+    params.add("NY",NY);
+    params.add("NZ",NZ);
+
+    Solver<Descriptor> solver(params);
+
+    //Solver initialization from the file
+    //solver.load_file("equili");
+
+	//Initialization
+	    //Density and phase field initialization
+	double rho_temp;
+	double phase_temp;
[…]

### 2011-04-18 Results/relative_velocity.py

@@ -13,6 +13,10 @@ def draw_velocities_new():
 
     fig=pylab.figure()
 
+    wang=genfromtxt("../Engauge/relative_wang.csv",delimiter=',',dtype=float)
+    pylab.semilogx(wang[:,0],wang[:,1],"ko",markersize=8,markerfacecolor="white")
+    
+    
     for counter,file in enumerate(files):
         
         mat=numpy.loadtxt(file)
@@ -33,8 +37,7 @@ def draw_velocities_new():
     velocities=[]
     
     #read the giavedoni data (not precise though)
-    wang=genfromtxt("../Engauge/relative_wang.csv",delimiter=',',dtype=float)
-
+  
     #print heil
     #print heil.shape
     
@@ -42,7 +45,6 @@ def draw_velocities_new():
     #ax=fig.add_subplot(111)
     #capillaries=numpy.array(velocities)*(2.0/3.0)/math.sqrt(8.0*0.04*0.04/9.0)
     
-    pylab.semilogx(wang[:,0],wang[:,1],"ko",markersize=8,markerfacecolor="white")
     #pylab.semilogx(capillary_theor,radiusses,'o-',markersize=7,linewidth=2)
     #pylab.loglog(giavedoni[:,0],giavedoni[:,1]/2.0,"bD-",linewidth=3,markersize=10)
     #pylab.loglog(capillary_theor,width_theor,"ys--",linewidth=3,markersize=10)
@@ -52,20 +54,20 @@ def draw_velocities_new():
     #pylab.ylim(ymin=0.01)
     #numpy.savetxt("capillary.dat",zip(capillaries,widths))
     
-    #pylab.xticks(fontsize=20)
-    #pylab.yticks(fontsize=20)
+    pylab.xticks(fontsize=20)
+    pylab.yticks(fontsize=20)
     
-    #pylab.ylabel(r'''$R_{diag},R_{axes}$''',fontsize=30)
-    #pylab.xlabel(r'''$Ca$''',fontsize=30)
+    pylab.ylabel(r'''$W$''',fontsize=30)
+    pylab.xlabel(r'''$Ca$''',fontsize=30)
     
     
     
     #labels=[r'''$H_{eff}='''+str(value-2)+r'''$''' for value in ny]
-    #leg=pylab.legend(["CPU results","Refined grid","Large body force","Heil","GPU results"],fancybox=True)
+    leg=pylab.legend(["Liu-Wang Simulations","Simulations"],loc=2,fancybox=True)
[…]

### 2011-04-18 Results/streamlines.py

@@ -22,7 +22,8 @@ def Analyze_Consequence():
     ax_zeros=[]
     diag_zeros=[]
     capillaries=[]
-    file_list=["Moderate/3/phase250000.vts","Moderate/7/phase250000.vts","SmallCapGrid/8/phase250000.vts"]
+    #file_list=["ModerateVelocity/3/phase240000.vts","Moderate/5/phase250000.vts","SmallCapGrid/8/phase250000.vts"]
+    file_list=["ModerateVelocity/3/phase240000.vts","Moderate/5/phase250000.vts"]
     for file in file_list:
         print file
         axis_zero, diag_zero, capillary=extract_profiles(file)
@@ -479,7 +480,7 @@ if __name__=="__main__":
     #draw_capillaries()
     #Analyze_Consequence() 
     #name="Velocity/10/phase240000.vts"    
-    name="Force0000002/8/phase250000.vts"    
+    name="VelocityModerate/10/phase240000.vts"    
     pyngl_streamline(name)    
     #extract_streamlines()
     #read_vtk_2d(name)    

### 2011-04-26 Text/paper.tex
Added Michal's changes.

@@
 \begin{abstract}
-Classical Bretherton problem describes the propagation of gas fingers through liquid media with
+The classical Bretherton problem describes the propagation of gas fingers through liquid media with
 thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
 complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
-the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
+the applicability and parameters choice for the two-dimensional case of the Bretherton problem (flow
 between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
-\cite{kuzmin-binary2d}. This
-work is the continuation of the previous work. It is focused on the three-dimensional capillaries
-simulations with square crosssections to
-validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
-number $0.1\leq Ca \leq 1.0$.  
-The flow is driven by a body force,
-and
+\cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate the applicability
+of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.1\leq Ca \leq 1.0$,
+we focus on the simulation of three-dimensional capillaries with square cross section.
+The flow is driven by a body force, and
 periodic boundary conditions are applied in the streamwise direction. The results show that the
 binary liquid model is able to capture a number of phenomena happening in three-dimensional
-capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
-on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
-used to simulate the Bretherton problem with good accuracy. 
+capillaries, such as the existence of a vortex in front of the bubble and the dependency
+of the bubble radii on the capillary number.  We conclude that lattice Boltzmann free energy binary
+liquid model can be used to simulate the Bretherton problem with good accuracy. 
 \end{abstract}
@@
 \end{equation}
-where $U_{\mathrm{bubble}}$ is the propagation bubble velocity, $\mu_{\mathrm{liq}}$ is the
+where $U_{\mathrm{bubble}}$ is the bubble's propagation velocity, $\mu_{\mathrm{liq}}$ is the
 kinematic liquid viscosity and $\gamma$ is the interfacial surface tension between gas and liquid.
@@
 The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
-is of significant interests for chemical industry as it is widely used in chemical monolith
-microreactors \cite{kreutzer-pressure-drop}. The large mass transfer can be achieved because of the
+is of significant interest for chemical industry as it is widely used in chemical monolith
+microreactors \cite{kreutzer-pressure-drop}. Large mass transfers can be achieved because of the
 large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. The heat transfer is
 also increased in comparison with one phase flows \cite{fukugata-levelset}. While it
-is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
-not possible to extrapolate it to the wider range of capillary and Reynolds numbers used in the
-chemical
-industry. Thus, the desire of consistent numerical simulations arises.
+is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, the
+results cannot be extrapolated to the wider range of capillary and Reynolds numbers used in the
+chemical industry.  Thus, the desire of consistent numerical simulations arises.
 
@@
 \citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the Bretherton
-analysis is valid only for small capillaries numbers $Ca\leq 0.003$ and deviates for larger ones.
+analysis is valid only for small capillary numbers $Ca\leq 0.003$ and deviates from actual
+measurements for larger ones.
 That is caused by complex interplay of the gravitational,
@@
 inertia effects
-are negligible for $Ca \leq 0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
-number from $0$ to $70$ for two-dimensional geometry flows. Later on, \citet{heil-bretherton}
+are negligible for $Ca \leq 0.05$ and have moderate impact for $Ca>0.05$ in the range of Reynolds
+number from $0$ to $70$ for two-dimensional flows. Later on, \citet{heil-bretherton}
 extended
-results for the flow between plates upto Reynolds number
+the results for the flow between plates upto Reynolds number
 $300$. The author indicates that while the influence of the established film thickness is
@@
 
-In comparison with the two-dimensional Bretherton flow  (circular tubes, flow between
-parallel
-plates), there is a vast number of experimental
-results available for the three-dimensional case as microchannels with square, triangular and
-rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
+In comparison with the two-dimensional Bretherton flow, there is a vast number of experimental
+results available for the three-dimensional case for microchannels with square, triangular and
+rectangular cross sections. For instance, \citet{cerro-bubble-train} performed a number of
 experiments for a
@@
 
-In comparison with two-dimensional geometries, the experimental works
+The experimental works
 \cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
@@
 \cite{heil-threedim,wong-films} that for rectangular
-or square shaped capillaries there is a transition for a
-certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
-symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
+or square-shaped capillaries there is a transition for a
+certain capillary number, where the bubble cross section changes from non-axisymmetric to axisymmetric.
+The non-axisymmetric case is attributed to the axial radius being 
 different from the
-diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
-the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
+diagonal radius and the bubble having a non-circular shape in the channel crossection. In this case
+the bubble shape mimics the shape of the square and looks like a square with rounded corners. The dependence
 of the diagonal and axial radii on the capillary number is shown in Fig. 
@@
 \end{figure}
-The transition number between non-axisymmetric case and symmetric is indicated in a number of
+The transition capillary number between non-axisymmetric case and symmetric one is indicated in a number of
 works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
@@
 the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. One of the example for the bubble radii
-dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
+radius of the droplet dependant on the capillary number. An example for the bubble radii
+dependence on the capillary number is presented in Fig. \ref{fig:heil:three:dim}.
 
@@
 slug and bubble cross sections and menisci appearance.
-\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
-square and rectangular shaped capillaries. They indicated the transition
+\citet{heil-threedim} performed three-dimensional simulations for circular-,
+square- and rectangular-shaped capillaries. They indicated the transition
 capillary number where the streamlines pattern
@@
 numbers, {\color{red} Fig. \ref{fig:streamlines:pattern} }. For the square channel the critical
-capillary number is $Ca=0.691$.  Authors also found an empirical correlation which
-allows to put
-different aspect ratio rectangular channels simulation results on one
-curve. It is also indicated that for certain aspect ratio microchannels $\alpha=\frac{a}{b}\geq
-2.04$ the interface can not make up the
+capillary number is $Ca=0.691$.  The authors also found an empirical correlation which
+makes it possible to put on one curve the results of simlations of rectengular channels with different aspect ratios.
+It is also indicated that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
+2.04$, the interface can not make up the
 axisymmetric case for any capillary number. \citet{wang-non-circular} performed numerical
@@
 non-circular
-shaped capillaries as square and triangular shaped capillaries. They also investigated the relative
+shaped capillaries such as square and triangular shaped capillaries. They also investigated the relative
 velocity of the bubble for
@@
 are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
-While the finite element methods solve the Bretherton flow as a free-surface problem with a sharp
-interface but without underlying details for gas, the lattice Boltzmann method which is a continuous
-interface method arises as a promising alternative for simulation of gas finger propagation. The
-continuous interface has more flexebility in simulations involving coalescence and/or droplet
-breakup.
+The finite element methods solve the Bretherton flow as a free-surface problem with a sharp
+interface but without underlying details for gas.  The lattice Boltzmann method on the other hand,
+as a continuous interface method provides more flexibility in simulations involving coalescence and/or
+droplet breakup, and thus arises as a promising alternative for simulation of gas finger propagation.
 
@@
 It was shown that the free-energy binary liquid model, which is a phase field method, simulates
-with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine the
-three-dimensional geometry simulations conducted by the free-energy binary liquid lattice Boltzmann
-method. We focus on the feasibility study of the numerical method for the microchannels with
-square crosssections.
+with reasonable accuracy the Bretherton/Taylor phenomena. The goal of this work is to examine 
+if similar conclusions can be drawn for three-dimensional flows in microchannels with square cross sections,
+simulated using the free-energy binary liquid lattice Boltzmann
+method.
 
@@
 the
-viscosity ratio. That is why the uniform density binary liquid model is suitable for this kind of
-simulations. The goal of this work is to do a feasibility study of the LBM
-binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
-problem. The work results are in good agreement with other simulations
-\cite{heil-threedim,wang-non-circular}.
+viscosity ratio, which is why the uniform density binary liquid model is suitable for this kind of
+simulations.
+
+% THIS HAS ALREADY BEEN STATED BEFORE
+% The goal of this work is to do a feasibility study of the LBM
+% binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
+% problem.
+
+% TODO: This sentence should probably be moved to a summary/conclusions section?
+The work results are in good agreement with other simulations \cite{heil-threedim,wang-non-circular}.
 
@@
 is limited to the low capillary number regime \cite{giavedoni-numerical}. The Shan-Chen model was
-used to simulate the Bretherton problem, which is sometimes referred to contain thermodynamically
+used to simulate the Bretherton problem, which is sometimes said to contain a thermodynamically
 inconsistent interface \cite{nourgaliev-breakup}.
@@
 \cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
-dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in Fig.
-\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
-The film thickness in this case is measured at the inlet. In lattice Boltzmann framework such
+dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in
+Fig.~\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
+The film thickness in this case is measured at the inlet. In the lattice Boltzmann framework such
 formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
 dynamic coupling of the inlet and outlet conditions \cite{giavedoni-numerical}. Instead, we
-proposed the numerical benchmark indicated in Fig. \ref{fig:lbm:benchmark}. The dimensions of the
+proposed the numerical benchmark indicated in Fig.~\ref{fig:lbm:benchmark}. The dimensions of the
 channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
@@
 the film stabilizes at the distances of $2.6-4.0$ diameters
-from the front tip depending on the Reynolds number. \citet{heil-threedim} measure the bubble radii
-at the distance as $5.5 H_{\mathrm{eff}}$ and indicate it to be sufficient for $Ca\leq 10$. In
-this work the conducted simulations capillary number is not larger than unity $Ca\leq 6$.  Thus, the
+from the front tip depending on the Reynolds number. \citet{heil-threedim} measured the bubble radii
+at the distance as $5.5 H_{\mathrm{eff}}$ and indicated it to be sufficient for $Ca\leq 10$. In
+this work the conducted simulations capillary number is not larger than unity $Ca\leq 6$.
+%%% TODO: ^ How is 6 not larger than unity?
+
+Thus, the
 film thickness is chosen to be measured in the
@@
 
-For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
-To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
-H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
-body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
+For simplicity, periodic boundary conditions are applied so that effectively a bubble train is simulated
+intead of a single bubble.  The mutual influence of bubbles on each other is minized by the use of a long
+channel ($15 H_{\mathrm{eff}}$). Periodic boundary conditions imply as well that the fluid is accelerated by a
+body force in the framework of the LBM and not with a pressure difference. Therefore, we can not
 address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
@@
 \includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
-\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
+\caption{The classical benchmark layout is presented. A semiinfinite gas bubble
 propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
@@
 \includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
-\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
+\caption{The lattice Boltzmann benchmark layout as used in present calculations. The dimensions of
 the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
 length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
-Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
+Periodic boundary conditions are applied in the streamwise direction. The flow is driven by a body
 force. \label{fig:lbm:benchmark}}
@@
 
-It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least as twice large
+It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least twice as large
 as the interface thickness. The interface thickness in the continuous formulation of the
@@
 resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to
-properly resolve the binary liquid microchannels bubbles motion. This size implies highly
-computational resource to be available for simulations.
+properly resolve the binary liquid microchannels bubbles motion. This size implies that relatvely large
+computational resources are required to perform the simulations.
 
-One can argue that the non-uniform grid can be used. However, the deposited film thickness is the
-function of the curvatures of the front bubble tip \cite{bretherton}. Therefore, the grid needs to
-be refined just at the interface between gas and liquid. It is complicated to achieve for the
+One can argue that a non-uniform grid could be used. However, the deposited film thickness is a
+function of the curvatures of the front bubble tip \cite{bretherton} and the grid would need to
+be refined just at the interface between gas and liquid. This is complicated to achieve for
 dynamic systems where a bubble moves in the streamwise direction.
+To reduce the computational overhead, we decided to take a simpler approach and exploit the
+inherent symmetry of the problem by simulating only a quarter of the channel (see Appendix
+\ref{append:sym} for details).
 
-Also, the binary liquid model is used. The model has an uniform density. A part of discussion is
+It should be noted that the binary liquid model we used has a uniform density. A part of discussion is
 addressed in the introduction that to fully address the Bretherton problem one needs to take into
@@
 
-Thus, having limitations of the computational memory and performance requirements the focus of the
-present work is a feasibility study of the binary liquid lattice Boltzmann method for the
-microchannel simulations with square crossections conducted in the moderate capillary number range
-$0.1 \leq Ca \leq 1.0$. We base our work on comparison with the already available studies.
+% AGAIN, THIS IS REPEATS WHAT HAS ALREADY BEEN SAID
+%Thus, having limitations of the computational memory and performance requirements the focus of the
+%present work is a feasibility study of the binary liquid lattice Boltzmann method for the
+%microchannel simulations with square crossections conducted in the moderate capillary number range
+%$0.1 \leq Ca \leq 1.0$. We base our work on comparison with the already available studies.
 
-To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
-\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
+The next section briefly explains the binary liquid lattice Boltzmann model. Then
 the steady-state approach, the dependency of the radii on the moderate capillary number values
@@
 information from one node to another in certain
-directions specified by the discrete velocity set.
+directions specified by a discrete velocity set.
 The LBE is typically implemented as follows \cite{ginzburg-boundary-main}:
@@
 $\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
-to change viscosity from the gas viscosity
+to smoothly change viscosity from the gas viscosity
 $\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
-$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
-accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
+$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ in a region when a
+phase gradient exists. The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
 A}{9}}$.
@@
 \label{sec:results}
-This section describes numerical simulations. One can refer to the appendix \ref{append:scaling} for
-initialization and scaling procedures. The simulations were conducted on different grids
-corresponding to fluid domain grids as $100 \times 100 \times 1500$, $160 \times 160 \times 1500$,
-$160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. The body forces were varied as
-$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$. The binary-liquid parameters were kept to be in
+This section describes numerical simulations. One can refer to the Appendix~\ref{append:scaling} for
+initialization and scaling procedures. The simulations were conducted on the following 
+fluid domain grids $100 \times 100 \times 1500$, $160 \times 160 \times 1500$,
+$160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. Body forces were varied as
+$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$. The binary-liquid model parameters were kept in
 the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
-and $\tau_{\mathrm{gas}}=0.7$ giving the gas-liquid dynamic viscosity ratio $10$. The relaxation
+and $\tau_{\mathrm{gas}}=0.7$ giving a gas-liquid dynamic viscosity ratio of $10$. The relaxation
 parameter for the phase field was $\tau_{\phi}=1.0$.
@@
 \label{sec:steady:state}
-We performed different simulations to understand a number of steps required for the problem to come
-on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
+We performed different simulations to understand a number of steps required for the simulation to
+settle down to a steady state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
 quarter of the channel with the initial
@@
 $220000$ and
-$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
+$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
 capillary number varies only in the 4th digit, radiuses are one percent accurate. 
@@
 \caption{Results for the steady-state case. One
+% TODO:structure?
 can see that $200000$ steps are enough to structure

### 2011-05-04 Text/paper.tex
Started Jos corrections.

@@
 \begin{abstract}
-The classical Bretherton problem describes the propagation of gas fingers through liquid media with
-thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
+The classical Bretherton problem describes the propagation of gas fingers through liquid media in a
+narrow channel with
+thin liquid films between bubbles and channel walls. The bubble shape and flow patterns are
 complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
-the applicability and parameters choice for the two-dimensional case of the Bretherton problem (flow
-between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
+the applicability and parameter choice for the two-dimensional case of the Bretherton problem (flow
+between parallel plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
 \cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate the applicability
 of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.1\leq Ca \leq 1.0$,
-we focus on the simulation of three-dimensional capillaries with square cross section.
+we focus on the simulation of three-dimensional capillaries with square cross sections.
 The flow is driven by a body force, and
@@
 The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
-narrow channels. Depending on the geometry it was found that the deposited film thickness
+narrow channels. Depending on the channel geometry it was found that the deposited film thickness
 is a complicated function of the capillary number $Ca$:
@@
 \end{equation}
-where $U_{\mathrm{bubble}}$ is the bubble's propagation velocity, $\mu_{\mathrm{liq}}$ is the
-kinematic liquid viscosity and $\gamma$ is the interfacial surface tension between gas and liquid.
+where $U_{\mathrm{bubble}}$ is the bubble velocity, $\mu_{\mathrm{liq}}$ is the
+dynamic liquid viscosity and $\gamma$ is the interfacial tension between gas and liquid.
 
@@
 \cite{bretherton,heil-bretherton}. 
-The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
+The problem of predicting flow patterns and associated mass transfers for Bretheron-type flows
 is of significant interest for chemical industry as it is widely used in chemical monolith
 microreactors \cite{kreutzer-pressure-drop}. Large mass transfers can be achieved because of the
-large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. The heat transfer is
-also increased in comparison with one phase flows \cite{fukugata-levelset}. While it
-is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, the
+large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. Heat transfer is
+also increased in comparison with single phase flow \cite{fukugata-levelset}. While it
+is possible to calculate the flow analytically for small capillary numbers and negligible Reynolds
+number \cite{bretherton}, the
 results cannot be extrapolated to the wider range of capillary and Reynolds numbers used in the
-chemical industry.  Thus, the desire of consistent numerical simulations arises.
+chemical industry.  Thus, the desire for consistent numerical simulations arises.
 
@@
 experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
-\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the Bretherton
+\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the analytical
+Bretherton
 analysis is valid only for small capillary numbers $Ca\leq 0.003$ and deviates from actual
 measurements for larger ones.
-That is caused by complex interplay of the gravitational,
+This is caused by a complex interplay of gravitational,
 interfacial, inertial and viscous forces \cite{gupta-review}. Historically, 
@@
 the results for the flow between plates upto Reynolds number
-$300$. The author indicates that while the influence of the established film thickness is
+$300$. This author indicates that while the influence on the established film thickness is
 insignificant
-($7$ percent from the film thickness measured at $Re=0$), the change in Reynolds number 
+($7$ percent deviation from the film thickness measured at $Re=0$), the change in Reynolds number 
 influences significantly the pressure distrubution and the flow field near the front bubble tip.
@@
 propagation of the semi-infinite finger in liquid media one needs to take into the account the
-viscous, gravitational, surface and inertia forces \cite{gupta-review}.
+viscous, gravitational, surface, and inertia forces \cite{gupta-review}.
 
-In comparison with the two-dimensional Bretherton flow, there is a vast number of experimental
-results available for the three-dimensional case for microchannels with square, triangular and
+In comparison with the two-dimensional Bretherton flows (circular tubes, flow between plates), there
+is a vast number of experimental
+results available for three-dimensional cases of microchannels with square, triangular and
 rectangular cross sections. For instance, \citet{cerro-bubble-train} performed a number of
@@
 bubble-train flow in capillaries of
-circular and square cross section for horizontal, upward and downward flows.
+circular and square cross sections for horizontal, upward and downward flows.
 \citet{shikazono-square} obtained
 experimental
-results for the deposition length dependency on the
+results for the deposition film thickness dependency on the
 capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
-shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
-based on the capillary number $Ca$ and the Weber number $We=Re\,Ca$.
+shaped mirochannels. They also presented the experimental correlation for the prediction of bubble
+radii
+based on the capillary number $Ca$ and the Weber number $We=Re\,Ca$, where the Reynolds number is
+defined as:
+\begin{equation}
+Re=\frac{U_{\mathrm{bubble}} H_{\mathrm{eff}}}{\nu_{\mathrm{liq}}},
+\end{equation}
+where $H_{\mathrm{eff}}$ is the channel height and $\nu_{\mathrm{liq}}$ is the kinematic liquid
+viscosity.
 
@@
 \cite{heil-threedim,wong-films} that for rectangular
-or square-shaped capillaries there is a transition for a
-certain capillary number, where the bubble cross section changes from non-axisymmetric to axisymmetric.
-The non-axisymmetric case is attributed to the axial radius being 
+or square-shaped capillaries there is a transition from an asymmetric to an axisymmetric shape for a
+certain capillary number.
+The non-axisymmetric shape is attributed to the wall normal radius (in Fig.
+\ref{fig:lbm:benchmark} the wall normal radius is measured along $x$ or $y$ axes) being 
 different from the
 diagonal radius and the bubble having a non-circular shape in the channel crossection. In this case
-the bubble shape mimics the shape of the square and looks like a square with rounded corners. The dependence
+the bubble shape mimics the shape of a rectangle and looks like a rectangle with rounded corners.
+From hereon we adapt the convention ``axial`` for the radius measured along the $x$ or $y$ axes,
+see Fig. \ref{fig:lbm:benchmark} \cite{heil-threedim}. The dependency
 of the diagonal and axial radii on the capillary number is shown in Fig. 
@@
 \includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
-\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
-capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
-number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
+\caption{\citet{heil-threedim} results for the variation of the deposited liquid for a range of
+capillary numbers. One can see the asymmetry between diagonal and axial diameters for the capillary
+numbers $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
 \end{figure}
-The transition capillary number between non-axisymmetric case and symmetric one is indicated in a number of
+The transition capillary number for square channels between non-axisymmetric case and axisymmetric
+one is indicated in a number of
 works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
@@
 the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. An example for the bubble radii
+radius of the bubble dependent on the capillary number. An example for the bubble radii
 dependence on the capillary number is presented in Fig. \ref{fig:heil:three:dim}.
@@
 polygonal capillaries and calculated bubble shapes for different
-slug and bubble cross sections and menisci appearance.
+slug and channel cross sections.
 \citet{heil-threedim} performed three-dimensional simulations for circular-,
-square- and rectangular-shaped capillaries. They indicated the transition
+square- and rectangular-shaped capillaries. They indicated a transition
 capillary number where the streamlines pattern
 changes from having a vortex in front of the bubble and not having it for larger capilllary
-numbers, {\color{red} Fig. \ref{fig:streamlines:pattern} }. For the square channel the critical
-capillary number is $Ca=0.691$.  The authors also found an empirical correlation which
-makes it possible to put on one curve the results of simlations of rectengular channels with different aspect ratios.
+numbers,  Fig. \ref{fig:streamlines:pattern}. For a square channel the critical
+capillary number for this phenomena is $Ca=0.691$.  \citet{heil-threedim} also found an empirical
+correlation which
+makes it possible to put on one curve the results of radii dependency on the capillary number
+simulations of rectangular channels with different aspect ratios.
 It is also indicated that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
-2.04$, the interface can not make up the
-axisymmetric case for any capillary number. \citet{wang-non-circular} performed numerical
-simulations by the VOF technique for the
-non-circular
-shaped capillaries such as square and triangular shaped capillaries. They also investigated the relative
-velocity of the bubble for
-different capillary numbers.
+2.04$, the interface does not become
+axisymmetric for any capillary number. \citet{wang-non-circular} performed numerical
+simulations using the Volume of Fluid (VOF) technique for
+capillaries with non-circular (square, triangular) cross sections. They also investigated the relative
+velocity of the bubble to the slug liquid velocity  for
+a number of capillary numbers.
 
 As it is mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
-phenomena are the Volume of Fluid (VOF) method \cite{wang-non-circular}, the level-set method
+phenomena are the VOF method \cite{wang-non-circular}, the level-set method
 \cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It is
-also indicated \cite{gupta-review}, that the new techniques which are still in the development stage
+also indicated \cite{gupta-review} that the new techniques which are still in the development stage
 are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
-The finite element methods solve the Bretherton flow as a free-surface problem with a sharp
-interface but without underlying details for gas.  The lattice Boltzmann method on the other hand,
-as a continuous interface method provides more flexibility in simulations involving coalescence and/or
+The finite element method solves the Bretherton flow as a free-surface problem with a sharp
+interface but without underlying details for the gas.  The lattice Boltzmann method on the other
+hand,
+as a continuous interface method, provides more flexibility in simulations involving coalescence
+and/or
 droplet breakup, and thus arises as a promising alternative for simulation of gas finger propagation.
 
-In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between plates.
+In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between parallel
+plates.
 It was shown that the free-energy binary liquid model, which is a phase field method, simulates
-with reasonable accuracy the Bretherton/Taylor phenomena. The goal of this work is to examine 
+with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine 
 if similar conclusions can be drawn for three-dimensional flows in microchannels with square cross sections,
@@
 method. 
+%As it was indicated by \citet{gupta-review} the flow in microchannels is a complex
+%interplay of viscous, gravitational, inertial and surface forces. For the sake of comparison, we
+%specifically design the numerical benchmark to avoid the influence of gravitational and inertial
+%forces.
 
@@
 ferrofluids \cite{kuzmin-aniso}, and multiphase flows
-\cite{swift,Shan-chen:extended}. The method is a particle method which allows to simulate physical
+\cite{swift,Shan-chen:extended}. The method is a particle method which allows one to simulate physical
 phenomena on the microscopic level. For instance, the introduction of the force or potential on the
@@
 simulates two liquids with the assumption of uniform overall
-density. While the classical Bretherton problem is stated for gas and liquid, which are of
+density. As it was indicated by \citet{gupta-review} the flow in microchannels is a complex
+interplay of viscous, gravitational, inertial and surface forces, and the classical Bretherton
+problem is stated for gas and liquid, which are of
 significantly
-different densities and viscosities, the parameters were carefully chosen as to neglect the inertia
-effects, see Section \ref{sec:numerical:benchmark}. Moreover, the results of
+different densities and viscosities, the parameters in our simulations were carefully chosen to
+avoid  inertia
+effects, see Section \ref{sec:numerical:benchmark}. Moreover, the numerical results of
 \citet{giavedoni-numerical} and \citet{heil-bretherton} show
@@
 simulations.
-
 % THIS HAS ALREADY BEEN STATED BEFORE
@@
 
-One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
+One should acknowledge the works of \citet{pagonabarraga-fingers} on menisci
 in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
@@
 found discrepancies with the classical Bretherton theory, which
-is limited to the low capillary number regime \cite{giavedoni-numerical}. The Shan-Chen model was
-used to simulate the Bretherton problem, which is sometimes said to contain a thermodynamically
+is limited to the low capillary number regime \cite{giavedoni-numerical}. They used the Shan-Chen model,
+ which is sometimes said to contain a thermodynamically
 inconsistent interface \cite{nourgaliev-breakup}.
@@
 the results section. The results cover the film thickness dependency on the capillary number, the
-relative velocity of the bubble and the vortex profiles. The paper
+relative to the slug liquid velocity of the bubble and the vortex profiles. The paper
 is concluded with a summary of the main findings.
@@
 dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in
-Fig.~\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
-The film thickness in this case is measured at the inlet. In the lattice Boltzmann framework such
+Fig.~\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid
+medium.
+The film thickness in this case is measured at the inlet. In the lattice Boltzmann framework such a
 formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
@@
 channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
-initialized bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
-the film stabilizes at the distances of $2.6-4.0$ diameters
+initial bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
+the film in the two-dimensional geometries stabilizes at distances of $2.6-4.0$ diameters
 from the front tip depending on the Reynolds number. \citet{heil-threedim} measured the bubble radii
-at the distance as $5.5 H_{\mathrm{eff}}$ and indicated it to be sufficient for $Ca\leq 10$. In
-this work the conducted simulations capillary number is not larger than unity $Ca\leq 6$.
-%%% TODO: ^ How is 6 not larger than unity?
-
+at the distance of $5.5 H_{\mathrm{eff}}$ and indicated it to be sufficient for $Ca\leq 10$. In
+this work the following relation holds for all conducted simulations: $0.05 \geq Ca \leq 6$.
 Thus, the
@@
 For simplicity, periodic boundary conditions are applied so that effectively a bubble train is simulated
-intead of a single bubble.  The mutual influence of bubbles on each other is minized by the use of a long
-channel ($15 H_{\mathrm{eff}}$). Periodic boundary conditions imply as well that the fluid is accelerated by a
+instead of a single bubble.  The mutual influence of bubbles on each other is minimized by the use
+of a long
+channel ($15 H_{\mathrm{eff}}$). Periodic boundary conditions imply as well that the fluid is driven
+by a
 body force in the framework of the LBM and not with a pressure difference. Therefore, we can not
-address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
-the gravity effects are negligible and the flow is in the horizontal direction driven by the body
-force. 
-\begin{figure}[ht]
+address simulations with upward and downward flows \cite{cerro-bubble-train} which require a
+simulataneouse imposement of pressure boundary conditions and the body force. The pressure boundary
+conditions for the binary liquid model is a challenge. Thus, it is assumed the flow is in the
+horizontal direction only and the upward/downward flows with pressure difference imposed are not
+simulated. 
+\begin{figure}[htb!]
 \includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
@@
 \end{figure}
-\begin{figure}[ht]
+\begin{figure}[htb!]
 \includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
@@
 free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
-resolved at least as $6-10$ nodes.
+resolved as at least $6-10$ nodes.
 
-\cite{heil-threedim} specify that the diagonal bubble radius is different from the
-axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
-$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal requirement for the film thickness to be
-resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to
-properly resolve the binary liquid microchannels bubbles motion. This size implies that relatvely large
+\citet{heil-threedim} specify that the diagonal bubble radius is different from the
+axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii are the same for larger values
+of
+$Ca$ with a value $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal
+requirement for the film thickness to be
+resolved, i.e. $6$ lattice nodes, one would obtain a grid size of $600\times 600 \times 9000 $ to
+properly resolve the binary liquid flow. This size implies that relatively
+large
 computational resources are required to perform the simulations.
 
-One can argue that a non-uniform grid could be used. However, the deposited film thickness is a
+One can argue that a non-uniform grid could be used. However, the film thickness is a
 function of the curvatures of the front bubble tip \cite{bretherton} and the grid would need to
@@
 
+% TODO: This paragraph should be revisited, somehow it reads a little weirdly..
 It should be noted that the binary liquid model we used has a uniform density. A part of discussion is
 addressed in the introduction that to fully address the Bretherton problem one needs to take into
-the account gravitational, viscous, inertial, and surface tension forces. However, 
-\citet{shikazono-square} indicated that for square shaped capillaries radii have the
-following dependency on the capillary number $Ca$ and the Weber number $We$:
-\begin{equation}
-\label{radii:experimental}
-\begin{aligned}
-&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
-&R_{axis}=
-\begin{cases}
-1, &R_{diag}>1\\
-R_{diag}, &R_{diag}\leq 1.
-\end{cases}
-\end{aligned}
-\end{equation}
-Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$
-implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from
-$R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$. The error of $3$ percent of the channel height is
-acceptable for current simulations where the Reynolds number is less than $10$.
+the account gravitational, viscous, inertial, and surface tension forces. Though many works
+\cite{heil-bretherton,heil-threedim,shikazono-square,kreutzer-taylor} indicate that the inertia
+effects are of significant importance, the results are of quite large deviations
+
+%However, 
+%\citet{shikazono-square} indicated that for square-shaped capillaries the bubble radii have the
+%following dependency on the capillary number $Ca$ and the Weber number $We$:
+%\begin{equation}
+%\label{radii:experimental}
+%\begin{aligned}
+%&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
+%&R_{axis}=
+%\begin{cases}
+%1, &R_{diag}>1\\
+%R_{diag}, &R_{diag}\leq 1.
+%\end{cases}
+%\end{aligned}
+%\end{equation}
+%Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$
+%implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from
+%$R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=1.0$. The error of $3$ percent of the channel height is
+%acceptable for current simulations where the Reynolds number is less than $10$.
 
@@
 
-The next section briefly explains the binary liquid lattice Boltzmann model. Then
-the steady-state approach, the dependency of the radii on the moderate capillary number values
-$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity, see Section
+The next section briefly explains the binary liquid lattice Boltzmann model.
+The steady-state approach, the dependency of the radii on the moderate capillary number values
+$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity are discussed in Section
 \ref{sec:results}.
@@
 based on a free-energy functional \cite{swift,landau}, and operates with two
-sets of populations: one to track the pressure and the velocity fields, and another to represent the
+sets of populations: one ($f_i$) to track the pressure and the velocity fields, and another ($g_i$) to represent the
 phase field $\phi$ indicating gas or liquid.
@@
 &g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
-\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), \leq i \leq Q-1\\
+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1 \leq i \leq Q-1\\
 &g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
@@
 where $\Gamma$ is the mobility parameter; the chemical potential
-$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
-tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($19$
+$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is a parameter related to the surface
+tension; $A$ is a parameter of the free-energy model; $Q$ is the number of directions ($19$
 for the $D3Q19$ model); the tensor
@@
 the sound speed parameter $c_s^2=1/3$. The bulk pressure
-is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. Velocity set is defined as:
+is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The discrete velocity set is defined as:
 \begin{equation} 
@@
 phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
-relaxation time is taken as linearly dependent on the relaxation
+relaxation time is taken to be linearly dependent on the relaxation
 times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
-$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
-to smoothly change viscosity from the gas viscosity
+$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This
+makes it possible to smoothly change viscosity from the gas viscosity
 $\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
@@
 non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
-number is presented for the range of moderate capillary numbers $0.1 \leq Ca 1.0$. Results are
+number is presented for the range of moderate capillary numbers $0.1 \leq Ca \leq 1.0$. Results are
 concluded with studies of the velocity pattern.
@@
 \label{sec:steady:state}
-We performed different simulations to understand a number of steps required for the simulation to
-settle down to a steady state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
-quarter of the channel with the initial
-width as $6.5$ lattice Boltzmann units together with the body force as
-$\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\mathrm{x}10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
+We performed different simulations to understand the number of steps required for the simulation to
+settle down to a steady state. The grid to be simulated is $52 \times 52 \times 750$ which simulates
+a quarter of the channel with the initial interface
+width of $6.5$ lattice Boltzmann units together with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
+\cdot 10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
 $220000$ and
 $240000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
-capillary number varies only in the 4th digit, radiuses are one percent accurate. 
+capillary number varies only in the 4th digit, the radii are one percent accurate. 
 \begin{table}
@@
 To further examine the steady-state the velocity in the streamwise direction is plotted. The values
-of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
+of the velocity are taken on the contour where $\phi=0$. This corresponds
  to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
 component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
-towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
+towards the reader, $z$ is the streamwise direction. As fluxes inside the bubble have
 clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
-examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
-to each other, then it's exactly the velocity with which the bubble propagates in the microchannel.
-For physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
+examine only the points corresponding to the front and the rear bubble tips. If one finds them equal
+to each other, then it is exactly the velocity with which the bubble propagates in the microchannel.
+For a physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
 \begin{figure}[ht]
@@
 The velocity pattern is known to change its behavior depending on the capillary number. For
-relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
-frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
-vortex in front of the bubble. The transition capillary number is indicated by
-\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
+relatively low capillary numbers $Ca<0.6$, in the frame comoving with the bubble there exists a vortex 
+in front of it \cite{heil-threedim}. For larger capillary numbers it is indicated that there is no
+vortex in front of the bubble. The transition capillary number specified by
+\citet{heil-threedim} for the rectangular channel is $Ca=0.691$. However, the
 distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
@@
 
-\section{Variation bubble length}
+\section{Bubble length variation}
 The work of \citet{wang-non-circular} shows the variation of the bubble
@@
 Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
-their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
+their simulations the non-symmetric shape was observed even for larger capillary numbers $Ca>4$ near
 the bubble front tip which after certain distance decay to have radii equal to each other. The
@@
 \ref{fig:capillary:comparison} is simulated with different techniques including different
-initialization, different grids as $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
+initialization, and different grids: $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
 \times 82 \time 2400$. All the results are consistent. Therefore for the capillary number $Ca\geq
-0.05$ the grid independence is achieved. The results are compared with results of
+0.05$ grid independence is achieved. The results are compared with the results of
 \citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
-correspondent references with the help of the program ``Engauge Digitizer''. The lattice Boltzmann
-can catch the transition between non-symmetric to axisymmetric shape of the bubble.
+correspondent references with the help of the program ``Engauge Digitizer''. We observe that
+lattice Boltzmann simulations exhibit 
+the transition between non-symmetric and axisymmetric shape of the bubble.
 
@@
 In comparison with
-the continuous method VOF \cite{wang-non-circular} present simulations show less noisy trend and
-more accurate transition.
+the continuous method VOF \cite{wang-non-circular} present simulations show a less noisy trend and
+a more accurate transition.
 
-For moderate capillary number $Ca\geq 0.2$ radii show less value than the literature results. That
-might be corresponding to the bubble train simulation or the influence of the binary-liquid model. 
+For moderate capillary numbers $Ca\geq 0.2$, the radii are smaller than the literature results. This
+may be caused by the bubble train simulation or the influence of the binary-liquid model. 
 \begin{figure}[ht]
@@
 characteristics of the mass flow is the normalized relative velocity dependence on the capillary
-number. Bubble moves always faster than surrounding liquid. The following quantity is defined as
+number. THe bubble always moves faster than the surrounding liquid. The following quantity is defined as
 relative velocity \cite{cerro-bubble-train}:
@@
 \end{equation}
-where $U_{\mathrm{bubble}}$ is the bubble interface velocity, $U_{ls}$ is the liquid superficial
+where $U_{\mathrm{bubble}}$ is the bubble interface velocity, and $U_{ls}$ is the liquid superficial
 velocity. As far as the steady state is achieved one can take the liquid superficial velocity
@@
 The relative velocity was calculated and compared with the results of \citet{wang-non-circular},
-see Fig. \ref{fig:relative:velocity}. One can see the agreement between simulations indicating that
+see Fig. \ref{fig:relative:velocity}. One can see agreement between the simulations indicating that
 the mass transfer coefficients can be captured accurately. 
@@
 
-The relative velocity is very important quantity in the mass flow characterization
-\cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and the
-inertia effects \citet{heil-bretherton} influence significantly the mass transfer coefficient. The
-scope of this work is the feasibility study of the lattice Boltzmann model to be applied for
-three-dimensional microchannels with square crosssections. The future work is planned to study mass
-transfer coefficient and solve associated advection-diffusion coefficient. However, the preliminary
+The relative velocity is an important quantity in mass flow characterization
+\cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and 
+inertia effects \citet{heil-bretherton} significantly influence the mass transfer coefficient. 
+Future work is planned to study this coefficient and calculate the associated advection-diffusion coefficient.
+ However, preliminary
 studies of the relative velocity show that the binary liquid model can be used for accurate mass
@@
 \section{Conclusion}
-The present work conducts the binary-liquid simulations for three-dimensional channels with square
-crosssections with the lattice Boltzman method. The
-method is a continuous interface method. By resolving properly the film thickness as twice as the
-interface thickness \cite{kuzmin-binary2d} the results were shown to be in consistence with the
-available in literature results. The film thickness dependency on the capillary number, the
-transition from the axisymmetric to symmetric case, the velocity pattern were investigated. While
-the goal of this work is the feasibility study, the more numerical studies are needed to
+The present work presents results of binary-liquid simulations for three-dimensional channels with square
+cross sections with the lattice Boltzman method. By resolving properly the film thickness as twice the
+interface thickness \cite{kuzmin-binary2d} the results were shown to be consistent with those
+available in the literature. The film thickness dependency on the capillary number, the
+transition from the asymmetric to symmetric case, and the velocity pattern were investigated. While
+the goal of this work is a feasibility study, more numerical studies are needed to
 understand the influence of the slug length, inertia effects, pressure distributions
-\cite{kreutzer-taylor,yue-mass} to design the microchannels. Results evedince that the lattice
+\cite{kreutzer-taylor,yue-mass} on the design of microchannels. Our results show that the lattice
 Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels.
@@
 \label{append:scaling}
-The thorough scaling procedure is described in the work \cite{kuzmin-binary2d}. Here only the
-outline is presented on how the simulations are initialized:
+The scaling procedure is thoroughly described in our previous work \cite{kuzmin-binary2d}. Here, we present
+only an outline of how the simulations are initialized:
 \begin{description}
@@
 bubble radius $R_{axial}$, the corresponding film thickness is $\delta=1-R_{axial}$. One can use
-correlations of \citet{shikazono-square} or of \citet{kreutzer-taylor}. It can be done either by
+correlations of \citet{shikazono-square} or of \citet{kreutzer-taylor}. It can also be done by
 taking correlations from numerical simulations by
 \citet{heil-bretherton} and \citet{giavedoni-numerical}.  
-\item[Grid choice] After specifying the film thickness, one needs to choose the associate
+\item[Grid choice] After specifying the film thickness, one needs to choose the
 number of nodes to resolve the film thickness. For binary liquid parameters used in simulations,
@@
 film thickness to be $2-2.5$ times larger than the interface thickness \cite{kuzmin-binary2d}.
-Giving the number of nodes to resolve the film thickness and giving the lattice Boltzmann
+Given the number of nodes to resolve the film thickness and given the lattice Boltzmann
 benchmark, see Section \ref{sec:numerical:benchmark}, one can obtain the grid dimensions.
 
-\item[Velocity] The Reynolds and the capillary numbers are supplied from the physical world. We
+\item[Velocity] The Reynolds and capillary numbers are supplied from the physical world. We
 accept the following definitions:
@@
 ratio $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}$, where $\mu_{\mathrm{gas}}$ is defined through
-the gas relaxation rate $\tau_{\mathrm{gas}}$. Parameter $\tau_{\mathrm{gas}}$ have constraints
-$\tau_{gas}\geq 0.5$, therefore constraining the linked parameter $\tau_{\mathrm{liq}}$. Moreover,
+the gas relaxation rate $\tau_{\mathrm{gas}}$. The parameter $\tau_{\mathrm{gas}}\geq 0.5$, therefore
+ constraining the linked parameter $\tau_{\mathrm{liq}}$. Moreover,
 the stability of the lattice Boltzmann scheme is known to deteriorate with the relaxation parameter
 $\tau_{\mathrm{liq},\mathrm{gas}}\approx 0.5$ \cite{kuzmin-trt-stability}. On the other hand the
-accuracy of the lattice Boltzmann liquid model is detiorated as
+accuracy of the lattice Boltzmann liquid model detiorates as
 $\Bigl(\tau_{\mathrm{liq},\mathrm{gas}}-\frac{1}{2}\Bigr)^2$ \cite{ginzburg-trt-simple-hydro}. Thus,
@@
 simultaneous change of $\tau_{\phi}$ with the change of parameters $k$ and $A$
-\cite{pagonabarraga-parameters}. Therefore, the recommendation is if Reynolds number is less than
+\cite{pagonabarraga-parameters}. Therefore, if Reynolds number is less than
 $50-100$ one can neglect the inertia effects and obtain the bubble velocity straight from the
@@
 \item[Body force] 
-It is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in simulations. The
-flow in the simulations is conducted by imposing the body force. Thus, the assumption is needed of
-how the body force is connected with the bubble velocity $U_{\mathrm{bubble}}$. We assume that the
-flow is close to planar Poiseuille flow with $U_{\mathrm{bubble}}$ being the maximum in the
+It is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in simulations. Since the
+flow in the simulations is driven by a body force, a connection between this force and the
+bubble velocity $U_{\mathrm{bubble}}$ has to be established. We assume that the
+flow is close to the planar Poiseuille flow with $U_{\mathrm{bubble}}$ being the maximum in the
 Poiseuille profile for the square shaped microchannels:
@@
 Though the Poiseuille profile assumption works reasonably well \cite{kuzmin-binary2d}, instead of
-calculating sums (\ref{poiseuille:square:profile})  one is suggested to start simulations with the
+calculating sums (\ref{poiseuille:square:profile}) we suggest to simply start simulations with the
 body force $\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-5}$ for grids of size $50\times 50 \times
@@
 $100000-300000$ time steps to reach steady-state, see Section \ref{sec:steady:state}. To design the
-simulations one needs to keep proportianality between parameters through equations above. For
+simulations one needs to keep proportionality between parameters through the equations above. For
 example if one knows the capillary number $Ca_1$ and the corresponding body force
@@
 \label{append:sym}
-To reduce a computational load in terms of memory and CPU power we performed the simulation of the
-quarter of the channel using the antisymmetric boundary conditions. In the macroscopic variables the
-mirror boundary conditions require to have same scalar fields. For velocity boundary conditions only
-tangential velocities are the same, perpendicular to boundary velocities have opposite
-signs. This implies the following in terms of the
-macroscopic variables:
+To reduce the computational load in terms of memory and CPU power by a factor of 4, we performed the simulation of 
+only a quarter of the channel using symmetric boundary conditions.  The scalar macroscopic variables
+at the boundary nodes have the same values as those of the adjacent fluid nodes.  The same applies
+for tangential velocities $U_{\tau}$, while velocities perpendicular  to the mirror boundary $U_n$ have opposite signs
+to that of the velocity of the fluid node.  This can be summarised as:
 \begin{equation}
@@
 \end{equation}
-where subscripts $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.
+where the subscripts $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.
 
 For the lattice Boltzmann model the incoming populations on the boundary node need to be specified.
-For generality we assume that the plane of the symmetry $\alpha=0$, where $\alpha$ is the
+For generality we assume that the symmetry plane is $\alpha=0$, where $\alpha$ is the
 direction $x$ or $y$. Then, for lattice Boltzmann populations the symmetric boundary conditions can
@@
 \begin{description}
-\item[I] The first step is to find the complimentary directions. The complimentary $j$-th direction
-to the $i$-th direction should satisfy three conditions:
+\item[I] Identify complementary directions. For the direction $i'$ to be complementary to the $i$-th
+direction, the following relations must hold:
 \begin{equation}
 \begin{aligned}
-&c_{i\alpha}=-c_{j\alpha}, \text{ where $\alpha=$ is the symmetry plane }\\
-&c_{i\beta}=-c_{j\beta}, \text{ where $\beta\neq\alpha$ }
+&c_{i\alpha}=-c_{i'\alpha}, \text{ where $\alpha=$ is the symmetry plane }\\
+&c_{i\beta}=-c_{i'\beta}, \text{ where $\beta\neq\alpha$ }
 \end{aligned}
@@
 
-Note that the complimentary directions for the directions parallel to the symmetry plane coincide
+Note that the complementary directions for the directions parallel to the symmetry plane coincide
 with the original direction.
 
-\item[II] The second step is taking the distribution populations from the nearest fluid node and
-apply them to the boundary node utilizing complimentary directions:
+\item[II] Take the distributions from the nearest fluid node and
+apply them to the boundary node utilizing the complementary directions:
 \begin{equation}
-f_{i,B}=f_{j,F},
+f_{i,B}=f_{i',F}.
 \end{equation}
-where $j$ is the index of the direction complimentary to the $i$-th direction.
 \end{description}
 
-By applying the procedure described above, one can obtain that the scalar fields and the tangential
-velocity are conserved, the normal flux is reversed. In terms of the lattice Boltzmann populations
-this condition improve the performance and memory requirements by a factor of $4$.
-
+It can easily be seen that the procedure described above conserves the scalar fields and the tangential
+velocity, while reversing the normal flux. 
 \bibliographystyle{unsrtnat}

### 2011-05-05 Text/paper.tex
Slow.. Corrections

@@
 of
-$Ca$ with a value $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal
+$Ca$ with a value $R_{diag}=R_{axis}=0.49 H_{\mathrm{eff}}$. Therefore, taking the minimal
 requirement for the film thickness to be
@@
 addressed in the introduction that to fully address the Bretherton problem one needs to take into
-the account gravitational, viscous, inertial, and surface tension forces. Though many works
-\cite{heil-bretherton,heil-threedim,shikazono-square,kreutzer-taylor} indicate that the inertia
-effects are of significant importance, the results are of quite large deviations
+the account gravitational, viscous, inertial, and surface tension forces.
+Among all the works there is no full agreement between results even in the limit of Reynolds number $0$.
+Moreover, as it was indicated by \citet{kreutzer-taylor} and by \citet{cerro-train} the pressure distribution, streamlines and bubble shape are strongly affected by the length of the liquid slug.
+Though we try to eliminate
 

### 2011-05-06 Text/paper.tex
Finally steady-state is fulfilled. Pulled data from the cluster.

@@
 addressed in the introduction that to fully address the Bretherton problem one needs to take into
-the account gravitational, viscous, inertial, and surface tension forces.
-Among all the works there is no full agreement between results even in the limit of Reynolds number $0$.
-Moreover, as it was indicated by \citet{kreutzer-taylor} and by \citet{cerro-train} the pressure distribution, streamlines and bubble shape are strongly affected by the length of the liquid slug.
-Though we try to eliminate
+the account gravitational, viscous, inertial, and surface tension forces. Among all the works there
+is no full 
+agreement between results even in the limit of Reynolds number $0$.  Moreover, as it was indicated
+by \citet{kreutzer-taylor} and by \citet{cerro-train} the pressure distribution, streamlines and
+bubble shape are strongly affected by the length of the liquid slug and the length of the bubble. 
+In current simulations we minimize the effects of bubbles train and inertia to have a least impact,
+as the largest Reynolds number is not bigger than $10$ and the lengths of a bubble and a channel are
+chosen to be as to minimize mutual effects of bubbles motion on each other and for the film
+thickness to stabilize. However, more thorough studies are needed in order to understand precisely
+the abovementioned effects.
 
@@
 force population which
-mimics the force behavior, Eq. \ref{full:navier:stokes}.
+mimics the force behavior.
 
@@
 \end{equation}
-where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
+where $\nu=c_s^2 (\tau-1/2)$ is the kinematic viscosity,
 $M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
@@
 (\ref{standard:implementation}) simulates the continuity equation and the Navier-Stokes equation.
-The lattice Boltzmann update rule for the second set $g_i$ simulates the phase propagation with the
+The lattice Boltzmann equation for the second set $g_i$ simulates the phase propagation with the
 supplied velocity from the Navier-Stokes equation.
@@
 \label{sec:results}
-This section describes numerical simulations. One can refer to the Appendix~\ref{append:scaling} for
+This section describes numerical simulations. One can refer to Appendix~\ref{append:scaling} for
 initialization and scaling procedures. The simulations were conducted on the following 
@@
 $160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. Body forces were varied as
-$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$. The binary-liquid model parameters were kept in
+$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$ lattice units. The binary-liquid model parameters
+were kept in
 the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
-and $\tau_{\mathrm{gas}}=0.7$ giving a gas-liquid dynamic viscosity ratio of $10$. The relaxation
+and $\tau_{\mathrm{gas}}=0.7$ giving a gas-liquid viscosity ratio of $10$. The relaxation
 parameter for the phase field was $\tau_{\phi}=1.0$.
@@
 \label{sec:steady:state}
-We performed different simulations to understand the number of steps required for the simulation to
+We performed different simulations to understand the number of time steps required for the
+simulation to
 settle down to a steady state. The grid to be simulated is $52 \times 52 \times 750$ which simulates
@@
 width of $6.5$ lattice Boltzmann units together with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\cdot 10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
-$220000$ and
-$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
+\cdot 10^{-6}$. The simulation was performed for $140000-250000$ iterations with the step of
+$10000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
 capillary number varies only in the 4th digit, the radii are one percent accurate. 
@@
 $140000$&$0.1879$&$0.8657$&$0.8708$\\
+$150000$&$0.1879$&$0.8652$&$0.8704$\\
 $160000$&$0.1878$&$0.8650$&$0.8702$\\
+$170000$&$0.1880$&$0.8648$&$0.8701$\\
 $180000$&$0.1882$&$0.8647$&$0.8700$\\
+$190000$&$0.1885$&$0.8646$&$0.8699$\\
 $200000$&$0.1893$&$0.8645$&$0.8698$\\
+$210000$&$0.1963$&$0.8645$&$0.8697$\\
 $220000$&$0.1925$&$0.8644$&$0.8697$\\
+$230000$&$0.1936$&$0.8644$&$0.8697$\\
 $240000$&$0.1961$&$0.8645$&$0.8697$\\
+$250000$&$0.1932$&$0.8645$&$0.8698$\\
 \hline
@@
 \caption{Results for the steady-state case. One
-% TODO:structure?
-can see that $200000$ steps are enough to structure
+can see that $200000$ steps are enough to shape
 the bubble and approach the steady state. The small noise in the capillary number is connected to

### 2011-05-06 Results/capillaries.py

@@ -131,10 +131,12 @@ def Analyze_Consequence():
     capillaries=[]
     file_list=[]
     os.chdir("Force0000002/4")
+    print "Current directory=", os.getcwd()
     for root,dirs,files in os.walk(os.getcwd()):
         for file in files:
             if file[0:5]=="phase":   
                 file_list.append(file)
+    print "File list=", file_list    
     for file in sorted(file_list):
         print os.getcwd()
         print file
@@ -189,10 +191,6 @@ def extract_profiles(name):
             phase_numpy[coory,coorz]=phase.GetTuple1(counter)
 
 
-    numpy.savetxt("vz.txt",vz)
-    numpy.savetxt("vy.txt",vy)
-    numpy.savetxt("phase.txt",phase_numpy)
-
     #parameters of the binary liquid model
     k=0.04
     a=0.04
@@ -218,8 +216,8 @@ def extract_profiles(name):
             counter=mid*dims[0]*dims[1]+coory*dims[0]+coorx
             phase_mid[coorx,coory]= phase.GetTuple1(counter)
     
-    pylab.figure()
-    pylab.imshow(phase_mid[1:,1:])
+    #pylab.figure()
+    #pylab.imshow(phase_mid[1:,1:])
     
     prof_axis=phase_mid[0,1:]
     prof_diag=numpy.diag(phase_mid[1:,1:])
@@ -229,18 +227,18 @@ def extract_profiles(name):
     axis_zero=Get_Zero(prof_axis)/(dims[0]-2.0)
     diag_zero=math.sqrt(2.0)*Get_Zero(prof_diag)/(dims[0]-2.0)
     
-    print "Velx",vx[0,mid]
-    print "Vely",vy[0,mid]
-    print "Velz",vz[0,mid]
+    print "Velx",vx[0,z2%dims[2]]
+    print "Vely",vy[0,z2%dims[2]]
+    print "Velz",vz[0,z2%dims[2]]
     
     #Calculation of the capillary number
-    capillary=vz[0,mid]*(2.0/3.0)/math.sqrt(8.0*k*a/9.0)    
+    capillary=vz[0,z2%dims[2]]*(2.0/3.0)/math.sqrt(8.0*k*a/9.0)    
     
     print axis_zero,diag_zero
     print "Capillary=",capillary
 
-    pylab.figure()
-    pylab.plot(numpy.diag(phase_mid[1:,1:]))
+    #pylab.figure()
[…]

### 2011-05-06 Conference/CFD2011/paper.tex
Finished Jos revision. Waiting for the feedback. Switching to rectangular channels. High performance issues and mass transfer.

@@
 \title{Lattice Boltzmann Microchannel Simulations via Binary Liquid Model}
-\paperID{CFD11-52}
+\paperID{CFD11-94}
 \author{Alexandr}{Kuzmin} %{forename}{surname}
@@
 finite element solution with previously published results.
-The solutions were obtained for circular and planar cases. 
+The solutions were obtained for circular and planar cases. Smooth particle hydrodynamics (SPH) is
+presented by \citet{huang-sph}. However except showing that the SPH is able to reproduce the
+gas-liquid interface, there is no thorough study of the deposited film thickness. 
 %The planar case results for the film
@@
 
-The experimental works supported by numerical simulations reveal interesting phenomena happening in
+The experimental works supported by numerical simulations reveal interesting phenomena occuring in
 three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
 or square shaped capillaries there is a transition for a
-certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
-symmetric case,
+certain capillary number, where the bubble crosssection changes from the non-axisymmetric to
+axisymmetric case,
 Fig. \ref{fig:sym:antisym}.
 From hereon, we limit ourselves to the case of microchannels with square crosssections.
-Non axisymmetric case is attributed to the case where the axial radius is different from the
+Non-axisymmetric case is attributed to the situation where the axial radius is different from the
 diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
@@
 \caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
-capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
+capillary number. One can see the asymmetry between diagonal ($r_d$) and axial $r_h$ radii for the
+capillary
 number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
 \end{figure}
-The transition number between non-axisymmetric case and symmetric is indicated in a number of
+The transition capillary number ($\widehat{Ca}$) between non-axisymmetric case and symmetric is
+indicated in a number of
 works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
@@
 the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
-radius of the droplet dependant on the capillary number. One of the example for the bubble radii
+radius of the droplet dependant on the capillary number. One example for the bubble radii
 dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.
@@
 square and rectangular shaped capillaries. Also, they investigated gravitational effects for
-circular shaped capillaries. They also found an empirical correlation which allows to put
+circular shaped capillaries. They also found an empirical correlation which allows one to place
 different  aspect ratio rectangular channels simulation results on one
@@
 \vskip .5em
-The lattice Boltzmann method has emerged as a successful method to simulate
+The lattice Boltzmann method (LBM) has emerged as a successful method to simulate
 a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
@@
 boundary conditions \cite{heil-threedim} for the film thickness to establish.  The geometry of
-the problem is also of concern. Most literature numerical simulations are done for a channel with
+the problem is also of concern. In the literature, most numerical simulations are done
+for a channel with
 circular cross section,
@@
 of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the interface velocity of the bubble center
-measured at the front meniscus.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta
+measured at the front meniscus. $N_{\mathrm{iter}}$ is the number of simulation time iterations. 
+$5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta
 (N_y-2)$ is the size of the film in
@@
 simulate gas finger/bubbles propagation in the microchannel. The model is able to catch not only
-the film widths but also the velocity patterns change. 
+the film thicknesses but also the change in velocity patterns. 
  An examination of the influence of grid resolution on the

### 2011-05-06 Text/paper.tex
Finished Jos revision. Waiting for the feedback. Switching to rectangular channels. High performance issues and mass transfer.

@@
 \usepackage{tabularx}
+\title{Binary-liquid lattice Boltzmann simulation of microchannels with square cross sections}
+\author{A.Kuzmin \and M.Januszewski \and D.Eskin \and F.Mostowfi \and J.Derksen}
 \begin{document}
+\maketitle
 \begin{abstract}
@@
 \includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
-\caption{\citet{heil-threedim} results for the variation of the deposited liquid for a range of
-capillary numbers. One can see the asymmetry between diagonal and axial diameters for the capillary
-numbers $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
+\caption{\citet{heil-threedim} results for the variation of the bubble radii for a range of
+capillary numbers for a square channel. One can see the asymmetry between diagonal and axial
+diameters for the capillary
+numbers $Ca\leq\widehat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
 \end{figure}
@@
 \begin{figure}[htb!]
-\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
-\caption{The classical benchmark layout is presented. A semiinfinite gas bubble
-propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
+\includegraphics*[bb=153 610 410 717,width=\textwidth]{Figures/benchmark_classical.eps} 
+\caption{The classical benchmark layout describes a semiinfinite gas bubble
+propagation through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
 free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
@@
 \begin{figure}[htb!]
-\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
+\includegraphics*[bb=152 470 415 713,width=\textwidth]{Figures/benchmark_lbm.eps}
 \caption{The lattice Boltzmann benchmark layout as used in present calculations. The dimensions of
@@
 agreement between results even in the limit of Reynolds number $0$.  Moreover, as it was indicated
-by \citet{kreutzer-taylor} and by \citet{cerro-train} the pressure distribution, streamlines and
-bubble shape are strongly affected by the length of the liquid slug and the length of the bubble. 
-In current simulations we minimize the effects of bubbles train and inertia to have a least impact,
-as the largest Reynolds number is not bigger than $10$ and the lengths of a bubble and a channel are
-chosen to be as to minimize mutual effects of bubbles motion on each other and for the film
-thickness to stabilize. However, more thorough studies are needed in order to understand precisely
-the abovementioned effects.
+by \citet{kreutzer-taylor} and by \citet{cerro-bubble-train} the pressure distribution, streamlines
+and bubble shape are strongly affected by the length of the liquid slug and the length of the
+bubble. In current simulations we minimize the effects of bubbles train and inertia to have a least
+impact, as the largest Reynolds number is not bigger than $10$ and the lengths of a bubble and a
+channel are chosen to be as to minimize mutual effects of bubbles motion on each other and for the
+film thickness to stabilize. However, more thorough studies are needed in order to understand
+precisely the abovementioned effects.
 
@@
 The steady-state approach, the dependency of the radii on the moderate capillary number values
-$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity are discussed in Section
+$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative to the liquid slug bubble velocity 
+are discussed in Section
 \ref{sec:results}.
@@
 settle down to a steady state. The grid to be simulated is $52 \times 52 \times 750$ which simulates
-a quarter of the channel with the initial interface
-width of $6.5$ lattice Boltzmann units together with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
+a quarter of the channel with the initial film thickness of $6.5$ lattice Boltzmann units together
+with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
 \cdot 10^{-6}$. The simulation was performed for $140000-250000$ iterations with the step of
 $10000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
-capillary number varies only in the 4th digit, the radii are one percent accurate. 
+capillary number variation is around $5$ percent, the radii variation is $0.1$ percent. 
 \begin{table}
@@
 of the velocity are taken on the contour where $\phi=0$. This corresponds
- to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
-component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
+ to the bubble interface. One can see in Fig. \ref{fig:velocity:contour} the contour plot and the
+corressponing streamwise velocity
+component. The crossection is a plane $x=0$, where $x$ points
 towards the reader, $z$ is the streamwise direction. As fluxes inside the bubble have
@@
 \subsection{Radii transition}
-Due to performance issues we can only access the
+Due to computer performance issues we can only access the
 transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
 is guaranteed. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
-were conducted. Surprisingly the transition occurs in the region computationally accessible for
+were conducted. One of the interesting things is that the transition occurs in the region
+computationally accessible for
 current simulations. In comparison with the results of the \citet{heil-threedim}
@@
 results are more close to the VOF continuous interface method simulation by
-\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and experimental dependency of
-\citet{shikazono-square}, see Table
-\ref{table:transition:results}. The transition happens for $Ca\approx 0.09$.  One can see
-two examples of non-symmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see Fig.
+\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and fall within $5$ percent radii variation on the
+capilllary number of experimental dependency of
+\citet{shikazono-square}. The transition for current simulations happens for $Ca\approx 0.09$, see
+Table
+\ref{table:transition:results}.  One can see
+two examples of non-axisymmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see
+Fig.
 \ref{fig:crosssections:sym}.  
 \begin{table}
-\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
+\begin{tabularx}{\textwidth}{|X|X|X|X|}%X|X|}
 \hline
-$Ca$&$R_{axis}$&$R_{diag}$&$Re$&$R_{axis,Wang}$&$R_{diag,Wang}$\\
+$Ca$&$R_{axis}$&$R_{diag}$&$Re$\\ %&$R_{axis,Wang}$&$R_{diag,Wang}$\\
 \hline
-$0.053$&$0.95$&$1.01$&$0.907$&$0.99$&$0.99$\\
-$0.078$&$0.93$&$0.95$&$1.062$&$0.96$&$0.96$\\
-$0.132$&$0.89$&$0.89$&$1.80$&$0.93$&$0.93$\\
+$0.053$&$0.95$&$1.01$&$0.907$\\%&$0.99$&$0.99$\\
+$0.078$&$0.93$&$0.95$&$1.062$\\%&$0.96$&$0.96$\\
+$0.102$&$0.91$&$0.91$&$1.392$\\
+$0.132$&$0.89$&$0.89$&$1.80$\\%&$0.93$&$0.93$\\
 \hline
@@
 \caption{Simulation results in terms of $R_{axis}$ and $R_{diag}$ for the transition region between
-non-symmetric and axisymmetric cases. $R_{axis,Wang}$ and $R_{diag,Wang}$ are experimental
-correlations by \citet{wang-non-circular}. \label{table:transition:results}}
+non-axisymmetric and axisymmetric cases. The transition occurs at $\widehat{Ca}=0.09$.
+\label{table:transition:results}}
 \end{table}
@@
 parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
-is asymmetric in comparison
-with the right one. The transition happens for $Ca\approx 0.09$.\label{fig:crosssections:sym}}
+is non-axisymmetric in comparison
+with the right one, i.e. $R_{diag}\neq R_{axis}$. The transition happens at
+$\widehat{Ca}=0.09$.\label{fig:crosssections:sym}}
 \end{figure}
 
-\section{Velocity pattern}
+\subsection{Velocity pattern}
 \label{sec:velocity}
 The velocity pattern is known to change its behavior depending on the capillary number. For
-relatively low capillary numbers $Ca<0.6$, in the frame comoving with the bubble there exists a vortex 
+relatively low capillary numbers $Ca<0.6$, in a reference frame moving with the bubble there
+exists a vortex 
 in front of it \cite{heil-threedim}. For larger capillary numbers it is indicated that there is no
 vortex in front of the bubble. The transition capillary number specified by
-\citet{heil-threedim} for the rectangular channel is $Ca=0.691$. However, the
+\citet{heil-threedim} for a square channel is $Ca=0.691$. However, the
 distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
@@
 One can
-see a clear transition between associated patterns. The transition capillary number
-\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
-transition occurs at $Ca=0.5-0.6$.
+see in Fig. \ref{fig:streamlines:pattern} a clear transition between associated patterns. The
+transition capillary number of
+\citet{heil-threedim} is slightly higher than the simulation results. In present simulations the
+transition occurs at $Ca=0.6-0.7$, see Fig. \ref{fig:streamlines:pattern}.
 \begin{figure}[ht]
@@
 
-\section{Bubble length variation}
+\subsection{Variation over bubble length}
 The work of \citet{wang-non-circular} shows the variation of the bubble
-thickness through the length of
+radii along the
 bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
 capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
-capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
+capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It agrees
 with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
@@
 \caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
-One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
+One can see that the diagonal radius shows the jump near the bubble rear tip. This agrees with the
 VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
@@
 
-Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
-their simulations the non-symmetric shape was observed even for larger capillary numbers $Ca>4$ near
-the bubble front tip which after certain distance decay to have radii equal to each other. The
+Another interesting phenomenon regarding the bubble shape was indicated by \citet{heil-threedim}. In
+their simulations the non-axisymmetric shape was observed even for larger capillary numbers $Ca>4$
+near
+the bubble front tip which after certain distance relaxes to have radii equal to each other. The
 bubble shape was calculated for the large capillary number $Ca=6.43$, see Fig.
@@
 $1$ percent and can be attributed to the linear interpolation error used for the interface
-calculation. 
+calculation. Therefore, the others do not observe the symmetry breakage for $Ca>4$. 
 \begin{figure}[ht]
 \includegraphics[width=\textwidth]{Figures/bubble_ca_large.eps}
-\caption{The bubble shape for the capillary number $Ca=6.43$ is presented.$z$ increases in the
+\caption{The bubble shape for the capillary number $Ca=6.43$ is presented. $z$ increases in the
 streamwise direction. One can see that near the front tip (large $z$) the diagonal radius exhibits
@@
 
-\section{Capillary number}
+\subsection{Capillary number}
 The purpose of this section is to compare radii dependence on the capillary number. The film
@@
 initialization, and different grids: $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
-\times 82 \time 2400$. All the results are consistent. Therefore for the capillary number $Ca\geq
+\times 82 \time 2400$. All the results are consistent. Therefore one can conclude that for the
+capillary number $Ca\geq
 0.05$ grid independence is achieved. The results are compared with the results of
 \citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
-correspondent references with the help of the program ``Engauge Digitizer''. We observe that
+corresponding references with the help of the program ``Engauge Digitizer''. We observe that
 lattice Boltzmann simulations exhibit 
-the transition between non-symmetric and axisymmetric shape of the bubble.
+the transition between non-axisymmetric and axisymmetric shape of the bubble.
 
 Present simulations show that the corresponding critical capillary number is close to
-$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular}. 
-In comparison with
-the continuous method VOF \cite{wang-non-circular} present simulations show a less noisy trend and
-a more accurate transition.
-
-For moderate capillary numbers $Ca\geq 0.2$, the radii are smaller than the literature results. This
-may be caused by the bubble train simulation or the influence of the binary-liquid model. 
+$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular} and is different from the
+work of \citet{heil-threedim}, $\widehat{Ca}=0.04$. However, in the range of the moderate
+capillary numbers the results are more close to the results of \citet{heil-threedim} and exhibit
+less noisy trend in radii values and a more accurate transition between non-axisymmetric ans
+axisymmetric cases than results of \cite{wang-non-circular}. 
+%For moderate capillary numbers $Ca\geq 0.2$, the radii are smaller than the literature results.
+%This
+%may be caused by the bubble train simulation or the influence of the binary-liquid model. 
 \begin{figure}[ht]
-\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_heil.eps}
-\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_wang.eps}\\
+\includegraphics[width=\textwidth]{Figures/capillaries_comparison_heil.eps}\\
+\includegraphics[width=\textwidth]{Figures/capillaries_comparison_wang.eps}\\
 \caption{The comparison for the axial and the diagonal radii
@@
 
-\section{Relative velocity}
-One of the differencies between the flow in tubes and square cappillaries is that the flow has
-different patterns. In square microchannels the main mass flow goes through corners
+\subsection{Relative velocity}
+One of the differencies between the flow in tubes and square capillaries is that the flow has
+different streamlines patterns. In square microchannels the main mass flow goes through corners
 \cite{heil-threedim,wang-non-circular}. This affects the mass flow significantly. One of the
 characteristics of the mass flow is the normalized relative velocity dependence on the capillary
-number. THe bubble always moves faster than the surrounding liquid. The following quantity is defined as
+number. The bubble always moves faster than the surrounding liquid. The following quantity is
+defined as
 relative velocity \cite{cerro-bubble-train}:
@@
 see Fig. \ref{fig:relative:velocity}. One can see agreement between the simulations indicating that
-the mass transfer coefficients can be captured accurately. 
+one of the mass transfer characteristic can be captured accurately. 
 \begin{figure}[ht]
@@
  However, preliminary
-studies of the relative velocity show that the binary liquid model can be used for accurate mass
-transfer characterization.
+studies of the relative velocity show that the binary liquid model can simulate accurately some
+characteristics of mass transfer.
 
@@
 interface thickness \cite{kuzmin-binary2d} the results were shown to be consistent with those
-available in the literature. The film thickness dependency on the capillary number, the
-transition from the asymmetric to symmetric case, and the velocity pattern were investigated. While
+available in the literature. Note that the literature results show some variations in results with
+no established benchmark which can be used to validate data. This work falls in the bulk of the
+literature data results. For instance, the  transition from the non-axisymmetric to the
+axisymmetric case is shown to be at $\widehat{Ca}=0.09$ which is close to results of
+\citet{wang-non-circular}. However, in the moderate range of parameters simulation results are
+close to the results of \citet{heil-threedim} and exhibit more accurate transition between
+non-axisymmetric and symmetric cases than results of \citet{wang-non-circular}. Overall, the film
+thickness dependency on the capillary number, the
+transition from the non-axisymmetric to symmetric case, and the velocity patterns were investigated.
+While
 the goal of this work is a feasibility study, more numerical studies are needed to
@@
 \cite{kreutzer-taylor,yue-mass} on the design of microchannels. Our results show that the lattice
-Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels.
+Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels. Results
+coincide with bulk literature results.
 
@@
 
-\item[Velocity] The Reynolds and capillary numbers are supplied from the physical world. We
-accept the following definitions:
+\item[Velocity] The Reynolds and capillary numbers are supplied from the physical world:
 \begin{equation}
@@
 The gas-liquid viscosity ratio in the current simulations was chosen as
-$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and the relaxation parameters were chosen to be
+$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=\frac{\nu_{\mathrm{liq}}}{\nu_{\mathrm{gas}}}=10$ and
+the relaxation parameters were chosen to be
 in the safe region, i.e. $\tau_{liq}=2.5$ and $\tau_{gas}=0.7$.
@@
 0}{\dfrac{16}{\pi^6} \dfrac{(-1)^i (-1)^j}{(2 i+1)(2 j+1)\bigl[\frac{(2 i
-+1)^2}{W_{\mathrm{eff}}}+\frac{(2 j+1)^2}{H_{\mathrm{eff}^2}}\bigr]}}}.
++1)^2}{W_{\mathrm{eff}}^2}+\frac{(2 j+1)^2}{H_{\mathrm{eff}}^2}\bigr]}}}.
 \end{equation}
@@
 $\frac{\mathrm{d}P_1}{\mathrm{d}z}$ for simulations conducted on the grid with the characteristic
-number of nodes $N_y^1$ , the body force for another simulation can be calculated as:
+number of nodes $N_{y1}$ , the body force for another simulation can be calculated as:
 \begin{equation}
@@
 It can easily be seen that the procedure described above conserves the scalar fields and the tangential
-velocity, while reversing the normal flux. 
+velocity, while reversing the normal flux. The plane of symmetry is located half-way between fluid
+and boundary nodes. 
 \bibliographystyle{unsrtnat}

### 2011-05-06 Results/velocity.py

@@ -121,7 +121,7 @@ def Analyze_Consequence():
     diag_zeros=[]
     capillaries=[]
     reynolds=[]
-    file_list=["Moderate/1/phase250000.vts","Moderate/3/phase250000.vts","Moderate/5/phase250000.vts","Moderate/7/phase250000.vts","Moderate/9/phase250000.vts"]
+    file_list=["Moderate/1/phase250000.vts","Moderate/3/phase250000.vts","Moderate/5/phase250000.vts","Moderate/6/phase250000.vts","Moderate/7/phase250000.vts","Moderate/9/phase250000.vts"]
                #"SmallCapGrid/8/phase250000.vts"]
     #file_list=["Velocity/"+str(dir)+"/phase240000.vts" for dir in range(1,11)]    
     #file_list=["Force0000002x82/2/phase180000.vts","Force0000002x82/4/phase180000.vts"] #,"Force0000005x52/8/phase250000.vts","Force0000005x52/10/phase250000.vts"]    

### 2011-05-15 Sailfish/analysis.py
Add the Saiflish simulation script and a simple script for data analysis.

@@ -1,0 +1,98 @@
+#!/usr/bin/python
+
+import math
+import sys
+import numpy as np
+from scipy.interpolate import interp1d
+from scipy.optimize import bisect, ridder, newton
+
+fname = sys.argv[1]
+
+def load(fname, num):
+    try:
+        return np.load('%s_%07d.npz' % (fname, num))
+    except:
+        return np.load('%s_%08d.npz' % (fname, num))
+
+prev_pos = None
+
+kappa = 0.04
+const_A = 0.04
+gamma = math.sqrt(8.0 * kappa * const_A / 9.0)
+tau_liq = 2.5
+mu_liq = (tau_liq - 0.5) / 3.0
+
+def measure(data, time_steps):
+    phi = data['phi'][1:-1,1:-1,:]
+
+    y = phi[-1,-1,:]
+    vx_y = data['v'][0,-2,-2,:]
+    max_x = phi.shape[-1] - 1
+    x = np.arange(0, phi.shape[-1])
+    f = interp1d(x, y)
+    f_vel = interp1d(x, vx_y)
+    idxs = np.where(y < 0)[0]
+
+    if idxs[0] > 0:
+        bub_x0 = ridder(f, max(idxs[0] - 20, 0), min(idxs[0] + 20, max_x))
+        bub_x1 = ridder(f, max(idxs[-1] - 20, 0), min(idxs[-1] + 20, max_x))
+        bub_len = bub_x1 - bub_x0
+        bub_mid_idx = idxs[len(idxs) / 2]
+    else:
+        end_idx = np.where(idxs - np.roll(idxs, -1) != -1)[0][0]
+        bub_x1 = ridder(f, max(idxs[end_idx] - 20, 0),
+                        min(idxs[end_idx] + 20, max_x))
+
+        try:
+            idx0 = idxs[end_idx+1]
+            bub_x0 = ridder(f, max(idx0 - 20, 0), min(idx0 + 20, max_x))
+            part2 = max_x+1 - bub_x0
+        except IndexError:
+            if f(0) * f(20) < 0:
+                bub_x0 = ridder(f, max(idx0 - 20, 0), min(idx0 + 20, max_x))
+            else:
+                bub_x0 = 0.0
+            part2 = 0.0
+
+        bub_len = bub_x1 + part2
+
+        if part2 > bub_x1:
+            bub_mid_idx = int(bub_x0 + bub_len / 2)
[…]

### 2011-05-15 Sailfish/bmc3d_mirror.py

@@ -1,0 +1,61 @@
+#!/usr/bin/python
+
+import numpy as np
+from sailfish import lbm, lb_binary
+from sailfish import geo
+
+import optparse
+from optparse import OptionGroup, OptionParser, OptionValueError
+
+class GeoFE(geo.LBMGeo3D):
+
+    wall_width = 2
+
+    def define_nodes(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+
+        wall_map = np.logical_or(hy < self.wall_width, hz < self.wall_width)
+        self.set_geo(wall_map, self.NODE_WALL)
+        self.set_mirror(1, 1)
+        self.set_mirror(2, 1)
+
+    def init_fields(self):
+        hz, hy, hx = np.mgrid[0:self.lat_nz, 0:self.lat_ny, 0:self.lat_nx]
+        width = self.sim.options.iwidth
+
+        self.sim.phi[:] = 1.0
+        self.sim.phi[np.logical_and(
+                       np.logical_and(hz >= width-1 + self.wall_width, hy >= width-1 + self.wall_width),
+                       np.logical_and(hx >= self.lat_nx/3.0 - 1.0, hx<2*self.lat_nx/3.0))] = -1.0
+        self.sim.rho[:] = 1.0
+
+class FESim(lb_binary.BinaryFluidFreeEnergy):
+    filename = 'fe_separation_3d'
+
+    def __init__(self, geo_class, defaults={}):
+        settings = {'verbose': True, 'lat_nx': 750, 'grid': 'D3Q19',
+                    'lat_ny': 52, 'lat_nz' : 52,
+                    'kappa': 0.04, 'Gamma':1.0, 'A': 0.04,
+                    'scr_scale': 1,
+                    'tau_a': 2.5, 'tau_b': 0.7, 'tau_phi': 2.0,
+                    'bc_wall': 'halfbb', 'bc_wall_grad_order': 1,
+                    'periodic_x': True, 'periodic_y': True,'periodic_z': True}
+        settings.update(defaults)
+
+        opts = []
+        opts.append(optparse.make_option('--force', dest='force', type='float', default=1e-5))
+        opts.append(optparse.make_option('--iwidth', dest='iwidth', type='int', default=5))
+
+        lb_binary.BinaryFluidFreeEnergy.__init__(self, geo_class, options=opts, defaults=settings)
+
[…]

### 2011-05-16 Text/paper.tex
Corrected capillary numbers.

@@
 \cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate the applicability
-of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.1\leq Ca \leq 1.0$,
+of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.05\leq Ca \leq
+6.0$,
 we focus on the simulation of three-dimensional capillaries with square cross sections.
@@
 The steady-state approach, the dependency of the radii on the moderate capillary number values
-$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative to the liquid slug bubble velocity 
+$0.05\leq Ca \leq 6.0$, the velocity pattern, and the relative to the liquid slug bubble velocity 
 are discussed in Section
@@
 non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
-number is presented for the range of moderate capillary numbers $0.1 \leq Ca \leq 1.0$. Results are
+number is presented for the range of moderate capillary numbers $0.05 \leq Ca \leq 6.0$. Results are
 concluded with studies of the velocity pattern.
@@
 body force $\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-5}$ for grids of size $50\times 50 \times
-1500$ and the capillary number $Ca\approx0.1-1.0$.  
+1500$ and the capillary number $Ca\approx0.1-6.0$.  
 \end{description}

### 2011-05-30 Text/paper.tex
Merging Schlumberger comments.

@@
 \usepackage{tabularx}
+\usepackage{overpic}
 \title{Binary-liquid lattice Boltzmann simulation of microchannels with square cross sections}
@@
 thin liquid films between bubbles and channel walls. The bubble shape and flow patterns are
-complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
-the applicability and parameter choice for the two-dimensional case of the Bretherton problem (flow
+complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently, we
+investigated
+an applicability and parameter selection for the two-dimensional case of the Bretherton problem
+(flow
 between parallel plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
-\cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate the applicability
+\cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate an
+applicability
 of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.05\leq Ca \leq
@@
 periodic boundary conditions are applied in the streamwise direction. The results show that the
-binary liquid model is able to capture a number of phenomena happening in three-dimensional
+binary liquid model is able to capture a number of phenomena occurin in three-dimensional
 capillaries, such as the existence of a vortex in front of the bubble and the dependency
@@
 
-For example, the deposited film thickness
+For example, the film thickness
 is proportional to $Ca^{2/3}$ in the range of small capillary numbers for circular channels
@@
 is of significant interest for chemical industry as it is widely used in chemical monolith
-microreactors \cite{kreutzer-pressure-drop}. Large mass transfers can be achieved because of the
+microreactors \cite{kreutzer-pressure-drop}. Intensive mass transfer can be achieved because of the
 large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. Heat transfer is
 also increased in comparison with single phase flow \cite{fukugata-levelset}. While it
-is possible to calculate the flow analytically for small capillary numbers and negligible Reynolds
+is possible to calculate the flow analytically for small capillary numbers and nearly zero Reynolds
 number \cite{bretherton}, the
-results cannot be extrapolated to the wider range of capillary and Reynolds numbers used in the
-chemical industry.  Thus, the desire for consistent numerical simulations arises.
+results cannot be extrapolated to the wide range of capillary and Reynolds numbers used in the
+chemical industry.  Thus, the need of consistent numerical simulations arises.
 
@@
 experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
-\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the analytical
+\citet{giavedoni-numerical,heil-bretherton}. All abovementioned investigations found that the 
 Bretherton
@@
 interfacial, inertial and viscous forces \cite{gupta-review}. Historically, 
-\citet{bretherton} indicated that inertia effects can be 
-neglected. \citet{giavedoni-numerical} suggested
+\citet{bretherton} neglected inertia effects. \citet{giavedoni-numerical} suggested
 that
@@
 extended
-the results for the flow between plates upto Reynolds number
-$300$. This author indicates that while the influence on the established film thickness is
+the results for the flow between plates up to $Re=300$. This author noted that while the
+influence of inertia on the established film thickness is
 insignificant
 ($7$ percent deviation from the film thickness measured at $Re=0$), the change in Reynolds number
-influences significantly the pressure distrubution and the flow field near the front bubble tip.
-\citet{cerro-bubble-train} show that the mass transfer is strongly affected by flow direction for
+significantly influences  the pressure distrubution and the flow field near the front bubble cap.
+\citet{cerro-bubble-train} showed that the mass transfer is strongly affected by flow direction for
 small capillary numbers in case of upward and downward flows. Thus, to fully describe the
-propagation of the semi-infinite finger in liquid media one needs to take into the account the
+propagation of the semi-infinite finger in liquid media one needs to take into account the
 viscous, gravitational, surface, and inertia forces \cite{gupta-review}.
@@
 is a vast number of experimental
-results available for three-dimensional cases of microchannels with square, triangular and
+results available for three-dimensional flows in microchannels with square, triangular and
 rectangular cross sections. For instance, \citet{cerro-bubble-train} performed a number of
@@
 experimental
-results for the deposition film thickness dependency on the
+results for the film thickness dependence on the
 capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
-shaped mirochannels. They also presented the experimental correlation for the prediction of bubble
+shaped mirochannels. They also presented the experimental correlation for  bubble
 radii
@@
 \cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
-wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
+wang-non-circular} showed interesting phenomena in three-dimensional geometries. It was found
 \cite{heil-threedim,wong-films} that for rectangular
@@
 different from the
-diagonal radius and the bubble having a non-circular shape in the channel crossection. In this case
+diagonal radius and the bubble having a non-circular shape in the channel crossection. In this case,
 the bubble shape mimics the shape of a rectangle and looks like a rectangle with rounded corners.
-From hereon we adapt the convention ``axial`` for the radius measured along the $x$ or $y$ axes,
-see Fig. \ref{fig:lbm:benchmark} \cite{heil-threedim}. The dependency
+Further we adapt the convention ``axial`` for the radius measured along the $x$ or $y$ axes,
+see Fig. \ref{fig:lbm:benchmark} \cite{heil-threedim}. The dependence
 of the diagonal and axial radii on the capillary number is shown in Fig. 
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
+\begin{overpic}[width=\textwidth]{Figures/capillary_width_heil.eps}
+\put(50,45){\includegraphics[width=0.4\textwidth]{Figures/inset.eps}}
+\end{overpic}
+%\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
 \caption{\citet{heil-threedim} results for the variation of the bubble radii for a range of
@@
 \end{figure}
-The transition capillary number for square channels between non-axisymmetric case and axisymmetric
-one is indicated in a number of
-works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
+The capillary number for square channels indicating transition between non-axisymmetric case to
+axisymmetric
+one is reported in a number of
+works ($\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
 $\widehat{Ca}=0.1$
-\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
+\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}). If the capillary
 number is larger
@@
 
-There are also a number of numerical works for the three-dimensional case. For instance, 
+There are also a number of numerical works on three-dimensional flows. For instance, 
 \citet{wong-films,wong-pressure} studied
@@
 \citet{heil-threedim} performed three-dimensional simulations for circular-,
-square- and rectangular-shaped capillaries. They indicated a transition
-capillary number where the streamlines pattern
-changes from having a vortex in front of the bubble and not having it for larger capilllary
-numbers,  Fig. \ref{fig:streamlines:pattern}. For a square channel the critical
-capillary number for this phenomena is $Ca=0.691$.  \citet{heil-threedim} also found an empirical
+square- and rectangular-shaped capillaries. They indicated a transition of flow regime in a liquid
+slug. Vortexes observed at lower capillary numbers disappear at the capillary number above a
+certain threshold,  Fig. \ref{fig:streamlines:pattern}. For a square channel the critical
+capillary number is $Ca=0.691$.  \citet{heil-threedim} also found an empirical
 correlation which
-makes it possible to put on one curve the results of radii dependency on the capillary number
+makes it possible to put on the same curve the results of radii dependency on the capillary number
 simulations of rectangular channels with different aspect ratios.
-It is also indicated that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
+It was also found that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
 2.04$, the interface does not become
@@
 simulations using the Volume of Fluid (VOF) technique for
-capillaries with non-circular (square, triangular) cross sections. They also investigated the relative
-velocity of the bubble to the slug liquid velocity  for
-a number of capillary numbers.
+capillaries with non-circular (square, triangular) cross sections. They also investigated the
+relative slug to bubble velocity for
+a range of capillary numbers.
 
-As it is mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
+As it was mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
 phenomena are the VOF method \cite{wang-non-circular}, the level-set method
-\cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It is
-also indicated \cite{gupta-review} that the new techniques which are still in the development stage
+\cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It was
+also indicated \cite{gupta-review} that the new techniques, which are still in the development
+stage,
 are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
 The finite element method solves the Bretherton flow as a free-surface problem with a sharp
-interface but without underlying details for the gas.  The lattice Boltzmann method on the other
-hand,
-as a continuous interface method, provides more flexibility in simulations involving coalescence
+interface but without underlying details for the gas.  The lattice Boltzmann method is a continuous
+interface method, and therefore provides more flexibility in simulations involving coalescence
 and/or
@@
 
-In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between parallel
+In our previous work \cite{kuzmin-binary2d}, we already analyzed the Bretherton flow between
+parallel
 plates.
-It was shown that the free-energy binary liquid model, which is a phase field method, simulates
-with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine 
-if similar conclusions can be drawn for three-dimensional flows in microchannels with square cross sections,
+It was shown that the free-energy binary liquid model, which is a phase field method, simulates the
+Bretherton/Taylor phenomenon with reasonable accuracy. The goal of this work is to examine 
+if similar conclusions can be drawn for three-dimensional flows in microchannels with square cross
+sections,
 simulated using the free-energy binary liquid lattice Boltzmann
@@
 simulates two liquids with the assumption of uniform overall
-density. As it was indicated by \citet{gupta-review} the flow in microchannels is a complex
+density. As it was indicated by \citet{gupta-review} a flow in microchannels is a complex
 interplay of viscous, gravitational, inertial and surface forces, and the classical Bretherton
@@
 the
-viscosity ratio, which is why the uniform density binary liquid model is suitable for this kind of
+viscosity ratio. Therefore the uniform density binary liquid model is suitable for this kind of
 simulations.
@@
 Boltzmann model is outlined. The simulation results for the three-dimensional case are presented in
-the results section. The results cover the film thickness dependency on the capillary number, the
-relative to the slug liquid velocity of the bubble and the vortex profiles. The paper
+the results section. That covers the film thickness dependency on the capillary number, the bubble
+to slug relative velocity and vortex profiles. The paper
 is concluded with a summary of the main findings.
@@
 \label{sec:numerical:benchmark}
-The main discussion here follows and is based on the two-dimensional benchmark approach
+The main discussion here is based on the two-dimensional benchmark approach
 \cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
@@
 medium.
-The film thickness in this case is measured at the inlet. In the lattice Boltzmann framework such a
+The film thickness in this case is measured at the inlet. In the lattice Boltzmann framework, such a
 formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
@@
 initial bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
-the film in the two-dimensional geometries stabilizes at distances of $2.6-4.0$ diameters
+the film in the two-dimensional geometry stabilizes at distances of $2.6-4.0$ diameters
 from the front tip depending on the Reynolds number. \citet{heil-threedim} measured the bubble radii
 at the distance of $5.5 H_{\mathrm{eff}}$ and indicated it to be sufficient for $Ca\leq 10$. In
-this work the following relation holds for all conducted simulations: $0.05 \geq Ca \leq 6$.
+this work, the following relation holds for all conducted simulations: $0.05 \geq Ca \leq 6$.
 Thus, the
@@
 
-For simplicity, periodic boundary conditions are applied so that effectively a bubble train is simulated
-instead of a single bubble.  The mutual influence of bubbles on each other is minimized by the use
-of a long
+For simplicity, periodic boundary conditions are applied. Therefore  a bubble
+train is simulated
+instead of a single bubble.  The mutual influence of neighbouring bubbles is minimized
+by using a long
 channel ($15 H_{\mathrm{eff}}$). Periodic boundary conditions imply as well that the fluid is driven
@@
 address simulations with upward and downward flows \cite{cerro-bubble-train} which require a
-simulataneouse imposement of pressure boundary conditions and the body force. The pressure boundary
-conditions for the binary liquid model is a challenge. Thus, it is assumed the flow is in the
+simulataneous imposement of pressure boundary conditions and the body force. The pressure boundary
+conditions for the binary liquid model is a challenge. Thus, we consider a flow in the
 horizontal direction only and the upward/downward flows with pressure difference imposed are not
@@
 \citet{heil-threedim} specify that the diagonal bubble radius is different from the
-axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii are the same for larger values
+axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.03$. The radii are the same for larger values
 of
@@
 % TODO: This paragraph should be revisited, somehow it reads a little weirdly..
-It should be noted that the binary liquid model we used has a uniform density. A part of discussion is
-addressed in the introduction that to fully address the Bretherton problem one needs to take into
-the account gravitational, viscous, inertial, and surface tension forces. Among all the works there
+It should be noted that the binary liquid model, we used, has a uniform density. A part of
+discussion is
+addressed in the introduction:  to describe the Bretherton problem one needs to take into
+account gravitational, viscous, inertial, and surface tension forces. Among all the works, there
 is no full 
@@
 simulation to
-settle down to a steady state. The grid to be simulated is $52 \times 52 \times 750$ which simulates
+settle down to a steady state. The grid  was $52 \times 52 \times 750$ which represents
 a quarter of the channel with the initial film thickness of $6.5$ lattice Boltzmann units together
 with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\cdot 10^{-6}$. The simulation was performed for $140000-250000$ iterations with the step of
-$10000$ iterations. The results are summarised in Table \ref{table:steady:state}. While the
+\cdot 10^{-6}$. The simulation was performed $250000$ iterations with the step of
+$10000$ iterations. The results for time iterations $140000-250000$ are summarised in Table
+\ref{table:steady:state}. While the
 capillary number variation is around $5$ percent, the radii variation is $0.1$ percent. 
@@
 
-To further examine the steady-state the velocity in the streamwise direction is plotted. The values
+To further examine the steady-state, the velocity in the streamwise direction is plotted. The values
 of the velocity are taken on the contour where $\phi=0$. This corresponds
@@
 The relative velocity was calculated and compared with the results of \citet{wang-non-circular},
-see Fig. \ref{fig:relative:velocity}. One can see agreement between the simulations indicating that
+see Fig. \ref{fig:relative:velocity}. One can see an agreement between the simulations indicating
+that
 one of the mass transfer characteristic can be captured accurately. 
@@
 \section{Conclusion}
-The present work presents results of binary-liquid simulations for three-dimensional channels with square
+This work presents results of binary-liquid simulations for three-dimensional channels with square
 cross sections with the lattice Boltzman method. By resolving properly the film thickness as twice the
 interface thickness \cite{kuzmin-binary2d} the results were shown to be consistent with those
-available in the literature. Note that the literature results show some variations in results with
+available in the literature. Note that the literature results show some variations with
 no established benchmark which can be used to validate data. This work falls in the bulk of the

### 2011-05-31 Text/paper.tex
Incorporated all Schlumberger corrections. Waiting for last Jos comments.

@@
-\documentclass{article}
-\usepackage[numbers,sort&compress]{natbib}
+\documentclass[preprint,12pt]{elsarticle}
+%\usepackage[numbers,sort&compress]{natbib}
+\usepackage{lineno}
+\journal{Chemical Engineering Journal}
+
 \usepackage{graphicx}
@@
 \usepackage{overpic}
+\begin{document}
+\begin{frontmatter}
 \title{Binary-liquid lattice Boltzmann simulation of microchannels with square cross sections}
-\author{A.Kuzmin \and M.Januszewski \and D.Eskin \and F.Mostowfi \and J.Derksen}
-\begin{document}
-\maketitle
+\author[uofa]{A.~Kuzmin\corref{cor1}}
+\cortext[cor1]{Corresponding author}
+\ead{kuzmin@ualberta.ca}
+\author[us]{M.~Januszewski}
+\ead{michalj@gmail.com}
+\author[schlum]{D.~Eskin}
+\ead{deskin@slb.com}
+\author[schlum]{F.~Mostowfi}
+\ead{fmostowfi@slb.com}
+\author[uofa]{J.J.~Derksen}
+\ead{jos@ualberta.ca}
+\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
+116 St, Edmonton, Alberta, T6G
+2V4 Canada}
+\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
+\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
+Canada}
 \begin{abstract}
@@
 \end{abstract}
+
+\begin{keyword}
+Taylor/Bretherton problem \sep Microchannel simulation \sep Multiphase flow \sep Lattice Boltzmann
+method
+\sep Binary liquid model \sep Flow in microchannels with square cross sections \sep Gravity driven
+%% keywords here, in the form: keyword \sep keyword
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+\end{keyword}
+
+\end{frontmatter}
+\linenumbers
 
@@
 \ref{fig:heil:three:dim}.
-\begin{figure}[ht]
+\begin{figure}[htb!]
 \begin{overpic}[width=\textwidth]{Figures/capillary_width_heil.eps}
@@
 with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
-\cdot 10^{-6}$. The simulation was performed $250000$ iterations with the step of
+\cdot 10^{-6}$. The simulation was performed for $250000$ iterations with the step of
 $10000$ iterations. The results for time iterations $140000-250000$ are summarised in Table
@@
 clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
-examine only the points corresponding to the front and the rear bubble tips. If one finds them equal
-to each other, then it is exactly the velocity with which the bubble propagates in the microchannel.
+examine only the points corresponding to the front and the rear bubble caps. If they are equal, then
+it is exactly the velocity with which the bubble propagates in the microchannel.
 For a physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
@@
 interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
-$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
-same velocities, which is actually the bubble velocity. For more details on velocity patterns
+$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear caps are propagating with nearly
+the
+same velocity, that is the actual bubble velocity. For more details on velocity patterns
 see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
@@
 \subsection{Radii transition}
-Due to computer performance issues we can only access the
+Due to computational power restrictions we can only access the
 transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
-is guaranteed. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
-were conducted. One of the interesting things is that the transition occurs in the region
-computationally accessible for
-current simulations. In comparison with the results of the \citet{heil-threedim}
+is achievable. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
+were conducted. In comparison with the results of the \citet{heil-threedim}
 ($\widehat{Ca}=0.04$) our
 results are more close to the VOF continuous interface method simulation by
-\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and fall within $5$ percent radii variation on the
+\citet{wang-non-circular} ($\widehat{Ca}=0.1$). Our data also fall within $5$ percent radii
+variation on the
 capilllary number of experimental dependency of
@@
 parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
-is non-axisymmetric in comparison
+is non-axisymmetric in contrast
 with the right one, i.e. $R_{diag}\neq R_{axis}$. The transition happens at
@@
 The velocity pattern is known to change its behavior depending on the capillary number. For
-relatively low capillary numbers $Ca<0.6$, in a reference frame moving with the bubble there
-exists a vortex 
-in front of it \cite{heil-threedim}. For larger capillary numbers it is indicated that there is no
+relatively low capillary numbers $Ca<0.6$, in a reference frame moving with the bubble a vortex 
+is observed \cite{heil-threedim}. For larger capillary numbers it is indicated that there is no
 vortex in front of the bubble. The transition capillary number specified by
 \citet{heil-threedim} for a square channel is $Ca=0.691$. However, the
-distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
+distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well as on
+the
 pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
-that the pressure is largely influenced by bubble frequency and slug distance. The present
+that the pressure is significantly influenced by bubble frequency and slug distance. The present
 simulations are conducted for bubble trains and certain differences in terms of changing streamline
@@
 transition capillary number of
-\citet{heil-threedim} is slightly higher than the simulation results. In present simulations the
+\citet{heil-threedim} is slightly higher than that in the simulation results. In present simulations
+the                                               
 transition occurs at $Ca=0.6-0.7$, see Fig. \ref{fig:streamlines:pattern}.
@@
 created
-before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
+in front of the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
 is a different value from $Ca=0.69$ \cite{heil-threedim}. However, we performed a number of
@@
 capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
-capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It agrees
-with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
+capillary numbers the diagonal radius exhibits a small jump near the rear bubble cap. That agrees
+with results of \citet{wang-non-circular}. However, our simulations show smoother behavior of
 the jump.  
@@
 \caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
-One can see that the diagonal radius shows the jump near the bubble rear tip. This agrees with the
+One can see that the diagonal radius shows the jump near the bubble rear cap. This agrees with the
 VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
@@
 near
-the bubble front tip which after certain distance relaxes to have radii equal to each other. The
+the bubble front cap which after certain distance relaxes to have radii equal to each other. The
 bubble shape was calculated for the large capillary number $Ca=6.43$, see Fig.
@@
 the bubble and it resembles the shape given by \citet{heil-threedim}, but the difference is around
-$1$ percent and can be attributed to the linear interpolation error used for the interface
-calculation. Therefore, the others do not observe the symmetry breakage for $Ca>4$. 
+$1$ percent and can be explained by error in the linear interpolation used for the
+interface
+calculation. Therefore, we do not observe the symmetry breakage for $Ca>4$. 
 \begin{figure}[ht]
@@
 \subsection{Capillary number}
-The purpose of this section is to compare radii dependence on the capillary number. The film
+The purpose of this section is to study dependence of the radii on the capillary number. The film
 thicknesses were extracted from the middle of the bubble. The data presented in Fig.
@@
 0.05$ grid independence is achieved. The results are compared with the results of
-\citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
+\citet{heil-threedim} and of \citet{wang-non-circular}. The data were extracted from the
 corresponding references with the help of the program ``Engauge Digitizer''. We observe that
@@
 Present simulations show that the corresponding critical capillary number is close to
-$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular} and is different from the
-work of \citet{heil-threedim}, $\widehat{Ca}=0.04$. However, in the range of the moderate
-capillary numbers the results are more close to the results of \citet{heil-threedim} and exhibit
+$\widehat{Ca}=0.1$ that agrees with the work of \citet{wang-non-circular} and is different from the
+results of \citet{heil-threedim}, $\widehat{Ca}=0.04$. However, in the range of the moderate
+capillary numbers our results are closer to the results of \citet{heil-threedim} and exhibit
 less noisy trend in radii values and a more accurate transition between non-axisymmetric ans
-axisymmetric cases than results of \cite{wang-non-circular}. 
+axisymmetric cases than results of \citet{wang-non-circular}. 
 %For moderate capillary numbers $Ca\geq 0.2$, the radii are smaller than the literature results.
@@
 One of the differencies between the flow in tubes and square capillaries is that the flow has
-different streamlines patterns. In square microchannels the main mass flow goes through corners
+different streamline patterns. In square microchannels, the main mass flow occurs through corners
 \cite{heil-threedim,wang-non-circular}. This affects the mass flow significantly. One of the
-characteristics of the mass flow is the normalized relative velocity dependence on the capillary
+characteristics of the mass flow is the normalized dependence of the relative velocity on the
+capillary
 number. The bubble always moves faster than the surrounding liquid. The following quantity is
@@
 velocity. As far as the steady state is achieved one can take the liquid superficial velocity
-exactly at the distance in the middle between bubbles. Then, the liquid superficial velocity is
+exactly at the distance in the middle between bubbles. The liquid superficial velocity is
 defined as:
@@
 that
-one of the mass transfer characteristic can be captured accurately. 
+one of the mass transport characteristics can be captured accurately. 
 \begin{figure}[ht]
@@
 
-The relative velocity is an important quantity in mass flow characterization
+The relative velocity is an important quantity of mass flow characterization
 \cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and 
-inertia effects \citet{heil-bretherton} significantly influence the mass transfer coefficient. 
-Future work is planned to study this coefficient and calculate the associated advection-diffusion coefficient.
- However, preliminary
-studies of the relative velocity show that the binary liquid model can simulate accurately some
-characteristics of mass transfer.
+inertia effects \citet{heil-bretherton} significantly influence  mass transfer. 
+Future work is planned to study this phenomenon and calculate the mass transfer coefficient by
+solving the advection-diffusion equation.
+% Preliminary
+%studies of the relative velocity show that the binary liquid model can simulate accurately some
+%characteristics of mass transfer.
 

### 2011-06-13 Results/rectangle.py
Started analyzing Rectangular results. Need to rescale them somehow.

@@ -1,0 +1,201 @@
+#!/usr/bin/python
+import pylab
+import os
+import numpy
+import vtk
+import math
+
+def Get_Zero(prof):
+    zero=0
+    #pylab.figure()
+    #pylab.plot(prof)
+    for counter in range(0, len(prof)):
+        if prof[counter]<0 and prof[counter+1]>=0:
+            zero=-(prof[counter]*(counter+1)-prof[counter+1]*counter)/(prof[counter+1]-prof[counter])
+    return zero+0.5
+
+
+def extract_profiles(name):
+    
+    gridreader = vtk.vtkXMLStructuredGridReader()
+    gridreader.SetFileName(name)
+    gridreader.Update()
+    
+    grid  = gridreader.GetOutput()
+    data  = grid.GetPointData()
+    points=grid.GetPoints()
+    dims  =grid.GetDimensions()
+    velocity=data.GetArray("Velocity")
+    phase=data.GetArray("Phase")
+    vz=numpy.zeros([dims[1],dims[2]])
+    vy=numpy.zeros([dims[1],dims[2]])
+    vx=numpy.zeros([dims[1],dims[2]])
+    phase_numpy=numpy.zeros([dims[1],dims[2]])
+    
+    print dims
+
+    for coory in range(0,dims[1]):
+        for coorz in range(0,dims[2]):
+            counter=coorz*dims[0]*dims[1]+coory*dims[0]
+            velx,vely,velz=velocity.GetTuple3(counter)
+            vz[coory,coorz]=velz
+            vy[coory,coorz]=vely
+            vx[coory,coorz]=velx
+            phase_numpy[coory,coorz]=phase.GetTuple1(counter)
+
+    #parameters of the binary liquid model
+    k=0.04
+    a=0.04
+    
+    center=phase_numpy[0,:]
+    z1 = numpy.min(numpy.where(center < 0.0))
+    z2 = numpy.max(numpy.where(center < 0.0))
+    if z1==0:
+        z2=numpy.min(numpy.where(center>0.0))+dims[2]
+        z1=numpy.max(numpy.where(center>0.0))
+    print z1,z2
+    
+    mid =((z1+z2)/2)%dims[2]
+    print mid
+
+    slug_mid=((z1+z2+dims[2])/2)%dims[2]
+    print "Slug_mid=",slug_mid
[…]

### 2011-06-14 Text/paper.tex
Added rectangular channels results.

@@
 \usepackage{overpic}
+\usepackage{tikz}
 \begin{document}
 \begin{frontmatter}
-\title{Binary-liquid lattice Boltzmann simulation of microchannels with square cross sections}
+\title{Three-dimensional binary-liquid lattice Boltzmann simulation of microchannels with square
+cross sections}
 \author[uofa]{A.~Kuzmin\corref{cor1}}
@@
 investigated
-an applicability and parameter selection for the two-dimensional case of the Bretherton problem
+an applicability and parameter selection for the two-dimensional Bretherton problem
 (flow
 between parallel plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
-\cite{kuzmin-binary2d}. This paper is a continuation of our previous work.  To validate an
-applicability
-of the LBM to Bretheron phenomena in the moderate range of the capillary number $0.05\leq Ca \leq
-6.0$,
-we focus on the simulation of three-dimensional capillaries with square cross sections.
+\cite{kuzmin-binary2d}. This paper is the continuation of our previous work with simulations of
+three-dimensional channels with square crosssections in the range of the capillary number $0.05\leq
+Ca \leq
+6.0$.
 The flow is driven by a body force, and
 periodic boundary conditions are applied in the streamwise direction. The results show that the
-binary liquid model is able to capture a number of phenomena occurin in three-dimensional
-capillaries, such as the existence of a vortex in front of the bubble and the dependency
-of the bubble radii on the capillary number.  We conclude that lattice Boltzmann free energy binary
+binary liquid model is able to correctly capture a number of phenomena occuring in three-dimensional
+capillaries, such as the existence of a vortex in front of the bubble and the way bubble radii
+depend on the capillary number.  We conclude that lattice Boltzmann free energy binary
 liquid model can be used to simulate the Bretherton problem with good accuracy. 
@@
 The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
-narrow channels. Depending on the channel geometry it was found that the deposited film thickness
+narrow channels. Depending on the channel geometry it was found \cite{gupta-review} that the 
+thickness of the liquid film between a bubble and channel walls
 is a complicated function of the capillary number $Ca$:
@@
 For example, the film thickness
-is proportional to $Ca^{2/3}$ in the range of small capillary numbers for circular channels
+is proportional to $Ca^{2/3}$ for small capillary numbers for circular channels
 \cite{bretherton,heil-bretherton}. 
-The problem of predicting flow patterns and associated mass transfers for Bretheron-type flows
+Predicting flow patterns and associated mass transfer for Bretheron-type flows
 is of significant interest for chemical industry as it is widely used in chemical monolith
@@
 large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. Heat transfer is
-also increased in comparison with single phase flow \cite{fukugata-levelset}. While it
+also enhanced in comparison with single phase flow \cite{fukugata-levelset}. While it
 is possible to calculate the flow analytically for small capillary numbers and nearly zero Reynolds
-number \cite{bretherton}, the
+number \cite{bretherton}, such
 results cannot be extrapolated to the wide range of capillary and Reynolds numbers used in the
@@
 
-The two-dimensional geometry (circular tubes, parallel plates) flows are studied extensively in
-experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
+Flows in two-dimensional geometries (circular tubes, parallel plates) have been studied extensively
+in the 
+experimental works of \citet{aussillous-deposition}, and \citet{cerro-bubble-train}, and the
+numerical works of
 \citet{giavedoni-numerical,heil-bretherton}. All abovementioned investigations found that the 
@@
 analysis is valid only for small capillary numbers $Ca\leq 0.003$ and deviates from actual
-measurements for larger ones.
+measurements for larger $Ca$.
 This is caused by a complex interplay of gravitational,
@@
 number from $0$ to $70$ for two-dimensional flows. Later on, \citet{heil-bretherton}
-extended
-the results for the flow between plates up to $Re=300$. This author noted that while the
+studied the flow between plates up to $Re=300$. This author noted that while the
 influence of inertia on the established film thickness is
@@
 experiments for a
-bubble-train flow in capillaries of
-circular and square cross sections for horizontal, upward and downward flows.
+bubble-train flow in capillaries of square cross sections for horizontal, upward and downward flows.
 \citet{shikazono-square} obtained
@@
 capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
-shaped mirochannels. They also presented the experimental correlation for  bubble
+shaped mirochannels. They also presented an experimental correlation for  bubble
 radii
@@
 wang-non-circular} showed interesting phenomena in three-dimensional geometries. It was found
-\cite{heil-threedim,wong-films} that for rectangular
-or square-shaped capillaries there is a transition from an asymmetric to an axisymmetric shape for a
+\cite{heil-threedim,wong-films} that for rectangular capillaries there is a transition from an
+asymmetric to an axisymmetric shape for a
 certain capillary number.
-The non-axisymmetric shape is attributed to the wall normal radius (in Fig.
+The non-axisymmetric shape describes the wall normal radius (in Fig.
 \ref{fig:lbm:benchmark} the wall normal radius is measured along $x$ or $y$ axes) being 
@@
 see Fig. \ref{fig:lbm:benchmark} \cite{heil-threedim}. The dependence
-of the diagonal and axial radii on the capillary number is shown in Fig. 
+of the diagonal and axial radii or a square channel on the capillary number is shown in Fig. 
 \ref{fig:heil:three:dim}.
@@
 \put(50,45){\includegraphics[width=0.4\textwidth]{Figures/inset.eps}}
+\put(20,20){\includegraphics[width=0.3\textwidth]{Figures/legend.eps}}
 \end{overpic}
@@
 \end{figure}
-The capillary number for square channels indicating transition between non-axisymmetric case to
-axisymmetric
-one is reported in a number of
+The capillary number for square channels at which transition between non-axisymmetric case to
+axisymmetric happens is reported in a number of
 works ($\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
@@
 square- and rectangular-shaped capillaries. They indicated a transition of flow regime in a liquid
-slug. Vortexes observed at lower capillary numbers disappear at the capillary number above a
+slug. Vortexes observed at lower capillary numbers disappear as the capillary number goes beyond a
 certain threshold,  Fig. \ref{fig:streamlines:pattern}. For a square channel the critical
@@
 correlation which
-makes it possible to put on the same curve the results of radii dependency on the capillary number
-simulations of rectangular channels with different aspect ratios.
+makes it possible to collapse results of radii dependency on the capillary number
+simulations of rectangular channels with different aspect ratios on a single curve.
 It was also found that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
 2.04$, the interface does not become
-axisymmetric for any capillary number. \citet{wang-non-circular} performed numerical
+circular for any capillary number. \citet{wang-non-circular} performed numerical
 simulations using the Volume of Fluid (VOF) technique for
@@
 
-As it was mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
+As was mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
 phenomena are the VOF method \cite{wang-non-circular}, the level-set method
@@
 The finite element method solves the Bretherton flow as a free-surface problem with a sharp
-interface but without underlying details for the gas.  The lattice Boltzmann method is a continuous
+interface but without gas.  The lattice Boltzmann method is a continuous
 interface method, and therefore provides more flexibility in simulations involving coalescence
@@
 
-The binary liquid free-energy LB model due to \citet{swift} we used
+The binary liquid free-energy LB model due to \citet{swift} that we used
 simulates two liquids with the assumption of uniform overall
-density. As it was indicated by \citet{gupta-review} a flow in microchannels is a complex
-interplay of viscous, gravitational, inertial and surface forces, and the classical Bretherton
+density. %As it was indicated by \citet{gupta-review} a flow in microchannels is a complex
+%interplay of viscous, gravitational, inertial and surface forces. 
+The classical Bretherton
 problem is stated for gas and liquid, which are of
 significantly
-different densities and viscosities, the parameters in our simulations were carefully chosen to
-avoid  inertia
-effects, see Section \ref{sec:numerical:benchmark}. Moreover, the numerical results of
+different density and viscosity. However, the numerical results of
 \citet{giavedoni-numerical} and \citet{heil-bretherton} show
 negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
-numbers. Thus, a major governing factor for microchannel flows is not the density ratio, but
+numbers, $Re<70$. The parameters in our simulations were carefully chosen to
+avoid  inertia
+effects, see Section \ref{sec:numerical:benchmark}. The maximum attained
+Reynolds number $Re<10$. Thus, inertia effects can be neglected. Therefore, a major governing
+factor for microchannel flows is not the density ratio, but
 the
-viscosity ratio. Therefore the uniform density binary liquid model is suitable for this kind of
+viscosity ratio. As a result, the uniform density binary liquid model is suitable for this kind of
 simulations.  
@@
 % TODO: This sentence should probably be moved to a summary/conclusions section?
-The work results are in good agreement with other simulations \cite{heil-threedim,wang-non-circular}.
+%The work results are in good agreement with other simulations
+%\cite{heil-threedim,wang-non-circular}.
 
@@
 the results section. That covers the film thickness dependency on the capillary number, the bubble
-to slug relative velocity and vortex profiles. The paper
+to slug relative velocity and vortex profiles for square microchannels. We also briefly cover flow
+simulations for rectangular channels. The paper
 is concluded with a summary of the main findings.
@@
 body force in the framework of the LBM and not with a pressure difference. Therefore, we can not
-address simulations with upward and downward flows \cite{cerro-bubble-train} which require a
-simulataneous imposement of pressure boundary conditions and the body force. The pressure boundary
-conditions for the binary liquid model is a challenge. Thus, we consider a flow in the
-horizontal direction only and the upward/downward flows with pressure difference imposed are not
-simulated. 
+address simulations with upward and downward flows \cite{cerro-bubble-train} which would require a
+simultaneous imposement of pressure boundary conditions and the body force. 
+%The pressure boundary
+%conditions for the binary liquid model is a challenge. Thus, we consider a flow in the
+%horizontal direction only and the upward/downward flows with pressure difference imposed are not
+%simulated. 
 \begin{figure}[htb!]
@@
 It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least twice as large
-as the interface thickness. The interface thickness in the continuous formulation of the
-free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
+as the interface thickness. The interface thickness occupies $3-5$ nodes due to the continuous
+interface formulation of the binary liquid model. Therefore, the
+film should be
 resolved as at least $6-10$ nodes.
 
-\citet{heil-threedim} specify that the diagonal bubble radius is different from the
-axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.03$. The radii are the same for larger values
+\citet{heil-threedim} and Fig. \ref{fig:heil:three:dim} specify that the diagonal bubble radius is
+different from the
+axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.03$. The radii are approximately the same for
+larger values
 of
@@
 To reduce the computational overhead, we decided to take a simpler approach and exploit the
-inherent symmetry of the problem by simulating only a quarter of the channel (see Appendix
+inherent symmetry of the problem by simulating only a quarter of the channel (see
 \ref{append:sym} for details).
@@
 % TODO: This paragraph should be revisited, somehow it reads a little weirdly..
-It should be noted that the binary liquid model, we used, has a uniform density. A part of
-discussion is
-addressed in the introduction:  to describe the Bretherton problem one needs to take into
+It should be noted that the binary liquid model we used has a uniform density. This issue was
+partially addressed in the introduction:  to describe the Bretherton problem one needs to take into
 account gravitational, viscous, inertial, and surface tension forces. Among all the works, there
 is no full 
-agreement between results even in the limit of Reynolds number $0$.  Moreover, as it was indicated
+agreement between results even in the limit of Reynolds number zero. Moreover,
+as was indicated
 by \citet{kreutzer-taylor} and by \citet{cerro-bubble-train} the pressure distribution, streamlines
 and bubble shape are strongly affected by the length of the liquid slug and the length of the
-bubble. In current simulations we minimize the effects of bubbles train and inertia to have a least
-impact, as the largest Reynolds number is not bigger than $10$ and the lengths of a bubble and a
-channel are chosen to be as to minimize mutual effects of bubbles motion on each other and for the
-film thickness to stabilize. However, more thorough studies are needed in order to understand
+bubble. In the current simulations we minimize the effects of bubbles train and inertia. Inertia
+effect are minimized as the largest Reynolds number is not bigger than $10$. The mutual
+effects of bubbles motion on each other is minimized by choosing lengths of a bubble and a
+channel large enough in comparision with the channel height. However, more thorough studies are
+needed in order to understand
 precisely the abovementioned effects.
@@
 
-The next section briefly explains the binary liquid lattice Boltzmann model.
-The steady-state approach, the dependency of the radii on the moderate capillary number values
-$0.05\leq Ca \leq 6.0$, the velocity pattern, and the relative to the liquid slug bubble velocity 
-are discussed in Section
-\ref{sec:results}.
+%The next section briefly explains the binary liquid lattice Boltzmann model.
+%The steady-state approach, the dependency of the radii on the moderate capillary number values
+%$0.05\leq Ca \leq 6.0$, the velocity pattern, and the relative to the liquid slug bubble velocity 
+%are discussed in Section
+%\ref{sec:results}.
 
@@
 relaxation parameter, $\omega_{\phi}$ is the phase relaxation parameter, and $F_i$ is the external
-force population which
-mimics the force behavior.
+force population responsible for force inclusion to the Navier-Stokes equation.
 
@@
 where $\nu=c_s^2 (\tau-1/2)$ is the kinematic viscosity,
-$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
+$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and
+$\tau_{\phi}=\frac{1}{\omega_{\phi}}$
 are the relaxation parameters of density and phase fields. The first equation of system
@@
 $\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
-$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ in a region when a
+$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ in a region where a
 phase gradient exists. The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
@@
 \label{sec:results}
-This section describes numerical simulations. One can refer to Appendix~\ref{append:scaling} for
+This section describes numerical simulations. We refer to \ref{append:scaling} for
 initialization and scaling procedures. The simulations were conducted on the following 
@@
 the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
-and $\tau_{\mathrm{gas}}=0.7$ giving a gas-liquid viscosity ratio of $10$. The relaxation
+and $\tau_{\mathrm{gas}}=0.7$ giving a gas-over-liquid viscosity ratio of $10$. The relaxation
 parameter for the phase field was $\tau_{\phi}=1.0$.
 
-First, we examine when the steady-state approach is
-achieved. Then, the critical capillary number is identified when the transition from the
+First, we examine when the steady-state is
+reached. Then, the critical capillary number is identified where the transition from the
 non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
@@
 We performed different simulations to understand the number of time steps required for the
-simulation to
+system to
 settle down to a steady state. The grid  was $52 \times 52 \times 750$ which represents
-a quarter of the channel with the initial film thickness of $6.5$ lattice Boltzmann units together
+a quarter of the channel with the initial film thickness of $6.5$ lattice units together
 with the body force $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
@@
 \ref{table:steady:state}. While the
-capillary number variation is around $5$ percent, the radii variation is $0.1$ percent. 
+capillary number variation is around $5$ percent, the radii variation is $0.1$ percent. Note, that
+the interface velocity is defined as the bubble tip velocity in the microchannel center. Multiphase
+lattice Boltzmann models is known to have spurious currents near the interface
+\cite{pooley-spurious,shan-spurious}. Therefore, the bubble interface velocity is influenced by
+spurious currents and capillary numbers show larger variation than radii. 
 \begin{table}
@@
 the bubble and approach the steady state. The small noise in the capillary number is connected to
-the identification of the interface. One needs to address the interpolation and the spurious
-currents existing in the system. All other parameters are relaxed to the
+the identification of the interface and spurious currents in the system. The spurious currents
+influence the bubble velocity identification which is taken at the front bubble tip. The binary
+liquid model is the continuous interface model. Thus one needs to interpolate phase function data to
+obtain the location of the bubble interface $\phi=0$ and its velocity. Thus, there is a larger
+deviation for capillary numbers. All other parameters are relaxed to the
 steady-state.\label{table:steady:state}}
@@
 of the velocity are taken on the contour where $\phi=0$. This corresponds
- to the bubble interface. One can see in Fig. \ref{fig:velocity:contour} the contour plot and the
+ to the bubble interface. That allows to check whether the bubble is propagating as a whole rigid
+body or it's shrinking or elongating. This is one of the characteristics of the steady-state.  One
+can see in Fig. \ref{fig:velocity:contour} the contour plot and the
 corressponing streamwise velocity
@@
 examine only the points corresponding to the front and the rear bubble caps. If they are equal, then
-it is exactly the velocity with which the bubble propagates in the microchannel.
+it is exactly the velocity with which the bubble propagates in the microchannel as a rigid body.
 For a physical analogy, one can imagine the rotating ball propagating in the streamwise direction.
+The front and rear ball points have the same streamwise velocity. 
 \begin{figure}[ht]
@@
 Due to computational power restrictions we can only access the
-transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
-is achievable. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
-were conducted. In comparison with the results of the \citet{heil-threedim}
+region of $Ca\geq 0.05$. In this region the proper resolution of the film thickness
+is achievable.  To identify the critical capillary number $\widehat{Ca}$ a number of
+simulations
+were conducted, see Table \ref{table:transition:results}. In comparison with the results of the
+\citet{heil-threedim}
 ($\widehat{Ca}=0.04$) our
-results are more close to the VOF continuous interface method simulation by
-\citet{wang-non-circular} ($\widehat{Ca}=0.1$). Our data also fall within $5$ percent radii
-variation on the
-capilllary number of experimental dependency of
-\citet{shikazono-square}. The transition for current simulations happens for $Ca\approx 0.09$, see
-Table
-\ref{table:transition:results}.  One can see
+results are closer to the VOF continuous interface method simulation by
+\citet{wang-non-circular} ($\widehat{Ca}=0.1$). The transition for current simulations
+happens for $Ca\approx 0.09$ calculated as linear interpolation between data presented in Table
+\ref{table:transition:results}. 
+%Our data also fall within $5$ percent radii
+%variation on the
+%capilllary number of experimental dependency of
+%\citet{shikazono-square}.   
+One can see
 two examples of non-axisymmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see
@@
 \caption{Simulation results in terms of $R_{axis}$ and $R_{diag}$ for the transition region between
-non-axisymmetric and axisymmetric cases. The transition occurs at $\widehat{Ca}=0.09$.
+non-axisymmetric and axisymmetric cases. The transition occurs at $\widehat{Ca}=0.09$, which is a
+linear interpolation between data with radii not equal and equal to each other. The grid for
+simulations was chosen as $50\times50\times1500$ for a quarter of a channel. That means that the
+corresponding interpolation error of determining radii is a half of the reverse grid number,
+$\frac{1}{2 N_x}=0.01$. Thus, if radii values are within $1$ percent we consider them equal to
+each other. We also included a corresponding Reynolds number to show that inertia effects can be
+neglected.
 \label{table:transition:results}}
@@
 \label{sec:velocity}
-The velocity pattern is known to change its behavior depending on the capillary number. For
+The liquid velocity pattern is known to change its behavior depending on the capillary number. For
 relatively low capillary numbers $Ca<0.6$, in a reference frame moving with the bubble a vortex 
@@
 vortex in front of the bubble. The transition capillary number specified by
-\citet{heil-threedim} for a square channel is $Ca=0.691$. However, the
-distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well as on
+\citet{heil-threedim} for a square channel as $Ca=0.691$. However, the
+distribution of vorticity strongly depends on the Reynolds number \cite{heil-bretherton}, as well as
+on
 the
@@
 One can
-see in Fig. \ref{fig:streamlines:pattern} a clear transition between associated patterns. The
-transition capillary number of
-\citet{heil-threedim} is slightly higher than that in the simulation results. In present simulations
-the                                               
-transition occurs at $Ca=0.6-0.7$, see Fig. \ref{fig:streamlines:pattern}.
+see in Fig. \ref{fig:streamlines:pattern} a clear transition between associated patterns. Among all
+different simulation runs with different grids and initial conditions, the
+transition occurs at $Ca=0.6-0.7$, see Fig. \ref{fig:streamlines:pattern}. The
+transition is identified between regimes where there exists and does not exist a vortex in the
+slug. Either all streamwise slug velocities have the same sign (no vortex), or have different signs
+(vortex). It is indicated by \citet{heil-bretherton,kreutzer-reactors} that the bubble train flow
+depends on the slugs and bubbles lengths. Thus, we expect the difference of bubble train simulations
+and simulations of a semi-infinite air finger propagation \cite{heil-threedim}. The transition also
+depends on the identification of the bubble velocity
+which is influenced by spurious currents and interpolation errors.
 \begin{figure}[ht]
@@
 created
-in front of the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
+in front of the bubble for $Ca=0.63$. For presented simulations the transition happens between
+$Ca=0.47$ and $Ca=0.63$, which
 is a different value from $Ca=0.69$ \cite{heil-threedim}. However, we performed a number of
 different types of simulations (different force, width initialization). The critical transition
-capillary number $\widehat{Ca}$ is not well imposed and varies from $\widehat{Ca}=0.6-0.7$.
+capillary number $\widehat{Ca}$ is not well fixed and varies from $\widehat{Ca}=0.6-0.7$.  
 \label{fig:streamlines:pattern}}
@@
 \begin{figure}[ht]
-\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_axis.eps}
-\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_diag.eps}\\
+\includegraphics[width=0.8\textwidth]{Figures/bubble_rad_axis.eps}\\
+\includegraphics[width=0.8\textwidth]{Figures/bubble_rad_diag.eps}\\
 \caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
@@
 initialization, and different grids: $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
-\times 82 \time 2400$. All the results are consistent. Therefore one can conclude that for the
+\times 82 \time 2400$. All our results are consistent. Therefore one can conclude that for the
 capillary number $Ca\geq
@@
 
-Present simulations show that the corresponding critical capillary number is close to
+The present simulations show that the corresponding critical capillary number is close to
 $\widehat{Ca}=0.1$ that agrees with the work of \citet{wang-non-circular} and is different from the
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Figures/capillaries_comparison_heil.eps}\\
-\includegraphics[width=\textwidth]{Figures/capillaries_comparison_wang.eps}\\
+\includegraphics[width=0.8\textwidth]{Figures/capillaries_comparison_heil.eps}\\
+\includegraphics[width=0.8\textwidth]{Figures/capillaries_comparison_wang.eps}\\
 \caption{The comparison for the axial and the diagonal radii
@@
 velocity. As far as the steady state is achieved one can take the liquid superficial velocity
-exactly at the distance in the middle between bubbles. The liquid superficial velocity is
+exactly at the cross section in the middle between bubbles. The liquid superficial velocity is
 defined as:
@@
 The relative velocity is an important quantity of mass flow characterization
-\cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and 
-inertia effects \citet{heil-bretherton} significantly influence  mass transfer. 
+\cite{kreutzer-taylor,yue-mass}. However, there are other important characteristics that
+significantly influence the mass transfer, i.e. the frequency of the bubbles \cite{kreutzer-taylor}
+and 
+inertia effects \cite{heil-bretherton}. 
 Future work is planned to study this phenomenon and calculate the mass transfer coefficient by
@@
 
+\subsection{Rectangular channel simulations}
+While the main focus of this paper is on the square microchannel simulations, we performed a number
+of simulations for rectangular channels with varying aspect ratio
+$\alpha=\frac{W}{H_{\mathrm{eff}}}$.\citet{heil-threedim} performed a number of simulations for the
+propagation of semi-infinite air finger in the microchannel. They indicated that the radii in $x$
+and $y$ directions depend on the capillary number and are increasing with the aspect ratio $\alpha$
+increases. The same tendenion can be seen for current simulation results, see Fig.
+\ref{fig:aspect:ratio}. One can see that radii do not coincide with each other. The same is
+indicated by \citet{heil-threedim} who mentioned that for radii to coincide one needs to have as
+long as $100$ shorter semi-axis sides bubble length. This is not possible to achieve with the
+current bubble train simulations.
+\begin{figure}[htb!]
+\includegraphics[width=\textwidth]{Figures/rectangle.eps}
+\caption{Radii in $x$ ($R_h$) and $y$ ($R_v$) directions for rectangular microchannels with
+different aspect ratios. Radii are rescaled on the shorter semi-axis value. One can see that with
+the increase of $\alpha$ the radii increase as well. However, due to small bubble lengths
+$R_h\neq R_v$ \cite{heil-threedim}. \label{fig:aspect:ratio}}
+\end{figure}
+
+They as well indicated that the simulations results with different aspect ratio can be
+put on one curve by calculating non-dimensional radius $s_{\infty}$ that scales different aspect
+ratio microchannels. It is based on the calculation of  $r_{\infty}$ (the radius that the air
+finger will have at an infinite distance from the bubble tip):
+\begin{equation}
+\begin{aligned}
+&\frac{r_{\infty}}{\frac{H_{\mathrm{eff}}}{2}}=\frac{\sqrt{\frac{Q}{\pi}}}{\frac{H_{\mathrm{eff}}}{2
+} }\\
+&s_{\infty}=\frac{r_{\infty}}{\frac{H_{\mathrm{eff}}}{2}}\alpha^{-1/2},
+\end{aligned}
+\end{equation}
+where $Q$ is the area occupied by air in the crossection. \citet{heil-threedim} indicated a simple
+physical interpretation of this phenomena:  providing that air is injected with the same
+velocity to the channels with different aspect ratio but the same area, the occupied area by bubble
+shouldn't depend on the aspect ratio. The current results, see Fig. \ref{fig:one:curve} show that
+this postulate can be extended for bubble train flows with different radii $R_h \neq R_v$ in
+comparison with the original work \cite{heil-threedim}. 
+\begin{figure}[htb!]
+\includegraphics[width=\textwidth]{Figures/onecurve.eps}\\
+\caption{The non-dimensional radius $s_{\infty}$ for different aspect ratio
+rectangular microchannels. One can see that all the curves coincide
+independently of the aspect ratio.\label{fig:one:curve}}
+\end{figure}
+
 \section{Conclusion}
@@
 interface thickness \cite{kuzmin-binary2d} the results were shown to be consistent with those
-available in the literature. Note that the literature results show some variations with
-no established benchmark which can be used to validate data. This work falls in the bulk of the
-literature data results. For instance, the  transition from the non-axisymmetric to the
+available in the literature. Note that the literature results show some variations. This work falls
+within the range of data presented in the literature. For instance, the  transition from the
+non-axisymmetric to the
 axisymmetric case is shown to be at $\widehat{Ca}=0.09$ which is close to results of
@@
 \cite{kreutzer-taylor,yue-mass} on the design of microchannels. Our results show that the lattice
-Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels. Results
-coincide with bulk literature results.
+Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels. 
+%Results
+%coincide with bulk literature results.
 
@@
 \label{append:scaling}
-The scaling procedure is thoroughly described in our previous work \cite{kuzmin-binary2d}. Here, we present
+The scaling procedure has been extensively  described in our previous work \cite{kuzmin-binary2d}.
+Here, we present
 only an outline of how the simulations are initialized:

### 2011-06-14 Results/rectangle.py

@@ -98,6 +98,8 @@ def extract_profiles(name):
     print Get_Zero(prof_axis_y)
     axis_zero_x=Get_Zero(prof_axis_x)/(dims[1]-2.0)
     axis_zero_y=Get_Zero(prof_axis_y)/(dims[1]-2.0)
+    area_indexes=numpy.where(phase_mid[1:,1:]<=0.0)
+    area=float(len(area_indexes[0]))/((dims[0]-2.0)*(dims[1]-2.0))
     #diag_zero=math.sqrt(2.0)*Get_Zero(prof_diag)/(dims[0]-2.0)
     
     
@@ -121,41 +123,50 @@ def extract_profiles(name):
     #print "Interface=",vz[0,z2%dims[2]]
     #print "Superficial=",superficial
     #return axis_zero,diag_zero,capillary,reynolds,vz[0,z2%dims[2]],superficial
-    return capillary,axis_zero_x,axis_zero_y
+    return capillary,axis_zero_x,axis_zero_y,2.0*math.sqrt(area/math.pi)
     
 def get_radii_rect12():
     dirs=['1','2','4','6','7','8','9','10']
     capillaries=[]
     axis_zeros_x=[]
     axis_zeros_y=[]
+    scales=[]
     for dir in dirs:
         name="Rectangle12/"+dir+"/phase240000.vts"
         print name
-        cap,axis_x,axis_y=extract_profiles(name)
+        cap,axis_x,axis_y,scale=extract_profiles(name)
         capillaries.append(cap)
         axis_zeros_x.append(axis_x)
         axis_zeros_y.append(axis_y)
+        scales.append(scale)
     pylab.figure(1)
     pylab.plot(capillaries,axis_zeros_x,'k-')
     pylab.plot(capillaries,axis_zeros_y,'k--')
-    numpy.savetxt("rectangle12.txt",zip(capillaries,axis_zeros_x,axis_zeros_y))
+    pylab.figure(2)
+    pylab.plot(capillaries,scales,'ks')
+    numpy.savetxt("rectangle12.txt",zip(capillaries,axis_zeros_x,axis_zeros_y,scales))
 
 def get_radii_rect14():
     dirs=['3','4','5','7','8','9']
     capillaries=[]
     axis_zeros_x=[]
     axis_zeros_y=[]
+    scales=[]
     for dir in dirs:
         name="Rectangle14/"+dir+"/phase240000.vts"
         print name
[…]

### 2011-06-20 Text/paper.tex
Jos modifications.

@@
 \begin{frontmatter}
-\title{Three-dimensional binary-liquid lattice Boltzmann simulation of microchannels with square
+\title{Three-dimensional binary-liquid lattice Boltzmann simulation of microchannels with rectangular
 cross sections}
@@
 \cite{kuzmin-binary2d}. This paper is the continuation of our previous work with simulations of
-three-dimensional channels with square crosssections in the range of the capillary number $0.05\leq
+three-dimensional channels with rectangular (mostly square) cross sections in the range of the capillary number $0.05\leq
 Ca \leq
@@
 results are closer to the VOF continuous interface method simulation by
-\citet{wang-non-circular} ($\widehat{Ca}=0.1$). The transition for current simulations
+\citet{wang-non-circular} ($\widehat{Ca}=0.1$).
+Due to linear approximation we consider a bubble to have a circular cross section if $R_{axis}$ and $R_{diag}$ differ by less than $1\%$.
+The transition for current simulations
 happens for $Ca\approx 0.09$ calculated as linear interpolation between data presented in Table
@@
 of simulations for rectangular channels with varying aspect ratio
-$\alpha=\frac{W}{H_{\mathrm{eff}}}$.\citet{heil-threedim} performed a number of simulations for the
+$\alpha=\frac{W}{H_{\mathrm{eff}}}=\frac{a_x}{a_y}$, where $a_x$ and $a_y$ are sides of the rectangular in $x$ and $y$ directions.\citet{heil-threedim} performed a number of simulations for the
 propagation of semi-infinite air finger in the microchannel. They indicated that the radii in $x$
-and $y$ directions depend on the capillary number and are increasing with the aspect ratio $\alpha$
-increases. The same tendenion can be seen for current simulation results, see Fig.
-\ref{fig:aspect:ratio}. One can see that radii do not coincide with each other. The same is
-indicated by \citet{heil-threedim} who mentioned that for radii to coincide one needs to have as
-long as $100$ shorter semi-axis sides bubble length. This is not possible to achieve with the
+and $y$ directions  are increasing with the aspect ratio $\alpha$
+increases. The same trend can be seen for the current simulation results, see Fig.
+\ref{fig:aspect:ratio}. One can see that diagonal and axial radii are not the same as they were in square channels for the capillary number range investigated here. The same is
+indicated by \citet{heil-threedim} who mentioned that for radii to coincide one needs to have really long bubbles, at least $100$ shorter semi-axis bubble length for the aspect ratio $\alpha=1.5$.
+This is not possible to achieve with the
 current bubble train simulations.
@@
 \includegraphics[width=\textwidth]{Figures/rectangle.eps}
-\caption{Radii in $x$ ($R_h$) and $y$ ($R_v$) directions for rectangular microchannels with
-different aspect ratios. Radii are rescaled on the shorter semi-axis value. One can see that with
-the increase of $\alpha$ the radii increase as well. However, due to small bubble lengths
-$R_h\neq R_v$ \cite{heil-threedim}. \label{fig:aspect:ratio}}
+\caption{Radii in $x$ ($R_x$) and $y$ ($R_y$) directions for rectangular microchannels with
+different aspect ratios. Radii are rescaled on the length of channel in $y$ direction. One can see that with
+the increase of $\alpha$ the radii increase as well. However, $R_x\neq R_y$. It was indicated by \citet{heil-threedim} one needs semi-infinite bubble to achieve $R_x=R_y$. \label{fig:aspect:ratio}}
 \end{figure}
@@
 physical interpretation of this phenomena:  providing that air is injected with the same
-velocity to the channels with different aspect ratio but the same area, the occupied area by bubble
-shouldn't depend on the aspect ratio. The current results, see Fig. \ref{fig:one:curve} show that
-this postulate can be extended for bubble train flows with different radii $R_h \neq R_v$ in
-comparison with the original work \cite{heil-threedim}. 
+velocity to the channels with different aspect ratio but the same area, the resulting occupied area by the bubble
+should not depend on the aspect ratio.
+That is quite interesting phenomena indicating that if bubbles are infinite long then they would have the same radius and will be circular.
+The current results, see Fig.
+\ref{fig:one:curve} show that
+this hypothesis can be extended for bubble train flows with different radii $R_h \neq R_v$. 
 \begin{figure}[htb!]

### 2011-06-20 Results/rectangle.py
Proper graphs.

@@ -233,11 +233,11 @@ def summarize_results():
     pylab.xlim(xmin=0.1,xmax=2.0)
     pylab.ylim(ymin=0.65,ymax=1.2)
     pylab.xlabel(r'''$Ca$''',fontsize=20)
-    pylab.ylabel(r'''$R_h,R_v$''',fontsize=20)
+    pylab.ylabel(r'''$R_x,R_y$''',fontsize=20)
     fig.subplots_adjust(left=0.15,bottom=0.15)
     pylab.xticks(fontsize=16)
     pylab.yticks(fontsize=16)
-    legs=[[r'''$R_v,\alpha=1.'''+name[1]+r'''$''',r'''$R_h,\alpha=1.'''+name[1]+r'''$'''] for name in files]
+    legs=[[r'''$R_y,\alpha=1.'''+name[1]+r'''$''',r'''$R_x,\alpha=1.'''+name[1]+r'''$'''] for name in files]
     legs=numpy.ravel(legs)
     pylab.legend(legs,fancybox=True)
     pylab.savefig('rectangle.eps',dpi=300,format="EPS")
@@ -269,6 +269,6 @@ if __name__=="__main__":
     #get_radii_rect14()
     #get_radii_rect16()
     #get_radii_rect18()
-    #summarize_results()
-    one_curve()
+    summarize_results()
+    #one_curve()
     pylab.show()

### 2011-09-12 Text/Submission/paper.tex

--- 2011-06-20 Text/paper.tex
@@
 \begin{figure}[htb!]
-\begin{overpic}[width=\textwidth]{Figures/capillary_width_heil.eps}
-\put(50,45){\includegraphics[width=0.4\textwidth]{Figures/inset.eps}}
-\put(20,20){\includegraphics[width=0.3\textwidth]{Figures/legend.eps}}
+\begin{overpic}[width=\textwidth]{capillary_width_heil.eps}
+\put(50,45){\includegraphics[width=0.4\textwidth]{inset.eps}}
+\put(20,20){\includegraphics[width=0.3\textwidth]{legend.eps}}
 \end{overpic}
@@
 \begin{figure}[htb!]
-\includegraphics*[bb=153 610 410 717,width=\textwidth]{Figures/benchmark_classical.eps} 
+\includegraphics*[bb=153 610 410 717,width=\textwidth]{benchmark_classical.eps} 
 \caption{The classical benchmark layout describes a semiinfinite gas bubble
@@
 \begin{figure}[htb!]
-\includegraphics*[bb=152 470 415 713,width=\textwidth]{Figures/benchmark_lbm.eps}
+\includegraphics*[bb=152 470 415 713,width=\textwidth]{benchmark_lbm.eps}
 \caption{The lattice Boltzmann benchmark layout as used in present calculations. The dimensions of
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
-\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
+\includegraphics[width=\textwidth]{velocity_interface_contour.eps}\\
+\includegraphics[width=\textwidth]{velocity_interface_values.eps}\\
 \caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
@@
 \begin{figure}[ht]
-\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca5.eps}
-\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca13.eps}\\
+\includegraphics[width=0.5\textwidth]{phase_crossection_ca5.eps}
+\includegraphics[width=0.5\textwidth]{phase_crossection_ca13.eps}\\
 \caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
@@
 \begin{figure}[ht]
-\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
-\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
+\includegraphics[angle=90,width=\textwidth]{stream_ca47.eps}\\
+\includegraphics[angle=90,width=\textwidth]{stream_ca63.eps}\\
 \caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
@@
 \begin{figure}[ht]
-\includegraphics[width=0.8\textwidth]{Figures/bubble_rad_axis.eps}\\
-\includegraphics[width=0.8\textwidth]{Figures/bubble_rad_diag.eps}\\
+\includegraphics[width=0.8\textwidth]{bubble_rad_axis.eps}\\
+\includegraphics[width=0.8\textwidth]{bubble_rad_diag.eps}\\
 \caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Figures/bubble_ca_large.eps}
+\includegraphics[width=\textwidth]{bubble_ca_large.eps}
 \caption{The bubble shape for the capillary number $Ca=6.43$ is presented. $z$ increases in the
@@
 \begin{figure}[ht]
-\includegraphics[width=0.8\textwidth]{Figures/capillaries_comparison_heil.eps}\\
-\includegraphics[width=0.8\textwidth]{Figures/capillaries_comparison_wang.eps}\\
+\includegraphics[width=0.8\textwidth]{capillaries_comparison_heil.eps}\\
+\includegraphics[width=0.8\textwidth]{capillaries_comparison_wang.eps}\\
 \caption{The comparison for the axial and the diagonal radii
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Figures/relative_velocity.eps}
+\includegraphics[width=\textwidth]{relative_velocity.eps}
 \caption{Relative velocity comparison between current simulations and simulations of
@@
 \begin{figure}[htb!]
-\includegraphics[width=\textwidth]{Figures/rectangle.eps}
+\includegraphics[width=\textwidth]{rectangle.eps}
 \caption{Radii in $x$ ($R_x$) and $y$ ($R_y$) directions for rectangular microchannels with
@@
 \begin{figure}[htb!]
-\includegraphics[width=\textwidth]{Figures/onecurve.eps}\\
+\includegraphics[width=\textwidth]{onecurve.eps}\\
 \caption{The non-dimensional radius $s_{\infty}$ for different aspect ratio

### 2011-09-22 Text/Submission/answers.tex
Started doing the referee report.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{bm}
\usepackage[sort&compress]{natbib}
\usepackage[normalem]{ulem}	% Part of the standard distribution

\begin{document}
It is our pleasure to address the reviewers' comments. Please consider our answers below.
\section{Reviewer I}

\textbf{Question:}
\begin{quotation}
When Re is very small, the authors neglect the density ratio of gas and liquid phase. I think more
reasons are still needed to explain the uniform density, especially on the interface. One of the
challenges for free-energy LB model is how to meet the Galilean invariance because of the density
gradient on the interface.
\end{quotation}

\textbf{Answer:} 
We added additional text explaining the choice of the binary-liquid model towards the simulation
of the Bretheron-Taylor phenomenon (see lines 99-112). The Galilean invariance is usually fixed by
third-order velocity terms in the equilibrium, see corresponding work of \citet{qian-galilean}.
However, the velocities in the current simulations are of order $0.001-0.01$ and we can neglect the
non-Galilean invariance.

\textbf{Question:}
\begin{quotation}
In Figures 7 and 8, it's better to give the reported curves by Liu and Wang and Hazel and Heil, so
that it's easy to find the agreement is well or not.
\end{quotation}

\textbf{Answer:} We redrew all the figures. For the figure 8 the data was extracted from the
figure and rescaled for comparison reasons.

\section{Reviewer II}

\textbf{Question:}
\begin{quotation}
The main quantity that the paper focuses on is the thickness of the film between the bubbles and the
wall. As shown by Figure 8 and as mentioned by the Authors in Section 4, the film thickness varies
along the bubble. However, the Authors do not explain the choice of the location of the measurement
of the film thickness. This choice is of primary importance when comparing results with those of
the
literature and in particular with works on finger propagation in which the film thickness is
measured
”at infinity”.
\end{quotation}

\textbf{Answer:} The explanations are added in lines 434-453.

\textbf{Question:}
\begin{quotation}
It may be appropriate to introduce Figure 8 on thickness variation before the main results on ﬁlm
thickness.
\end{quotation}

\textbf{Answer:} We agree and moved the section before the results on the capillary number
dependency.

\textbf{Question:}
\begin{quotation}
The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
Reynolds
number effects on the film thickness for capillary numbers larger than 0.003” is in contradiction
with
the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
relatively
wide range of Reynolds numbers.”
\end{quotation}

\textbf{Answer:} We elaborate the Reynolds number influence in lines 34-43.

\textbf{Question:}
\begin{quotation}
Although inertia has been shown to have a minor effect on ﬁlm thickness, inertia does influence
some flow variables in the Bretherton’s problem (see Ref. 3 for instance). To check whether inertia
is important or not, and to discuss the assumption of uniform density, the Authors should provide
the reader with the Reynolds numbers associated with each simulation. This is also required for the
consistency of Figure 7 in which results are compared with other works of the literature.
\end{quotation}

\textbf{Answer:} Thank you for pointing this out. We included the discussions of the Reynolds
effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
the capillary range regime and the grid depedency study.

\textbf{Question:}
\begin{quotation}
In this work, the diphasic flow is driven by a body force. The body force is therefore the control
parameter in the simulations. The Authors should give more detail about this body force. Indeed,
although the estimation of the body force from the pressure gradient is well described, it is not
clear
where this body force appears in the governing equations (Equations 2, 3 and 4). Does this body
force correspond to the variable $F_i$ in Equation 2? Should not this force or its equivalent appear
in
the Navier-Stokes equations (Equation 4)?
\end{quotation}

\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
how it
mimics the force behavior.

\textbf{Question:}
\begin{quotation}
I would suggest to calculate the Bond number to check if the shapes of the bubbles and the main
results of the paper are not inﬂuenced.
\end{quotation}

\textbf{Answer:} 
The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
current simulation equals to $7$, which is less than the number for which the shape of the bubble is
affected according to \cite{zheng-large-ratio}.

\textbf{Question:}
\begin{quotation}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estimations of the body force to apply and of the film thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a different way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.
\end{quotation}

\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
about initialization and this was taken into account by reducing repetitive text and
introducing step-by-step explanations, see Section about initialization (lines $259-305$). We also
added
the
recommendations for controlling parameters in the same section.  The
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann method are refered to the non-dimensionalization procedure. Therefore, we
decided to
keep suggestions for controlling and initialization of simulations as informative guidelines for the
LBM community.

\textbf{Question:}
\begin{quotation}
 Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
obtained
under the assumption of an axisymmetric Poiseuille flow. However, among the ”Key words”, one can
read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
that are done in this work. The Authors should discuss this choice of estimating the pressure
gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
in Equation 5).
\end{quotation}

\textbf{Answer:} The factor $1/8$ in Eq. (5) is taken for the maximum velocity for the Poiseille
flow between
plates. Our assumption is that the bubble travels with the constant velocity which equals to the
maximum velocity of the Poiseuille profile between plates. Factor $1/12$ comes from the averaging
of the flow rate. However, bubble moves faster than the liquid, i.e. the average liquid velocity.
That is why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
well.

\textbf{Question:}
\begin{quotation}
The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
journal of applied science. The Authors would improve signiﬁcantly the manuscript by defining more
rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
A suggestion is to emphasize within an array the links between physical parameters and the model
parameters. The variables $i$, $f_i^{eq}$ , $g_i$ should be deﬁned and the Equations for $f$ and $g$
in Equation
3 should be described (which equation corresponds to continuity equation...). Equation 2 could
perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
physical variable f of Equation 3. The notation $i = 1 \div 8$ is unusual, it would be clearer to
use a
classical notation such as $1 \leq i \leq 8$. The sound speed $c_s^2$ should be deﬁned just after
its
introduction in line 107.
\end{quotation}

\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the
physical parameters with the lattice Botlzmann parameters (lines $149-160$). As well more
explanations are
put towards the
interface surface tension description (lines $136-139$). The reference to the
continuity and the Navier-Stokes equations are added (lines $140-143$). The equation (2) is not
rewritten in terms of
generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
governing equations. The notations have been corrected.

\textbf{Question:}
\begin{quotation}
The name of the Section ”Lattice Boltzmann benchmark” lets the reader think that this Section
would describe the numerical procedure (assumptions, choice of the geometry and the grid, boundary
conditions, control parameters, initialization of the scheme). This is not the case because this
Section
develops the challenges and the difficulties of using the Lattice Boltzmann method to bubble ﬂows.
This Section should be renamed or reorganized.
\end{quotation}

\textbf{Answer:} We renamed the section to ``Approach to benchmark''.

\textbf{Question:}
\begin{quotation}
The Authors should make a distinction between variables in lattice units and dimensional or non
dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y - 2$ where $N_y$ is
the
number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
is in
contradiction with the use of $H_{eff}$ in Equation 5. The Authors should define each variable
without
any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
improvements
would fit well in the Subsection named ”The nondimensionalization and initialization procedure”
which does not in fact introduce much nondimensionalisation procedure.
\end{quotation}

\textbf{Answer:} We added a decription paragraph about the approach to units (lines 149-160). In
fact, all equations are the same in the
lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
preferred units as soon as the capillary number is the same in the LBM and the physical world.

\textbf{Question:}
\begin{quotation}
Some notations are unclear. For instance, the film thickness is denoted as $h_{\infty}$ in Equation
7 whereas
it is denoted $\delta$ in most of the paper. The notations w or $\delta_{sim}$ are also found before
line 381 or in
Table 3 respectively. The thickness of the interface is denoted $\xi$ in Figure 1 but the same
variable
is
denoted $5\xi$ in Table 1. The speed of the bubble is denoted $U_{\mathrm{bubble}}$ in some places
and $U_b$ in others.
The variable $Uin$ Equation 5 does not seem to be deﬁned.
\end{quotation}

\textbf{Answer:} We thoroughly corrected all the inconsistencies.

\textbf{Question:}
\begin{quotation}
Three figures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
unusual
for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
which are not essential.
\end{quotation}

\textbf{Answer:} We eliminated unnecessary pictures and references to them.

\textbf{Question:} 
\begin{quotation}
In Figure 7, the points of data reported from the numerical work of Ref. 3 exhibit a surprisingly
noisy trend that corresponds more to an experimental study. The Authors should explain how they
obtain this data and plot them.
\end{quotation}

\textbf{Answer:} We added a description to the figure caption explaining data acquisition.

\textbf{Question:} 
\begin{quotation}
The plots of Figure 8 are of less quality than the other figures of the paper.
\end{quotation}

\textbf{Answer:} Unfortunately, it was not possible to extract data as a picture from the paper by
\citet{sehgal-microchannel}. Thus, the curves were extracted from the graphs and represented in a
new figure scaled to the present simulations for comparison reasons.

\textbf{Question:}
\begin{quotation}
The work of Yang seems to be close from the present study since it involves the
solution of Bretherton
flow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
differences between their work and this paper.
\end{quotation}

\textbf{Answer:} We outlined the differences in the introduction section (lines $92-100$).

\textbf{Question:} It would be interesting if the Authors could mention a study in which a level-set
method is used in
order to complete the bibliography.

\textbf{Answer:} We added references to the introduction section [17-19].

\textbf{Questions:} 
\begin{quotation}
A secondary quantity of interest is the bubble speed or in a non-dimensional
form the capillary
number. This quantity is measured at the center of the bubbles. The Authors should describe how
this quantity as well as the center of the bubble are measured.
\end{quotation}

\textbf{Answer:} Thank you for pointing that out. In the classical Bretherton problem
$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
results are the same.

\textbf{Question:} 
\begin{quotation}
 If possible, the Authors should present results showing that steady state is reached. It would
also
be interesting if the Authors show some patterns of the flow by plotting for instance the vector
ﬁeld
of velocity around the bubbles.
\end{quotation}

\textbf{Answer:} We added a steady-state study (lines $334-341$). We added an additional
paragraph which shows the change of flow pattern for the
streamlines as it is indicated in literature (lines $463-471$, Figure 8).

\textbf{Question:}
\begin{quotation}
In the Abstract, the Authors should give the range of real capillary numbers they measured instead
of the targeted values ($Ca_{lit}$).
\end{quotation}

\textbf{Answer:} We changed it to the measured capillary numbers.

\textbf{Question:} 
\begin{quotation}
The Authors should discuss more their results (Sections 5-4, 5-5, 5-6). Did the Authors measure the
pressure drops at the front and rear interfaces of the bubbles?
\end{quotation}

\textbf{Answer:} We added additional information about results to the corresponding sections. The
pressure drop is not measured as they are influenced largely by Reynolds number according to
\citet{heil-bretherton}.

\textbf{Question:}
\begin{quotation}
The Authors mention the shooting method. It is not clear if this method is really used or just
mentioned as a part of the discussion.
\end{quotation}

\textbf{Answer:} This is a part of discussion. The corresponding
sentence is added (lines $459-461$).

\textbf{Question:}
\begin{quotation}
It seems more appropriate to introduce the variables of the Bretherton’s problem as well as to plot
Figure 1 elsewhere than in the Section ”Results”.
\end{quotation}

\textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
separated it from the numerical benchmark layout.

\textbf{Question:}
\begin{quotation}
Minor corrections...
\end{quotation}

\textbf{Answer:} All the corrections are taken into the account.

\bibliographystyle{unsrtnat}
\bibliography{answers}

\end{document}

### 2011-09-22 Text/Submission/paper.tex
Started doing the referee report.

@@
 significantly
-different density and viscosity. However, the numerical results of
+different density and viscosity. {\color{red} The system is simulated in a right manner if all
+non-dimensional parameters are matched in real physical system and simulation system.
+However, one needs to consider the leading non-dimensional numbers which are responsible for
+physical phenomenon. Even, the binary-liquid model cannot match the gas-liquid density ratio, we
+consider it as a non-leading parameter. First of all, the numerical
+results of
 \citet{giavedoni-numerical} and \citet{heil-bretherton} show
@@
 effects, see Section \ref{sec:numerical:benchmark}. The maximum attained
-Reynolds number $Re<10$. Thus, inertia effects can be neglected. Therefore, a major governing
+Reynolds number $Re<10$. The second reason is that density ratio is involved while describing a
+bubble shape change. There is a corresponding number defining the shape change, Bond number. Bond
+number is $Bo=\frac{Ca\,Re}{Fr^2}$. For real microchannel systems the Bond number is less than
+$Bo<0.1$ and we can neglect a bubble shape change.}
+Therefore, a major governing
 factor for microchannel flows is not the density ratio, but

### 2011-09-22 Engauge/draw.py

@@ -1,0 +1,112 @@
+#!/usr/bin/python
+import numpy
+import pylab
+
+def paint_curves():
+
+    wang=numpy.genfromtxt("wang_axial.csv",delimiter=',',dtype=float)
+    print wang
+    fig=pylab.figure()
+    wang2=numpy.array(wang)
+    wang2[:,0]=0.0
+    print wang
+    ind=numpy.where(wang2>1.0)
+    print ind
+    #ind[1]=ind[1]+1
+    wang[ind]=-1.0
+    max_length=max(wang[:,0])
+    wang[:,0]=max_length-wang[:,0]
+    #pylab.plot(wang[:,0],wang[:,3][::-1],"kv",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0],wang[:,3],"kv",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0],wang[:,1],"ko",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0],wang[:,2],"ks",markersize=8) #,markerfacecolor="white")
+
+    #leg=pylab.legend([r'''$Ca=1.0$''',r'''$Ca=0.3$''',r'''$Ca=0.08$'''],loc=3,fancybox=True)
+   
+    pylab.ylim(ymin=0.0)
+    
+    #for counter,file in enumerate(files):
+        
+    #    mat=numpy.loadtxt(file)
+        
+        #print mat
+    #    pylab.semilogx(mat[:,0],numpy.divide((mat[:,4]-mat[:,5]),mat[:,4]),"kv",markersize=8)
+    
+    
+    #capillary_theor=[0.905,0.986,1.08,1.19,1.33,1.49]
+    #radiusses=[0.783,0.778,0.772,0.764,0.753,0.747]
+   
+    #style=["bo","rH","c<","y>","bs","g^"]
+    #color=["b","r","c","y","b","g"]
+    #style_diff=["b-","r:","c-.","y--","b^","g<"]
+    
+    
+    pylab.xticks(fontsize=20)
+    pylab.yticks(fontsize=20)
+    
+    pylab.ylabel(r'''$R_{axis}$''',fontsize=30)
+    pylab.xlabel(r'''$z$''',fontsize=30)
+    
+    #legtext = leg.get_texts() # all the text.Text instance in the legend
+    #for text in legtext:
+    #    text.set_fontsize(20) 
[…]

### 2011-09-23 Text/Submission/answers.tex
Almost finished a referee report.

@@
 \end{quotation}
+\textbf{Answer:} We added additional figures.
 
-\textbf{Answer:} We redrew all the figures. For the figure 8 the data was extracted from the
-figure and rescaled for comparison reasons.
+\textbf{Question:}
+\begin{quotation}
+In Figure 9, the variation of bubble radii in the range of capillary number 0.05<=Ca<=0.6 is
+still different from the results of Hazel and Heil and of Liu and Wang. The critical capillary
+number of Liu and Wang is close to 0.8 and the critical capillary number of Hazel and Heil is close
+to 0.04, but the value of the authors is 0.1. The authors should analyze the reason of the
+difference.
+\end{quotation}
+\textbf{Answer:} The reviewer is right that the critical capillary number of Liu and Wang is close
+to 0.8 as it can be seen from their digitized numbers. However, Liu and Wang state in their paper
+that the critical capillary number is 0.1. That means they used weaker criterion for the axial
+radius to be equal to the diagonal radius. Therefore we consider that the critical number for
+transition is between 0.04 and 0.1. The criterion is not well defined and depends on the flow
+conditions. For example, Hazel and Heil simulate the air finger propagation with the bubble length
+of more than 10 diameters of the channel. In our case, we simulate the bubble train where each
+bubble has a certain variation of the film thickness over the bubble length. We chose to measure the
+film thickness in the middle of the bubble. There is an assumption that the bubble length is long
+enough (3 diameters) for thickness to coincide with the air finger film thicknesses of the infinite
+length \cite{cerro-space}. However, in the case of smaller capillary the film thickness variation
+is quite large. Thus, more defined criteria should be given in papers in order to validate our
+results. Moreover, some formulations like the propagation of the semiinfinite air finger is not
+achievable due to computational resources constraints.
+
+\textbf{Question:}
+\begin{quotation}
+In Figure 10, there is a deviation of the relative velocity between the present and published data.
+Can the authors extend the range of capillary numbers as conducted by Liu and Wang?
+\end{quotation}
+\textbf{Answer:} Unfortunately that requires months with the uniform grids formulations of the
+lattice Boltzmann method or some grid refinement needed. We didn't have aim to go towards smaller
+capillary numbers as they nicely covered by original work of \citet{bretherton}.
+
+\textbf{Question:}
+\begin{quotation}
+Check some errors, e.g., on Page 7 Line 137 "0.05>=Ca<=6", on Page 17 Line 300, "82×82×1500 and
+82×82", and on Page 24 References 18 and 20.
+\end{quotation}
+\textbf{Answer:} Thank you for pointing that. All the errors are corrected. Reference 20 is
+available online and is not published in journals.
 
@@
 \begin{quotation}
-The main quantity that the paper focuses on is the thickness of the film between the bubbles and the
-wall. As shown by Figure 8 and as mentioned by the Authors in Section 4, the film thickness varies
-along the bubble. However, the Authors do not explain the choice of the location of the measurement
-of the film thickness. This choice is of primary importance when comparing results with those of
-the
-literature and in particular with works on finger propagation in which the film thickness is
-measured
-”at infinity”.
+Generally speaking, the motion of a drop in a channel is a function of channel shape, drop size,
+ratio between driving force and surface tension, density ratio, and viscosity ratio.  For the flow
+of a long gas bubble (Taylor-Bretherton) the density and viscosity ratios are nearly zero, making it
+possible to fit the film thickness as a function of a single dimensionless number Ca.  In the
+problem studied in this paper, the bubble is long, the density ratio is unity and the viscosity
+ratio is 0.1 (gas over liquid).  As such, while the film thickness in this system is also reduced to
+a function of the Capillary number, I do not think that the system considered by the authors is a
+Taylor-Bretherton system, as the density and viscosity ratios are different.
+
+The authors used the sensitivity of the film thickness on Re to justify that the sensitivity on
+density ratio is low (the first paragraph of Page 6).  However, I do not see the one-to-one
+connection between Re and the density ratio.  Are there any direct numerical/experimental evidence
+that density ratio has small effect on the film thickness, or evidence that Re is solely (or nearly
+solely) determined by the density ratio?
 \end{quotation}
 
-\textbf{Answer:} The explanations are added in lines 434-453.
+\textbf{Answer:}  The reviewer is right about not a full compliance with the classical
+Bretherton/Taylor problem. We added more explanations (lines 99-112). The system is governed with
+the non-dimensional numbers. Mainly, apart from geometrical dimensions the system can be fully
+descisribed by the following set of parameters: Re, Ca and Bo (Eo), where the Bond number Bo is
+taken into the account density ratio and is mainly involved in problems for a bubble change.
+However, the Bond number in simulations and real experimental setups is low. That's why we consider
+it as a non-leading non-dimensional parameter.
 
@@
 \begin{quotation}
-It may be appropriate to introduce Figure 8 on thickness variation before the main results on ﬁlm
-thickness.
+Aside from this critique, it looks like the numerical resolution needs some clarifications.  While
+some numbers and discussions were provided, they seem to be in conflict with each other.  For
+example, in the discussion given in Page 8 and top of Page 9, the authors stated that the minimum
+resolution needed to resolve the film is 600x600x9000, and, if a quarter of the domain is simulated,
+a resolution of 300x300x9000 is needed.  However, the domain sizes presented later are all smaller,
+e.g. 100x100x1500, 160x160x1500, 160x160x2400, 200x200x3000 (Page 12).  How were these grid
+resolutions selected, and what is the effect of grid resolution on the result?
 \end{quotation}
 
-\textbf{Answer:} We agree and moved the section before the results on the capillary number
-dependency.
+\textbf{Answer:} As an example, the corresponding grid $600\times600\times9000$ was taken to resolve
+$\delta=0.49 H_{\mathrm{eff}}$ for $Ca=0.03$ from reported simulations of \citet{heil-threedim}.
+However, luckily for us for current simulations the value of the film thickness was higher
+than $0.49 H_{\mathrm{eff}}$. That's why the computational resources demand is lower and we used
+smaller grids. The simulation for $600\times 600 \times 9000$ is too much computationally demanding.
+To give account for numbers one simulation with the grid $200\times200\times 3000$ is obtained after
+$24$ days on $32$ CPUs. The values of these grids were chosen to resolve the film thickness for the
+smallest capillary number (film thickness is twice bigger than the interface thickness). We've done
+inextensive grid resolution study (see Fig. 9) to support our results.  We clarified
+this issue on lines 168-171.
 
@@
 \begin{quotation}
-The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
-Reynolds
-number effects on the film thickness for capillary numbers larger than 0.003” is in contradiction
-with
-the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
-relatively
-wide range of Reynolds numbers.”
+A couple of minor issues are listed below:
+Page 4, figure 1 - I suppose the vertical axis is $r_d$ and $r_h$ normalized by $H_eff$?  The axis
+label perhaps should be clarified.
+Page 5, line 18 - missing "from" between "capillary number" and "simulations".
+Page 7, line 19 - "0.05 >= Ca <= 6" should be "0.05 <= Ca <= 6".
+Page 10 - On the lattice Boltzmann free energy model, what are the boundary conditions for f and g
+populations?  Were they simple bounce-back?  Is contact angle implemented in the method?
+Page 12, line 13 - "k = A = 0.004, 0.04".  I do not understand this notation - does this mean k is
+always equal to A and their values should be between 0.004 and 0.04?
+Page 17, line 17-19 - The text says that Figure 9 contains data with different initializations and
+different grids.  However, the distinction between data sets cannot be viewed from the figure. 
+Perhaps different symbols should be used so that the readers can tell one data set from another?
+Page 17, line 40 - "differencies" should be "differences".
+Page 19, line 25 - ". the film thickness as twice the interface thickness".  Please see the previous
+comment on grid resolution.
+Page 20, line 19 - Again, the notation used for the range of A and k is rather unusual.
+Page 29, line 46 - ". relaxed after to the axis radius" - perhaps a typo in the sentence?
 \end{quotation}
 
-\textbf{Answer:} We elaborate the Reynolds number influence in lines 34-43.
+\textbf{Answer:} Thank you for improving the article. All the corrections are taken into the
+account. We also added more explanations to the text to address the following questions:
+\begin{itemize}
+ \item For $f$ and $g$ populations the wall was simulated using the simple bounce-back rule. We
+showed before that results do not depend on the wettability of the wall
+\cite{kuzmin-binary2d}. Thus, the walls were chosen with neutral wettability. 
+ \item We chose $k$ and $A$ equal to each other as this choice conserves the film thickness ($5$
+grid spacings). $k=A=0.04$ is taken for most simulations. $k=A=0.004$ is taken for simulation with 
+$Ca=6$. Usually for stability issues the choice of $k=A$ is a good one. The values of $k$ and $A$
+can vary in different limits. However, the stability is drastically reduced for $k=A=0.25$ or if
+$k=A \ll 1$. The corresponding text is added (lines 219-221).
+\item Figure 9 shows the results grid independency. It's exactly the reason why we didn't
+distinguish different initializations and grid numbers as proper designed simulations should not
+depend on it. Corresponding text is added (lines 316-317).
+\end{itemize}
 
@@
 \begin{quotation}
-Although inertia has been shown to have a minor effect on ﬁlm thickness, inertia does influence
-some flow variables in the Bretherton’s problem (see Ref. 3 for instance). To check whether inertia
-is important or not, and to discuss the assumption of uniform density, the Authors should provide
-the reader with the Reynolds numbers associated with each simulation. This is also required for the
-consistency of Figure 7 in which results are compared with other works of the literature.
+Finally, I have a question on how the initial distributions of f and g were determined?  Did the
+simulation start with uniform distributions of f and g populations and they "phase separate" to form
+the bubble?
 \end{quotation}
 
-\textbf{Answer:} Thank you for pointing this out. We included the discussions of the Reynolds
-effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
-the capillary range regime and the grid depedency study.
+\textbf{Answer:} Initial distributions $f$ and $g$ were taken through equilibrium distribution
+functions, which were constructed from initial phase and density fields. Density was taken uniform
+with the value $1$ across the domain. Bubble was initialized as the rectangular parallelepiped in
+the middle domain. After that the system evolves with time and interplay between surface tension
+and viscous forces form the bubble. It was shown \cite{kuzmin-binary2d} that the final steady-state
+result does not depend on the initialization.
+
+Reviewer III:
+\textbf{Question:}
+\begin{quotation}
+Check some spelling errors, also in the abstract;
+\end{quotation}
+
+\textbf{Answer:} We thoroughly corrected all found mistakes.
 
@@
 \begin{quotation}
-In this work, the diphasic flow is driven by a body force. The body force is therefore the control
-parameter in the simulations. The Authors should give more detail about this body force. Indeed,
-although the estimation of the body force from the pressure gradient is well described, it is not
-clear
-where this body force appears in the governing equations (Equations 2, 3 and 4). Does this body
-force correspond to the variable $F_i$ in Equation 2? Should not this force or its equivalent appear
-in
-the Navier-Stokes equations (Equation 4)?
+Improve Figure quality: Figs. 4 and above all 5 are hard to read and understand. The ticks should
+be the same height, those in Fig. 5 are hardly visible.
 \end{quotation}
 
-\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
-of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
-inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
-how it
-mimics the force behavior.
+\textbf{Answer:} We constructed figures with the same size ticks, title and labels. For better
+readability we rearranged figures to be of larger size.
 
@@
 \begin{quotation}
-I would suggest to calculate the Bond number to check if the shapes of the bubbles and the main
-results of the paper are not inﬂuenced.
+The bibliography should be improved: when introducing LB, I think that more hints should be given
+to the applications, for instance to the case of magnetically-driven vaporization in ferrofluids
+studies or regarding multiphase approach based on pseudopotentials.
 \end{quotation}
 
-\textbf{Answer:} 
-The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
-current simulation equals to $7$, which is less than the number for which the shape of the bubble is
-affected according to \cite{zheng-large-ratio}.
+\textbf{Answer:} We expanded the introduction for industry applications and discussion of
+multiphase models (lines 117-123).
+\textbf{Question:}
+\begin{quotation}
+Did the Authors perform any comparison with pseudopotential multiphase approaches? I am
+especially referring to recent advances of pseudopotentials, such as the case of doubly-attractive
+potentials and to recent papers regarding LB multiphase simulation across scales.
+\end{quotation}
+
+\textbf{Answer:} Unfortunately, to our best knowledge there is only one study of the pseudopotential
+model towards the Bretherton-Taylor phenomenon. The study is in 2D and cannot be applicable to this
+case. However we outlined the approaches for pseudopotential models (lines 117-123).
 
 \textbf{Question:}
-\begin{quotation}
-The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
-estimations of the body force to apply and of the film thickness to choose grid resolution. The
-presentation
-of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
-again described in a different way in Section 5-4. Similar equations for the calculation of the
-pressure
-gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
-into one unless they explain their differences. In fact, it would probably be straightforward not to
-detail the calculation of estimated variables since they are intermediary variables which allow the
-design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
-say they control and impose directly the body force without describing all the estimation procedure.
-\end{quotation}
-
-\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
-about initialization and this was taken into account by reducing repetitive text and
-introducing step-by-step explanations, see Section about initialization (lines $259-305$). We also
-added
-the
-recommendations for controlling parameters in the same section.  The
-dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
-Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
-the lattice Boltzmann method are refered to the non-dimensionalization procedure. Therefore, we
-decided to
-keep suggestions for controlling and initialization of simulations as informative guidelines for the
-LBM community.
-
-\textbf{Question:}
-\begin{quotation}
- Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
-obtained
-under the assumption of an axisymmetric Poiseuille flow. However, among the ”Key words”, one can
-read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
-that are done in this work. The Authors should discuss this choice of estimating the pressure
-gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
-in Equation 5).
-\end{quotation}
-
-\textbf{Answer:} The factor $1/8$ in Eq. (5) is taken for the maximum velocity for the Poiseille
-flow between
-plates. Our assumption is that the bubble travels with the constant velocity which equals to the
-maximum velocity of the Poiseuille profile between plates. Factor $1/12$ comes from the averaging
-of the flow rate. However, bubble moves faster than the liquid, i.e. the average liquid velocity.
-That is why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
-well.
-
-\textbf{Question:}
-\begin{quotation}
-The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
-journal of applied science. The Authors would improve signiﬁcantly the manuscript by defining more
-rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
-lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
-A suggestion is to emphasize within an array the links between physical parameters and the model
-parameters. The variables $i$, $f_i^{eq}$ , $g_i$ should be deﬁned and the Equations for $f$ and $g$
-in Equation
-3 should be described (which equation corresponds to continuity equation...). Equation 2 could
-perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
-physical variable f of Equation 3. The notation $i = 1 \div 8$ is unusual, it would be clearer to
-use a
-classical notation such as $1 \leq i \leq 8$. The sound speed $c_s^2$ should be deﬁned just after
-its
-introduction in line 107.
-\end{quotation}
-
-\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the
-physical parameters with the lattice Botlzmann parameters (lines $149-160$). As well more
-explanations are
-put towards the
-interface surface tension description (lines $136-139$). The reference to the
-continuity and the Navier-Stokes equations are added (lines $140-143$). The equation (2) is not
-rewritten in terms of
-generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
-governing equations. The notations have been corrected.
-
-\textbf{Question:}
-\begin{quotation}
-The name of the Section ”Lattice Boltzmann benchmark” lets the reader think that this Section
-would describe the numerical procedure (assumptions, choice of the geometry and the grid, boundary
-conditions, control parameters, initialization of the scheme). This is not the case because this
-Section
-develops the challenges and the difficulties of using the Lattice Boltzmann method to bubble ﬂows.
-This Section should be renamed or reorganized.
-\end{quotation}
-
-\textbf{Answer:} We renamed the section to ``Approach to benchmark''.
-
-\textbf{Question:}
-\begin{quotation}
-The Authors should make a distinction between variables in lattice units and dimensional or non
-dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y - 2$ where $N_y$ is
-the
-number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
-is in
-contradiction with the use of $H_{eff}$ in Equation 5. The Authors should define each variable
-without
-any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
-improvements
-would fit well in the Subsection named ”The nondimensionalization and initialization procedure”
-which does not in fact introduce much nondimensionalisation procedure.
-\end{quotation}
-
-\textbf{Answer:} We added a decription paragraph about the approach to units (lines 149-160). In
-fact, all equations are the same in the
-lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
-preferred units as soon as the capillary number is the same in the LBM and the physical world.
-
-\textbf{Question:}
-\begin{quotation}
-Some notations are unclear. For instance, the film thickness is denoted as $h_{\infty}$ in Equation
-7 whereas
-it is denoted $\delta$ in most of the paper. The notations w or $\delta_{sim}$ are also found before
-line 381 or in
-Table 3 respectively. The thickness of the interface is denoted $\xi$ in Figure 1 but the same
-variable
-is
-denoted $5\xi$ in Table 1. The speed of the bubble is denoted $U_{\mathrm{bubble}}$ in some places
-and $U_b$ in others.
-The variable $Uin$ Equation 5 does not seem to be deﬁned.
-\end{quotation}
-
-\textbf{Answer:} We thoroughly corrected all the inconsistencies.
-
-\textbf{Question:}
-\begin{quotation}
-Three figures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
-unusual
-for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
-which are not essential.
-\end{quotation}
-
-\textbf{Answer:} We eliminated unnecessary pictures and references to them.
-
-\textbf{Question:} 
-\begin{quotation}
-In Figure 7, the points of data reported from the numerical work of Ref. 3 exhibit a surprisingly
-noisy trend that corresponds more to an experimental study. The Authors should explain how they
-obtain this data and plot them.
-\end{quotation}
-
-\textbf{Answer:} We added a description to the figure caption explaining data acquisition.
-
-\textbf{Question:} 
-\begin{quotation}
-The plots of Figure 8 are of less quality than the other figures of the paper.
-\end{quotation}
-
-\textbf{Answer:} Unfortunately, it was not possible to extract data as a picture from the paper by
-\citet{sehgal-microchannel}. Thus, the curves were extracted from the graphs and represented in a
-new figure scaled to the present simulations for comparison reasons.
-
-\textbf{Question:}
-\begin{quotation}
-The work of Yang seems to be close from the present study since it involves the
-solution of Bretherton
-flow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
-differences between their work and this paper.
-\end{quotation}
-
-\textbf{Answer:} We outlined the differences in the introduction section (lines $92-100$).
-
-\textbf{Question:} It would be interesting if the Authors could mention a study in which a level-set
-method is used in
-order to complete the bibliography.
 
@@
 \bibliographystyle{unsrtnat}
-\bibliography{answers}
+\bibliography{paper}
 

### 2011-09-23 Text/Submission/paper.tex
Almost finished a referee report.

@@
 diameters for the capillary
-numbers $Ca\leq\widehat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
+numbers $Ca\leq\widehat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. $r_h$ and $r_d$ are scaled on
+the half side length $H_{\mathrm{eff}}/2$. \label{fig:heil:three:dim}}
 \end{figure}
@@
 makes it possible to collapse results of radii dependency on the capillary number
-simulations of rectangular channels with different aspect ratios on a single curve.
+for rectangular channels with different aspect ratios on a single curve.
 It was also found that for microchannels with a certain aspect ratio $\alpha=\frac{a}{b}\geq
@@
 microscopic level allows to restore multiphase macroscopic equations \cite{swift,
-Shan-chen:extended}.
+Shan-chen:extended}. {\color{red} Nowadays the method is finding its way to more practical
+applications, for example magnetically driven droplet break-up and vaporization
+\cite{falcucci-vapour}, mixing in cross-channel flow \cite{jos-mixing}, etc. }
 
@@
 simulates two liquids with the assumption of uniform overall
-density. %As it was indicated by \citet{gupta-review} a flow in microchannels is a complex
+density. 
+ %As it was indicated by \citet{gupta-review} a flow in microchannels is a complex
 %interplay of viscous, gravitational, inertial and surface forces. 
@@
 simulations.
+
+{\color{red} Note that other multiphase models can be used which can account for different
+densities, as the pseudopotential based approach \cite{Sbragaglia-Succi} or color model
+\cite{swift}. In the case of the pseudopotential based approach, one needs to use the
+binary-liquid formulation as to avoid possible condensation/evaporation \cite{pooley-mrt-filling}
+and to perform Laplace law simulations to obtain the surface tension as it is not given explicitly.
+For the case of the color model, one needs to use more complicated schemes to reconstruct the
+surface. All other finding of the paper can be applicable for all the abovementioned models.}
 % THIS HAS ALREADY BEEN STATED BEFORE
@@
 at the distance of $5.5 H_{\mathrm{eff}}$ and indicated it to be sufficient for $Ca\leq 10$. In
-this work, the following relation holds for all conducted simulations: $0.05 \geq Ca \leq 6$.
+this work, the following relation holds for all conducted simulations: $0.05 \leq Ca \leq 6$.
 Thus, the
@@
 large
-computational resources are required to perform the simulations.
+computational resources are required to perform the simulations. {\color{red} However, as it will
+be shown later the film thickness in simulations is larger than the example number ($0.49
+H_{\mathrm{eff}}$ for $Ca\approx 0.03$). Therefore, actual grids are smaller (indicated in
+Section \ref{sec:results}) than the computationally demanding grid $600 \times 600 \times 9000$.}
 
@@
 phase gradient exists. The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
-A}{9}}$.
+A}{9}}$. {\color{red} The walls were simulated using the bounce-back rule \cite{yu}. The walls were
+taken with neutral wettability as results do not depend on the wettability \cite{kuzmin-binary2d}.}
 
@@
 $160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. Body forces were varied as
-$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$ lattice units. The binary-liquid model parameters
+$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$ lattice units.{\color{red} The binary-liquid model
+parameters
 were kept in
-the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
+the stable region $k=A=0.04$ for most simulations and $k=A=0.004$ for simulations with $Ca=6.70$. 
+Note that the choice of $k=A$ guarantee that the film thickness is $5$ grid spacings, which is a
+good choice for stability issues. } The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
 and $\tau_{\mathrm{gas}}=0.7$ giving a gas-over-liquid viscosity ratio of $10$. The relaxation
@@
 \end{table}
-\begin{figure}[ht]
-\includegraphics[width=0.5\textwidth]{phase_crossection_ca5.eps}
-\includegraphics[width=0.5\textwidth]{phase_crossection_ca13.eps}\\
+\begin{figure}[htb!]
+\includegraphics[width=0.8\textwidth]{phase_crossection_ca5.eps}\\
+\includegraphics[width=0.8\textwidth]{phase_crossection_ca13.eps}\\
 \caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
-(left) and $Ca=1.13$ (right). Crosssections are rescaled on the half channel height. Other
+(top) and $Ca=1.13$ (bottom). Crosssections are rescaled on the half channel height. Other
 parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
@@
 the jump.  
-\begin{figure}[ht]
-\includegraphics[width=0.8\textwidth]{bubble_rad_axis.eps}\\
-\includegraphics[width=0.8\textwidth]{bubble_rad_diag.eps}\\
-\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
+\begin{figure}[htb!]
+\includegraphics[width=0.45\textwidth]{bubble_rad_axis.eps}\hfill
+\includegraphics[width=0.45\textwidth]{bubble_wang_axis.eps}\\
+\includegraphics[width=0.45\textwidth]{bubble_rad_diag.eps}\hfill
+\includegraphics[width=0.45\textwidth]{bubble_wang_diag.eps}\\
+\caption{Simulations radii variations (left)  along the bubble for different capillary numbers in
+the plane $x=0$.
 One can see that the diagonal radius shows the jump near the bubble rear cap. This agrees with the
-VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
+VOF simulations of \citet{wang-non-circular} (right).Coordinate $z$ increases in the streamwise
+direction. The discretization of \citet{wang-non-circular} data was performed with ''Engauge
+Digitizer``.
 \label{fig:bubble:variation:capillaries}}
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{bubble_ca_large.eps}
-\caption{The bubble shape for the capillary number $Ca=6.43$ is presented. $z$ increases in the
+\includegraphics[width=0.9\textwidth]{bubble_ca_large.eps}\\
+\includegraphics[width=0.9\textwidth]{bubble_heil_ca10.eps}
+\caption{The bubble shape for the capillary number $Ca=6.43$ is presented (top). $z$ increases in
+the
 streamwise direction. One can see that near the front tip (large $z$) the diagonal radius exhibits
-a jump which is relaxed after to the axis radius. Even this resembles
-the shape of the curves obtained by \citet{heil-threedim}, but the
+a jump which is relaxed to the axis radius (small $z$).{\color{red} Even this resembles
+the shape of the curves obtained by \citet{heil-threedim} (bottom, opposite streamwise direction),
+but the
 difference between radii is around $1$ percent and can be attributed to
-the linear interpolation used in the calculation of the interface.\label{fig:bubble:ca:large}}
+the linear interpolation used in the calculation of the interface.}\label{fig:bubble:ca:large}}
 \end{figure}
@@
 initialization, and different grids: $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
-\times 82 \time 2400$. All our results are consistent. Therefore one can conclude that for the
+\times 82 \times 2400$. All our results are consistent {\color{red} and show the grid and
+initialization independency}. Therefore one can conclude that for the
 capillary number $Ca\geq
@@
 \subsection{Relative velocity}
-One of the differencies between the flow in tubes and square capillaries is that the flow has
+One of the differences between the flow in tubes and square capillaries is that the flow has
 different streamline patterns. In square microchannels, the main mass flow occurs through corners
@@
 number of nodes to resolve the film thickness. For binary liquid parameters used in simulations,
-$A=0.004,0.04$ and $k=0.004,0.04$ the
+$k=A=0.04$ and $k=A=0.004$ the
 interface is spread over approximately $5$ lattice units. To obtain grid independent results one

### 2011-09-23 Engauge/draw.py

@@ -5,25 +5,32 @@ import pylab
 def paint_curves():
 
     wang=numpy.genfromtxt("wang_axial.csv",delimiter=',',dtype=float)
-    print wang
+    #print wang
     fig=pylab.figure()
     wang2=numpy.array(wang)
     wang2[:,0]=0.0
-    print wang
+    #print wang
     ind=numpy.where(wang2>1.0)
-    print ind
+    #print ind
     #ind[1]=ind[1]+1
     wang[ind]=-1.0
     max_length=max(wang[:,0])
     wang[:,0]=max_length-wang[:,0]
+    print numpy.where(wang[:,1]>0)
+    shift1=wang[max(numpy.where(wang[:,1]>0.1)[0]),0]
+    shift2=wang[max(numpy.where(wang[:,2]>0.1)[0]),0]
+    shift3=wang[max(numpy.where(wang[:,3]>0.1)[0]),0]
+    
+    #print shift1, shift2,shift3
+    
     #pylab.plot(wang[:,0],wang[:,3][::-1],"kv",markersize=8) #,markerfacecolor="white")
-    pylab.plot(wang[:,0],wang[:,3],"kv",markersize=8) #,markerfacecolor="white")
-    pylab.plot(wang[:,0],wang[:,1],"ko",markersize=8) #,markerfacecolor="white")
-    pylab.plot(wang[:,0],wang[:,2],"ks",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0]-shift3,wang[:,3],"kv",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0]-shift1,wang[:,1],"ko",markersize=8) #,markerfacecolor="white")
+    pylab.plot(wang[:,0]-shift2,wang[:,2],"ks",markersize=8) #,markerfacecolor="white")
 
-    #leg=pylab.legend([r'''$Ca=1.0$''',r'''$Ca=0.3$''',r'''$Ca=0.08$'''],loc=3,fancybox=True)
-   
-    pylab.ylim(ymin=0.0)
+    leg=pylab.legend([r'''$Ca=1.0$''',r'''$Ca=0.3$''',r'''$Ca=0.08$'''],loc=1,fancybox=True)
+    pylab.ylim(ymin=0.1,ymax=1.1)
+    pylab.xlim(xmin=0.0)
     
     #for counter,file in enumerate(files):
         
[…]

### 2011-09-24 Text/Submission/answers.tex
Finished draft answers for the referee.

@@
 \begin{document}
-It is our pleasure to address the reviewers' comments. Please consider our answers below.
+It is our pleasure to address the reviewers' comments. Please consider our answers below. All corrections in the manuscript are marked by red.
 \section{Reviewer I}
+Question and Answers:
 
-\textbf{Question:}
-\begin{quotation}
-When Re is very small, the authors neglect the density ratio of gas and liquid phase. I think more
+\textbf{Question:} When Re is very small, the authors neglect the density ratio of gas and liquid phase. I think more
 reasons are still needed to explain the uniform density, especially on the interface. One of the
@@
 gradient on the interface.
-\end{quotation}
 
-\textbf{Answer:} 
-We added additional text explaining the choice of the binary-liquid model towards the simulation
+\textbf{Answer:}  We added additional text explaining the choice of the binary-liquid model towards the simulation
 of the Bretheron-Taylor phenomenon (see lines 99-112). The Galilean invariance is usually fixed by
@@
 
-\textbf{Question:}
-\begin{quotation}
-In Figures 7 and 8, it's better to give the reported curves by Liu and Wang and Hazel and Heil, so
+\textbf{Question:} In Figures 7 and 8, it's better to give the reported curves by Liu and Wang and Hazel and Heil, so
 that it's easy to find the agreement is well or not.
-\end{quotation}
+
 \textbf{Answer:} We added additional figures.
 
-\textbf{Question:}
-\begin{quotation}
-In Figure 9, the variation of bubble radii in the range of capillary number 0.05<=Ca<=0.6 is
+\textbf{Question:} In Figure 9, the variation of bubble radii in the range of capillary number 0.05<=Ca<=0.6 is
 still different from the results of Hazel and Heil and of Liu and Wang. The critical capillary
@@
 difference.
-\end{quotation}
+
 \textbf{Answer:} The reviewer is right that the critical capillary number of Liu and Wang is close
@@
 
-\textbf{Question:}
-\begin{quotation}
-In Figure 10, there is a deviation of the relative velocity between the present and published data.
+\textbf{Question:} In Figure 10, there is a deviation of the relative velocity between the present and published data.
 Can the authors extend the range of capillary numbers as conducted by Liu and Wang?
-\end{quotation}
+
 \textbf{Answer:} Unfortunately that requires months with the uniform grids formulations of the
@@
 
-\textbf{Question:}
-\begin{quotation}
-Check some errors, e.g., on Page 7 Line 137 "0.05>=Ca<=6", on Page 17 Line 300, "82×82×1500 and
+\textbf{Question:} Check some errors, e.g., on Page 7 Line 137 "0.05>=Ca<=6", on Page 17 Line 300, "82×82×1500 and
 82×82", and on Page 24 References 18 and 20.
-\end{quotation}
+
 \textbf{Answer:} Thank you for pointing that. All the errors are corrected. Reference 20 is
@@
 \section{Reviewer II}
+Questions and Answers:
 
-\textbf{Question:}
-\begin{quotation}
-Generally speaking, the motion of a drop in a channel is a function of channel shape, drop size,
+\textbf{Question:} Generally speaking, the motion of a drop in a channel is a function of channel shape, drop size,
 ratio between driving force and surface tension, density ratio, and viscosity ratio.  For the flow
@@
 a function of the Capillary number, I do not think that the system considered by the authors is a
-Taylor-Bretherton system, as the density and viscosity ratios are different.
-
-The authors used the sensitivity of the film thickness on Re to justify that the sensitivity on
+Taylor-Bretherton system, as the density and viscosity ratios are different. The authors used the sensitivity of the film thickness on Re to justify that the sensitivity on
 density ratio is low (the first paragraph of Page 6).  However, I do not see the one-to-one
@@
 solely) determined by the density ratio?
-\end{quotation}
 
@@
 
-\textbf{Question:}
-\begin{quotation}
-Aside from this critique, it looks like the numerical resolution needs some clarifications.  While
+\textbf{Question:} Aside from this critique, it looks like the numerical resolution needs some clarifications.  While
 some numbers and discussions were provided, they seem to be in conflict with each other.  For
@@
 resolutions selected, and what is the effect of grid resolution on the result?
-\end{quotation}
 
@@
 
-\textbf{Question:}
-\begin{quotation}
-A couple of minor issues are listed below:
+\textbf{Question:} A couple of minor issues are listed below:
 Page 4, figure 1 - I suppose the vertical axis is $r_d$ and $r_h$ normalized by $H_eff$?  The axis
-label perhaps should be clarified.
-Page 5, line 18 - missing "from" between "capillary number" and "simulations".
-Page 7, line 19 - "0.05 >= Ca <= 6" should be "0.05 <= Ca <= 6".
+label perhaps should be clarified.\\
+Page 5, line 18 - missing "from" between "capillary number" and "simulations".\\
+Page 7, line 19 - "0.05 >= Ca <= 6" should be "0.05 <= Ca <= 6".\\
 Page 10 - On the lattice Boltzmann free energy model, what are the boundary conditions for f and g
-populations?  Were they simple bounce-back?  Is contact angle implemented in the method?
+populations?  Were they simple bounce-back?  Is contact angle implemented in the method?\\
 Page 12, line 13 - "k = A = 0.004, 0.04".  I do not understand this notation - does this mean k is
-always equal to A and their values should be between 0.004 and 0.04?
+always equal to A and their values should be between 0.004 and 0.04?\\
 Page 17, line 17-19 - The text says that Figure 9 contains data with different initializations and
 different grids.  However, the distinction between data sets cannot be viewed from the figure. 
-Perhaps different symbols should be used so that the readers can tell one data set from another?
-Page 17, line 40 - "differencies" should be "differences".
+Perhaps different symbols should be used so that the readers can tell one data set from another?\\
+Page 17, line 40 - "differencies" should be "differences".\\
 Page 19, line 25 - ". the film thickness as twice the interface thickness".  Please see the previous
-comment on grid resolution.
-Page 20, line 19 - Again, the notation used for the range of A and k is rather unusual.
+comment on grid resolution.\\
+Page 20, line 19 - Again, the notation used for the range of A and k is rather unusual.\\
 Page 29, line 46 - ". relaxed after to the axis radius" - perhaps a typo in the sentence?
-\end{quotation}
 
@@
 
-\textbf{Question:}
-\begin{quotation}
-Finally, I have a question on how the initial distributions of f and g were determined?  Did the
+\textbf{Question:} Finally, I have a question on how the initial distributions of f and g were determined?  Did the
 simulation start with uniform distributions of f and g populations and they "phase separate" to form
 the bubble?
-\end{quotation}
 
@@
 
-Reviewer III:
-\textbf{Question:}
-\begin{quotation}
-Check some spelling errors, also in the abstract;
-\end{quotation}
+\section{Reviewer III}:
+
+Questions and Answers:
+
+\textbf{Question:} Check some spelling errors, also in the abstract;
 
@@
 
-\textbf{Question:} 
-\begin{quotation}
-Improve Figure quality: Figs. 4 and above all 5 are hard to read and understand. The ticks should
+\textbf{Question:}  Improve Figure quality: Figs. 4 and above all 5 are hard to read and understand. The ticks should
 be the same height, those in Fig. 5 are hardly visible.
-\end{quotation}
 
@@
 
-\textbf{Question:} 
-\begin{quotation}
-The bibliography should be improved: when introducing LB, I think that more hints should be given
+\textbf{Question:}  The bibliography should be improved: when introducing LB, I think that more hints should be given
 to the applications, for instance to the case of magnetically-driven vaporization in ferrofluids
 studies or regarding multiphase approach based on pseudopotentials.
-\end{quotation}
 
@@
 multiphase models (lines 117-123).
-\textbf{Question:}
-\begin{quotation}
-Did the Authors perform any comparison with pseudopotential multiphase approaches? I am
+
+\textbf{Question:} Did the Authors perform any comparison with pseudopotential multiphase approaches? I am
 especially referring to recent advances of pseudopotentials, such as the case of doubly-attractive
 potentials and to recent papers regarding LB multiphase simulation across scales.
-\end{quotation}
 
@@
 model towards the Bretherton-Taylor phenomenon. The study is in 2D and cannot be applicable to this
-case. However we outlined the approaches for pseudopotential models (lines 117-123).
+case. However, we outlined the approaches for pseudopotential models (lines 117-123) and all the paper findings are applicable for pseudopotential models as well.
 
-\textbf{Question:}
+\textbf{Question:} Could the Authors give more hints on the nature of the walls? Are they hydrophilic or hydrophobic and how do the Authors implement such a nature?
 
-\textbf{Answer:} We added references to the introduction section [17-19].
-
-\textbf{Questions:} 
-\begin{quotation}
-A secondary quantity of interest is the bubble speed or in a non-dimensional
-form the capillary
-number. This quantity is measured at the center of the bubbles. The Authors should describe how
-this quantity as well as the center of the bubble are measured.
-\end{quotation}
-
-\textbf{Answer:} Thank you for pointing that out. In the classical Bretherton problem
-$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
-results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
-results are the same.
-
-\textbf{Question:} 
-\begin{quotation}
- If possible, the Authors should present results showing that steady state is reached. It would
-also
-be interesting if the Authors show some patterns of the flow by plotting for instance the vector
-ﬁeld
-of velocity around the bubbles.
-\end{quotation}
-
-\textbf{Answer:} We added a steady-state study (lines $334-341$). We added an additional
-paragraph which shows the change of flow pattern for the
-streamlines as it is indicated in literature (lines $463-471$, Figure 8).
-
-\textbf{Question:}
-\begin{quotation}
-In the Abstract, the Authors should give the range of real capillary numbers they measured instead
-of the targeted values ($Ca_{lit}$).
-\end{quotation}
-
-\textbf{Answer:} We changed it to the measured capillary numbers.
-
-\textbf{Question:} 
-\begin{quotation}
-The Authors should discuss more their results (Sections 5-4, 5-5, 5-6). Did the Authors measure the
-pressure drops at the front and rear interfaces of the bubbles?
-\end{quotation}
-
-\textbf{Answer:} We added additional information about results to the corresponding sections. The
-pressure drop is not measured as they are influenced largely by Reynolds number according to
-\citet{heil-bretherton}.
-
-\textbf{Question:}
-\begin{quotation}
-The Authors mention the shooting method. It is not clear if this method is really used or just
-mentioned as a part of the discussion.
-\end{quotation}
-
-\textbf{Answer:} This is a part of discussion. The corresponding
-sentence is added (lines $459-461$).
-
-\textbf{Question:}
-\begin{quotation}
-It seems more appropriate to introduce the variables of the Bretherton’s problem as well as to plot
-Figure 1 elsewhere than in the Section ”Results”.
-\end{quotation}
-
-\textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
-separated it from the numerical benchmark layout.
-
-\textbf{Question:}
-\begin{quotation}
-Minor corrections...
-\end{quotation}
-
-\textbf{Answer:} All the corrections are taken into the account.
-
+\textbf{Answer:} The walls are neutral. It was shown previously \cite{kuzmin-binary2d} that for the Bretherton phenomenon when the film thickness is resolved the nature of the walls is not involved.
 \bibliographystyle{unsrtnat}

### 2011-09-24 Text/Submission/paper.tex
Finished draft answers for the referee.

@@
 
-{\color{red} Note that other multiphase models can be used which can account for different
+{\color{red} Note that other multiphase lattice Boltzmann models can be used which can account for different
 densities, as the pseudopotential based approach \cite{Sbragaglia-Succi} or color model
@@
 For the case of the color model, one needs to use more complicated schemes to reconstruct the
-surface. All other finding of the paper can be applicable for all the abovementioned models.}
+surface. All other findings of the manuscript are applicable for all above mentioned models.}
 % THIS HAS ALREADY BEEN STATED BEFORE

### 2011-09-25 Text/paper.tex
Fix typos.

@@
 thin liquid films between bubbles and channel walls. The bubble shape and flow patterns are
-complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently, we
+complicated functions of the capillary number $Ca$ and Reynolds number $Re$. Recently, we
 investigated
@@
 periodic boundary conditions are applied in the streamwise direction. The results show that the
-binary liquid model is able to correctly capture a number of phenomena occuring in three-dimensional
+binary liquid model is able to correctly capture a number of phenomena occurring in three-dimensional
 capillaries, such as the existence of a vortex in front of the bubble and the way bubble radii
@@
 \cite{bretherton,heil-bretherton}. 
-Predicting flow patterns and associated mass transfer for Bretheron-type flows
+Predicting flow patterns and associated mass transfer for Bretherton-type flows
 is of significant interest for chemical industry as it is widely used in chemical monolith
@@
 ($7$ percent deviation from the film thickness measured at $Re=0$), the change in Reynolds number
-significantly influences  the pressure distrubution and the flow field near the front bubble cap.
+significantly influences  the pressure distribution and the flow field near the front bubble cap.
 \citet{cerro-bubble-train} showed that the mass transfer is strongly affected by flow direction for
@@
 capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
-shaped mirochannels. They also presented an experimental correlation for  bubble
+shaped microchannels. They also presented an experimental correlation for  bubble
 radii
@@
 different from the
-diagonal radius and the bubble having a non-circular shape in the channel crossection. In this case,
+diagonal radius and the bubble having a non-circular shape in the channel cross section. In this case,
 the bubble shape mimics the shape of a rectangle and looks like a rectangle with rounded corners.
@@
 address simulations with upward and downward flows \cite{cerro-bubble-train} which would require a
-simultaneous imposement of pressure boundary conditions and the body force. 
+simultaneous imposition of pressure boundary conditions and the body force. 
 %The pressure boundary
@@
 \includegraphics*[bb=153 610 410 717,width=\textwidth]{Figures/benchmark_classical.eps} 
-\caption{The classical benchmark layout describes a semiinfinite gas bubble
+\caption{The classical benchmark layout describes a semi-infinite gas bubble
 propagation through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
@@
 effects of bubbles motion on each other is minimized by choosing lengths of a bubble and a
-channel large enough in comparision with the channel height. However, more thorough studies are
+channel large enough in comparison with the channel height. However, more thorough studies are
 needed in order to understand
@@
 \cdot 10^{-6}$. The simulation was performed for $250000$ iterations with the step of
-$10000$ iterations. The results for time iterations $140000-250000$ are summarised in Table
+$10000$ iterations. The results for time iterations $140000-250000$ are summarized in Table
 \ref{table:steady:state}. While the
@@
 can see in Fig. \ref{fig:velocity:contour} the contour plot and the
-corressponing streamwise velocity
-component. The crossection is a plane $x=0$, where $x$ points
+corresponding streamwise velocity
+component. The cross section is a plane $x=0$, where $x$ points
 towards the reader, $z$ is the streamwise direction. As fluxes inside the bubble have
@@
 \includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca13.eps}\\
-\caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
-(left) and $Ca=1.13$ (right). Crosssections are rescaled on the half channel height. Other
+\caption{Cross sections of the phase field in the middle of the bubble for $Ca=0.053$
+(left) and $Ca=1.13$ (right). Cross sections are rescaled on the half channel height. Other
 parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
@@
 \end{equation}
-where $Q$ is the area occupied by air in the crossection. \citet{heil-threedim} indicated a simple
+where $Q$ is the area occupied by air in the cross section. \citet{heil-threedim} indicated a simple
 physical interpretation of this phenomena:  providing that air is injected with the same
@@
 This work presents results of binary-liquid simulations for three-dimensional channels with square
-cross sections with the lattice Boltzman method. By resolving properly the film thickness as twice the
+cross sections with the lattice Boltzmann method. By resolving properly the film thickness as twice the
 interface thickness \cite{kuzmin-binary2d} the results were shown to be consistent with those
@@
 $\tau_{\mathrm{liq},\mathrm{gas}}\approx 0.5$ \cite{kuzmin-trt-stability}. On the other hand the
-accuracy of the lattice Boltzmann liquid model detiorates as
+accuracy of the lattice Boltzmann liquid model deteriorates as
 $\Bigl(\tau_{\mathrm{liq},\mathrm{gas}}-\frac{1}{2}\Bigr)^2$ \cite{ginzburg-trt-simple-hydro}. Thus,
@@
 for tangential velocities $U_{\tau}$, while velocities perpendicular  to the mirror boundary $U_n$ have opposite signs
-to that of the velocity of the fluid node.  This can be summarised as:
+to that of the velocity of the fluid node.  This can be summarized as:
 \begin{equation}
